### 2024-03-06 main.tex
1st commit, algorithms and no literature

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{algorithm}

\newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
\newcommand{\criticalpath}{\mathcal{P}}
\newcommand{\parents}[1]{\,\Pi_{#1}}
\newcommand{\children}[1]{\,C_{#1}}
\newcommand{\cluster}{\,\mathcal{S}}
\newcommand{\daghetpart}{\algo{DagHetPart}\xspace}
\newcommand{\dagmem}{\algo{DagHetMem}\xspace}

\title{Dynamic Scheduling }

\begin{document}
    \maketitle

    \section{Model}

    \begin{table}[h]
        \begin{center}
            \begin{tabular}{rl}
                \hline
                \textbf{Symbol} & \textbf{Meaning}  \\
                \hline
                $G = (V, E)$  & Workflow graph, set of vertices (tasks) and edges  \\
                $\parents{u}$, $\children{u}$ & Parents of a task $u$, children of a task $u$ \\
                $m_u$& Memory weight of task $u$ \\
                $w_u$   & Workload of  task $u$  (normalized execution time)    \\
                $c_{u,v}$   & Communication volume along the edge $(u,v)\in E$ \\
                $F$, $\mathcal{F}$ & A partitioning function and the partition it creates \\
                $V_i$ & Block number $i$\\ %\wrt~some $F$   \\
                $\cluster$, $k$   & Computing system and its number of processors   \\
                $p_j$, proc($V_i$)  & Processor number $j$, processor of block $V_i$ \\
                $M_j$, $s_j$   & Memory size and speed of processor $p_j$   \\
                $\beta$ & Bandwidth in the compute system  \\
                $\bottomlevel{u}$   & Bottom weight of task $u$   \\
                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$ \\
                $\Gamma = (\mathcal{V}, \mathcal{E})$ & Quotient graph, its vertices and its edges  \\
                $r_u$, $r_{V_i}$  & Memory requirement of  task $u$ and of block $V_i$   \\
                $r_{\max}$   & Maximum memory requirement in a workflow   \\
                $\criticalpath$ & Critical path in a workflow  \\
                \hline
            \end{tabular}
        \end{center}
        \caption{Notation.} \label{tabnotation}
    \end{table}

    \paragraph{Workflow-related changes}

    \begin{itemize}
        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.

    \end{itemize}

    The following changes are not a part of this article's scope:

    \begin{itemize}
        \item The workflow structure changes: edges or tasks come in or leave.
    \end{itemize}

    \paragraph{Execution environment-related changes }

    \begin{itemize}
        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.

    \end{itemize}

    The following changes are not a part of this article's scope:

    \begin{itemize}
        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
    \end{itemize}

    \subsection{Time of changes }

    We consider discrete time in seconds.
    The time point(s) at which the changes happen is unambiguously defined.

    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
    The start time of the source task in the workflow is zero.
    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$

    \subsection{Changes and knowledge horizon - important questions TBA}

    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.

    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
    However, how do we model the following:
    \begin{itemize}
        \item Do we know the new weights of currently running tasks and tasks that have not yet started? This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
        \item How many times of change do we model - one per workflow run, or multiple?
        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
    \end{itemize}

    \section{Related work}

    In the limited time I spent looking, I found no paper addressing the exact same problem.

    Wang et al.~\cite{wang2019dynamic} proposes a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
    Particles are possible solution in the solution space.
    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
    Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
    However, the approach is for clouds.

    Daniels et al.

    \section{Algorithms}

    \subsection{Various useful approaches}

    Bader et al. developed a system that identifies task similarities.
    Given a workflow, it returns tasks similar to a given task.
    We can use this system in our modelling and say that we know the new weights for finished tasks and will modify future tasks in accordance to their similarity.
    For example, if a task finished 2 times later than expected, then all future similar tasks will be assumed to finish 2 times later.

    \subsection{Local search}
    One strategy that can be used is local search, a similar approach as with swaps.
    The starting point could be a previously valid solution (for example, one by HetPart).
    In the first step, the solution would need to become valid again.
    Then, we repartition blocks on the critical path  and merge blocks in attempt to improve makespan.
    Finally, we swap blocks between processors to finish makespan improvement.

    \subsubsection{Restore Validity}
    We~(Algorithm~\ref{alg:RestoreValidity}) start with the previously partitioned and assigned quotient tree $\Gamma$,
    and the new weights on tasks and edges.
    We first recompute the memory requirements of all blocks.
    Then we identify all blocks that have become invalidated and exceed the memory of their assigned processor
    (Line~\ref{line:damaged}).
    For all these blocks, we try out ``border'' tasks, that is, tasks, that have neighbours in neighbouring blocks
    (Line~\ref{line:neighb}).
    We tentatively move them (creating a temporary quotient DAG $\Gamma'$) and assess the impact of this move on the
    makespan(Line~\ref{line:makespan}).
    The condition is, of course, that the neighboring block now does not exceed the memory of its own processor.
    The task that delivers the best makespan in then moved in Line~\ref{line:movebest}.
    The entire process is being repeated until the block fits into the memory of the processor again.
    We return the modified quotient graph $\Gamma$ that contains the mapping to processors.

    \begin{algorithm}[h!]
        \caption{Restore Validity}
        \label{alg:RestoreValidity}
        \begin{algorithmic}[1]
            \Procedure{Restore Validity}{$\Gamma$, $W$ }\\
            \Comment{Input: initial schedule as quotient tree $\Gamma$, new  weights $W$}
            \State $\Gamma \gets $ \textsc{QuotientDAG}($\Gamma$, $W$ ); \Comment Quotient DAG with new weights

            \State $D \gets \emptyset$ \Comment {Damaged blocks that require action}
            \For{$\nu_i \in \Gamma$}
                \State $r_{\nu_m} \gets W$
                \Comment{Recompute the new memory requirements of blocks in the quotient DAG}
            \EndFor
            \For{$\nu_i \in \Gamma$}
                \label{line:damaged}
                \If{$r_{\nu_i} > M_{proc(\nu_i)}$}
                    \State $D \gets D \cup \{\nu_i\}$
                \EndIf
            \EndFor

            \For{$\nu_i \in D$}
                \Comment Restore the validity of each damaged block
                \While{$r_{\nu_i} > M_{proc(\nu_i)}$}
                    \State $MS_{min}\gets \infty; v_{min} \gets NULL; \nu_{min} \gets NULL$
                    \Comment Moving $v_{min}$ to $\nu_{min}$ gives $MS_{min}$
                    \For{$v \in \nu_i$}
                        \For{$u \in ( \Pi_v \cup C_v) \cap (\Pi_{\nu_i} \cup C_{\nu_i})$}   \label{line:neighb}
                        \State $\Gamma' \gets \Gamma \mid \nu(v) == \nu(u)$
                        \If{ $r_{\nu(u)} > M_{proc(\nu(u))}$ and \textsc{Makespan}($\Gamma'$)$\leq MS_{min}$}\label{line:makespan}
                        \State $MS_{min} \gets $\textsc{Makespan}($\Gamma'$); $v_{min} \gets v$; $\nu_{min} \gets \nu(u);$
                        \EndIf
                        \EndFor
                    \EndFor
                    \State  $\Gamma \gets \Gamma \mid \nu(v_{min}) == \nu_{min}$ \label{line:movebest}
                \EndWhile
            \EndFor
            \State return $\Gamma$
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}

    \subsubsection{Repartition and Merge}

    In the next step, we center our attention around changes in the execution environments and how the changes in
    the workflow reflect on it.

    First, we try to further partition blocks that lie on the critical path of the workflow.
    We first compute the critical path (Line~\ref{line:cp}) and then, as long as there are free processors in the
    execution environment(Line~\ref{line:whilefree}, assumed to be sorted in descending order of memory sizes),
    we partition the blocks on the critical path(Line~\ref{line:repartition}).
    We try to partition each block into 2 smaller blocks.
    However, due to the nature of the partitioner, more can be created.
    We assume that the resulting blocks(Line~\ref{line:repartition}) are given back sorted by their memory requirement.
    We then assign them to free processors as long as the blocks fit and there are free processors.
    If even one part does not fit(Line~\ref{line:ifnotfits}), then we discard the entire partition and try partitioning
    the next block on the critical path.

    Not only partitioning (introducing more parallelism) can lead to a makespan improvement.
    Shorter runtimes of tasks can justify merging some blocks together.
    For all pairs of blocks that are neighbours (Line~\ref{Line:neighbblocks}), we tentatively merge them and assign
    the result to the processor of the first block.
    If the merged block fits on this processor and the merged quotient graph has a smaller overall makespan, then we save
    the merge and proceed with other blocks.

    \begin{algorithm}[h!]
        \caption{Repartition and Merge}
        \label{alg:RepartitionMerge}
        \begin{algorithmic}[1]
            \Procedure{Repartition and Merge}{$\Gamma$, $\cluster$ }\\
            \Comment{Input: quotient tree $\Gamma$, execution environment $\cluster$}
            \State $CP \gets $\textsc{Criticalpath}($\Gamma$) \label{line:cp}
            \State $Free \gets \{ p \in \cluster \mid p$ is free$\}$
            \While{$\mid Free \mid \neq 0$}\label{line:whilefree}
            \For{$\nu \in CP$}
                \State $\{ \nu_1, \nu_2,\dots\}\gets$\textsc{Partition}($\nu$,2) \label{line:repartition}
                \For{$\nu' \in \{ \nu_1, \nu_2,\dots\}$}
                    \If{$r_\nu' \leq Free.$head()}
                        \State proc($\nu'$)$\gets Free.$head();
                    \Else \label{line:ifnotfits}
                    \State \textsc{Merge}$\nu' \in \{ \nu_1, \nu_2,\dots\}$; break;
                    \Comment Merge the parts back and leave this partitioning, as it does not fit
                    \EndIf
                \EndFor
            \EndFor
            \EndWhile

            \For{$\nu_1, \nu_2 \in \Gamma \mid \nu_1 \in ( \Pi_{\nu_2} \cup C_{\nu_2})$} \label{line:neighbblocks}

            \State $\{\Gamma', \nu_m\} \gets $ \textsc{Merge}($\nu_1, \nu_2$); proc($\nu_m$)$\gets $ proc($\nu_1$);
            \State $\mu \gets $\textsc{Makespan}($\Gamma'$)
            \If{$r_{\nu_m} \leq M_{proc(\nu_m)}$ and $\mu \leq $ \textsc{Makespan}($\Gamma$)}
                \State $\Gamma \gets \Gamma'$
            \EndIf
            \EndFor

            \State return $\Gamma$
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}

    \subsubsection{Swaps}

    Then a process similar to Step 4 in \daghetpart can be executed: the newly rearranged blocks are swapped until no
    improvement in makespan can be achieved.

    \subsection{Limitations}

    We assume that moving a task from one processor to another costs nothing, when in reality it requires a copy operation that can take time at least equal to the sum of the weights of all its input files divided by the bandwidth.

    \section{Interface}

    To trigger a renewed computation of a schedule, the scheduler needs the following informations:
    \begin{itemize}
        \item The timestamp of the change (amount of seconds passed in the system)
        \item Per changed task:
        \begin{itemize}
            \item Task name (and id, if available)
            \item The new runtime
            \item The new memory requirement
        \end{itemize}

        Tasks, for which this information is not delivered are assumed unchanged.
        Previously given information will be used to determine their runtime and memory requirement.
        \item Per changed file transfer (edge in the workflow graph)

        \begin{itemize}
            \item The source and sink task names (to identify the edge)
            \item The new edge weight (size of files transferred)
        \end{itemize}

        \item  Per changed processor:
        \begin{itemize}
            \item Processor name and number (both required, because we have several processors with the same name)
            \item If exited or entered

        \end{itemize}

        Processors, for which no such information is delivered, are assumed to be unchanged.
        Processor exit time is assumed to be the time of the change. This means that all tasks that were scheduled there, have ran.
        However, currently running tasks need to be restarted on another processor and no future tasks cannot be scheduled on the exited processor.
        All task data is assumed to be retrievable from an exited processor without additional time requirements to do so.
    \end{itemize}

    Based on the timestamp, the renewed runtimes of tasks and renewed processor speeds, the scheduler can identify tasks that are finished, running or not started.
    The output of the scheduler is the new schedule: an assignment of each of the (not yet executed) tasks to a processor.

    Assumptions:

    \begin{itemize}
        \item No tasks can disappear, no new tasks can appear during an execution. Same holds for edges.
        \item The processor cannot change characteristics. It can only fail (exit) or appear (enter).

    \end{itemize}

    To be Discussed: The runtime of the scheduler itself is assumed to be zero, for the ease of simulating.

    \bibliographystyle{alpha}
    \bibliography{references}
\end{document}

### 2024-03-07 main.tex
ACM template for icpp. Minor issues with reference format, but otherwise functional

@@
-\documentclass{article}
-
-% Language setting
-% Replace `english' with e.g. `spanish' to change the document language
-\usepackage[english]{babel}
-
-% Set page size and margins
-% Replace `letterpaper' with `a4paper' for UK/EU standard size
-\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
-
-% Useful packages
+\documentclass[sigconf,review,anonymous]{acmart}
+\usepackage{algorithmicx}
 \usepackage{amsmath}
 \usepackage{graphicx}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{algpseudocode}
-\usepackage{algorithmicx}
 \usepackage{algorithm}
 
+%%
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+    \providecommand\BibTeX{{%
+        Bib\TeX}}}
+
+\setcopyright{acmlicensed}
+\copyrightyear{2024}
+\acmYear{2024}
+\acmDOI{XXXXXXX.XXXXXXX}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[ICPP '24]{53rd International Conference on Parallel Processing}{August 12-15, 2024}{Gotland, Sweden}
+\acmISBN{978-1-4503-XXXX-X/18/06}
+
+\newcommand{\algo}[1]{\textsc{#1}}
 \newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
@@
 \newcommand{\dagmem}{\algo{DagHetMem}\xspace}
+%%
+%% end of the preamble, start of the body of the document source.
+\begin{document}
 
-\title{Dynamic Scheduling }
+%%
+%% The "title" command has an optional parameter,
+%% allowing the author to define a "short title" to be used in page headers.
+    \title{Adaptive Scheduling of Scientific Workflows}
 
-\begin{document}
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+    \author{Svetlana Kulagina}
+    \email{svetlana.kulagina@hu-berlin.de}
+    \orcid{0000-0002-2108-9425}
+    \affiliation{%
+        \institution{Humboldt Universitaet zu Berlin}
+        \streetaddress{Unter den Linden 6}
+        \city{Berlin}
+        \country{Germany}
+        \postcode{10099}
+    }
+
+       \author{Henning Meyerhenke}
+       \affiliation{%
+           \institution{Humboldt Universitaet zu Berlin}
+           \city{Berlin}
+           \country{Germany}
+       }
+       \email{meyerhenke@hu-berlin.de}
+
+    \author{Anne Benoit}
+    \affiliation{%
+        \institution{ENS Lyon}
+        \city{Lyon}
+        \country{France}
+    }
+    \email{Anne.Benoit@ens-lyon.fr}
+
+%%
+%% By default, the full list of authors will be used in the page
+%% headers. Often, this list is too long, and will overlap
+%% other information printed in the page headers. This command allows
+%% the author to define a more concise list
+%% of authors' names for this purpose.
+    \renewcommand{\shortauthors}{Trovato et al.}
+
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
+    \begin{abstract}
+       Todo: reinsert CCSXML concepts
+    \end{abstract}
+
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+    \keywords{Scheduling, Adaptive Scheduling, DAG}
+    \received{15 April 2024}
+    \received[revised]{12 March 2009}
+    \received[accepted]{5 June 2009}
+
+%%
+%% This command processes the author and affiliation and title
+%% information and builds the first part of the formatted document.
     \maketitle
+
+    \section{Introduction}
 
@@
 
-    Then a process similar to Step 4 in \daghetpart can be executed: the newly rearranged blocks are swapped until no
+    Then a process similar to Step 4 in %\daghetpart
+    can be executed: the newly rearranged blocks are swapped until no
     improvement in makespan can be achieved.
@@
 
-    We assume that moving a task from one processor to another costs nothing, when in reality it requires a copy operation that can take time at least equal to the sum of the weights of all its input files divided by the bandwidth.
+    We assume that moving a task from one processor to another costs nothing, when in reality it requires a copy
+    operation that can take time at least equal to the sum of the weights of all its input files divided by the bandwidth.
 
-    \section{Interface}
+    \bibliographystyle{ACM-Reference-Format}
+    \bibliography{references}
 
-    To trigger a renewed computation of a schedule, the scheduler needs the following informations:
-    \begin{itemize}
-        \item The timestamp of the change (amount of seconds passed in the system)
-        \item Per changed task:
-        \begin{itemize}
-            \item Task name (and id, if available)
-            \item The new runtime
-            \item The new memory requirement
-        \end{itemize}
-
-        Tasks, for which this information is not delivered are assumed unchanged.
-        Previously given information will be used to determine their runtime and memory requirement.
-        \item Per changed file transfer (edge in the workflow graph)
-
-        \begin{itemize}
-            \item The source and sink task names (to identify the edge)
-            \item The new edge weight (size of files transferred)
-        \end{itemize}
-
-        \item  Per changed processor:
-        \begin{itemize}
-            \item Processor name and number (both required, because we have several processors with the same name)
-            \item If exited or entered
-
-        \end{itemize}
-
-        Processors, for which no such information is delivered, are assumed to be unchanged.
-        Processor exit time is assumed to be the time of the change. This means that all tasks that were scheduled there, have ran.
-        However, currently running tasks need to be restarted on another processor and no future tasks cannot be scheduled on the exited processor.
-        All task data is assumed to be retrievable from an exited processor without additional time requirements to do so.
-    \end{itemize}
-
-    Based on the timestamp, the renewed runtimes of tasks and renewed processor speeds, the scheduler can identify tasks that are finished, running or not started.
-    The output of the scheduler is the new schedule: an assignment of each of the (not yet executed) tasks to a processor.
-
-    Assumptions:
-
-    \begin{itemize}
-        \item No tasks can disappear, no new tasks can appear during an execution. Same holds for edges.
-        \item The processor cannot change characteristics. It can only fail (exit) or appear (enter).
-
-    \end{itemize}
-
-    To be Discussed: The runtime of the scheduler itself is assumed to be zero, for the ease of simulating.
-
-    \bibliographystyle{alpha}
-    \bibliography{references}
 \end{document}
+\endinput
+%%
+%% End of file `sample-sigconf.tex'.

### 2024-03-07 main.tex
First three papers described. WIP for Section 3 Related Work

@@
 \usepackage{amsmath}
-\usepackage{graphicx}
 %\usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 \usepackage{algorithm}
+\usepackage{xspace}
 
@@
 \acmISBN{978-1-4503-XXXX-X/18/06}
+
+\include{macros}
 
@@
     However, the approach is for clouds.
+    \xspace
 
     Daniels et al.
+
+    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling heuristic
+    for clouds.
+    Based on a 3-objective cost model, their greedy scheduling algorithm schedules tasks into machines.
+    The authors use provenance data to make scheduling decisions.
+    The cost model is a set of linear equations computed in the simulation environment, it represents the cost of an execution based on the criteria.
+    In the algorithms, the authors test out 4 scenarios - one preferring each criteria, and a balanced one.
+    The algorithm merely chooses the best virtual machine for each next task based on the cost given by the model.
+    An additional algorithm combines several cloud activities (task executions) into one to improve the cost, so that each
+    execution is not too small (utilize the granularity factor, the smallest entity of payment of the cloud producer).
+    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
+    They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
+
+    Rahman~\etal~\cite{rahman2013} propose a scheduling heuristic for grids that proposes mapping of tasks to machines by calculating
+    the critical path in the graph dynamically at every step.
+    They call it the dynamic critical path (DCP).
+    For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
+    on the start time of a task (differing by the slack this task has).
+    All tasks on this critical path have the same earliest and latest start times, because they cannot be ddelayed.
+
+    The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
+    If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum execution time for
+    the task is chosen.
+    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
+    The authors evaluate their approaches on random workflows of the size up to 300 tasks.
+
+    The authors provide an overview over (simpler) scheduling heuristics.
+    For example, GRASP (generally randomized adaptive search procedure) conducts a number of iterations to search an optimal
+    solution for mapping tasks on machines.
+    A solution is generated at each step, and the best solution is kept at the end.
+    The search terminates when a certain termination criterion is reached.
+    It generates better results than other algorithms, because it explores the whole solution space.
+
+    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
+    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
+    The scenario contains AND- and OR-branches, where AND-branches indicate acitivities that need to be executed in parallel.
+
+    The heuristic first partitions the set into local schedules by using affinity matrices to determine similar
+    activities and group them together.
+    Then they physically allocate these partitions to groups of disaster responders and tasks within this group.
+    They define a constraint system for that.
+    The dynamic part deals with changes and distinguishes between retriable and compensation acitivities.
+    The heuristic calculates a new execution path with these tasks.
+    In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
 
@@
 
-            \For{$\nu_1, \nu_2 \in \Gamma \mid \nu_1 \in ( \Pi_{\nu_2} \cup C_{\nu_2})$} \label{line:neighbblocks}
-
+            \For{$\nu_1, \nu_2 \in \Gamma \mid \nu_1 \in ( \Pi_{\nu_2} \cup C_{\nu_2})$}
+                \label{line:neighbblocks}
             \State $\{\Gamma', \nu_m\} \gets $ \textsc{Merge}($\nu_1, \nu_2$); proc($\nu_m$)$\gets $ proc($\nu_1$);

### 2024-03-10 main.tex
More related work

@@
 
-    Daniels et al.
+    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
+    on a single machine.
+    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
+    that these runtime changes are unrelated to each other).
+    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
+    performance measure $\phi$.
+    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
 
@@
     In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
+
+    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
+    The aim is to minimize the makespan, and the experiments were conducted on a single wotkflow with 10 tasks.
+    The procesdure involves building a first (static) schedule, priodic resource monitoring and rescheduling the remaining
+    tasks.
+    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
+    For each resuorce group, there is an own scheduler, and an overall global scheduler responsible for distributing
+    tasks to groups.
+    The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
+    The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
+
+    The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
+    THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
+    They then go through tasks and map than to such processors that minimize the earliest start times of this task's
+    successors.
+    To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
+    average communication and computation costs fir the dependent tasks.
+
+    The rescheduling is being triggered when either a load on a resource increases over a threshold, or if a new resource
+    is added.
+    The algorithm produces a new mapping from scrath, and this mapping is being accepted if the resulting maespan is
+    smaller than the previously predicted one.
 

### 2024-03-11 main.tex
description of baseline (HEFT with memories)

@@
 
-    \subsection{Various useful approaches}
+    \subsection{Baseline: HEFT with memories}
 
-    Bader et al. developed a system that identifies task similarities.
-    Given a workflow, it returns tasks similar to a given task.
-    We can use this system in our modelling and say that we know the new weights for finished tasks and will modify future tasks in accordance to their similarity.
-    For example, if a task finished 2 times later than expected, then all future similar tasks will be assumed to finish 2 times later.
+    \algo{HEFT} is one of the most popular scheduling heuristics(TODO cite).
+    This algorithm schedules tasks on heterogeneous computing systems with the aim to minimize
+    the workflow's overall completion time.
+    The classical formulation of \algo{HEFT} does not respect memory boundaries, so we suggest a
+    slightly changed version that also takes memory sizes into account.
 
-    \subsection{Local search}
+    \algo{HEFT} first assigns all tasks their ranks on all processors.
+    This rank includes the bottom level of the task, including the maximum cost of its successors on the same processor.
+    Then, for each task, they are being assigned to the processor with the best rank.
+    Finally, the algorithm calculates the order of execution of tasks on each processor, depending on their dependencies.
+
+    We only change the assignment to processor part.
+    Before assigning a new task there, we calculate the memory requirement of already assigned tasks.
+    If the remaining memory is big enough to host the task, we assign it there, otherwise we try the next-best processor
+    rank-wise.
+
+    \skug{Even easier: calculate the memory requirement of already assigned tasks by summing them up. This will lead
+    to an overestimation of memory requirement.}
+
+    \subsection{New heuristic: Local search}
     One strategy that can be used is local search, a similar approach as with swaps.

### 2024-03-14 main.tex
new HEFT-based heuristic

@@
 \newcommand{\dagmem}{\algo{DagHetMem}\xspace}
+
+\newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
+\newcommand{\hmey}[1]{{\color{red}[HM: #1]}}
+\newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
 %%
@@
     Then, for each task, they are being assigned to the processor with the best rank.
+    The algorithm looks to minimize the earliest finish time of the task when it chooses the processor with the best rank.
     Finally, the algorithm calculates the order of execution of tasks on each processor, depending on their dependencies.
@@
     to an overestimation of memory requirement.}
+
+    \subsection{New heuristic: HEFT with minimal memory traversal}
+
+    This heuristic changes both the rank calculation step, and the assignment step.
+
+    Before calculating ranks, we calculate the optimal memory traversal over the entire graph.
+    We then calculate ranks as bottom levels on each processor, but adding some small $\epsilon$ on tasks that have the
+    same bottom level, but are higher in the memory traversal order.
+    So, for a traversal $v_1 \rightarrow v_2 \rightarrow v_3$, if both $v_2$ and $v_3$ have the bottom level of $100$,
+    $v_2$ will get the rank of $100+\epsilon$.
+
+    Then we sort the tasks ascending according to their rank.
+    We then look to minimize not the earliest finishing time of the task itself, but the earliest starting time of
+    its \skug{heaviest} successor.
+    The term heaviest is currently not defined exactly.
+    The intuition is to try to optimize for that successor that hold the schedule up the most.
+
+    We then do a procedure similar to our FillUpFit intuition.
+    We traverse the graph along the optimal memory traversal.
+    We keep track of the processor that we last used.
+    With each new task $v$, we check if it has the best rank on the same processor.
+    If so, and if the processor's memory is big enough, we fit the task there.
+
+    If the memory is too full or if $v$ has a better rank elsewhere, we calculate earliest starting times
+    of all its successors on all other processors (we assume that not only $v$ itself goes on that processor, but also
+    all its successors).
+    We then choose the processor that gives the best earliest starting time for the  \skug{heaviest} successor
+    and assign $v$ there.
 

### 2024-03-21 main.tex
temp new heuristic

@@
 \newcommand{\dagmem}{\algo{DagHetMem}\xspace}
+
+\newcommand{\MM}{M}
+\newcommand{\MC}{MC}
+\newcommand{\rt}{rt}
+\newcommand{\curM}{curM}
+\newcommand{\curC}{curC}
+\newcommand{\PD}{PD}
 
@@
 
+\section{Proposal of a new heuristic with slightly refined model}
+
+The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
+but favor processor reuse as is done in HEFT.
+
+\subsection{Model}
+
+For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
+For an  edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
+
+For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds: 
+$\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
+We can decide to evict some data from the main memory if we are sending the data
+to another processor; it then stays in the communication buffer until it has been sent.
+
+We keep track of the current ready time of each processor and each communication
+channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
+are set to~$0$.
+
+We also keep track of the current memory usage, $\curM_j$ and $\curC_j$,
+on respectively the processor memory and communication buffer. 
+Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
+that are in $\MM_j$ but may be evicted to be communicated, if 
+more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
+They correspond to some $c_{u,v}$'s.
+
+\subsection{Heuristics}
+
+We consider two variants of HEFT accounting for memory usage, which only
+differ in the order they consider tasks to be scheduled.
+
+HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
+level is defined as
+$$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
+ (the max is 0 if there is no outgoing edge).
+
+HEFTM-MM orders tasks as dictated by MinMem.
+\AB{Actually, while writing,
+I remember that this might be complicated to compute MinMem for the
+entire graph, so this might not be a feasible variant, or needs to be updated?}
+
+Then, the idea is to pick the next free task in the given order,
+and greedily assign it to a processor, by trying all possible options
+and keeping the most promising one.
+
+\medskip
+\noindent{\em Tentative assignment of task~$v$ on $p_j$.}\\
+{\bf Step 1.} First, we need to check that for all predecessors~$u$ of~$v$ that are mapped
+on~$p_j$, the data $c_{u,v}$ is still in the memory of~$p_j$, 
+i.e., $c_{u,v}\in PD_j$. Otherwise, the finish time is set to~$+\infty$ (invalid choice).
+
+\smallskip
+\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
+$$Res = \curM_j  + m_v + \sum_{u \in Pred(v), u\notin T(p_j)}  \{c_{u,v}\}
++ \sum_{w\in Succ(v)}  \{c_{v,w}\} - \MM_j .$$
+
+$T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
+already in the memory of~$p_j$. However, the files from the 
+other predecessor tasks must be loaded in memory before executing task~$v$,
+as well as $m_v$ and the data generated for all successor tasks.
+$Res$ is then checking whether there was enough memory; if it is positive, 
+it means that we have exceeded the memory of~$p_j$ with this tentative 
+assignment.
+
+In this case ($Res >0$), we  try evicting
+some data from memory so that we have enough memory to execute task~$v$. 
+We need to evict at least $Res$ data. For now, we propose a greedy approach,
+evicting the largest files of $\PD_j$ until $Res$ data has been evicted. 
+\AB{I can further detail if needed, moving data of $\PD_j$ from $\curM_j$
+to $\curC_j$, and removing it from $\PD_j$...}
+If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
+we set the finish time to~$+\infty$ (invalid choice).
+
+\smallskip
+\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$. Its starting time $st_v$
+is dictated by the maximum between $rt_j$, and all communications that
+must be orchestrated from predecessor tasks $u\notin T(p_j)$. Its
+finish time on $p_j$ is then 
+$FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
+
+\medskip
+\noindent{\em Assignment of task~$v$.}\\
+Once we have computed all finish times for task~$v$, 
+we keep the minimum $FT(v,p_j)$ and assign task~$v$
+to processor~$p_j$, updating $\MM_j$ and $\MC_j$ as needed,
+as well as all the $rt$ values for $p_j$ and the communication channels.
+
+\section{Conclusion}
+
     \bibliographystyle{ACM-Reference-Format}

### 2024-03-22 main.tex
temp

@@
 but favor processor reuse as is done in HEFT. 
+Furthermore, this would allow us to handle variability on the fly, by updating
+the bottom levels if some parameters vary, and computing the schedule
+only for the near future...
 
@@
 
+ HEFTM-BLC: from the study of the fork (see below), it seems important
+ to also account for the size of the data as input of a task,
+ to give more priority at tasks with potential large incoming communications.
+ For each task, we compute its modified bottom level: 
+ $$blc(u) = bl(u) + \max_{(v,u)\in E} c_{v,u}   . $$
+
 HEFTM-MM orders tasks as dictated by MinMem.
@@
 We need to evict at least $Res$ data. For now, we propose a greedy approach,
-evicting the largest files of $\PD_j$ until $Res$ data has been evicted. 
+evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
+in order to avoid costly communications. 
+\AB{FYI We initially discussed  evicting the largest files, but this leads to 
+large communications and does not seem efficient after all... Maybe we can think of another 
+approach that would take into account both data size and bottom level...}
 \AB{I can further detail if needed, moving data of $\PD_j$ from $\curM_j$
@@
 
+\subsection{The fork}
+We look at the behavior of these heuristics on a fork graph,
+where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
+to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
+
+Without memory, this problem is NP-complete; this is equivalent
+to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
+and with two processors. Half of the tasks must be sent to the processor
+on which $T_0$ is not executed, and the optimal makespan is 
+$w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
+
+However, with an infinite number of identical processors, it can be
+solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$; 
+the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
+while the remaining $n-k$ tasks are executed locally (try all values of $k$).
+
+With heterogeneous processors, it is probably NP-complete again
+because we could ensure that there are only two processors fast enough
+and get back to the 2-partition...
+
+We also had an example where evicting large files first in step 2
+can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
+$f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
+imposes that we free one unit of memory before executing one
+of the tasks... Actually the new version with BLC would start 
+considering $T_2$ and be fine in this case...
+
+\AB{Can we prove that we have (maybe) a 2-approximation, 
+at least for the fork? What worst-case can we think of? }
+
 \section{Conclusion}

### 2024-03-22 main.tex
notes on small examples

--- 2024-03-21 main.tex
@@
 Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
-that are in $\MM_j$ but may be evicted to be communicated, if 
+that are in the memory of size $\MM_j$ but may be evicted to be communicated, if 
 more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
@@
 
+\subsection{Approximation}
+\hmey{Rough notes:}
+Let's use a fork to see how the algorithm behaves and if it provides some approximation.
+Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
+
 \section{Conclusion}

### 2024-03-22 main.tex

--- 2024-03-22 main.tex
@@
 
+\subsection{Approximation}
+\hmey{Rough notes:}
+Let's use a fork to see how the algorithm behaves and if it provides some approximation.
+Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
+
 \subsection{The fork}

### 2024-04-14 main.tex
changed order of subsections
described the implementation in detail (WIP!), taking a break

@@
 \subsection{Heuristics}
+    Like the original HEFT, our heuristic consistst of two steps: first, computing task ranks,
+    and second, assigning tasks to processors in the order defined in the first step.
+    We consider three variants of HEFT accounting for memory usage, which only
+    differ in the order they consider tasks to be scheduled.
 
-We consider two variants of HEFT accounting for memory usage, which only
-differ in the order they consider tasks to be scheduled.
+\subsubsection{Step 1: calculate task ranks}
 
@@
 HEFTM-MM orders tasks as dictated by MinMem.
-\AB{Actually, while writing,
-I remember that this might be complicated to compute MinMem for the
-entire graph, so this might not be a feasible variant, or needs to be updated?}
+
+\subsubsection{Task assignment}
 
@@
 \noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
-$$Res = \curM_j  + m_v + \sum_{u \in Pred(v), u\notin T(p_j)}  \{c_{u,v}\}
-+ \sum_{w\in Succ(v)}  \{c_{v,w}\} - \MM_j .$$
+\[Res = \curM_j  - m_v + \sum_{u \in Pred(v), u\notin T(p_j)}  \{c_{u,v}\}
+- \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]
 
@@
 as well as $m_v$ and the data generated for all successor tasks.
-$Res$ is then checking whether there was enough memory; if it is positive, 
+$Res$ is then checking whether there was enough memory; if it is negative,
 it means that we have exceeded the memory of~$p_j$ with this tentative 
@@
 
-In this case ($Res >0$), we  try evicting
+In this case ($Res <0$), we  try evicting
 some data from memory so that we have enough memory to execute task~$v$. 
-We need to evict at least $Res$ data. For now, we propose a greedy approach,
-evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
+We need to evict at least $Res$ data.
+For now, we propose a greedy approach, evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
 in order to avoid costly communications. 
@@
 approach that would take into account both data size and bottom level...}
-\AB{I can further detail if needed, moving data of $\PD_j$ from $\curM_j$
-to $\curC_j$, and removing it from $\PD_j$...}
+While tentatively evicting files, we remove them from the list of pending memories and move them into a list
+    of memories pending in the communication buffer.
+We keep track of the available buffer size, too.
+
 If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
+    or if while doing so we exceeded the size of the available buffer,
 we set the finish time to~$+\infty$ (invalid choice).
@@
 $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
+    \skug {TODO PUT THE FORMULA HERE}
 
@@
 we keep the minimum $FT(v,p_j)$ and assign task~$v$
-to processor~$p_j$, updating $\MM_j$ and $\MC_j$ as needed,
+to processor~$p_j$.
+In detail, we:
+
+    \textbullet
+    Evict the memories that need to be evicted to free the memory. We remove these memories from pending memories
+        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
+
+    \textbullet For every predecessor of~$v$ that has been assigned to another processor, we
+
+    \textbullet We compute the correct amount of available memory for $p_j$ (for when the task is done).
+For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
+    the incoming edge, also freeing the same amount of available memory (increasing $CurM_j$).
+For each successor, on the other hand, we add the edge weights to pending memories and reduce $CurM_j$ by the corresponding
+    amount.
+
+    \textbullet
+We adjust ready times on communication buffers $rt_{j, j'}$ for every predecessor processor $u$ of $v$: we increase them by the
+    communication time $c(v, u) / \beta$.
+
+    updating $\MM_j$ and $\MC_j$ as needed,
 as well as all the $rt$ values for $p_j$ and the communication channels.
-
-\subsection{Approximation}
-\hmey{Rough notes:}
-Let's use a fork to see how the algorithm behaves and if it provides some approximation.
-Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
 

### 2024-04-15 main.tex
further corrected and improved the description of the implementation

@@
     of memories pending in the communication buffer.
-We keep track of the available buffer size, too.
+We keep track of the available buffer size, too - each time a file gets moved into the pending in buffer, the available buffer size is reduced by its weight.
 
@@
 \smallskip
-\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$. Its starting time $st_v$
+\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
+    Its starting time $st_v$
 is dictated by the maximum between $rt_j$, and all communications that
-must be orchestrated from predecessor tasks $u\notin T(p_j)$. Its
-finish time on $p_j$ is then 
+must be orchestrated from predecessor tasks $u\notin T(p_j)$.
+    The starting time is then
+    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
+    Its finish time on $p_j$ is then
 $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
-    \skug {TODO PUT THE FORMULA HERE}
 
@@
 
-    \textbullet For every predecessor of~$v$ that has been assigned to another processor, we
+    \textbullet For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
+    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
+    communication time $c( u,v) / \beta$.
 
@@
     amount.
-
-    \textbullet
-We adjust ready times on communication buffers $rt_{j, j'}$ for every predecessor processor $u$ of $v$: we increase them by the
-    communication time $c(v, u) / \beta$.
 

### 2024-04-16 main.tex
described vanilla HEFT without mems, corrected the description of heuristics

@@
 
-    \section{Algorithms}
+\section{Proposal of a new heuristic with slightly refined model}
 
-    \subsection{Baseline: HEFT with memories}
+The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
+but favor processor reuse as is done in HEFT. 
+Furthermore, this would allow us to handle variability on the fly, by updating
+the bottom levels if some parameters vary, and computing the schedule
+only for the near future...
 
-    \algo{HEFT} is one of the most popular scheduling heuristics(TODO cite).
-    This algorithm schedules tasks on heterogeneous computing systems with the aim to minimize
-    the workflow's overall completion time.
-    The classical formulation of \algo{HEFT} does not respect memory boundaries, so we suggest a
-    slightly changed version that also takes memory sizes into account.
+\subsection{Model}
 
-    \algo{HEFT} first assigns all tasks their ranks on all processors.
-    This rank includes the bottom level of the task, including the maximum cost of its successors on the same processor.
-    Then, for each task, they are being assigned to the processor with the best rank.
-    The algorithm looks to minimize the earliest finish time of the task when it chooses the processor with the best rank.
-    Finally, the algorithm calculates the order of execution of tasks on each processor, depending on their dependencies.
+For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
+For an  edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
 
-    We only change the assignment to processor part.
-    Before assigning a new task there, we calculate the memory requirement of already assigned tasks.
-    If the remaining memory is big enough to host the task, we assign it there, otherwise we try the next-best processor
-    rank-wise.
+For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds: 
+$\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
+We can decide to evict some data from the main memory if we are sending the data
+to another processor; it then stays in the communication buffer until it has been sent.
 
-    \skug{Even easier: calculate the memory requirement of already assigned tasks by summing them up. This will lead
-    to an overestimation of memory requirement.}
+We keep track of the current ready time of each processor and each communication
+channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
+are set to~$0$.
 
-    \subsection{New heuristic: HEFT with minimal memory traversal}
+We also keep track of the currently available memory, $availM_j$ and $availC_j$,
+on respectively the processor memory and communication buffer. 
+Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
+that are in the memory of size $\MM_j$ but may be evicted to be communicated, if 
+more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
+They correspond to some $c_{u,v}$'s.
 
-    This heuristic changes both the rank calculation step, and the assignment step.
+\subsection{Baseline: original HEFT without memories}
 
-    Before calculating ranks, we calculate the optimal memory traversal over the entire graph.
-    We then calculate ranks as bottom levels on each processor, but adding some small $\epsilon$ on tasks that have the
-    same bottom level, but are higher in the memory traversal order.
-    So, for a traversal $v_1 \rightarrow v_2 \rightarrow v_3$, if both $v_2$ and $v_3$ have the bottom level of $100$,
-    $v_2$ will get the rank of $100+\epsilon$.
+    Original HEFT does not consider memory sizes.
+    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
+    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.
 
-    Then we sort the tasks ascending according to their rank.
-    We then look to minimize not the earliest finishing time of the task itself, but the earliest starting time of
-    its \skug{heaviest} successor.
-    The term heaviest is currently not defined exactly.
-    The intuition is to try to optimize for that successor that hold the schedule up the most.
+    HEFT works in two steps.
+    In the first step, it calculates the ranks of tasks by computing their non-increasing bottom levels.
+    The bottom level of a task is defined as
+    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
+    (the max is 0 if there is no outgoing edge).
+    The tasks are sorted by non-decreasing ranks.
 
-    We then do a procedure similar to our FillUpFit intuition.
-    We traverse the graph along the optimal memory traversal.
-    We keep track of the processor that we last used.
-    With each new task $v$, we check if it has the best rank on the same processor.
-    If so, and if the processor's memory is big enough, we fit the task there.
+    In the second step, the algorithm iterates over the ranks and tries to assign the task to the processor where it
+    has the earliest finish time.
+    We tentatively assign each task to each processor.
+    The task's starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
+    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
+    The starting time is then
+    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
+    Its finish time on $p_j$ is then
+    $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
 
-    If the memory is too full or if $v$ has a better rank elsewhere, we calculate earliest starting times
-    of all its successors on all other processors (we assume that not only $v$ itself goes on that processor, but also
-    all its successors).
-    We then choose the processor that gives the best earliest starting time for the  \skug{heaviest} successor
-    and assign $v$ there.
+    Once we have computed all finish times for task~$v$,
+    we keep the minimum $FT(v,p_j)$ and assign task~$v$
+    to processor~$p_j$.
+
+    \textit{Assignment to processor}
+    When assigning the task, we set the ready time of the processor~$j$ $rt_j$ to the finish time of the task.
+    For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
+    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
+    communication time $c( u,v) / \beta$.
+
+\subsection{Heuristics}
+    Like the original HEFT, our heuristic consistst of two steps: first, computing task ranks,
+    and second, assigning tasks to processors in the order defined in the first step.
+    We consider three variants of HEFT accounting for memory usage, which only
+    differ in the order they consider tasks to be scheduled.
+
+\subsubsection{Step 1: calculate task ranks}
+
+HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
+level is defined as
+$$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
+ (the max is 0 if there is no outgoing edge).
+
+ HEFTM-BLC: from the study of the fork (see below), it seems important
+ to also account for the size of the data as input of a task,
+ to give more priority at tasks with potential large incoming communications.
+ For each task, we compute its modified bottom level: 
+ $$blc(u) = bl(u) + \max_{(v,u)\in E} c_{v,u}   . $$
+
+HEFTM-MM orders tasks as dictated by MinMem.
+
+\subsubsection{Task assignment}
+
+Then, the idea is to pick the next free task in the given order,
+and greedily assign it to a processor, by trying all possible options
+and keeping the most promising one.
+
+\medskip
+\noindent{\em Tentative assignment of task~$v$ on $p_j$.}\\
+{\bf Step 1.} First, we need to check that for all predecessors~$u$ of~$v$ that are mapped
+on~$p_j$, the data $c_{u,v}$ is still in the memory of~$p_j$, 
+i.e., $c_{u,v}\in PD_j$. Otherwise, the finish time is set to~$+\infty$ (invalid choice).
+
+\smallskip
+\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
+\[Res = availM_j  - m_v - \sum_{u \in \Pi(v), u\notin T(p_j)}  \{c_{u,v}\}
+- \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]
+
+$T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
+already in the memory of~$p_j$. However, the files from the 
+other predecessor tasks must be loaded in memory before executing task~$v$,
+as well as $m_v$ and the data generated for all successor tasks.
+$Res$ is then checking whether there was enough memory; if it is negative,
+it means that we have exceeded the memory of~$p_j$ with this tentative 
+assignment.
+
+In this case ($Res <0$), we  try evicting
+some data from memory so that we have enough memory to execute task~$v$. 
+We need to evict at least $Res$ data.
+For now, we propose a greedy approach, evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
+in order to avoid costly communications. 
+\AB{FYI We initially discussed  evicting the largest files, but this leads to 
+large communications and does not seem efficient after all... Maybe we can think of another 
+approach that would take into account both data size and bottom level...}
+While tentatively evicting files, we remove them from the list of pending memories and move them into a list
+    of memories pending in the communication buffer.
+We keep track of the available buffer size, too - each time a file gets moved into the pending in buffer, the available buffer size is reduced by its weight.
+
+If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
+    or if while doing so we exceeded the size of the available buffer,
+we set the finish time to~$+\infty$ (invalid choice).
+
+\smallskip
+\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
+    Its starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
+must be orchestrated from predecessor tasks $u\notin T(p_j)$.
+    The starting time is then
+    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} } \]
+    Its finish time on $p_j$ is then
+$FT(v,p_j) = ST(v, p_j) + \frac{w_v}{s_j}$.
+
+\medskip
+\noindent{\em Assignment of task~$v$.}\\
+Once we have computed all finish times for task~$v$, 
+we keep the minimum $FT(v,p_j)$ and assign task~$v$
+to processor~$p_j$.
+In detail, we:
+    \begin{itemize}
+        \item  Evict the file memories that correspond to edge weights that need to be evicted to free the memory.
+        We remove these files from pending memories
+        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
+        \item    Calculate the new $availM_j$ on the processor.
+        We subtract the weights of all incoming files from predecessors assigned to the same processor,
+        and add the weights of outgoing files generated by the currently assigned task.
+        \item  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
+        communication buffers $rt_{j', j}$ for the processor~$j'$that the predecessor $u$ has been assigned to: we increase them by the
+        communication time $c( u,v) / \beta$.
+        We also remove the incoming files from either the pending memories or pending data in buffers of these other
+        processors, and increase the available memories or available buffer sizes on these processors.
+        \item We compute the correct amount of available memory for $p_j$ (for when the task is done).
+        For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
+        the incoming edge, also freeing the same amount of available memory (increasing $availM_j$).
+        For each successor, on the other hand, we add the edge weights to pending memories and reduce $availM_j$ by the corresponding
+        amount.
+    \end{itemize}
+
+\subsection{The fork}
+We look at the behavior of these heuristics on a fork graph,
+where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
+to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
+
+Without memory, this problem is NP-complete; this is equivalent
+to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
+and with two processors. Half of the tasks must be sent to the processor
+on which $T_0$ is not executed, and the optimal makespan is 
+$w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
+
+However, with an infinite number of identical processors, it can be
+solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$; 
+the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
+while the remaining $n-k$ tasks are executed locally (try all values of $k$).
+
+With heterogeneous processors, it is probably NP-complete again
+because we could ensure that there are only two processors fast enough
+and get back to the 2-partition...
+
+We also had an example where evicting large files first in step 2
+can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
+$f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
+imposes that we free one unit of memory before executing one
+of the tasks... Actually the new version with BLC would start 
+considering $T_2$ and be fine in this case...
+
+\AB{Can we prove that we have (maybe) a 2-approximation, 
+at least for the fork? What worst-case can we think of? }
+
+\subsection{Approximation}
+\hmey{Rough notes:}
+Let's use a fork to see how the algorithm behaves and if it provides some approximation.
+Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
+
+\section{Conclusion}
+
+\section{Appendix}
 
@@
 
-    \subsection{Limitations}
+    \subsection{New heuristic: HEFT with minimal memory traversal}
 
-    We assume that moving a task from one processor to another costs nothing, when in reality it requires a copy
-    operation that can take time at least equal to the sum of the weights of all its input files divided by the bandwidth.
+    This heuristic changes both the rank calculation step, and the assignment step.
 
-\section{Proposal of a new heuristic with slightly refined model}
+    Before calculating ranks, we calculate the optimal memory traversal over the entire graph.
+    We then calculate ranks as bottom levels on each processor, but adding some small $\epsilon$ on tasks that have the
+    same bottom level, but are higher in the memory traversal order.
+    So, for a traversal $v_1 \rightarrow v_2 \rightarrow v_3$, if both $v_2$ and $v_3$ have the bottom level of $100$,
+    $v_2$ will get the rank of $100+\epsilon$.
 
-The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
-but favor processor reuse as is done in HEFT. 
-Furthermore, this would allow us to handle variability on the fly, by updating
-the bottom levels if some parameters vary, and computing the schedule
-only for the near future...
+    Then we sort the tasks ascending according to their rank.
+    We then look to minimize not the earliest finishing time of the task itself, but the earliest starting time of
+    its \skug{heaviest} successor.
+    The term heaviest is currently not defined exactly.
+    The intuition is to try to optimize for that successor that hold the schedule up the most.
 
-\subsection{Model}
+    We then do a procedure similar to our FillUpFit intuition.
+    We traverse the graph along the optimal memory traversal.
+    We keep track of the processor that we last used.
+    With each new task $v$, we check if it has the best rank on the same processor.
+    If so, and if the processor's memory is big enough, we fit the task there.
 
-For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
-For an  edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
-
-For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds: 
-$\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
-We can decide to evict some data from the main memory if we are sending the data
-to another processor; it then stays in the communication buffer until it has been sent.
-
-We keep track of the current ready time of each processor and each communication
-channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
-are set to~$0$.
-
-We also keep track of the current memory usage, $\curM_j$ and $\curC_j$,
-on respectively the processor memory and communication buffer. 
-Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
-that are in the memory of size $\MM_j$ but may be evicted to be communicated, if 
-more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
-They correspond to some $c_{u,v}$'s.
-
-\subsection{Heuristics}
-    Like the original HEFT, our heuristic consistst of two steps: first, computing task ranks,
-    and second, assigning tasks to processors in the order defined in the first step.
-    We consider three variants of HEFT accounting for memory usage, which only
-    differ in the order they consider tasks to be scheduled.
-
-\subsubsection{Step 1: calculate task ranks}
-
-HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
-level is defined as
-$$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
- (the max is 0 if there is no outgoing edge).
-
- HEFTM-BLC: from the study of the fork (see below), it seems important
- to also account for the size of the data as input of a task,
- to give more priority at tasks with potential large incoming communications.
- For each task, we compute its modified bottom level: 
- $$blc(u) = bl(u) + \max_{(v,u)\in E} c_{v,u}   . $$
-
-HEFTM-MM orders tasks as dictated by MinMem.
-
-\subsubsection{Task assignment}
-
-Then, the idea is to pick the next free task in the given order,
-and greedily assign it to a processor, by trying all possible options
-and keeping the most promising one.
-
-\medskip
-\noindent{\em Tentative assignment of task~$v$ on $p_j$.}\\
-{\bf Step 1.} First, we need to check that for all predecessors~$u$ of~$v$ that are mapped
-on~$p_j$, the data $c_{u,v}$ is still in the memory of~$p_j$, 
-i.e., $c_{u,v}\in PD_j$. Otherwise, the finish time is set to~$+\infty$ (invalid choice).
-
-\smallskip
-\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
-\[Res = \curM_j  - m_v + \sum_{u \in Pred(v), u\notin T(p_j)}  \{c_{u,v}\}
-- \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]
-
-$T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
-already in the memory of~$p_j$. However, the files from the 
-other predecessor tasks must be loaded in memory before executing task~$v$,
-as well as $m_v$ and the data generated for all successor tasks.
-$Res$ is then checking whether there was enough memory; if it is negative,
-it means that we have exceeded the memory of~$p_j$ with this tentative 
-assignment.
-
-In this case ($Res <0$), we  try evicting
-some data from memory so that we have enough memory to execute task~$v$. 
-We need to evict at least $Res$ data.
-For now, we propose a greedy approach, evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
-in order to avoid costly communications. 
-\AB{FYI We initially discussed  evicting the largest files, but this leads to 
-large communications and does not seem efficient after all... Maybe we can think of another 
-approach that would take into account both data size and bottom level...}
-While tentatively evicting files, we remove them from the list of pending memories and move them into a list
-    of memories pending in the communication buffer.
-We keep track of the available buffer size, too - each time a file gets moved into the pending in buffer, the available buffer size is reduced by its weight.
-
-If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
-    or if while doing so we exceeded the size of the available buffer,
-we set the finish time to~$+\infty$ (invalid choice).
-
-\smallskip
-\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
-    Its starting time $st_v$
-is dictated by the maximum between $rt_j$, and all communications that
-must be orchestrated from predecessor tasks $u\notin T(p_j)$.
-    The starting time is then
-    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
-    Its finish time on $p_j$ is then
-$FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
-
-\medskip
-\noindent{\em Assignment of task~$v$.}\\
-Once we have computed all finish times for task~$v$, 
-we keep the minimum $FT(v,p_j)$ and assign task~$v$
-to processor~$p_j$.
-In detail, we:
-
-    \textbullet
-    Evict the memories that need to be evicted to free the memory. We remove these memories from pending memories
-        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
-
-    \textbullet For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
-    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
-    communication time $c( u,v) / \beta$.
-
-    \textbullet We compute the correct amount of available memory for $p_j$ (for when the task is done).
-For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
-    the incoming edge, also freeing the same amount of available memory (increasing $CurM_j$).
-For each successor, on the other hand, we add the edge weights to pending memories and reduce $CurM_j$ by the corresponding
-    amount.
-
-    updating $\MM_j$ and $\MC_j$ as needed,
-as well as all the $rt$ values for $p_j$ and the communication channels.
-
-\subsection{The fork}
-We look at the behavior of these heuristics on a fork graph,
-where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
-to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
-
-Without memory, this problem is NP-complete; this is equivalent
-to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
-and with two processors. Half of the tasks must be sent to the processor
-on which $T_0$ is not executed, and the optimal makespan is 
-$w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
-
-However, with an infinite number of identical processors, it can be
-solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$; 
-the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
-while the remaining $n-k$ tasks are executed locally (try all values of $k$).
-
-With heterogeneous processors, it is probably NP-complete again
-because we could ensure that there are only two processors fast enough
-and get back to the 2-partition...
-
-We also had an example where evicting large files first in step 2
-can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
-$f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
-imposes that we free one unit of memory before executing one
-of the tasks... Actually the new version with BLC would start 
-considering $T_2$ and be fine in this case...
-
-\AB{Can we prove that we have (maybe) a 2-approximation, 
-at least for the fork? What worst-case can we think of? }
-
-\subsection{Approximation}
-\hmey{Rough notes:}
-Let's use a fork to see how the algorithm behaves and if it provides some approximation.
-Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
-
-\section{Conclusion}
+    If the memory is too full or if $v$ has a better rank elsewhere, we calculate earliest starting times
+    of all its successors on all other processors (we assume that not only $v$ itself goes on that processor, but also
+    all its successors).
+    We then choose the processor that gives the best earliest starting time for the  \skug{heaviest} successor
+    and assign $v$ there.
 

### 2024-04-22 main.tex
corrected the blc definition

@@
  For each task, we compute its modified bottom level: 
- $$blc(u) = bl(u) + \max_{(v,u)\in E} c_{v,u}   . $$
+ $$blc(u) = w_u + \max_{(u,w)\in E} \{c_{u,w} + blc(w)\} + \max_{(v,u)\in E} c_{v,u}   . $$
+
+    \skug{avoid having mixed ranks, when the memory size of the lower task is not taken into account}
 

### 2024-04-24 main.tex
described retracing of a schedule

@@
 
+\subsection{Retracing the effects of change on an existing schedule}
+    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
+    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
+    they can lead to a later finishing time (\eg if some tasks longer and delay other tasks), or they can have no effect (\eg if new processors
+    joined the cluster, but the old schedule did not account for them).
+    To assess the impact, we need to retrace the schedule.
+
+    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
+    entire schedule.
+
+    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
+    is a topologial ordering.
+    We then repeat steps similar to those we did during tentative assignment in our heuristic, except we do not choose a processor
+    anymore, but rather see if the current one still fits.
+
+    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from our heuristic.
+    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ or $c_{v,w}$ from successors $w$,
+    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements from other tasks).
+    If originally,$Res$ was positive (no files were evicted from memory into the communication buffer), then it has to stay this way -
+    otherwise evicted files can invalidate next tasks.
+    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
+    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
+    We update the $availM_j$ and $availMC_j$ according to the new memory constraints.
+
+    Then we can re-calculate the finish time of the task on its processor like in Step 3.
+    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
+    (after delayed previous tasks), and changed communication buffer availability.
+
+    Then, after having updated the processor's values, we move on to the next task.
+
 \subsection{The fork}

### 2024-04-29 main.tex
first draft draft for the dynamic scenario

@@
                 $\parents{u}$, $\children{u}$         & Parents of a task $u$, children of a task $u$            \\
-                $m_u$& Memory weight of task $u$ \\
+                $m_u$                                 & Memory weight of task $u$                                \\
                 $w_u$                                 & Workload of task $u$  (normalized execution time)          \\
@@
                 $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
-                $V_i$ & Block number $i$\\ %\wrt~some $F$   \\
+                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
                 $\cluster$, $k$                    & Computing system and its number of processors           \\
@@
 
+    \subsection{Dynamic Scenario}
+
+    In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
+    This information may include, memory usage, runtime, graph structures, or the status of the underlying infrastructure.
+    In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
+    By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
+    Also the underlying infrastructure can change during the workflow execution.
+    Examples are processor failures, node recoveries, or acquisition of new nodes.
+    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
+    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.
+
+    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{}.
+    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
+    %For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
+    %Bader~et~al.~ report a prediction error between 13\% and 17\% for their method, showing an exponential task error distribution.
+    % @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
+    Such a dynamic execution environment necessitates for a dynamic scheduling method where the schedule can be recomputed during the workflow execution.
+
     \subsection{Retracing the effects of change on an existing schedule}
@@
         \begin{algorithmic}[1]
-            \Procedure{Restore Validity}{$\Gamma$, $W$ }\\
+            \Procedure{Restore Validity}{$\Gamma$, $W$ }
+                \\
                 \Comment{Input: initial schedule as quotient tree $\Gamma$, new weights $W$}
@@
                         \For{$v \in \nu_i$}
-                        \For{$u \in ( \Pi_v \cup C_v) \cap (\Pi_{\nu_i} \cup C_{\nu_i})$}   \label{line:neighb}
+                            \For{$u \in ( \Pi_v \cup C_v) \cap (\Pi_{\nu_i} \cup C_{\nu_i})$}
+                                \label{line:neighb}
                                 \State $\Gamma' \gets \Gamma \mid \nu(v) == \nu(u)$
-                        \If{ $r_{\nu(u)} > M_{proc(\nu(u))}$ and \textsc{Makespan}($\Gamma'$)$\leq MS_{min}$}\label{line:makespan}
+                                \If{ $r_{\nu(u)} > M_{proc(\nu(u))}$ and \textsc{Makespan}($\Gamma'$)$\leq MS_{min}$}
+                                    \label{line:makespan}
                                     \State $MS_{min} \gets $\textsc{Makespan}($\Gamma'$); $v_{min} \gets v$; $\nu_{min} \gets \nu(u);$
@@
         \begin{algorithmic}[1]
-            \Procedure{Repartition and Merge}{$\Gamma$, $\cluster$ }\\
+            \Procedure{Repartition and Merge}{$\Gamma$, $\cluster$ }
+                \\
                 \Comment{Input: quotient tree $\Gamma$, execution environment $\cluster$}
@@
                 \State $Free \gets \{ p \in \cluster \mid p$ is free$\}$
-            \While{$\mid Free \mid \neq 0$}\label{line:whilefree}
+                \While{$\mid Free \mid \neq 0$}
+                    \label{line:whilefree}
                     \For{$\nu \in CP$}
@@
                                 \State proc($\nu'$)$\gets Free.$head();
-                    \Else \label{line:ifnotfits}
+                            \Else
+                                \label{line:ifnotfits}
                                 \State \textsc{Merge}$\nu' \in \{ \nu_1, \nu_2,\dots\}$; break;

### 2024-05-02 main.tex
added the correct template

@@
-\documentclass[sigconf,review,anonymous]{acmart}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+%\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
 \usepackage{algorithmicx}
-\usepackage{amsmath}
 %\usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-    \providecommand\BibTeX{{%
-        Bib\TeX}}}
-
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[ICPP '24]{53rd International Conference on Parallel Processing}{August 12-15, 2024}{Gotland, Sweden}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \include{macros}
@@
 \newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
-%%
-%% end of the preamble, start of the body of the document source.
+
 \begin{document}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-    \title{Adaptive Scheduling of Scientific Workflows}
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-    \author{Svetlana Kulagina}
-    \email{svetlana.kulagina@hu-berlin.de}
-    \orcid{0000-0002-2108-9425}
-    \affiliation{%
-        \institution{Humboldt Universitaet zu Berlin}
-        \streetaddress{Unter den Linden 6}
-        \city{Berlin}
-        \country{Germany}
-        \postcode{10099}
+    \title{Adaptive Scheduling of Scientific Workflows\\
+%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
+%should not be used}
+    \thanks{Identify applicable funding agency here. If none, delete this.}
     }
 
-    \author{Henning Meyerhenke}
-    \affiliation{%
-        \institution{Humboldt Universitaet zu Berlin}
-        \city{Berlin}
-        \country{Germany}
-    }
-    \email{meyerhenke@hu-berlin.de}
+%\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%\textit{name of organization (of Aff.)}\\
+%City, Country \\
+%email address or ORCID}
+%\and
+%\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%\textit{name of organization (of Aff.)}\\
+%City, Country \\
+%email address or ORCID} }
 
-    \author{Anne Benoit}
-    \affiliation{%
-        \institution{ENS Lyon}
-        \city{Lyon}
-        \country{France}
-    }
-    \email{Anne.Benoit@ens-lyon.fr}
+    \maketitle
 
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-    \renewcommand{\shortauthors}{Trovato et al.}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
     \begin{abstract}
-        Todo: reinsert CCSXML concepts
+        TODO: Insert abstract
     \end{abstract}
 
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-    \keywords{Scheduling, Adaptive Scheduling, DAG}
-    \received{15 April 2024}
-    \received[revised]{12 March 2009}
-    \received[accepted]{5 June 2009}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-    \maketitle
+    \begin{IEEEkeywords}
+        Scheduling, Adaptive Scheduling, DAG
+    \end{IEEEkeywords}
 
@@
     imposes that we free one unit of memory before executing one
-    of the tasks... Actually the new version with BLC would start
-    considering $T_2$ and be fine in this case...
+    of the tasks\ldots Actually the new version with BLC would start
+    considering $T_2$ and be fine in this case\ldots
 
@@
 
-    \bibliographystyle{ACM-Reference-Format}
+    \bibliographystyle{IEEEtran}
     \bibliography{references}
@@
 \end{document}
-\endinput
-%%
-%% End of file `sample-sigconf.tex'.

### 2024-05-02 dynamic-scenario.tex
removed unrelated text

--- 2024-04-24 main.tex
@@
-\documentclass[sigconf,review,anonymous]{acmart}
-\usepackage{algorithmicx}
+%! Author = kulagins
+%! Date = 02.05.24
+
+% Preamble
+\documentclass[11pt]{article}
+
+% Packages
 \usepackage{amsmath}
-%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-\usepackage{xspace}
 
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-    \providecommand\BibTeX{{%
-        Bib\TeX}}}
-
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[ICPP '24]{53rd International Conference on Parallel Processing}{August 12-15, 2024}{Gotland, Sweden}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-\include{macros}
-
-\newcommand{\algo}[1]{\textsc{#1}}
-\newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
-\newcommand{\criticalpath}{\mathcal{P}}
-\newcommand{\parents}[1]{\,\Pi_{#1}}
-\newcommand{\children}[1]{\,C_{#1}}
-\newcommand{\cluster}{\,\mathcal{S}}
-\newcommand{\daghetpart}{\algo{DagHetPart}\xspace}
-\newcommand{\dagmem}{\algo{DagHetMem}\xspace}
-
-\newcommand{\MM}{M}
-\newcommand{\MC}{MC}
-\newcommand{\rt}{rt}
-\newcommand{\curM}{curM}
-\newcommand{\curC}{curC}
-\newcommand{\PD}{PD}
-
-\newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
-\newcommand{\hmey}[1]{{\color{red}[HM: #1]}}
-\newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
-%%
-%% end of the preamble, start of the body of the document source.
+% Document
 \begin{document}
-
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-    \title{Adaptive Scheduling of Scientific Workflows}
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-    \author{Svetlana Kulagina}
-    \email{svetlana.kulagina@hu-berlin.de}
-    \orcid{0000-0002-2108-9425}
-    \affiliation{%
-        \institution{Humboldt Universitaet zu Berlin}
-        \streetaddress{Unter den Linden 6}
-        \city{Berlin}
-        \country{Germany}
-        \postcode{10099}
-    }
-
-       \author{Henning Meyerhenke}
-       \affiliation{%
-           \institution{Humboldt Universitaet zu Berlin}
-           \city{Berlin}
-           \country{Germany}
-       }
-       \email{meyerhenke@hu-berlin.de}
-
-    \author{Anne Benoit}
-    \affiliation{%
-        \institution{ENS Lyon}
-        \city{Lyon}
-        \country{France}
-    }
-    \email{Anne.Benoit@ens-lyon.fr}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-    \renewcommand{\shortauthors}{Trovato et al.}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
-    \begin{abstract}
-       Todo: reinsert CCSXML concepts
-    \end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-    \keywords{Scheduling, Adaptive Scheduling, DAG}
-    \received{15 April 2024}
-    \received[revised]{12 March 2009}
-    \received[accepted]{5 June 2009}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-    \maketitle
-
-    \section{Introduction}
-
-    \section{Model}
-
-    \begin{table}[h]
-        \begin{center}
-            \begin{tabular}{rl}
-                \hline
-                \textbf{Symbol} & \textbf{Meaning}  \\
-                \hline
-                $G = (V, E)$  & Workflow graph, set of vertices (tasks) and edges  \\
-                $\parents{u}$, $\children{u}$ & Parents of a task $u$, children of a task $u$ \\
-                $m_u$& Memory weight of task $u$ \\
-                $w_u$   & Workload of  task $u$  (normalized execution time)    \\
-                $c_{u,v}$   & Communication volume along the edge $(u,v)\in E$ \\
-                $F$, $\mathcal{F}$ & A partitioning function and the partition it creates \\
-                $V_i$ & Block number $i$\\ %\wrt~some $F$   \\
-                $\cluster$, $k$   & Computing system and its number of processors   \\
-                $p_j$, proc($V_i$)  & Processor number $j$, processor of block $V_i$ \\
-                $M_j$, $s_j$   & Memory size and speed of processor $p_j$   \\
-                $\beta$ & Bandwidth in the compute system  \\
-                $\bottomlevel{u}$   & Bottom weight of task $u$   \\
-                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$ \\
-                $\Gamma = (\mathcal{V}, \mathcal{E})$ & Quotient graph, its vertices and its edges  \\
-                $r_u$, $r_{V_i}$  & Memory requirement of  task $u$ and of block $V_i$   \\
-                $r_{\max}$   & Maximum memory requirement in a workflow   \\
-                $\criticalpath$ & Critical path in a workflow  \\
-                \hline
-            \end{tabular}
-        \end{center}
-        \caption{Notation.} \label{tabnotation}
-    \end{table}
 
@@
 
-    \section{Related work}
-
-    In the limited time I spent looking, I found no paper addressing the exact same problem.
-
-    Wang et al.~\cite{wang2019dynamic} proposes a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
-    Particles are possible solution in the solution space.
-    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
-    Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
-    However, the approach is for clouds.
-    \xspace
-
-    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
-    on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
-    that these runtime changes are unrelated to each other).
-    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
-    performance measure $\phi$.
-    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
-
-    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling heuristic
-    for clouds.
-    Based on a 3-objective cost model, their greedy scheduling algorithm schedules tasks into machines.
-    The authors use provenance data to make scheduling decisions.
-    The cost model is a set of linear equations computed in the simulation environment, it represents the cost of an execution based on the criteria.
-    In the algorithms, the authors test out 4 scenarios - one preferring each criteria, and a balanced one.
-    The algorithm merely chooses the best virtual machine for each next task based on the cost given by the model.
-    An additional algorithm combines several cloud activities (task executions) into one to improve the cost, so that each
-    execution is not too small (utilize the granularity factor, the smallest entity of payment of the cloud producer).
-    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
-    They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
-
-    Rahman~\etal~\cite{rahman2013} propose a scheduling heuristic for grids that proposes mapping of tasks to machines by calculating
-    the critical path in the graph dynamically at every step.
-    They call it the dynamic critical path (DCP).
-    For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
-    on the start time of a task (differing by the slack this task has).
-    All tasks on this critical path have the same earliest and latest start times, because they cannot be ddelayed.
-
-    The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
-    If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum execution time for
-    the task is chosen.
-    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
-    The authors evaluate their approaches on random workflows of the size up to 300 tasks.
-
-    The authors provide an overview over (simpler) scheduling heuristics.
-    For example, GRASP (generally randomized adaptive search procedure) conducts a number of iterations to search an optimal
-    solution for mapping tasks on machines.
-    A solution is generated at each step, and the best solution is kept at the end.
-    The search terminates when a certain termination criterion is reached.
-    It generates better results than other algorithms, because it explores the whole solution space.
-
-    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
-    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
-    The scenario contains AND- and OR-branches, where AND-branches indicate acitivities that need to be executed in parallel.
-
-    The heuristic first partitions the set into local schedules by using affinity matrices to determine similar
-    activities and group them together.
-    Then they physically allocate these partitions to groups of disaster responders and tasks within this group.
-    They define a constraint system for that.
-    The dynamic part deals with changes and distinguishes between retriable and compensation acitivities.
-    The heuristic calculates a new execution path with these tasks.
-    In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
-
-    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
-    The aim is to minimize the makespan, and the experiments were conducted on a single wotkflow with 10 tasks.
-    The procesdure involves building a first (static) schedule, priodic resource monitoring and rescheduling the remaining
-    tasks.
-    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
-    For each resuorce group, there is an own scheduler, and an overall global scheduler responsible for distributing
-    tasks to groups.
-    The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
-    The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
-
-    The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
-    THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
-    They then go through tasks and map than to such processors that minimize the earliest start times of this task's
-    successors.
-    To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
-    average communication and computation costs fir the dependent tasks.
-
-    The rescheduling is being triggered when either a load on a resource increases over a threshold, or if a new resource
-    is added.
-    The algorithm produces a new mapping from scrath, and this mapping is being accepted if the resulting maespan is
-    smaller than the previously predicted one.
-
-\section{Proposal of a new heuristic with slightly refined model}
-
-The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
-but favor processor reuse as is done in HEFT. 
-Furthermore, this would allow us to handle variability on the fly, by updating
-the bottom levels if some parameters vary, and computing the schedule
-only for the near future...
-
-\subsection{Model}
-
-For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
-For an  edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
-
-For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds: 
-$\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
-We can decide to evict some data from the main memory if we are sending the data
-to another processor; it then stays in the communication buffer until it has been sent.
-
-We keep track of the current ready time of each processor and each communication
-channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
-are set to~$0$.
-
-We also keep track of the currently available memory, $availM_j$ and $availC_j$,
-on respectively the processor memory and communication buffer. 
-Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
-that are in the memory of size $\MM_j$ but may be evicted to be communicated, if 
-more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
-They correspond to some $c_{u,v}$'s.
-
-\subsection{Baseline: original HEFT without memories}
-
-    Original HEFT does not consider memory sizes.
-    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
-    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.
-
-    HEFT works in two steps.
-    In the first step, it calculates the ranks of tasks by computing their non-increasing bottom levels.
-    The bottom level of a task is defined as
-    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
-    (the max is 0 if there is no outgoing edge).
-    The tasks are sorted by non-decreasing ranks.
-
-    In the second step, the algorithm iterates over the ranks and tries to assign the task to the processor where it
-    has the earliest finish time.
-    We tentatively assign each task to each processor.
-    The task's starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
-    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
-    The starting time is then
-    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
-    Its finish time on $p_j$ is then
-    $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
-
-    Once we have computed all finish times for task~$v$,
-    we keep the minimum $FT(v,p_j)$ and assign task~$v$
-    to processor~$p_j$.
-
-    \textit{Assignment to processor}
-    When assigning the task, we set the ready time of the processor~$j$ $rt_j$ to the finish time of the task.
-    For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
-    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
-    communication time $c( u,v) / \beta$.
-
-\subsection{Heuristics}
-    Like the original HEFT, our heuristic consistst of two steps: first, computing task ranks,
-    and second, assigning tasks to processors in the order defined in the first step.
-    We consider three variants of HEFT accounting for memory usage, which only
-    differ in the order they consider tasks to be scheduled.
-
-\subsubsection{Step 1: calculate task ranks}
-
-HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
-level is defined as
-$$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
- (the max is 0 if there is no outgoing edge).
-
- HEFTM-BLC: from the study of the fork (see below), it seems important
- to also account for the size of the data as input of a task,
- to give more priority at tasks with potential large incoming communications.
- For each task, we compute its modified bottom level: 
- $$blc(u) = w_u + \max_{(u,w)\in E} \{c_{u,w} + blc(w)\} + \max_{(v,u)\in E} c_{v,u}   . $$
-
-    \skug{avoid having mixed ranks, when the memory size of the lower task is not taken into account}
-
-HEFTM-MM orders tasks as dictated by MinMem.
-
-\subsubsection{Task assignment}
-
-Then, the idea is to pick the next free task in the given order,
-and greedily assign it to a processor, by trying all possible options
-and keeping the most promising one.
-
-\medskip
-\noindent{\em Tentative assignment of task~$v$ on $p_j$.}\\
-{\bf Step 1.} First, we need to check that for all predecessors~$u$ of~$v$ that are mapped
-on~$p_j$, the data $c_{u,v}$ is still in the memory of~$p_j$, 
-i.e., $c_{u,v}\in PD_j$. Otherwise, the finish time is set to~$+\infty$ (invalid choice).
-
-\smallskip
-\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
-\[Res = availM_j  - m_v - \sum_{u \in \Pi(v), u\notin T(p_j)}  \{c_{u,v}\}
-- \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]
-
-$T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
-already in the memory of~$p_j$. However, the files from the 
-other predecessor tasks must be loaded in memory before executing task~$v$,
-as well as $m_v$ and the data generated for all successor tasks.
-$Res$ is then checking whether there was enough memory; if it is negative,
-it means that we have exceeded the memory of~$p_j$ with this tentative 
-assignment.
-
-In this case ($Res <0$), we  try evicting
-some data from memory so that we have enough memory to execute task~$v$. 
-We need to evict at least $Res$ data.
-For now, we propose a greedy approach, evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
-in order to avoid costly communications. 
-\AB{FYI We initially discussed  evicting the largest files, but this leads to 
-large communications and does not seem efficient after all... Maybe we can think of another 
-approach that would take into account both data size and bottom level...}
-While tentatively evicting files, we remove them from the list of pending memories and move them into a list
-    of memories pending in the communication buffer.
-We keep track of the available buffer size, too - each time a file gets moved into the pending in buffer, the available buffer size is reduced by its weight.
-
-If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
-    or if while doing so we exceeded the size of the available buffer,
-we set the finish time to~$+\infty$ (invalid choice).
-
-\smallskip
-\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
-    Its starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
-must be orchestrated from predecessor tasks $u\notin T(p_j)$.
-    The starting time is then
-    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} } \]
-    Its finish time on $p_j$ is then
-$FT(v,p_j) = ST(v, p_j) + \frac{w_v}{s_j}$.
-
-\medskip
-\noindent{\em Assignment of task~$v$.}\\
-Once we have computed all finish times for task~$v$, 
-we keep the minimum $FT(v,p_j)$ and assign task~$v$
-to processor~$p_j$.
-In detail, we:
-    \begin{itemize}
-        \item  Evict the file memories that correspond to edge weights that need to be evicted to free the memory.
-        We remove these files from pending memories
-        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
-        \item    Calculate the new $availM_j$ on the processor.
-        We subtract the weights of all incoming files from predecessors assigned to the same processor,
-        and add the weights of outgoing files generated by the currently assigned task.
-        \item  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
-        communication buffers $rt_{j', j}$ for the processor~$j'$that the predecessor $u$ has been assigned to: we increase them by the
-        communication time $c( u,v) / \beta$.
-        We also remove the incoming files from either the pending memories or pending data in buffers of these other
-        processors, and increase the available memories or available buffer sizes on these processors.
-        \item We compute the correct amount of available memory for $p_j$ (for when the task is done).
-        For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
-        the incoming edge, also freeing the same amount of available memory (increasing $availM_j$).
-        For each successor, on the other hand, we add the edge weights to pending memories and reduce $availM_j$ by the corresponding
-        amount.
-    \end{itemize}
-
-\subsection{Retracing the effects of change on an existing schedule}
-    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
-    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
-    they can lead to a later finishing time (\eg if some tasks longer and delay other tasks), or they can have no effect (\eg if new processors
-    joined the cluster, but the old schedule did not account for them).
-    To assess the impact, we need to retrace the schedule.
-
-    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
-    entire schedule.
-
-    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
-    is a topologial ordering.
-    We then repeat steps similar to those we did during tentative assignment in our heuristic, except we do not choose a processor
-    anymore, but rather see if the current one still fits.
-
-    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from our heuristic.
-    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ or $c_{v,w}$ from successors $w$,
-    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements from other tasks).
-    If originally,$Res$ was positive (no files were evicted from memory into the communication buffer), then it has to stay this way -
-    otherwise evicted files can invalidate next tasks.
-    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
-    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
-    We update the $availM_j$ and $availMC_j$ according to the new memory constraints.
-
-    Then we can re-calculate the finish time of the task on its processor like in Step 3.
-    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
-    (after delayed previous tasks), and changed communication buffer availability.
-
-    Then, after having updated the processor's values, we move on to the next task.
-
-\subsection{The fork}
-We look at the behavior of these heuristics on a fork graph,
-where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
-to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
-
-Without memory, this problem is NP-complete; this is equivalent
-to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
-and with two processors. Half of the tasks must be sent to the processor
-on which $T_0$ is not executed, and the optimal makespan is 
-$w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
-
-However, with an infinite number of identical processors, it can be
-solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$; 
-the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
-while the remaining $n-k$ tasks are executed locally (try all values of $k$).
-
-With heterogeneous processors, it is probably NP-complete again
-because we could ensure that there are only two processors fast enough
-and get back to the 2-partition...
-
-We also had an example where evicting large files first in step 2
-can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
-$f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
-imposes that we free one unit of memory before executing one
-of the tasks... Actually the new version with BLC would start 
-considering $T_2$ and be fine in this case...
-
-\AB{Can we prove that we have (maybe) a 2-approximation, 
-at least for the fork? What worst-case can we think of? }
-
-\subsection{Approximation}
-\hmey{Rough notes:}
-Let's use a fork to see how the algorithm behaves and if it provides some approximation.
-Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
-
-\section{Conclusion}
-
-\section{Appendix}
-
-    \subsection{New heuristic: Local search}
-    One strategy that can be used is local search, a similar approach as with swaps.
-    The starting point could be a previously valid solution (for example, one by HetPart).
-    In the first step, the solution would need to become valid again.
-    Then, we repartition blocks on the critical path  and merge blocks in attempt to improve makespan.
-    Finally, we swap blocks between processors to finish makespan improvement.
-
-    \subsubsection{Restore Validity}
-    We~(Algorithm~\ref{alg:RestoreValidity}) start with the previously partitioned and assigned quotient tree $\Gamma$,
-    and the new weights on tasks and edges.
-    We first recompute the memory requirements of all blocks.
-    Then we identify all blocks that have become invalidated and exceed the memory of their assigned processor
-    (Line~\ref{line:damaged}).
-    For all these blocks, we try out ``border'' tasks, that is, tasks, that have neighbours in neighbouring blocks
-    (Line~\ref{line:neighb}).
-    We tentatively move them (creating a temporary quotient DAG $\Gamma'$) and assess the impact of this move on the
-    makespan(Line~\ref{line:makespan}).
-    The condition is, of course, that the neighboring block now does not exceed the memory of its own processor.
-    The task that delivers the best makespan in then moved in Line~\ref{line:movebest}.
-    The entire process is being repeated until the block fits into the memory of the processor again.
-    We return the modified quotient graph $\Gamma$ that contains the mapping to processors.
-
-    \begin{algorithm}[h!]
-        \caption{Restore Validity}
-        \label{alg:RestoreValidity}
-        \begin{algorithmic}[1]
-            \Procedure{Restore Validity}{$\Gamma$, $W$ }\\
-            \Comment{Input: initial schedule as quotient tree $\Gamma$, new  weights $W$}
-            \State $\Gamma \gets $ \textsc{QuotientDAG}($\Gamma$, $W$ ); \Comment Quotient DAG with new weights
-
-            \State $D \gets \emptyset$ \Comment {Damaged blocks that require action}
-            \For{$\nu_i \in \Gamma$}
-                \State $r_{\nu_m} \gets W$
-                \Comment{Recompute the new memory requirements of blocks in the quotient DAG}
-            \EndFor
-            \For{$\nu_i \in \Gamma$}
-                \label{line:damaged}
-                \If{$r_{\nu_i} > M_{proc(\nu_i)}$}
-                    \State $D \gets D \cup \{\nu_i\}$
-                \EndIf
-            \EndFor
-
-            \For{$\nu_i \in D$}
-                \Comment Restore the validity of each damaged block
-                \While{$r_{\nu_i} > M_{proc(\nu_i)}$}
-                    \State $MS_{min}\gets \infty; v_{min} \gets NULL; \nu_{min} \gets NULL$
-                    \Comment Moving $v_{min}$ to $\nu_{min}$ gives $MS_{min}$
-                    \For{$v \in \nu_i$}
-                        \For{$u \in ( \Pi_v \cup C_v) \cap (\Pi_{\nu_i} \cup C_{\nu_i})$}   \label{line:neighb}
-                        \State $\Gamma' \gets \Gamma \mid \nu(v) == \nu(u)$
-                        \If{ $r_{\nu(u)} > M_{proc(\nu(u))}$ and \textsc{Makespan}($\Gamma'$)$\leq MS_{min}$}\label{line:makespan}
-                        \State $MS_{min} \gets $\textsc{Makespan}($\Gamma'$); $v_{min} \gets v$; $\nu_{min} \gets \nu(u);$
-                        \EndIf
-                        \EndFor
-                    \EndFor
-                    \State  $\Gamma \gets \Gamma \mid \nu(v_{min}) == \nu_{min}$ \label{line:movebest}
-                \EndWhile
-            \EndFor
-            \State return $\Gamma$
-            \EndProcedure
-        \end{algorithmic}
-    \end{algorithm}
-
-    \subsubsection{Repartition and Merge}
-
-    In the next step, we center our attention around changes in the execution environments and how the changes in
-    the workflow reflect on it.
-
-    First, we try to further partition blocks that lie on the critical path of the workflow.
-    We first compute the critical path (Line~\ref{line:cp}) and then, as long as there are free processors in the
-    execution environment(Line~\ref{line:whilefree}, assumed to be sorted in descending order of memory sizes),
-    we partition the blocks on the critical path(Line~\ref{line:repartition}).
-    We try to partition each block into 2 smaller blocks.
-    However, due to the nature of the partitioner, more can be created.
-    We assume that the resulting blocks(Line~\ref{line:repartition}) are given back sorted by their memory requirement.
-    We then assign them to free processors as long as the blocks fit and there are free processors.
-    If even one part does not fit(Line~\ref{line:ifnotfits}), then we discard the entire partition and try partitioning
-    the next block on the critical path.
-
-    Not only partitioning (introducing more parallelism) can lead to a makespan improvement.
-    Shorter runtimes of tasks can justify merging some blocks together.
-    For all pairs of blocks that are neighbours (Line~\ref{Line:neighbblocks}), we tentatively merge them and assign
-    the result to the processor of the first block.
-    If the merged block fits on this processor and the merged quotient graph has a smaller overall makespan, then we save
-    the merge and proceed with other blocks.
-
-    \begin{algorithm}[h!]
-        \caption{Repartition and Merge}
-        \label{alg:RepartitionMerge}
-        \begin{algorithmic}[1]
-            \Procedure{Repartition and Merge}{$\Gamma$, $\cluster$ }\\
-            \Comment{Input: quotient tree $\Gamma$, execution environment $\cluster$}
-            \State $CP \gets $\textsc{Criticalpath}($\Gamma$) \label{line:cp}
-            \State $Free \gets \{ p \in \cluster \mid p$ is free$\}$
-            \While{$\mid Free \mid \neq 0$}\label{line:whilefree}
-            \For{$\nu \in CP$}
-                \State $\{ \nu_1, \nu_2,\dots\}\gets$\textsc{Partition}($\nu$,2) \label{line:repartition}
-                \For{$\nu' \in \{ \nu_1, \nu_2,\dots\}$}
-                    \If{$r_\nu' \leq Free.$head()}
-                        \State proc($\nu'$)$\gets Free.$head();
-                    \Else \label{line:ifnotfits}
-                    \State \textsc{Merge}$\nu' \in \{ \nu_1, \nu_2,\dots\}$; break;
-                    \Comment Merge the parts back and leave this partitioning, as it does not fit
-                    \EndIf
-                \EndFor
-            \EndFor
-            \EndWhile
-
-            \For{$\nu_1, \nu_2 \in \Gamma \mid \nu_1 \in ( \Pi_{\nu_2} \cup C_{\nu_2})$}
-                \label{line:neighbblocks}
-                \State $\{\Gamma', \nu_m\} \gets $ \textsc{Merge}($\nu_1, \nu_2$); proc($\nu_m$)$\gets $ proc($\nu_1$);
-                \State $\mu \gets $\textsc{Makespan}($\Gamma'$)
-                \If{$r_{\nu_m} \leq M_{proc(\nu_m)}$ and $\mu \leq $ \textsc{Makespan}($\Gamma$)}
-                    \State $\Gamma \gets \Gamma'$
-                \EndIf
-            \EndFor
-
-            \State return $\Gamma$
-            \EndProcedure
-        \end{algorithmic}
-    \end{algorithm}
-
-    \subsubsection{Swaps}
-
-    Then a process similar to Step 4 in %\daghetpart
-    can be executed: the newly rearranged blocks are swapped until no
-    improvement in makespan can be achieved.
-
-    \subsection{New heuristic: HEFT with minimal memory traversal}
-
-    This heuristic changes both the rank calculation step, and the assignment step.
-
-    Before calculating ranks, we calculate the optimal memory traversal over the entire graph.
-    We then calculate ranks as bottom levels on each processor, but adding some small $\epsilon$ on tasks that have the
-    same bottom level, but are higher in the memory traversal order.
-    So, for a traversal $v_1 \rightarrow v_2 \rightarrow v_3$, if both $v_2$ and $v_3$ have the bottom level of $100$,
-    $v_2$ will get the rank of $100+\epsilon$.
-
-    Then we sort the tasks ascending according to their rank.
-    We then look to minimize not the earliest finishing time of the task itself, but the earliest starting time of
-    its \skug{heaviest} successor.
-    The term heaviest is currently not defined exactly.
-    The intuition is to try to optimize for that successor that hold the schedule up the most.
-
-    We then do a procedure similar to our FillUpFit intuition.
-    We traverse the graph along the optimal memory traversal.
-    We keep track of the processor that we last used.
-    With each new task $v$, we check if it has the best rank on the same processor.
-    If so, and if the processor's memory is big enough, we fit the task there.
-
-    If the memory is too full or if $v$ has a better rank elsewhere, we calculate earliest starting times
-    of all its successors on all other processors (we assume that not only $v$ itself goes on that processor, but also
-    all its successors).
-    We then choose the processor that gives the best earliest starting time for the  \skug{heaviest} successor
-    and assign $v$ there.
-
-    \bibliographystyle{ACM-Reference-Format}
-    \bibliography{references}
-
 \end{document}
-\endinput
-%%
-%% End of file `sample-sigconf.tex'.

### 2024-05-02 local-search-herustic.tex
removed unrelated text

--- 2024-04-24 main.tex
@@
-\documentclass[sigconf,review,anonymous]{acmart}
-\usepackage{algorithmicx}
+%! Author = kulagins
+%! Date = 02.05.24
+
+% Preamble
+\documentclass[11pt]{article}
+
+% Packages
 \usepackage{amsmath}
-%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-\usepackage{xspace}
 
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-    \providecommand\BibTeX{{%
-        Bib\TeX}}}
-
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[ICPP '24]{53rd International Conference on Parallel Processing}{August 12-15, 2024}{Gotland, Sweden}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-\include{macros}
-
-\newcommand{\algo}[1]{\textsc{#1}}
-\newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
-\newcommand{\criticalpath}{\mathcal{P}}
-\newcommand{\parents}[1]{\,\Pi_{#1}}
-\newcommand{\children}[1]{\,C_{#1}}
-\newcommand{\cluster}{\,\mathcal{S}}
-\newcommand{\daghetpart}{\algo{DagHetPart}\xspace}
-\newcommand{\dagmem}{\algo{DagHetMem}\xspace}
-
-\newcommand{\MM}{M}
-\newcommand{\MC}{MC}
-\newcommand{\rt}{rt}
-\newcommand{\curM}{curM}
-\newcommand{\curC}{curC}
-\newcommand{\PD}{PD}
-
-\newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
-\newcommand{\hmey}[1]{{\color{red}[HM: #1]}}
-\newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
-%%
-%% end of the preamble, start of the body of the document source.
+% Document
 \begin{document}
-
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-    \title{Adaptive Scheduling of Scientific Workflows}
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-    \author{Svetlana Kulagina}
-    \email{svetlana.kulagina@hu-berlin.de}
-    \orcid{0000-0002-2108-9425}
-    \affiliation{%
-        \institution{Humboldt Universitaet zu Berlin}
-        \streetaddress{Unter den Linden 6}
-        \city{Berlin}
-        \country{Germany}
-        \postcode{10099}
-    }
-
-       \author{Henning Meyerhenke}
-       \affiliation{%
-           \institution{Humboldt Universitaet zu Berlin}
-           \city{Berlin}
-           \country{Germany}
-       }
-       \email{meyerhenke@hu-berlin.de}
-
-    \author{Anne Benoit}
-    \affiliation{%
-        \institution{ENS Lyon}
-        \city{Lyon}
-        \country{France}
-    }
-    \email{Anne.Benoit@ens-lyon.fr}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-    \renewcommand{\shortauthors}{Trovato et al.}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
-    \begin{abstract}
-       Todo: reinsert CCSXML concepts
-    \end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-    \keywords{Scheduling, Adaptive Scheduling, DAG}
-    \received{15 April 2024}
-    \received[revised]{12 March 2009}
-    \received[accepted]{5 June 2009}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-    \maketitle
-
-    \section{Introduction}
-
-    \section{Model}
-
-    \begin{table}[h]
-        \begin{center}
-            \begin{tabular}{rl}
-                \hline
-                \textbf{Symbol} & \textbf{Meaning}  \\
-                \hline
-                $G = (V, E)$  & Workflow graph, set of vertices (tasks) and edges  \\
-                $\parents{u}$, $\children{u}$ & Parents of a task $u$, children of a task $u$ \\
-                $m_u$& Memory weight of task $u$ \\
-                $w_u$   & Workload of  task $u$  (normalized execution time)    \\
-                $c_{u,v}$   & Communication volume along the edge $(u,v)\in E$ \\
-                $F$, $\mathcal{F}$ & A partitioning function and the partition it creates \\
-                $V_i$ & Block number $i$\\ %\wrt~some $F$   \\
-                $\cluster$, $k$   & Computing system and its number of processors   \\
-                $p_j$, proc($V_i$)  & Processor number $j$, processor of block $V_i$ \\
-                $M_j$, $s_j$   & Memory size and speed of processor $p_j$   \\
-                $\beta$ & Bandwidth in the compute system  \\
-                $\bottomlevel{u}$   & Bottom weight of task $u$   \\
-                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$ \\
-                $\Gamma = (\mathcal{V}, \mathcal{E})$ & Quotient graph, its vertices and its edges  \\
-                $r_u$, $r_{V_i}$  & Memory requirement of  task $u$ and of block $V_i$   \\
-                $r_{\max}$   & Maximum memory requirement in a workflow   \\
-                $\criticalpath$ & Critical path in a workflow  \\
-                \hline
-            \end{tabular}
-        \end{center}
-        \caption{Notation.} \label{tabnotation}
-    \end{table}
-
-    \paragraph{Workflow-related changes}
-
-    \begin{itemize}
-        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
-        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
-
-    \end{itemize}
-
-    The following changes are not a part of this article's scope:
-
-    \begin{itemize}
-        \item The workflow structure changes: edges or tasks come in or leave.
-    \end{itemize}
-
-    \paragraph{Execution environment-related changes }
-
-    \begin{itemize}
-        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
-        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
-
-    \end{itemize}
-
-    The following changes are not a part of this article's scope:
-
-    \begin{itemize}
-        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
-    \end{itemize}
-
-    \subsection{Time of changes }
-
-    We consider discrete time in seconds.
-    The time point(s) at which the changes happen is unambiguously defined.
-
-    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
-    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
-    The start time of the source task in the workflow is zero.
-    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
-
-    \subsection{Changes and knowledge horizon - important questions TBA}
-
-    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
-
-    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
-    However, how do we model the following:
-    \begin{itemize}
-        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
-This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
-        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
-        \item How many times of change do we model - one per workflow run, or multiple?
-        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
-    \end{itemize}
-
-    \section{Related work}
-
-    In the limited time I spent looking, I found no paper addressing the exact same problem.
-
-    Wang et al.~\cite{wang2019dynamic} proposes a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
-    Particles are possible solution in the solution space.
-    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
-    Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
-    However, the approach is for clouds.
-    \xspace
-
-    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
-    on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
-    that these runtime changes are unrelated to each other).
-    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
-    performance measure $\phi$.
-    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
-
-    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling heuristic
-    for clouds.
-    Based on a 3-objective cost model, their greedy scheduling algorithm schedules tasks into machines.
-    The authors use provenance data to make scheduling decisions.
-    The cost model is a set of linear equations computed in the simulation environment, it represents the cost of an execution based on the criteria.
-    In the algorithms, the authors test out 4 scenarios - one preferring each criteria, and a balanced one.
-    The algorithm merely chooses the best virtual machine for each next task based on the cost given by the model.
-    An additional algorithm combines several cloud activities (task executions) into one to improve the cost, so that each
-    execution is not too small (utilize the granularity factor, the smallest entity of payment of the cloud producer).
-    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
-    They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
-
-    Rahman~\etal~\cite{rahman2013} propose a scheduling heuristic for grids that proposes mapping of tasks to machines by calculating
-    the critical path in the graph dynamically at every step.
-    They call it the dynamic critical path (DCP).
-    For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
-    on the start time of a task (differing by the slack this task has).
-    All tasks on this critical path have the same earliest and latest start times, because they cannot be ddelayed.
-
-    The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
-    If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum execution time for
-    the task is chosen.
-    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
-    The authors evaluate their approaches on random workflows of the size up to 300 tasks.
-
-    The authors provide an overview over (simpler) scheduling heuristics.
-    For example, GRASP (generally randomized adaptive search procedure) conducts a number of iterations to search an optimal
-    solution for mapping tasks on machines.
-    A solution is generated at each step, and the best solution is kept at the end.
-    The search terminates when a certain termination criterion is reached.
-    It generates better results than other algorithms, because it explores the whole solution space.
-
-    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
-    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
-    The scenario contains AND- and OR-branches, where AND-branches indicate acitivities that need to be executed in parallel.
-
-    The heuristic first partitions the set into local schedules by using affinity matrices to determine similar
-    activities and group them together.
-    Then they physically allocate these partitions to groups of disaster responders and tasks within this group.
-    They define a constraint system for that.
-    The dynamic part deals with changes and distinguishes between retriable and compensation acitivities.
-    The heuristic calculates a new execution path with these tasks.
-    In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
-
-    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
-    The aim is to minimize the makespan, and the experiments were conducted on a single wotkflow with 10 tasks.
-    The procesdure involves building a first (static) schedule, priodic resource monitoring and rescheduling the remaining
-    tasks.
-    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
-    For each resuorce group, there is an own scheduler, and an overall global scheduler responsible for distributing
-    tasks to groups.
-    The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
-    The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
-
-    The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
-    THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
-    They then go through tasks and map than to such processors that minimize the earliest start times of this task's
-    successors.
-    To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
-    average communication and computation costs fir the dependent tasks.
-
-    The rescheduling is being triggered when either a load on a resource increases over a threshold, or if a new resource
-    is added.
-    The algorithm produces a new mapping from scrath, and this mapping is being accepted if the resulting maespan is
-    smaller than the previously predicted one.
-
-\section{Proposal of a new heuristic with slightly refined model}
-
-The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
-but favor processor reuse as is done in HEFT. 
-Furthermore, this would allow us to handle variability on the fly, by updating
-the bottom levels if some parameters vary, and computing the schedule
-only for the near future...
-
-\subsection{Model}
-
-For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
-For an  edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
-
-For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds: 
-$\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
-We can decide to evict some data from the main memory if we are sending the data
-to another processor; it then stays in the communication buffer until it has been sent.
-
-We keep track of the current ready time of each processor and each communication
-channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
-are set to~$0$.
-
-We also keep track of the currently available memory, $availM_j$ and $availC_j$,
-on respectively the processor memory and communication buffer. 
-Furthermore, $\PD_j$ is a priority queue with the {\em pending data} 
-that are in the memory of size $\MM_j$ but may be evicted to be communicated, if 
-more memory is needed on~$p_j$. They are ordered by non-decreasing size. 
-They correspond to some $c_{u,v}$'s.
-
-\subsection{Baseline: original HEFT without memories}
-
-    Original HEFT does not consider memory sizes.
-    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
-    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.
-
-    HEFT works in two steps.
-    In the first step, it calculates the ranks of tasks by computing their non-increasing bottom levels.
-    The bottom level of a task is defined as
-    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
-    (the max is 0 if there is no outgoing edge).
-    The tasks are sorted by non-decreasing ranks.
-
-    In the second step, the algorithm iterates over the ranks and tries to assign the task to the processor where it
-    has the earliest finish time.
-    We tentatively assign each task to each processor.
-    The task's starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
-    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
-    The starting time is then
-    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
-    Its finish time on $p_j$ is then
-    $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
-
-    Once we have computed all finish times for task~$v$,
-    we keep the minimum $FT(v,p_j)$ and assign task~$v$
-    to processor~$p_j$.
-
-    \textit{Assignment to processor}
-    When assigning the task, we set the ready time of the processor~$j$ $rt_j$ to the finish time of the task.
-    For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
-    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
-    communication time $c( u,v) / \beta$.
-
-\subsection{Heuristics}
-    Like the original HEFT, our heuristic consistst of two steps: first, computing task ranks,
-    and second, assigning tasks to processors in the order defined in the first step.
-    We consider three variants of HEFT accounting for memory usage, which only
-    differ in the order they consider tasks to be scheduled.
-
-\subsubsection{Step 1: calculate task ranks}
-
-HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
-level is defined as
-$$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
- (the max is 0 if there is no outgoing edge).
-
- HEFTM-BLC: from the study of the fork (see below), it seems important
- to also account for the size of the data as input of a task,
- to give more priority at tasks with potential large incoming communications.
- For each task, we compute its modified bottom level: 
- $$blc(u) = w_u + \max_{(u,w)\in E} \{c_{u,w} + blc(w)\} + \max_{(v,u)\in E} c_{v,u}   . $$
-
-    \skug{avoid having mixed ranks, when the memory size of the lower task is not taken into account}
-
-HEFTM-MM orders tasks as dictated by MinMem.
-
-\subsubsection{Task assignment}
-
-Then, the idea is to pick the next free task in the given order,
-and greedily assign it to a processor, by trying all possible options
-and keeping the most promising one.
-
-\medskip
-\noindent{\em Tentative assignment of task~$v$ on $p_j$.}\\
-{\bf Step 1.} First, we need to check that for all predecessors~$u$ of~$v$ that are mapped
-on~$p_j$, the data $c_{u,v}$ is still in the memory of~$p_j$, 
-i.e., $c_{u,v}\in PD_j$. Otherwise, the finish time is set to~$+\infty$ (invalid choice).
-
-\smallskip
-\noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
-\[Res = availM_j  - m_v - \sum_{u \in \Pi(v), u\notin T(p_j)}  \{c_{u,v}\}
-- \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]
-
-$T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
-already in the memory of~$p_j$. However, the files from the 
-other predecessor tasks must be loaded in memory before executing task~$v$,
-as well as $m_v$ and the data generated for all successor tasks.
-$Res$ is then checking whether there was enough memory; if it is negative,
-it means that we have exceeded the memory of~$p_j$ with this tentative 
-assignment.
-
-In this case ($Res <0$), we  try evicting
-some data from memory so that we have enough memory to execute task~$v$. 
-We need to evict at least $Res$ data.
-For now, we propose a greedy approach, evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
-in order to avoid costly communications. 
-\AB{FYI We initially discussed  evicting the largest files, but this leads to 
-large communications and does not seem efficient after all... Maybe we can think of another 
-approach that would take into account both data size and bottom level...}
-While tentatively evicting files, we remove them from the list of pending memories and move them into a list
-    of memories pending in the communication buffer.
-We keep track of the available buffer size, too - each time a file gets moved into the pending in buffer, the available buffer size is reduced by its weight.
-
-If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
-    or if while doing so we exceeded the size of the available buffer,
-we set the finish time to~$+\infty$ (invalid choice).
-
-\smallskip
-\noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
-    Its starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
-must be orchestrated from predecessor tasks $u\notin T(p_j)$.
-    The starting time is then
-    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} } \]
-    Its finish time on $p_j$ is then
-$FT(v,p_j) = ST(v, p_j) + \frac{w_v}{s_j}$.
-
-\medskip
-\noindent{\em Assignment of task~$v$.}\\
-Once we have computed all finish times for task~$v$, 
-we keep the minimum $FT(v,p_j)$ and assign task~$v$
-to processor~$p_j$.
-In detail, we:
-    \begin{itemize}
-        \item  Evict the file memories that correspond to edge weights that need to be evicted to free the memory.
-        We remove these files from pending memories
-        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
-        \item    Calculate the new $availM_j$ on the processor.
-        We subtract the weights of all incoming files from predecessors assigned to the same processor,
-        and add the weights of outgoing files generated by the currently assigned task.
-        \item  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
-        communication buffers $rt_{j', j}$ for the processor~$j'$that the predecessor $u$ has been assigned to: we increase them by the
-        communication time $c( u,v) / \beta$.
-        We also remove the incoming files from either the pending memories or pending data in buffers of these other
-        processors, and increase the available memories or available buffer sizes on these processors.
-        \item We compute the correct amount of available memory for $p_j$ (for when the task is done).
-        For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
-        the incoming edge, also freeing the same amount of available memory (increasing $availM_j$).
-        For each successor, on the other hand, we add the edge weights to pending memories and reduce $availM_j$ by the corresponding
-        amount.
-    \end{itemize}
-
-\subsection{Retracing the effects of change on an existing schedule}
-    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
-    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
-    they can lead to a later finishing time (\eg if some tasks longer and delay other tasks), or they can have no effect (\eg if new processors
-    joined the cluster, but the old schedule did not account for them).
-    To assess the impact, we need to retrace the schedule.
-
-    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
-    entire schedule.
-
-    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
-    is a topologial ordering.
-    We then repeat steps similar to those we did during tentative assignment in our heuristic, except we do not choose a processor
-    anymore, but rather see if the current one still fits.
-
-    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from our heuristic.
-    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ or $c_{v,w}$ from successors $w$,
-    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements from other tasks).
-    If originally,$Res$ was positive (no files were evicted from memory into the communication buffer), then it has to stay this way -
-    otherwise evicted files can invalidate next tasks.
-    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
-    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
-    We update the $availM_j$ and $availMC_j$ according to the new memory constraints.
-
-    Then we can re-calculate the finish time of the task on its processor like in Step 3.
-    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
-    (after delayed previous tasks), and changed communication buffer availability.
-
-    Then, after having updated the processor's values, we move on to the next task.
-
-\subsection{The fork}
-We look at the behavior of these heuristics on a fork graph,
-where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
-to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
-
-Without memory, this problem is NP-complete; this is equivalent
-to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
-and with two processors. Half of the tasks must be sent to the processor
-on which $T_0$ is not executed, and the optimal makespan is 
-$w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
-
-However, with an infinite number of identical processors, it can be
-solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$; 
-the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
-while the remaining $n-k$ tasks are executed locally (try all values of $k$).
-
-With heterogeneous processors, it is probably NP-complete again
-because we could ensure that there are only two processors fast enough
-and get back to the 2-partition...
-
-We also had an example where evicting large files first in step 2
-can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
-$f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
-imposes that we free one unit of memory before executing one
-of the tasks... Actually the new version with BLC would start 
-considering $T_2$ and be fine in this case...
-
-\AB{Can we prove that we have (maybe) a 2-approximation, 
-at least for the fork? What worst-case can we think of? }
-
-\subsection{Approximation}
-\hmey{Rough notes:}
-Let's use a fork to see how the algorithm behaves and if it provides some approximation.
-Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
-
-\section{Conclusion}
-
-\section{Appendix}
 
@@
 
-    \subsection{New heuristic: HEFT with minimal memory traversal}
-
-    This heuristic changes both the rank calculation step, and the assignment step.
-
-    Before calculating ranks, we calculate the optimal memory traversal over the entire graph.
-    We then calculate ranks as bottom levels on each processor, but adding some small $\epsilon$ on tasks that have the
-    same bottom level, but are higher in the memory traversal order.
-    So, for a traversal $v_1 \rightarrow v_2 \rightarrow v_3$, if both $v_2$ and $v_3$ have the bottom level of $100$,
-    $v_2$ will get the rank of $100+\epsilon$.
-
-    Then we sort the tasks ascending according to their rank.
-    We then look to minimize not the earliest finishing time of the task itself, but the earliest starting time of
-    its \skug{heaviest} successor.
-    The term heaviest is currently not defined exactly.
-    The intuition is to try to optimize for that successor that hold the schedule up the most.
-
-    We then do a procedure similar to our FillUpFit intuition.
-    We traverse the graph along the optimal memory traversal.
-    We keep track of the processor that we last used.
-    With each new task $v$, we check if it has the best rank on the same processor.
-    If so, and if the processor's memory is big enough, we fit the task there.
-
-    If the memory is too full or if $v$ has a better rank elsewhere, we calculate earliest starting times
-    of all its successors on all other processors (we assume that not only $v$ itself goes on that processor, but also
-    all its successors).
-    We then choose the processor that gives the best earliest starting time for the  \skug{heaviest} successor
-    and assign $v$ there.
-
-    \bibliographystyle{ACM-Reference-Format}
-    \bibliography{references}
-
 \end{document}
-\endinput
-%%
-%% End of file `sample-sigconf.tex'.

### 2024-05-02 main.tex
removed unrelated text

@@
     \end{table}
-
-    \paragraph{Workflow-related changes}
-
-    \begin{itemize}
-        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
-        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
-
-    \end{itemize}
-
-    The following changes are not a part of this article's scope:
-
-    \begin{itemize}
-        \item The workflow structure changes: edges or tasks come in or leave.
-    \end{itemize}
-
-    \paragraph{Execution environment-related changes }
-
-    \begin{itemize}
-        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
-        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
-
-    \end{itemize}
-
-    The following changes are not a part of this article's scope:
-
-    \begin{itemize}
-        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
-    \end{itemize}
-
-    \subsection{Time of changes }
-
-    We consider discrete time in seconds.
-    The time point(s) at which the changes happen is unambiguously defined.
-
-    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
-    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
-    The start time of the source task in the workflow is zero.
-    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
-
-    \subsection{Changes and knowledge horizon - important questions TBA}
-
-    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
-
-    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
-    However, how do we model the following:
-    \begin{itemize}
-        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
-This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
-        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
-        \item How many times of change do we model - one per workflow run, or multiple?
-        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
-    \end{itemize}
 
@@
 
-    \section{Appendix}
-
-    \subsection{New heuristic: Local search}
-    One strategy that can be used is local search, a similar approach as with swaps.
-    The starting point could be a previously valid solution (for example, one by HetPart).
-    In the first step, the solution would need to become valid again.
-    Then, we repartition blocks on the critical path and merge blocks in attempt to improve makespan.
-    Finally, we swap blocks between processors to finish makespan improvement.
-
-    \subsubsection{Restore Validity}
-    We~(Algorithm~\ref{alg:RestoreValidity}) start with the previously partitioned and assigned quotient tree $\Gamma$,
-    and the new weights on tasks and edges.
-    We first recompute the memory requirements of all blocks.
-    Then we identify all blocks that have become invalidated and exceed the memory of their assigned processor
-    (Line~\ref{line:damaged}).
-    For all these blocks, we try out ``border'' tasks, that is, tasks, that have neighbours in neighbouring blocks
-    (Line~\ref{line:neighb}).
-    We tentatively move them (creating a temporary quotient DAG $\Gamma'$) and assess the impact of this move on the
-    makespan(Line~\ref{line:makespan}).
-    The condition is, of course, that the neighboring block now does not exceed the memory of its own processor.
-    The task that delivers the best makespan in then moved in Line~\ref{line:movebest}.
-    The entire process is being repeated until the block fits into the memory of the processor again.
-    We return the modified quotient graph $\Gamma$ that contains the mapping to processors.
-
-    \begin{algorithm}[h!]
-        \caption{Restore Validity}
-        \label{alg:RestoreValidity}
-        \begin{algorithmic}[1]
-            \Procedure{Restore Validity}{$\Gamma$, $W$ }
-                \\
-                \Comment{Input: initial schedule as quotient tree $\Gamma$, new weights $W$}
-                \State $\Gamma \gets $ \textsc{QuotientDAG}($\Gamma$, $W$ ); \Comment Quotient DAG with new weights
-
-                \State $D \gets \emptyset$ \Comment {Damaged blocks that require action}
-                \For{$\nu_i \in \Gamma$}
-                    \State $r_{\nu_m} \gets W$
-                    \Comment{Recompute the new memory requirements of blocks in the quotient DAG}
-                \EndFor
-                \For{$\nu_i \in \Gamma$}
-                    \label{line:damaged}
-                    \If{$r_{\nu_i} > M_{proc(\nu_i)}$}
-                        \State $D \gets D \cup \{\nu_i\}$
-                    \EndIf
-                \EndFor
-
-                \For{$\nu_i \in D$}
-                    \Comment Restore the validity of each damaged block
-                    \While{$r_{\nu_i} > M_{proc(\nu_i)}$}
-                        \State $MS_{min}\gets \infty; v_{min} \gets NULL; \nu_{min} \gets NULL$
-                        \Comment Moving $v_{min}$ to $\nu_{min}$ gives $MS_{min}$
-                        \For{$v \in \nu_i$}
-                            \For{$u \in ( \Pi_v \cup C_v) \cap (\Pi_{\nu_i} \cup C_{\nu_i})$}
-                                \label{line:neighb}
-                                \State $\Gamma' \gets \Gamma \mid \nu(v) == \nu(u)$
-                                \If{ $r_{\nu(u)} > M_{proc(\nu(u))}$ and \textsc{Makespan}($\Gamma'$)$\leq MS_{min}$}
-                                    \label{line:makespan}
-                                    \State $MS_{min} \gets $\textsc{Makespan}($\Gamma'$); $v_{min} \gets v$; $\nu_{min} \gets \nu(u);$
-                                \EndIf
-                            \EndFor
-                        \EndFor
-                        \State  $\Gamma \gets \Gamma \mid \nu(v_{min}) == \nu_{min}$ \label{line:movebest}
-                    \EndWhile
-                \EndFor
-                \State return $\Gamma$
-            \EndProcedure
-        \end{algorithmic}
-    \end{algorithm}
-
-    \subsubsection{Repartition and Merge}
-
-    In the next step, we center our attention around changes in the execution environments and how the changes in
-    the workflow reflect on it.
-
-    First, we try to further partition blocks that lie on the critical path of the workflow.
-    We first compute the critical path (Line~\ref{line:cp}) and then, as long as there are free processors in the
-    execution environment(Line~\ref{line:whilefree}, assumed to be sorted in descending order of memory sizes),
-    we partition the blocks on the critical path(Line~\ref{line:repartition}).
-    We try to partition each block into 2 smaller blocks.
-    However, due to the nature of the partitioner, more can be created.
-    We assume that the resulting blocks(Line~\ref{line:repartition}) are given back sorted by their memory requirement.
-    We then assign them to free processors as long as the blocks fit and there are free processors.
-    If even one part does not fit(Line~\ref{line:ifnotfits}), then we discard the entire partition and try partitioning
-    the next block on the critical path.
-
-    Not only partitioning (introducing more parallelism) can lead to a makespan improvement.
-    Shorter runtimes of tasks can justify merging some blocks together.
-    For all pairs of blocks that are neighbours (Line~\ref{Line:neighbblocks}), we tentatively merge them and assign
-    the result to the processor of the first block.
-    If the merged block fits on this processor and the merged quotient graph has a smaller overall makespan, then we save
-    the merge and proceed with other blocks.
-
-    \begin{algorithm}[h!]
-        \caption{Repartition and Merge}
-        \label{alg:RepartitionMerge}
-        \begin{algorithmic}[1]
-            \Procedure{Repartition and Merge}{$\Gamma$, $\cluster$ }
-                \\
-                \Comment{Input: quotient tree $\Gamma$, execution environment $\cluster$}
-                \State $CP \gets $\textsc{Criticalpath}($\Gamma$) \label{line:cp}
-                \State $Free \gets \{ p \in \cluster \mid p$ is free$\}$
-                \While{$\mid Free \mid \neq 0$}
-                    \label{line:whilefree}
-                    \For{$\nu \in CP$}
-                        \State $\{ \nu_1, \nu_2,\dots\}\gets$\textsc{Partition}($\nu$,2) \label{line:repartition}
-                        \For{$\nu' \in \{ \nu_1, \nu_2,\dots\}$}
-                            \If{$r_\nu' \leq Free.$head()}
-                                \State proc($\nu'$)$\gets Free.$head();
-                            \Else
-                                \label{line:ifnotfits}
-                                \State \textsc{Merge}$\nu' \in \{ \nu_1, \nu_2,\dots\}$; break;
-                                \Comment Merge the parts back and leave this partitioning, as it does not fit
-                            \EndIf
-                        \EndFor
-                    \EndFor
-                \EndWhile
-
-                \For{$\nu_1, \nu_2 \in \Gamma \mid \nu_1 \in ( \Pi_{\nu_2} \cup C_{\nu_2})$}
-                    \label{line:neighbblocks}
-                    \State $\{\Gamma', \nu_m\} \gets $ \textsc{Merge}($\nu_1, \nu_2$); proc($\nu_m$)$\gets $ proc($\nu_1$);
-                    \State $\mu \gets $\textsc{Makespan}($\Gamma'$)
-                    \If{$r_{\nu_m} \leq M_{proc(\nu_m)}$ and $\mu \leq $ \textsc{Makespan}($\Gamma$)}
-                        \State $\Gamma \gets \Gamma'$
-                    \EndIf
-                \EndFor
-
-                \State return $\Gamma$
-            \EndProcedure
-        \end{algorithmic}
-    \end{algorithm}
-
-    \subsubsection{Swaps}
-
-    Then a process similar to Step 4 in %\daghetpart
-    can be executed: the newly rearranged blocks are swapped until no
-    improvement in makespan can be achieved.
-
-    \subsection{New heuristic: HEFT with minimal memory traversal}
-
-    This heuristic changes both the rank calculation step, and the assignment step.
-
-    Before calculating ranks, we calculate the optimal memory traversal over the entire graph.
-    We then calculate ranks as bottom levels on each processor, but adding some small $\epsilon$ on tasks that have the
-    same bottom level, but are higher in the memory traversal order.
-    So, for a traversal $v_1 \rightarrow v_2 \rightarrow v_3$, if both $v_2$ and $v_3$ have the bottom level of $100$,
-    $v_2$ will get the rank of $100+\epsilon$.
-
-    Then we sort the tasks ascending according to their rank.
-    We then look to minimize not the earliest finishing time of the task itself, but the earliest starting time of
-    its \skug{heaviest} successor.
-    The term heaviest is currently not defined exactly.
-    The intuition is to try to optimize for that successor that hold the schedule up the most.
-
-    We then do a procedure similar to our FillUpFit intuition.
-    We traverse the graph along the optimal memory traversal.
-    We keep track of the processor that we last used.
-    With each new task $v$, we check if it has the best rank on the same processor.
-    If so, and if the processor's memory is big enough, we fit the task there.
-
-    If the memory is too full or if $v$ has a better rank elsewhere, we calculate earliest starting times
-    of all its successors on all other processors (we assume that not only $v$ itself goes on that processor, but also
-    all its successors).
-    We then choose the processor that gives the best earliest starting time for the  \skug{heaviest} successor
-    and assign $v$ there.
-
     \bibliographystyle{IEEEtran}

### 2024-05-02 dynamic-scenario.tex
more unrelated text

@@
 
+    \subsection{Dynamic Scenario}
+
+    In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
+    This information may include, memory usage, runtime, graph structures, or the status of the underlying infrastructure.
+    In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
+    By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
+    Also the underlying infrastructure can change during the workflow execution.
+    Examples are processor failures, node recoveries, or acquisition of new nodes.
+    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
+    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.
+
+    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{}.
+    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
+%For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
+%Bader~et~al.~ report a prediction error between 13\% and 17\% for their method, showing an exponential task error distribution.
+% @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
+    Such a dynamic execution environment necessitates for a dynamic scheduling method where the schedule can be recomputed during the workflow execution.
+
+    \subsection{Retracing the effects of change on an existing schedule}
+    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
+    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
+    they can lead to a later finishing time (\eg if some tasks longer and delay other tasks), or they can have no effect (\eg if new processors
+    joined the cluster, but the old schedule did not account for them).
+    To assess the impact, we need to retrace the schedule.
+
+    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
+    entire schedule.
+
+    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
+    is a topologial ordering.
+    We then repeat steps similar to those we did during tentative assignment in our heuristic, except we do not choose a processor
+    anymore, but rather see if the current one still fits.
+
+    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from our heuristic.
+    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ or $c_{v,w}$ from successors $w$,
+    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements from other tasks).
+    If originally,$Res$ was positive (no files were evicted from memory into the communication buffer), then it has to stay this way -
+    otherwise evicted files can invalidate next tasks.
+    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
+    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
+    We update the $availM_j$ and $availMC_j$ according to the new memory constraints.
+
+    Then we can re-calculate the finish time of the task on its processor like in Step 3.
+    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
+    (after delayed previous tasks), and changed communication buffer availability.
+
+    Then, after having updated the processor's values, we move on to the next task.
+
 \end{document}

### 2024-05-02 main.tex
more unrelated text

@@
 
-    \subsection{Dynamic Scenario}
-
-    In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
-    This information may include, memory usage, runtime, graph structures, or the status of the underlying infrastructure.
-    In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
-    By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
-    Also the underlying infrastructure can change during the workflow execution.
-    Examples are processor failures, node recoveries, or acquisition of new nodes.
-    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
-    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.
-
-    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{}.
-    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
-%For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
-%Bader~et~al.~ report a prediction error between 13\% and 17\% for their method, showing an exponential task error distribution.
-% @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
-    Such a dynamic execution environment necessitates for a dynamic scheduling method where the schedule can be recomputed during the workflow execution.
-
-    \subsection{Retracing the effects of change on an existing schedule}
-    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
-    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
-    they can lead to a later finishing time (\eg if some tasks longer and delay other tasks), or they can have no effect (\eg if new processors
-    joined the cluster, but the old schedule did not account for them).
-    To assess the impact, we need to retrace the schedule.
-
-    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
-    entire schedule.
-
-    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
-    is a topologial ordering.
-    We then repeat steps similar to those we did during tentative assignment in our heuristic, except we do not choose a processor
-    anymore, but rather see if the current one still fits.
-
-    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from our heuristic.
-    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ or $c_{v,w}$ from successors $w$,
-    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements from other tasks).
-    If originally,$Res$ was positive (no files were evicted from memory into the communication buffer), then it has to stay this way -
-    otherwise evicted files can invalidate next tasks.
-    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
-    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
-    We update the $availM_j$ and $availMC_j$ according to the new memory constraints.
-
-    Then we can re-calculate the finish time of the task on its processor like in Step 3.
-    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
-    (after delayed previous tasks), and changed communication buffer availability.
-
-    Then, after having updated the processor's values, we move on to the next task.
-
     \subsection{The fork}

### 2024-05-02 main.tex
WIP. various schanges, incl results

@@
 \usepackage{xspace}
+\usepackage{hyperref}
+\usepackage{numprint}
 
@@
 \newcommand{\cluster}{\,\mathcal{S}}
-\newcommand{\daghetpart}{\algo{DagHetPart}\xspace}
-\newcommand{\dagmem}{\algo{DagHetMem}\xspace}
+
+\newcommand{\heftmm}{\algo{HEFTM\_MM}\xspace}
+\newcommand{\heftbl}{\algo{HEFTM\_BL}\xspace}
+\newcommand{\heftblc}{\algo{HEFTM\_BLC}\xspace}
 
@@
 
+    \section{Experimental evaluation}
+    \subsection{Setup}
+    \label{sec:setup}
+
+    All algorithms are implemented in C++ and compiled with g++ (v.11.2.0).
+    The experiments are executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
+    and CentOS 8 as OS.
+    Code, input data, and experiment scripts are available for review under \url{TODO:insert link}.
+
+    Next, we describe the set of workflows used in the evaluation and then the clusters on which the
+    workflows are scheduled.
+
+    \subsubsection{Workflow instances}
+    The input data for the experiments consists of two sets of workflows: real-world workflows
+    obtained from~\cite{ewels2020nf} and workflows obtained by simulating real-world workflows
+    with the WFGen generator~\cite{COLEMAN202216}.
+    First, we discuss how the graph topology is generated,
+    then we focus on the weights associated to tasks and edges.
+
+    \paragraph{Workflow graphs}
+    For real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
+    respective repository and transformed into .dot format using the nextflow option ``-with-dag''.
+    The resulting DAG contains many pseudo-tasks that are only internal representations in nextflow
+    (and not actual tasks); that is why we removed them.
+
+    For the simulated workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
+    the desired number of tasks.
+    The model workflows are described on the WFGen website, and we used the following ones:
+    1000Genome, BLAST, BWA, Epigenomics, Montage, Seismology, and SoyKB.
+    Other models could not be generated without errors.
+%
+    As number of tasks, we use: 200, \numprint{1000}, \numprint{2000}, \numprint{4000}, \numprint{8000}, \numprint{10000},
+    \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
+    We divide the workflows into three groups by size: small ones with up to \numprint{8000} tasks, middle
+    ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.
+    Note that for some workflow models, such as SoyKB or Montage, only a subset of the workflow sizes could be generated,
+    because the workflow generator either took a disproportionally long time or yielded errors.
+
+    Overall, this yields four workflow types, denoted by real, small, mid, and big.
+
+    \paragraph{Generation of task and edge weights}
+    For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
+    The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
+    Each row corresponds to an execution of one task on one cluster node.
+    Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
+    to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
+    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
+    size of files that the task sends to all its children.
+    For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
+    To avoid underestimation, we take the maximum among all values from different runs on the same cluster node.
+    Not all tasks have historical runtime data stored in the tables.
+    In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
+    For two more, around 40\% of tasks have no historical runtime data stored.
+    Hence, in the absence of historical data about a task, we give it a weight of~1.
+%
+    Because the historical data contains absolute measured values and the cluster node information is a relative value,
+    we normalize all values extracted from the historical data by the smallest one.
+    This way, task memory weights fit into the memory of the cluster nodes.
+    Additionally, this way, the tasks without historical data receive less insignificant values compared to tasks with historical data.
+
+    For the simulated workflows, we generate random execution times and memory weights for each task as well as
+    edge weights for each precedence constraint.
+    We generate uniformly distributed values between $1$ and $10$ for edge weights,
+    $1$ and $1000$ for the workloads,
+    and $1$ and $192$ for memory weights.
+    When doing so, we try to mimic the weights observed in the historical data, hence \eg the low lower bounds for the
+    workloads.
+
+    \subsubsection{Target computing systems}
+    To fully benefit from the historical data, the  {\em default} experimental environment
+    that we consider is a cluster based on the same six
+    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
+    We set the number of each kind of node to six, thus having 36 processors in total. % the whole cluster.
+
+        Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
+        ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines; ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and large memory
+        (see Table~\ref{tab:procs} in Appendix~\ref{sub:appx-cluster-config}~\cite{daghetpart_full_version}).
+    Note that, by nature,
+    memory sizes are normalized values, relative to each other, too.
+    Therefore, we additionally normalize memory weights of real-world workflows
+    to the maximum size of 192 to make sure they fit.
+    For simulated workflows, we increase memory sizes proportionally until the task
+    with the biggest memory requirement still
+    has a processor it could be executed on.
+
+    To test various settings, we also vary the cluster configuration and consider
+    variants of the default cluster presented above:\\
+    $\bullet$ {\em Small} and {\em large} clusters. While the default cluster consists of 36 nodes (6 of each kind), we
+    consider a small cluster with one processor of each kind ($6$ processors in total), and a
+    large cluster with ten processors of each kind ($60$ processors in total). \skug{leaving the big cluster inside for
+    the case we do want to include it}\\
+    $\bullet$ {\em Fat-and-thin} cluster. In this cluster, the fastest processors have the smallest amount of memory (thin),
+    while the slowest processors have the largest memories (fat).
+    The overall amount of memory in the cluster, as well as the overall speed of all processors remain the same as in the
+            {\em normal} cluster.
+%\end{itemize}
+
+    \subsection{Results}
+    \subsubsection{Normal cluster}
+
+    On the normal cluster, the baseline produces invalid results in $40$\% of small workflows and $66.6$\% of middle-sized
+    and big workflows.
+    In most cases, the failures of the baseline can be traced to choosing the processor with not enough memory when several
+    options are available.
+    Due to not taking memories into account, the baseline chooses a processor with already full memory, while there can
+    be another processor with the same speed that still has enough memory.
+    In this case, \heftbl produces the same makespan while producing a valid solution.
+    The heuristics produce valid results by default and also manage to find a solution in all cases (on all workflow sizes).
+
+    On the normal cluster and among small workflows, the makespans produced by \heftbl and \heftblc are $1$\% worse than the baseline,
+    \heftmm $19$\% worse.
+    For middle-sized ones \heftmm is $12.3$\% worse, the rest is also $1-2$\% worse.
+    For the big workflows however, the results of \heftbl and \heftblc are $<1$\% worse and \heftmm is only $9$\% worse.
+    The heuristics pay off more on larger workflows.
+    \subsubsection{Memory usage}
+    \subsubsection{Impact of cluster size}
+    On a cluster with only 6 processors (1 of each kind), the baseline makes more failures: from $43.4$\%  (small workflows)
+    to $88.9$\% (mid-sized) and $94.4$\% (big workflows).
+    These values also become more unattainable: the makespans heuristics produce are worse by $3-12$\% (\heftbl) to $12-52$\%
+    (\heftmm).
+    However, these comparisons are less expressive, because most of baseline solutions are invalid.
+    \skug{TODO: add numbers about failures on the small cluster with the fixed bug}
+
+    \subsubsection{Impact of cluster composition}
+    On the {\em fat-and-thin} cluster, the baseline produces $60$\% invalid results on small workflows, $83.3$\% on
+    middle-sized and $88.9$\% on big workflows, and therefore performs worse than on the normal cluster.
+
+    The relative makespans are $14-29$\% worse for small workflows, $22-40$\% worse on middle-sized workflows and $16-35$\%
+    on big workflows.
+    \wrt failures, all heuristics again manage to schedule all workflows.
+    \subsubsection{Behavior of workflow types when scaling workflow size}
+    \subsubsection{Running times}
+    The geometric means of runtimes of all heuristics on small workflows lie between $0.9$ and $2.2$ seconds.
+    For middle-sized workflows, the runtimes for \heftbl and \heftblc are close to $2$ minutes, while \heftmm
+    takes $5.1$ minutes.
+    For the big workflows, again, \heftbl and \heftblc show a better runtime ($5$ and $5.4$ minutes, respectively).
+    \heftmm takes $13.9$ minutes on average.
+    \subsubsection{Summary}
+
     \section{Conclusion}

### 2024-05-02 main.tex
various changes in first sections

@@
 
+    Only mapping tasks to processors is not enough.
+    Reusing processors is important to achieve good makespans.
+
+    HEFT is one of the most popular heuristics (cite a review).
+    However, existing HEFT variations do not take memory sizes into consideration (\skug{check in related work if true!}).
+
+    In this paper, we formulate the problem that takes memories into consideration.
+    We propose three HEFT-based heuristics that take memory size into account: \heftbl, \heftblc, and \heftmm.
+    The difference is the way they order tasks for scheduling.
+    We compare our heuristics to a baseline HEFT variant that does not take memory sizes into account.
+
+    We find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
+
     \section{Model}
+
+    \skug{TODO:unify the notation with the table!!}
+
+    For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
+    For an edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
+
+    For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds:
+    $\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
+    We can decide to evict some data from the main memory if we are sending the data
+    to another processor; it then stays in the communication buffer until it has been sent.
+
+    We keep track of the current ready time of each processor and each communication
+    channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
+    are set to~$0$.
+
+    We also keep track of the currently available memory, $availM_j$ and $availC_j$,
+    on respectively the processor memory and communication buffer.
+    Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
+    that are in the memory of size $\MM_j$ but may be evicted to be communicated, if
+    more memory is needed on~$p_j$. They are ordered by non-decreasing size.
+    They correspond to some $c_{u,v}$'s.
 
@@
     \section{Related work}
+    \skug{problem: this is written for the old heuristic - parititioning-based heuristic to solvea dynamic problem! it is irrelevant to heft}
 
@@
     only for the near future...
-
-    \subsection{Model}
-
-    For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
-    For an edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
-
-    For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds:
-    $\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
-    We can decide to evict some data from the main memory if we are sending the data
-    to another processor; it then stays in the communication buffer until it has been sent.
-
-    We keep track of the current ready time of each processor and each communication
-    channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
-    are set to~$0$.
-
-    We also keep track of the currently available memory, $availM_j$ and $availC_j$,
-    on respectively the processor memory and communication buffer.
-    Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
-    that are in the memory of size $\MM_j$ but may be evicted to be communicated, if
-    more memory is needed on~$p_j$. They are ordered by non-decreasing size.
-    They correspond to some $c_{u,v}$'s.
 

### 2024-05-05 dynamic-scenario.tex
related work - 1st draft, WIP

@@
 
+    \section{Related work}
+    \skug{problem: this is written for the old heuristic - parititioning-based heuristic to solvea dynamic problem! it is irrelevant to heft}
+
+    In the limited time I spent looking, I found no paper addressing the exact same problem.
+
+    Wang et al.~\cite{wang2019dynamic} proposes a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
+    Particles are possible solution in the solution space.
+    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
+    Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
+    However, the approach is for clouds.
+    \xspace
+
+    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
+    on a single machine.
+    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
+    that these runtime changes are unrelated to each other).
+    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
+    performance measure $\phi$.
+    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
+
+    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling heuristic
+    for clouds.
+    Based on a 3-objective cost model, their greedy scheduling algorithm schedules tasks into machines.
+    The authors use provenance data to make scheduling decisions.
+    The cost model is a set of linear equations computed in the simulation environment, it represents the cost of an execution based on the criteria.
+    In the algorithms, the authors test out 4 scenarios - one preferring each criteria, and a balanced one.
+    The algorithm merely chooses the best virtual machine for each next task based on the cost given by the model.
+    An additional algorithm combines several cloud activities (task executions) into one to improve the cost, so that each
+    execution is not too small (utilize the granularity factor, the smallest entity of payment of the cloud producer).
+    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
+    They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
+
+    Rahman~\etal~\cite{rahman2013} propose a scheduling heuristic for grids that proposes mapping of tasks to machines by calculating
+    the critical path in the graph dynamically at every step.
+    They call it the dynamic critical path (DCP).
+    For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
+    on the start time of a task (differing by the slack this task has).
+    All tasks on this critical path have the same earliest and latest start times, because they cannot be ddelayed.
+
+    The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
+    If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum execution time for
+    the task is chosen.
+    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
+    The authors evaluate their approaches on random workflows of the size up to 300 tasks.
+
+    The authors provide an overview over (simpler) scheduling heuristics.
+    For example, GRASP (generally randomized adaptive search procedure) conducts a number of iterations to search an optimal
+    solution for mapping tasks on machines.
+    A solution is generated at each step, and the best solution is kept at the end.
+    The search terminates when a certain termination criterion is reached.
+    It generates better results than other algorithms, because it explores the whole solution space.
+
+    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
+    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
+    The scenario contains AND- and OR-branches, where AND-branches indicate acitivities that need to be executed in parallel.
+
+    The heuristic first partitions the set into local schedules by using affinity matrices to determine similar
+    activities and group them together.
+    Then they physically allocate these partitions to groups of disaster responders and tasks within this group.
+    They define a constraint system for that.
+    The dynamic part deals with changes and distinguishes between retriable and compensation acitivities.
+    The heuristic calculates a new execution path with these tasks.
+    In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
+
+    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
+    The aim is to minimize the makespan, and the experiments were conducted on a single wotkflow with 10 tasks.
+    The procesdure involves building a first (static) schedule, priodic resource monitoring and rescheduling the remaining
+    tasks.
+    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
+    For each resuorce group, there is an own scheduler, and an overall global scheduler responsible for distributing
+    tasks to groups.
+    The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
+    The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
+
+    The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
+    THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
+    They then go through tasks and map than to such processors that minimize the earliest start times of this task's
+    successors.
+    To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
+    average communication and computation costs fir the dependent tasks.
+
+    The rescheduling is being triggered when either a load on a resource increases over a threshold, or if a new resource
+    is added.
+    The algorithm produces a new mapping from scrath, and this mapping is being accepted if the resulting maespan is
+    smaller than the previously predicted one.
 \end{document}

### 2024-05-05 main.tex
related work - 1st draft, WIP

@@
 
+    %%% CONTEXT %%%
     Only mapping tasks to processors is not enough.
@@
 
+    %%% MOTIVATION %%%
     HEFT is one of the most popular heuristics (cite a review).
@@
 
+    %%% CONTRIBUTION %%%
     In this paper, we formulate the problem that takes memories into consideration.
@@
 
-    \section{Related work}
-    \skug{problem: this is written for the old heuristic - parititioning-based heuristic to solvea dynamic problem! it is irrelevant to heft}
+    \section{Related work: new}
 
-    In the limited time I spent looking, I found no paper addressing the exact same problem.
+    \paragraph{HEFT-based algorithms}
 
-    Wang et al.~\cite{wang2019dynamic} proposes a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
-    Particles are possible solution in the solution space.
-    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
-    Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
-    However, the approach is for clouds.
-    \xspace
+    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} has become a base for multiple heuristics in workflow (~\cite{SHI2006665, SANDOKJI2019482}) scheduling
+    and in cloud-oriented environments(\cite{samadi2018eheft}).
+    HEFT is even combined with reinforcement learning techniques(\cite{yano2022cqga}).
 
-    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
-    on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
-    that these runtime changes are unrelated to each other).
-    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
-    performance measure $\phi$.
-    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
+    HEFT-based or list-based heuristics are more applicable to very large workflows in comparison to greedy
+    or exact algorithms(~\cite{de2012provenance},~\cite{rahman2013}).
 
-    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling heuristic
-    for clouds.
-    Based on a 3-objective cost model, their greedy scheduling algorithm schedules tasks into machines.
-    The authors use provenance data to make scheduling decisions.
-    The cost model is a set of linear equations computed in the simulation environment, it represents the cost of an execution based on the criteria.
-    In the algorithms, the authors test out 4 scenarios - one preferring each criteria, and a balanced one.
-    The algorithm merely chooses the best virtual machine for each next task based on the cost given by the model.
-    An additional algorithm combines several cloud activities (task executions) into one to improve the cost, so that each
-    execution is not too small (utilize the granularity factor, the smallest entity of payment of the cloud producer).
-    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
-    They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
+    \paragraph{Memory-aware scheduling algorithms}
+    Incorporating processor memory sizes into the scheduling problem is another layer of complexity in a scheduling problem.
 
-    Rahman~\etal~\cite{rahman2013} propose a scheduling heuristic for grids that proposes mapping of tasks to machines by calculating
-    the critical path in the graph dynamically at every step.
-    They call it the dynamic critical path (DCP).
-    For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
-    on the start time of a task (differing by the slack this task has).
-    All tasks on this critical path have the same earliest and latest start times, because they cannot be ddelayed.
+    Different models of memory available on processors and memory requirements of tasks.
+    The most straightforward memory representations assume no memory requirement on tasks of the workflow and single
+    memories on each processor(\eg~\cite{rodriguez2019exploration}).
+    In an assumed dual-memory systems(\cite{herrmann2014memory}) a processor can have access to a memory of two different kinds,
+    and each task can be executed on one sort of memory.
+    Cloud-oriented model can include costs associated with memory usage(\cite{liang2020memory}).
 
-    The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
-    If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum execution time for
-    the task is chosen.
-    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
-    The authors evaluate their approaches on random workflows of the size up to 300 tasks.
-
-    The authors provide an overview over (simpler) scheduling heuristics.
-    For example, GRASP (generally randomized adaptive search procedure) conducts a number of iterations to search an optimal
-    solution for mapping tasks on machines.
-    A solution is generated at each step, and the best solution is kept at the end.
-    The search terminates when a certain termination criterion is reached.
-    It generates better results than other algorithms, because it explores the whole solution space.
-
-    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
-    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
-    The scenario contains AND- and OR-branches, where AND-branches indicate acitivities that need to be executed in parallel.
-
-    The heuristic first partitions the set into local schedules by using affinity matrices to determine similar
-    activities and group them together.
-    Then they physically allocate these partitions to groups of disaster responders and tasks within this group.
-    They define a constraint system for that.
-    The dynamic part deals with changes and distinguishes between retriable and compensation acitivities.
-    The heuristic calculates a new execution path with these tasks.
-    In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
-
-    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
-    The aim is to minimize the makespan, and the experiments were conducted on a single wotkflow with 10 tasks.
-    The procesdure involves building a first (static) schedule, priodic resource monitoring and rescheduling the remaining
-    tasks.
-    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
-    For each resuorce group, there is an own scheduler, and an overall global scheduler responsible for distributing
-    tasks to groups.
-    The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
-    The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
-
-    The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
-    THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
-    They then go through tasks and map than to such processors that minimize the earliest start times of this task's
-    successors.
-    To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
-    average communication and computation costs fir the dependent tasks.
-
-    The rescheduling is being triggered when either a load on a resource increases over a threshold, or if a new resource
-    is added.
-    The algorithm produces a new mapping from scrath, and this mapping is being accepted if the resulting maespan is
-    smaller than the previously predicted one.
+    The algorithm presented by Yao\etal~\cite{yao2022memory} looks similar to our heuristics on the first glance, but the
+    model the authors adopt differs from the one assumed in this paper.
+    The biggest difference is in the way processor memory is represented.
+    Each processor has an own internal memory and all processors share a common external one.
+    The internal (local) memory is used to store the task files.
+    The external memory is used to store evicted files to make room for the execution of a task on a processor.
+    However, this marks the decisive difference to our approach: once the files have been evicted into the communication
+    buffer in our approach, they can only be transferred to another processor.
+    In~\cite{yao2022memory} however, the same processor can evict files and then take them back.
+    This makes the approach not directly comparable with ours.
+    Other differences in the model include edge weights in the workflow graph.
+    Each edges in~\cite{yao2022memory} has two weights - the size of the files transferred along it,
+    and the time of communication along this edge, while we assume a singular weight that represents file size and also
+    determines the transfer time (along with the bandwidth).
+    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
+    They assume that the {\em minimum} local memory should be bigger than the memory requirement of the largest task.
 

### 2024-05-05 main.tex
unified the model

@@
 
-    \skug{TODO:unify the notation with the table!!}
+    We first describe the target applications, which are (large scientific) workflows,
+    in Section~\ref{sec.mod.work}.  Next, we define the execution
+    environment, a heterogeneous system (in terms of processor speed and memory size),
+    in Section~\ref{sec.mod.plat}.
 
-    For a task~$u\in V$, we have a memory usage~$m_u$ and execution time~$w_u$.
-    For an edge $(u,v)\in E$, we have a data of size $c_{u,v}$.
+    \subsection{Workflow}
+    \label{sec.mod.work}
+    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
+    $E$ is a set of directed edges of the form $e=(u,v)$, with $u,v\in V$, expressing precedence constraints between tasks.
+    Each task~$u \in V$  is performing $w_u$ operations, and it also
+    requires some amount of memory to be executed, denoted as~$m_u$.
+    Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.
 
-    For each processor $P_j$, we have a speed~$s_j$, and also two memory bounds:
-    $\MM_j$ the processor memory, and $\MC_j$, the size of the communication buffer.
+    Hence, the total  memory requirement for the execution of task~$u$ consists of the input files
+    (size of the files to be received from the parents),
+    the output files (size of the files to be sent to the children),
+    and the memory size~$m_u$:
+    \[
+        r_u = m_u + \sum_{v:(v,u)\in E}c_{v,u} + \sum_{v:(u,v)\in E} c_{u,v}.
+    \]
+
+    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
+    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without parents is called a {\it source task}.
+    The children tasks of~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
+    $ \children{u} = \{v \in V: (u,v) \in E\}$. A task without children is called a {\it target task}.
+    Each task may have multiple parents and children.
+
+    \subsection{Execution environment}
+    \label{sec.mod.plat}
+
+    The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
+    consists of $k$ processors $p_1, \dots, p_k$.
+    Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory size $M_j$, a communication
+    buffer size $MC_j$ and a speed~$s_j$.
     We can decide to evict some data from the main memory if we are sending the data
     to another processor; it then stays in the communication buffer until it has been sent.
+    The execution time of a single task~$u\in V$ on a processor~$p_j$ is expressed as $\frac{w_u}{s_j}$.
+    We assume that all processors are connected with the same bandwidth~$\beta$.
 
     We keep track of the current ready time of each processor and each communication
-    channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$. Initially, all the ready times
-    are set to~$0$.
-
+    channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$.
+    Initially, all the ready times are set to~$0$.
     We also keep track of the currently available memory, $availM_j$ and $availC_j$,
@@
     They correspond to some $c_{u,v}$'s.
+
+    We use the \algo{memDag} algorithm from~\cite{KAYAASLAN20181} to compute
+    the memory requirement by transforming the workflow into a series-parallel graph,
+    and then finding the traversal that leads to the minimum memory consumption.
 
@@
                 $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
-                $M_j$, $s_j$                               & Memory size and speed of processor $p_j$                          \\
+                $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer and speed of processor $p_j$                          \\
                 $\beta$                     & Bandwidth in the compute system                                \\
@@
                 $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
-                $r_{\max}$                       & Maximum memory requirement in a workflow                              \\
-                $\criticalpath$                       & Critical path in a workflow                              \\
                 \hline

### 2024-05-06 main.tex
Added Results:memory usage

@@
     \subsubsection{Memory usage}
+    Minimizing memory usage is an interesting measure of the quality of the solution produced by a heuristic.
+    We studied the average peak memory consumption in our cluster.
+    We looked at the maximum memory usage on each processor during a task execution, and averaged it over all processors
+    in the execution environment.
+
+    The results adhere to the intuition: \heftmm produces consistently smaller memory usage among all the heuristics.
+    So, for small workflows, the average memory occupation by \heftbl and \heftblc is $11.8$\% and $11.7$\%, respectively,
+    while \heftmm manged to occupy $10.2$\% of memories.
+    For middle-sized workflows, \heftbl and \heftblc occupied $27.8$\% and $28.2$\%, respectively, while \heftmm occupied
+    $21$\%.
+    For the big workflows, \heftmm occupied $20.9$\% against $33.7$\% by \heftbl and $27.6$\% by \heftblc.
+    Interestingly, for big workflows, \heftmm occupied on average even a little less that for the middle-sized ones.
+    This means that the bound of exploiting parallelization has been achieved for workflows around the size of middle-sized ones,
+    and for the big workflows, the makespan is not improved by ``spreading'' the workflow more.
+
     \subsubsection{Impact of cluster size}

### 2024-05-07 main.tex
description of 3 pc cluster

@@
     $\bullet$ {\em Small} and {\em large} clusters. While the default cluster consists of 36 nodes (6 of each kind), we
-    consider a small cluster with one processor of each kind ($6$ processors in total), and a
+    consider a small cluster with three processors of each kind ($18$ processors in total), and a
     large cluster with ten processors of each kind ($60$ processors in total). \skug{leaving the big cluster inside for
@@
     \subsubsection{Impact of cluster size}
-    On a cluster with only 6 processors (1 of each kind), the baseline makes more failures: from $43.4$\%  (small workflows)
-    to $88.9$\% (mid-sized) and $94.4$\% (big workflows).
-    These values also become more unattainable: the makespans heuristics produce are worse by $3-12$\% (\heftbl) to $12-52$\%
-    (\heftmm).
-    However, these comparisons are less expressive, because most of baseline solutions are invalid.
-    \skug{TODO: add numbers about failures on the small cluster with the fixed bug}
+    On a cluster with only $18$ processors ($3$ of each kind), the baseline makes more failures: from $43.8$\%  (small workflows)
+    to $77.8$\% (mid-sized) and $88.9$\% (big workflows).
+    At the same time, each heuristic is unable to schedule one big workflow and one middle-sized one.
+    On the workflows that they are able to schedule, \heftbl and \heftblc produce makespans that are $<1$\% bigger than
+    those of the baseline on small and middle-sized workflows.
+    On the big workflows, they produce makespans that are $3.5$\% and $2.3$\% bigger respectively.
+    The performance of \heftmm improves more with workflow size: it goes from producing $16$\% bigger makespans on small
+    workflows, to $8$\% on middle-sized ones, to $7$\% on big workflows.
+
+   Like in the normal cluster, \heftmm again consistently shows smaller peak memory usage in the cluster.
+    It utilizes on average $15$\% of memory on small workflows ($17.7$\% for \heftbl and \heftblc),
+    $31$\% on middle-sied workflows ($45.2$\% and $44.9$\% on \heftbl and \heftblc, respectively),
+    and $34$\% on big workflows ($50.1$\% and $50.2$\% for \heftbl and \heftblc).
 

### 2024-05-07 main.tex
model edited a bit

@@
     Hence, the total  memory requirement for the execution of task~$u$ consists of the input files
-    (size of the files to be received from the parents),
-    the output files (size of the files to be sent to the children),
+    (total size of the files to be received from the parents),
+    the output files (total size of the files to be sent to the children),
     and the memory size~$m_u$:
@@
     consists of $k$ processors $p_1, \dots, p_k$.
-    Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory size $M_j$, a communication
-    buffer size $MC_j$ and a speed~$s_j$.
+    Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$, a communication
+    buffer of size $MC_j$ and a speed~$s_j$.
     We can decide to evict some data from the main memory if we are sending the data
@@
     We also keep track of the currently available memory, $availM_j$ and $availC_j$,
-    on respectively the processor memory and communication buffer.
+    on the processor memory and communication buffer, respectively.
     Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
-    that are in the memory of size $\MM_j$ but may be evicted to be communicated, if
-    more memory is needed on~$p_j$. They are ordered by non-decreasing size.
-    They correspond to some $c_{u,v}$'s.
+    that are in the memory of size $\MM_j$ but may be evicted to be communicated if
+    more memory is needed on~$p_j$. They are ordered by non-decreasing size and
+    correspond to some $c_{u,v}$.
 
-    We use the \algo{memDag} algorithm from~\cite{KAYAASLAN20181} to compute
-    the memory requirement by transforming the workflow into a series-parallel graph,
-    and then finding the traversal that leads to the minimum memory consumption.
+    We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
+    the memory requirement; it transforms the workflow into a series-parallel graph
+    and then finds the traversal that leads to the minimum memory consumption.
 
@@
                 $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
-                $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer and speed of processor $p_j$                          \\
+                $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer size, and speed of proc.\ $p_j$                          \\
                 $\beta$                     & Bandwidth in the compute system                                \\
@@
         \end{center}
-        \caption{Notation.} \label{tabnotation}
+        \caption{Notation} \label{tabnotation}
     \end{table}

### 2024-05-07 main.tex
rel. work edited, needs more work

@@
 \usepackage{numprint}
+\usepackage{todonotes}
 
@@
 %should not be used}
-    \thanks{Identify applicable funding agency here. If none, delete this.}
+    % \thanks{Identify applicable funding agency here. If none, delete this.}
     }
@@
 
-    The algorithm presented by Yao\etal~\cite{yao2022memory} looks similar to our heuristics on the first glance, but the
+    The algorithm presented by Yao \etal~\cite{yao2022memory} looks similar to our heuristics on the first glance, but the
     model the authors adopt differs from the one assumed in this paper.
     The biggest difference is in the way processor memory is represented.
-    Each processor has an own internal memory and all processors share a common external one.
+    In their model, each processor has an own internal memory and all processors share a common external one.
     The internal (local) memory is used to store the task files.
     The external memory is used to store evicted files to make room for the execution of a task on a processor.
-    However, this marks the decisive difference to our approach: once the files have been evicted into the communication
+    Yet, and this marks the decisive difference to our approach: once the files have been evicted into the communication
     buffer in our approach, they can only be transferred to another processor.
-    In~\cite{yao2022memory} however, the same processor can evict files and then take them back.
+    In~\cite{yao2022memory}, in turn, the same processor can evict files and then take them back.
     This makes the approach not directly comparable with ours.
     Other differences in the model include edge weights in the workflow graph.
-    Each edges in~\cite{yao2022memory} has two weights - the size of the files transferred along it,
-    and the time of communication along this edge, while we assume a singular weight that represents file size and also
+    Each edge in~\cite{yao2022memory} has two weights -- the size of the files transferred along it,
+    and the time of communication along this edge, while we assume a single weight that represents file size and also
     determines the transfer time (along with the bandwidth).
@@
     They assume that the {\em minimum} local memory should be bigger than the memory requirement of the largest task.
+
+\todo[inline]{TBD (HM): Still, wouldn't an empirical comparison make sense from a reviewer's point of view?}
 

### 2024-05-07 main.tex
first picture about the normal cluster

--- 2024-05-07 main.tex
@@
 
-    On the normal cluster and among small workflows, the makespans produced by \heftbl and \heftblc are $1$\% worse than the baseline,
-    \heftmm $19$\% worse.
-    For middle-sized ones \heftmm is $12.3$\% worse, the rest is also $1-2$\% worse.
-    For the big workflows however, the results of \heftbl and \heftblc are $<1$\% worse and \heftmm is only $9$\% worse.
+    The excess makespans produced by the heuristics are shown in Figure~\ref{fig:excess-ms}.
+    On the normal cluster and among small workflows, the makespans produced by \heftbl and \heftblc are virtually indistinguishable
+    from those of the baseline (no change and $0.4$\% change), while \heftmm produces  $18$\% worse makespans.
+    For middle-sized ones \heftmm is $9$\% worse,  \heftblc $0.9$\% worse.
+    For the big workflows however, while results of \heftbl and \heftblc are still barely worse ($<1$\% worse),
+     \heftmm is only $5.8$\% worse.
     The heuristics pay off more on larger workflows.
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=0.495\columnwidth] {images/ms-relative-3groups}
+        \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm for different . Smaller is better.}
+       \label{fig:excess-ms}
+        \vspace{-0.3cm}
+    \end{figure}
     \subsubsection{Memory usage}

### 2024-05-15 main.tex
reinserted text about dynamic cases, marked fullness and polishedness on chapters

@@
     \begin{abstract}
+        \skug{Fullness score (how much text is available): 0-none to 5 - everything.
+
+        Polishedness scroe (how well-written is the chapter): 0 messy - 5 very clean.
+        }
         TODO: Insert abstract
@@
 
-    \section{Introduction}
+    \section{Introduction: \skug{Full: 0, Polished: 0}}
 
@@
 
-    \section{Model}
+    \section{Model: \skug{Full: 4, polished: 3}}
 
@@
 
-    \section{Related work: new}
+    \paragraph{Workflow-related changes}
+
+    \begin{itemize}
+        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
+        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
+
+    \end{itemize}
+
+    The following changes are not a part of this article's scope:
+
+    \begin{itemize}
+        \item The workflow structure changes: edges or tasks come in or leave.
+    \end{itemize}
+
+    \paragraph{Execution environment-related changes }
+
+    \begin{itemize}
+        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
+        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
+
+    \end{itemize}
+
+    The following changes are not a part of this article's scope:
+
+    \begin{itemize}
+        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
+    \end{itemize}
+
+    \subsection{Time of changes }
+
+    We consider discrete time in seconds.
+    The time point(s) at which the changes happen is unambiguously defined.
+
+    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
+    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
+    The start time of the source task in the workflow is zero.
+    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
+
+    \subsection{Changes and knowledge horizon - important questions TBA}
+
+    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
+
+    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
+    However, how do we model the following:
+    \begin{itemize}
+        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
+This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
+        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
+        \item How many times of change do we model - one per workflow run, or multiple?
+        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
+    \end{itemize}
+
+    \section{Related work: \skug{Full: 4, Polished: 0}}
 
@@
 
-    \section{Proposal of a new heuristic with slightly refined model}
+    \paragraph{Dynamic case}
+    Wang et al.~\cite{wang2019dynamic} proposes a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
+    Particles are possible solution in the solution space.
+    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
+    Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
+    However, the approach is for clouds.
+    \xspace
+
+    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
+    on a single machine.
+    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
+    that these runtime changes are unrelated to each other).
+    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
+    performance measure $\phi$.
+    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
+
+    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling heuristic
+    for clouds.
+    Based on a 3-objective cost model, their greedy scheduling algorithm schedules tasks into machines.
+    The authors use provenance data to make scheduling decisions.
+    The cost model is a set of linear equations computed in the simulation environment, it represents the cost of an execution based on the criteria.
+    In the algorithms, the authors test out 4 scenarios - one preferring each criteria, and a balanced one.
+    The algorithm merely chooses the best virtual machine for each next task based on the cost given by the model.
+    An additional algorithm combines several cloud activities (task executions) into one to improve the cost, so that each
+    execution is not too small (utilize the granularity factor, the smallest entity of payment of the cloud producer).
+    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
+    They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
+
+    Rahman~\etal~\cite{rahman2013} propose a scheduling heuristic for grids that proposes mapping of tasks to machines by calculating
+    the critical path in the graph dynamically at every step.
+    They call it the dynamic critical path (DCP).
+    For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
+    on the start time of a task (differing by the slack this task has).
+    All tasks on this critical path have the same earliest and latest start times, because they cannot be ddelayed.
+
+    The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
+    If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum execution time for
+    the task is chosen.
+    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
+    The authors evaluate their approaches on random workflows of the size up to 300 tasks.
+
+    The authors provide an overview over (simpler) scheduling heuristics.
+    For example, GRASP (generally randomized adaptive search procedure) conducts a number of iterations to search an optimal
+    solution for mapping tasks on machines.
+    A solution is generated at each step, and the best solution is kept at the end.
+    The search terminates when a certain termination criterion is reached.
+    It generates better results than other algorithms, because it explores the whole solution space.
+
+    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
+    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
+    The scenario contains AND- and OR-branches, where AND-branches indicate acitivities that need to be executed in parallel.
+
+    The heuristic first partitions the set into local schedules by using affinity matrices to determine similar
+    activities and group them together.
+    Then they physically allocate these partitions to groups of disaster responders and tasks within this group.
+    They define a constraint system for that.
+    The dynamic part deals with changes and distinguishes between retriable and compensation acitivities.
+    The heuristic calculates a new execution path with these tasks.
+    In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
+
+    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
+    The aim is to minimize the makespan, and the experiments were conducted on a single wotkflow with 10 tasks.
+    The procesdure involves building a first (static) schedule, priodic resource monitoring and rescheduling the remaining
+    tasks.
+    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
+    For each resuorce group, there is an own scheduler, and an overall global scheduler responsible for distributing
+    tasks to groups.
+    The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
+    The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
+
+    The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
+    THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
+    They then go through tasks and map than to such processors that minimize the earliest start times of this task's
+    successors.
+    To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
+    average communication and computation costs fir the dependent tasks.
+
+    The rescheduling is being triggered when either a load on a resource increases over a threshold, or if a new resource
+    is added.
+    The algorithm produces a new mapping from scrath, and this mapping is being accepted if the resulting maespan is
+    smaller than the previously predicted one.
+
+    \section{Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
 
@@
 
+    \subsection{Dynamic Scenario}
+
+    In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
+    This information may include, memory usage, runtime, graph structures, or the status of the underlying infrastructure.
+    In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
+    By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
+    Also the underlying infrastructure can change during the workflow execution.
+    Examples are processor failures, node recoveries, or acquisition of new nodes.
+    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
+    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.
+
+    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{}.
+    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
+%For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
+%Bader~et~al.~ report a prediction error between 13\% and 17\% for their method, showing an exponential task error distribution.
+% @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
+    Such a dynamic execution environment necessitates for a dynamic scheduling method where the schedule can be recomputed during the workflow execution.
+
+    \subsection{Retracing the effects of change on an existing schedule}
+    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
+    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
+    they can lead to a later finishing time (\eg if some tasks longer and delay other tasks), or they can have no effect (\eg if new processors
+    joined the cluster, but the old schedule did not account for them).
+    To assess the impact, we need to retrace the schedule.
+
+    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
+    entire schedule.
+
+    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
+    is a topologial ordering.
+    We then repeat steps similar to those we did during tentative assignment in our heuristic, except we do not choose a processor
+    anymore, but rather see if the current one still fits.
+
+    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from our heuristic.
+    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ or $c_{v,w}$ from successors $w$,
+    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements from other tasks).
+    If originally,$Res$ was positive (no files were evicted from memory into the communication buffer), then it has to stay this way -
+    otherwise evicted files can invalidate next tasks.
+    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
+    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
+    We update the $availM_j$ and $availMC_j$ according to the new memory constraints.
+
+    Then we can re-calculate the finish time of the task on its processor like in Step 3.
+    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
+    (after delayed previous tasks), and changed communication buffer availability.
+
+    Then, after having updated the processor's values, we move on to the next task.
+
     \subsection{Approximation}
@@
 
-    \section{Experimental evaluation}
-    \subsection{Setup}
+    \section{Experimental evaluation: \skug{Full: 3, Polished: 3}}
+    \subsection{Setup: \skug{Full: 5, Polished: 4}}
     \label{sec:setup}
@@
 
-    \subsection{Results}
+    \subsection{Results \skug{Full: 3, Polished: 2}}
     \subsubsection{Normal cluster}
@@
 
-    \section{Conclusion}
+    \section{Conclusion  \skug{Full: 0, Polished: 0}}
 

### 2024-07-24 main.tex
literature search for competitor

@@
 
-        Polishedness scroe (how well-written is the chapter): 0 messy - 5 very clean.
+        Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
         }
@@
 
-    \paragraph{HEFT-based algorithms}
+    We discuss relevant scheduling approaches that either reuse processors, or respect memory requirements of the processors.
 
-    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} has become a base for multiple heuristics in workflow (~\cite{SHI2006665, SANDOKJI2019482}) scheduling
-    and in cloud-oriented environments(\cite{samadi2018eheft}).
-    HEFT is even combined with reinforcement learning techniques(\cite{yano2022cqga}).
+    \subsection{Early list schedulers with unlimited processors}
+    An entire cluster of works on list schedulers has been carried out as early as the 90s.
+    They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
+    with the speed of 1.
 
-    HEFT-based or list-based heuristics are more applicable to very large workflows in comparison to greedy
-    or exact algorithms(~\cite{de2012provenance},~\cite{rahman2013}).
+    The \textit{task duplication}-based approaches note that sometimes running a task twice on different machines can
+    help reduce the makespan by saving communication costs.
+    The two approaches are scheduling with the partial duplication~(SPD), and withe the full duplication~(SFD).
+    For a join task (a task whose incoming degree is larger than its outgoing degree), SPD finds a critical immediate
+    parent (the one that gives the largest start time to the join task) and duplicate only it.
+    SFD duplicates all parents of a join node.
+    \cite{dfrn1997} is a task duplication algorithm.
+    The authors duplicate first (create copies of all parent tasks) and then reduce (remove) the ones that can be removed without harming
+    the makespan.
+    The critical path fast duplication algorithm \texttt{CPFD}~\cite{5727760} classifies tasks into three categories: a critical
+    path task, and in-branch task, or out-branch task.
+    It schedules critical path tasks first, then in-branch tasks.
 
-    \paragraph{Memory-aware scheduling algorithms}
-    Incorporating processor memory sizes into the scheduling problem is another layer of complexity in a scheduling problem.
+    Linear clustering~\cite{KWOK1999381} acts on critical paths in the workflow.
+    It assigns the current critical path to one processor, removes all these tasks from the workflow, recomputes the critical
+    path and repeats the procedure.
+    \texttt{Heaviest node first} algorithm~\cite{SHIRAZI1990222} assigns the tasks level by level, and in each level,
+    schedules the heaviest (with largest computation time)  ones first.
 
-    Different models of memory available on processors and memory requirements of tasks.
-    The most straightforward memory representations assume no memory requirement on tasks of the workflow and single
-    memories on each processor(\eg~\cite{rodriguez2019exploration}).
-    In an assumed dual-memory systems(\cite{herrmann2014memory}) a processor can have access to a memory of two different kinds,
-    and each task can be executed on one sort of memory.
+    \subsection{Static list schedulers, especially HEFT-based algorithms}
+
+    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
+    It and all its successors consist of two phases: task ordering, and task assignment.
+    In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
+    and then schedule tasks in the order of these priorities.
+    The modifications of HEFT revolve around the way the priorities of tasks are computed, and the logic
+    of processor assignment.
+    All such algorithms assume a heterogeneous execution environment.
+
+    So, during task prioritization phase~\cite{sulaiman2021hybrid} compute the standard deviation of the computation cost
+    (between processors) and add it to the mean value to account for the differences between processor speeds.
+    In the processor choice phase, they duplicate the entry task and the longest parent tasks during idle times on the processor.
+
+    \cite{alebrahim2017task} compute a the bottom level based on the difference of execution times on
+    the fastest and the slowest processors, divided by the speedup (speeds ratio) of these two processors.
+    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
+    They choose the processor with the lowest execution time, and cross over to other processors sometimes.
+    They build upon~\cite{shetti2013optimization}.
+
+    PEFT (Predict earliest finish time) algorithm~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
+    Cost Table OCT.
+    OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task, if this
+    processor is chosen for this task.
+    Ranking is based on OCT values.
+    Processor choice stage minimizes the Optimistic EFT, which is EFT plus the the longest path to the exit node for each task.
+
+    The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
+    comparison to HEFT.
+    It combines the standard deviation with the communication cost weight on the tasks.
+    In the second stage, the algorithm duplicates the entry task if there is a need for it.
+
+    The TSHCS (Task Scheduling for Heterogeneous Comupting Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
+    by adding randomized decisions to the second phase.
+    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
+    produces the lowest finish time.
+
+    The \texttt{SDC} algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
+    average execution cost in its weight.
+    The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT) that
+    additionally notes how large the communication between current node and its children will be on the
+    average provided that it is scheduled on the current processor.
+
+    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement
+    learning techniques(\cite{yano2022cqga}).
+
+    \subsection{Memory-aware scheduling algorithms}
+    Respecting processor memories adds a constraint to a scheduling problem.
+    Therefore, only specifically memory-targeted algorithms address this issue.
+    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
+    is formulated and addressed in the algorithm.
+
+    Different models of memory available on processors and memory requirements of tasks have been presented.
+
+    Marchal~et~al.~\cite{marchal2018parallel} assume a memory model where each processor has an individual memory available.
+    Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
+    A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a DAG is provided,
+    as well as an ILP solution to the scheduling problem.
+    The memory model requires deleting all input data upon starting of the task and adding all output files there.
+
+    In an assumed dual-memory systems(\cite{herrmann2014memory}) a processor can have access to a memory of two different
+    kinds (red or blue), and each task can be executed on only one sort of memory.
+    The communications happen only between these two kinds of processors (communications withing each group are ignored).
+    The authors then formulate an ILP solution for this problem formulation.
+
+    The algorithm presented by Yao \etal~\cite{yao2022memory}, each processor has an own internal memory and all
+    processors share a common external one. The internal (local) memory is used to store the task files.
+    The external memory is used to store evicted files to make room for the execution of a task on a processor.
+    All processors, including the original one, can access these files.
+    Each edge in~\cite{yao2022memory} has two weights -- the size of the files transferred along it,
+    and the time of communication along this edge.
+    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
+
+    \cite{ding2024ils} assumes connected processors with individual limited memories.
+    The collective set of memories forms the global memory, to which each processor has access, however the access time
+    to global memory is different.
+    Each memory access in the graph is modelled as a memory access token on the task, while the edges have no weights.
+    The solved problem is how to allocate initial input data in processor memories so that the overall
+    execution is minimized and the memories are not exceeded.
+    They propose an integer linear programming model.
+    %that minimizes the length of the critical path, including a greedy initial solution.
+
+    In~\cite{rodriguez2019exploration}, the authors assume memory requirement on tasks represented as tiles.
+    Each processor has individual memories to process the task, but only the shared memories store the tiles containing
+    memory tiles occupied by memory tiles.
+
     Cloud-oriented model can include costs associated with memory usage(\cite{liang2020memory}).
 
-    The algorithm presented by Yao \etal~\cite{yao2022memory} looks similar to our heuristics on the first glance, but the
-    model the authors adopt differs from the one assumed in this paper.
-    The biggest difference is in the way processor memory is represented.
-    In their model, each processor has an own internal memory and all processors share a common external one.
-    The internal (local) memory is used to store the task files.
-    The external memory is used to store evicted files to make room for the execution of a task on a processor.
-    Yet, and this marks the decisive difference to our approach: once the files have been evicted into the communication
-    buffer in our approach, they can only be transferred to another processor.
-    In~\cite{yao2022memory}, in turn, the same processor can evict files and then take them back.
-    This makes the approach not directly comparable with ours.
-    Other differences in the model include edge weights in the workflow graph.
-    Each edge in~\cite{yao2022memory} has two weights -- the size of the files transferred along it,
-    and the time of communication along this edge, while we assume a single weight that represents file size and also
-    determines the transfer time (along with the bandwidth).
-    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
-    They assume that the {\em minimum} local memory should be bigger than the memory requirement of the largest task.
+    \subsection{Dynamic/adaptive algorithms}
 
-\todo[inline]{TBD (HM): Still, wouldn't an empirical comparison make sense from a reviewer's point of view?}
+    DVR HEFT~\cite{SANDOKJI2019482} uses an almost unchanged HEFT algorithm in their static step, executing three slightly
+    varying variants of task weighting and choosing the variant that gives the best overall makespan.
+    In the dynamic phase, they receive new tasks and schedule them on either idle processors or those processors that give them
+    the earliest finish time.
+    Task failures are not covered.
 
-    \paragraph{Dynamic case}
-    Wang et al.~\cite{wang2019dynamic} proposes a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
+    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
+    by calculating the critical path in the graph dynamically at every step.
+    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
+    %on the start time of a task (differing by the slack this task has).
+    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
+    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
+    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
+    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
+    %execution time for the task is chosen.
+    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
+    The approach is evaluated on random workflows of the size up to 300 tasks.
+
+    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
+    The procedure involves building a first (static) schedule, periodic resource monitoring and rescheduling the remaining
+    tasks.
+    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
+    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
+    tasks to groups.
+    The static heuristic is HEFT with earliest start time as priority.
+    Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
+    is smaller than the previous one.
+    The experiments were conducted on a single workflow with 10 tasks.
+%
+    % The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
+    %The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
+
+    %The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
+    %THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
+    %They then go through tasks and map than to such processors that minimize the earliest start times of this task's
+    %successors.
+    %To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
+    %average communication and computation costs fir the dependent tasks.
+    %
+    %The rescheduling is triggered when either a load on a resource increases over a threshold, or if a new resource
+    %is added.
+
+    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
+    but constrained by cost.
+
+    Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
     Particles are possible solution in the solution space.
     However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
-    Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
-    However, the approach is for clouds.
-    \xspace
+    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
+
+    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling algorithm
+    for clouds.
+    They solve a set of linear equations that represent the cost of an execution based on the criteria.
+    The authors test out 4 scenarios - one preferring each criteria, and a balanced one.
+    The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
+    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
+    %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
+ %   The authors use provenance data to make scheduling decisions.
 
@@
 
-    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling heuristic
-    for clouds.
-    Based on a 3-objective cost model, their greedy scheduling algorithm schedules tasks into machines.
-    The authors use provenance data to make scheduling decisions.
-    The cost model is a set of linear equations computed in the simulation environment, it represents the cost of an execution based on the criteria.
-    In the algorithms, the authors test out 4 scenarios - one preferring each criteria, and a balanced one.
-    The algorithm merely chooses the best virtual machine for each next task based on the cost given by the model.
-    An additional algorithm combines several cloud activities (task executions) into one to improve the cost, so that each
-    execution is not too small (utilize the granularity factor, the smallest entity of payment of the cloud producer).
-    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
-    They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
+    \subsection{Other notable works}
 
-    Rahman~\etal~\cite{rahman2013} propose a scheduling heuristic for grids that proposes mapping of tasks to machines by calculating
-    the critical path in the graph dynamically at every step.
-    They call it the dynamic critical path (DCP).
-    For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
-    on the start time of a task (differing by the slack this task has).
-    All tasks on this critical path have the same earliest and latest start times, because they cannot be ddelayed.
+    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
+    They utilize task duplication when creating the clusters (grains).
+    Their scheduler then maps clusters to processors.
+    They assume unlimited processors with the speed of 1.
+    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
+    to it as possible.
+    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
+    %release times.
 
-    The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
-    If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum execution time for
-    the task is chosen.
-    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
-    The authors evaluate their approaches on random workflows of the size up to 300 tasks.
-
-    The authors provide an overview over (simpler) scheduling heuristics.
-    For example, GRASP (generally randomized adaptive search procedure) conducts a number of iterations to search an optimal
-    solution for mapping tasks on machines.
+    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
+    to search for an optimal solution for mapping tasks on machines.
     A solution is generated at each step, and the best solution is kept at the end.
@@
     These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
-    The scenario contains AND- and OR-branches, where AND-branches indicate acitivities that need to be executed in parallel.
+    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
+    The heuristic first determines similar activities and groups them together.
+    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
+    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
+    The heuristic calculates a new execution path with these tasks.
 
-    The heuristic first partitions the set into local schedules by using affinity matrices to determine similar
-    activities and group them together.
-    Then they physically allocate these partitions to groups of disaster responders and tasks within this group.
-    They define a constraint system for that.
-    The dynamic part deals with changes and distinguishes between retriable and compensation acitivities.
-    The heuristic calculates a new execution path with these tasks.
-    In general, this heuristic is not as related as it looks, because of very specific workflow and task structure.
+    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
+    (processor) speed heterogeneity.
+    Each accelerator has its own memory that can be zero.
+    Each accelerator's characteristics depend on the task it runs and are not fixed.
 
-    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
-    The aim is to minimize the makespan, and the experiments were conducted on a single wotkflow with 10 tasks.
-    The procesdure involves building a first (static) schedule, priodic resource monitoring and rescheduling the remaining
-    tasks.
-    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
-    For each resuorce group, there is an own scheduler, and an overall global scheduler responsible for distributing
-    tasks to groups.
-    The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
-    The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
+    \cite{meng2018traffic} investigate scheduling on multi-core chips.
+    Their model is far from ours.
 
-    The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
-    THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
-    They then go through tasks and map than to such processors that minimize the earliest start times of this task's
-    successors.
-    To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
-    average communication and computation costs fir the dependent tasks.
-
-    The rescheduling is being triggered when either a load on a resource increases over a threshold, or if a new resource
-    is added.
-    The algorithm produces a new mapping from scrath, and this mapping is being accepted if the resulting maespan is
-    smaller than the previously predicted one.
+    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
+    They prioritize tasks that have failed before or are well-predictable.
 
@@
 
-    \bibliographystyle{IEEEtran}
+    \bibliographystyle{apalike}
     \bibliography{references}

### 2024-09-30 main.tex
abstract - version 0.0

@@
     \begin{abstract}
-        \skug{Fullness score (how much text is available): 0-none to 5 - everything.
 
-        Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
-        }
-        TODO: Insert abstract
+        Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
+        They are often represented as directed acyclic graphs (DAGs),
+        where vertices correspond to tasks and edges represent the dependencies between them.
+       % Since these graphs are often large in both the number of tasks and their resource requirements, it is important
+       % to schedule them efficiently on parallel or distributed compute systems.
+        Typically, each task requires a
+        certain amount of memory to be executed and needs to communicate data to its successor tasks.
+        The goal is thus to execute the workflow  as fast as possible (i.e., to minimize its makespan) while satisfying the memory constraints.  %and using parallelism to speed up the execution.
+
+        Hence, we investigate adaptive memory-aware scheduling  of DAG-shaped workflows onto heterogeneous platforms where
+        each processor can have a different speed and a different memory size.
+        We propose a two-step HEFT-inspired heuristic.
+        Its first step is to compute the weights of the tasks.
+        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
+        and weights along the optimal memory traversal.
+        In the second step, we try assigning each task to each processors and execute the assignment that
+        is feasible with regard to memory size and gives the earliest finishing time to the task.
+        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
+        assign a task to a processor.
+        We suggest two eviction strategies - largest files first and smallest files first.
+
+        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
+        ones with real task and edge weights with up to 30,000 tasks shows that
+        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
+        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
+        on small one.
+        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
+        average memory utlization by $t\%$.
+
     \end{abstract}
@@
     \section{Introduction: \skug{Full: 0, Polished: 0}}
+
+    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
+
+    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
+    }
+    TODO: Insert abstract
 
@@
 
-    \section{Related work: \skug{Full: 4, Polished: 0}}
+    \section{Related work: \skug{Full: 5, Polished: 4}}
 

### 2024-10-02 main.tex
new title and two todos

@@
 
-    \title{Adaptive Scheduling of Scientific Workflows\\
+    \title{Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\
 %{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
     \subsubsection{Workflow instances}
+    \skug{Todo: describe the workflows the same asin the abstract}
     The input data for the experiments consists of two sets of workflows: real-world workflows
@@
     \subsubsection{Target computing systems}
+    \skug{Check that the description of speeds and memories is plausible and consistent with experimental setup. esp check for 19200 -> 192 GB.}
     To fully benefit from the historical data, the  {\em default} experimental environment
@@
     \subsection{Results \skug{Full: 3, Polished: 2}}
-    \subsubsection{Normal cluster}
+    \subsubsection{Static experiments}
+    \paragraph{Normal cluster}
 

### 2024-10-03 main.tex
tentative new version of abstract

--- 2024-07-24 main.tex
@@
     \begin{abstract}
+
+        %Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
+        Scientific workflows are often represented as directed acyclic graphs (DAGs),
+        where vertices correspond to tasks and edges represent the dependencies between them.
+        Typically, each task requires a
+        certain amount of memory to be executed and needs to communicate data to its successor tasks.
+        The goal is generally to execute the workflow  as fast as possible (i.e., to minimize its makespan), 
+        while satisfying the memory constraints.  
+        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows onto 
+        heterogeneous platforms, where each processor can have a different speed and a different memory size.
+        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that accounts for memory, and 
+        that includes eviction strategies for cases when it might be beneficial to remove some data from memory
+	in order to have enough memory to execute other tasks.
+%
+	Furthermore, while HEFT assumes a perfect knowledge of the execution time and memory usage
+	of each task, the actual values might differ upon execution. Thus, we propose an adaptive
+	scheduling strategy, where a schedule is recomputed when there has been a significant variation in terms
+	of execution time or memory. 
+%	
+	The scheduler has been closely integrated with a runtime system, allowing us to perform a thorough
+	experimental evaluation on real-world workflows. The runtime system warns the scheduler when 
+	the task parameters have changed, and a schedule can be recomputed on the fly. The memory-aware
+	strategy allows us to schedule task graphs that would run out of memory with a state-of-the-art
+	scheduler, and the adaptive setting allows us to significantly reduce the makespan.
+
+%                when tentatively assigning tasks, in order to 
+%        Its first step is to compute the weights of the tasks.
+%        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
+%        and weights along the optimal memory traversal.
+%        In the second step, we try assigning each task to each processors and execute the assignment that
+%        is feasible with regard to memory size and gives the earliest finishing time to the task.
+%        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
+%        assign a task to a processor.
+%        We suggest two eviction strategies - largest files first and smallest files first.
+
+%        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
+%        ones with real task and edge weights with up to 30,000 tasks shows that
+%        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
+%        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
+%        on small one.
+%        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
+%        average memory utlization by $t\%$.
+
+    \end{abstract}
+
+%    \begin{IEEEkeywords}
+%        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
+%    \end{IEEEkeywords}
+
+    \section{Introduction: \skug{Full: 0, Polished: 0}}
+
     \skug{Fullness score (how much text is available): 0-none to 5 - everything.
@@
     TODO: Insert abstract
-    \end{abstract}
-
-    \begin{IEEEkeywords}
-        Scheduling, Adaptive Scheduling, DAG
-    \end{IEEEkeywords}
-
-    \section{Introduction: \skug{Full: 0, Polished: 0}}
 

### 2024-10-03 main.tex
very few minor edits in abstract

@@
         while satisfying the memory constraints.  
-        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows onto 
+        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows on 
         heterogeneous platforms, where each processor can have a different speed and a different memory size.
-        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that accounts for memory, and 
-        that includes eviction strategies for cases when it might be beneficial to remove some data from memory
+        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that (in contrast to the original) accounts for memory and 
+        includes eviction strategies for cases when it might be beneficial to remove some data from memory
 	in order to have enough memory to execute other tasks.
 %
-	Furthermore, while HEFT assumes a perfect knowledge of the execution time and memory usage
+	Furthermore, while HEFT assumes perfect knowledge of the execution time and memory usage
 	of each task, the actual values might differ upon execution. Thus, we propose an adaptive

### 2024-10-04 main.tex
temp
started some updates, done for tonight

@@
 
-    \section{Introduction: \skug{Full: 0, Polished: 0}}
+    \section{Introduction} %: \skug{Full: 0, Polished: 0}}
 
-    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
-
-    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
-    }
-    TODO: Insert abstract
+%    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
+%
+%    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
+%    }
+%    TODO: Insert abstract
 
@@
 
-    \section{Model: \skug{Full: 4, polished: 3}}
+    \section{Model} %: \skug{Full: 4, polished: 3}}
 
-    We first describe the target applications, which are (large scientific) workflows,
+    We first describe the target applications, which are (large scientific) workflows for which we
+    do not have a perfect a priori knowlodge,
     in Section~\ref{sec.mod.work}.  Next, we define the execution
     environment, a heterogeneous system (in terms of processor speed and memory size),
-    in Section~\ref{sec.mod.plat}.
+    in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
+    Section~\ref{sec.mod.pb}. The key notations are summarized in Table~\ref{tabnotation}.
 
@@
     Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.
-
-    Hence, the total  memory requirement for the execution of task~$u$ consists of the input files
+   Note that $m_u$ is the total memory usage
+    of a task during its execution, including input and output files currently being read and written, 
+    and hence the total memory requirement  for the execution of task~$u$ consists of the maximum
+    between the input files
     (total size of the files to be received from the parents),
     the output files (total size of the files to be sent to the children),
-    and the memory size~$m_u$:
+    and the total memory size~$m_u$ (usually achieving the maximum):
     \[
-        r_u = m_u + \sum_{v:(v,u)\in E}c_{v,u} + \sum_{v:(u,v)\in E} c_{u,v}.
+        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
     \]
+\AB{I transformed the sum into a max, correct? Or maybe we could assume $r_u=m_u$? }
 
@@
     Each task may have multiple parents and children.
+
+Furthermore, we place ourself in a context where we do not have a perfect knowledge 
+of the task parameters ($w_u$ and $m_u$) before they start their execution,
+but only estimates. \AB{Add motivation: related work with variable task durations for instance...}
+Hence, scheduling decisions are made on these estimated parameters, and
+may be reconsidered at runtime when a task starts its execution and we know
+its exact parameters.
 
@@
 
-    \paragraph{Workflow-related changes}
+\subsection{Optimization problem}
+\label{sec.mod.pb}
 
-    \begin{itemize}
-        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
-        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
+The goal is to find a schedule of the DAG~$G$ onto the $k$ processors, 
+so that the makespan (total execution time) is minimized while
+respecting memory constraints. If a processor runs out of memory to execute
+a task, the schedule is said to be {\em invalid}.
 
-    \end{itemize}
+Since tasks are subject to variability, we aim at minimizing the actual makespan
+achieved at the end of the execution, while decisions may be taken building
+on the estimated task parameters.
 
-    The following changes are not a part of this article's scope:
-
-    \begin{itemize}
-        \item The workflow structure changes: edges or tasks come in or leave.
-    \end{itemize}
-
-    \paragraph{Execution environment-related changes }
-
-    \begin{itemize}
-        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
-        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
-
-    \end{itemize}
-
-    The following changes are not a part of this article's scope:
-
-    \begin{itemize}
-        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
-    \end{itemize}
-
-    \subsection{Time of changes }
-
-    We consider discrete time in seconds.
-    The time point(s) at which the changes happen is unambiguously defined.
-
-    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
-    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
-    The start time of the source task in the workflow is zero.
-    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
-
-    \subsection{Changes and knowledge horizon - important questions TBA}
-
-    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
-
-    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
-    However, how do we model the following:
-    \begin{itemize}
-        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
+%    \paragraph{Workflow-related changes}
+%
+%    \begin{itemize}
+%        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
+%        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
+%
+%    \end{itemize}
+%
+%    The following changes are not a part of this article's scope:
+%
+%    \begin{itemize}
+%        \item The workflow structure changes: edges or tasks come in or leave.
+%    \end{itemize}
+%
+%    \paragraph{Execution environment-related changes }
+%
+%
+%    \begin{itemize}
+%        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
+%        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
+%
+%    \end{itemize}
+%
+%    The following changes are not a part of this article's scope:
+%
+%    \begin{itemize}
+%        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
+%    \end{itemize}
+%
+%    \subsection{Time of changes }
+%
+%    We consider discrete time in seconds.
+%    The time point(s) at which the changes happen is unambiguously defined.
+%
+%    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
+%    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
+%    The start time of the source task in the workflow is zero.
+%    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
+%
+%    \subsection{Changes and knowledge horizon - important questions TBA}
+%
+%    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
+%
+%    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
+%    However, how do we model the following:
+%    \begin{itemize}
+%        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
 This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
-        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
-        \item How many times of change do we model - one per workflow run, or multiple?
-        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
-    \end{itemize}
+%        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
+%        \item How many times of change do we model - one per workflow run, or multiple?
+%        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
+%    \end{itemize}
 
@@
 
-    \section{Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
+    \section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
+\label{sec.heuristics}
 
-    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
-    but favor processor reuse as is done in HEFT.
-    Furthermore, this would allow us to handle variability on the fly, by updating
-    the bottom levels if some parameters vary, and computing the schedule
-    only for the near future...
+%    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
+%    but favor processor reuse as is done in HEFT.
+%    Furthermore, this would allow us to handle variability on the fly, by updating
+%    the bottom levels if some parameters vary, and computing the schedule
+%    only for the near future...
+
+In order to be able to easily adapt to variability of task parameters, we design
+variants of HEFT that account for memory usage and aim at minimizing the makespan. 
+First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic that does not account for the memory
+(and hence, may return invalid schedules that will not be able to run on the platform
+by running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
+heuristics, including eviction strategies to move some data in communication buffers
+in case there is not enough memory available on some processors.
 
     \subsection{Baseline: original HEFT without memories}
+\label{sec.heft}
 
@@
 
-    HEFT works in two steps.
-    In the first step, it calculates the ranks of tasks by computing their non-increasing bottom levels.
+    HEFT works in two stages.
+    In the first stage, it computes the ranks of tasks by computing their non-increasing bottom levels.
     The bottom level of a task is defined as
@@
 
-    In the second step, the algorithm iterates over the ranks and tries to assign the task to the processor where it
+    In the second stage, the algorithm iterates over the ranks and tries to assign the task to the processor where it
     has the earliest finish time.
-    We tentatively assign each task to each processor.
-    The task's starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
+    We tentatively assign each task~$v$ to each processor~$j$.
+    The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$, 
+    and all communications that
     must be orchestrated from predecessor tasks $u\notin T(p_j)$.
-    The starting time is then
-    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]
-    Its finish time on $p_j$ is then
+    The starting time is then:
+
+ {\footnotesize{   \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]}}
+
+    Finally, its finish time on $p_j$ is 
     $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.
@@
 
-    \subsection{Heuristics}
-    Like the original HEFT, our heuristic consistst of two steps: first, computing task ranks,
-    and second, assigning tasks to processors in the order defined in the first step.
-    We consider three variants of HEFT accounting for memory usage, which only
-    differ in the order they consider tasks to be scheduled.
+    \subsection{Memory-aware heuristics}
+    \label{sec.heftm}
+    Like the original HEFT, the memory-aware versions of HEFT consist of two stages: 
+    first, they compute the task ranks,
+    and second, they assign tasks to processors in the order defined in the first step.
+    We consider three variants of HEFT accounting for memory usage (HEFTM), which only
+    differ in the order they consider tasks to be scheduled in the first stage.
 
-    \subsubsection{Step 1: calculate task ranks}
+\smallskip
+    \noindent{\bf Compute task ranks. }
 
-    HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
+We design three variants of memory-aware HEFT: 
+\begin{itemize}
+\item    HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
     level is defined as
@@
 
-    HEFTM-BLC: from the study of the fork (see below), it seems important
-    to also account for the size of the data as input of a task,
-    to give more priority at tasks with potential large incoming communications.
-    For each task, we compute its modified bottom level:
+\item    HEFTM-BLC %: from the study of the fork (see below), it seems important
+   %  to also account for the size of the data as input of a task,
+    is  giving more priority at tasks with potential large incoming communications, 
+    hence aiming at clearing the memory used by files as soon as possible, 
+    to have more free memory for remaining tasks to be executed on the processor. 
+    Therefore, for each task, we compute a modified bottom level accounting for communications: 
     $$blc(u) = w_u + \max_{(u,w)\in E} \{c_{u,w} + blc(w)\} + \max_{(v,u)\in E} c_{v,u}   . $$
 
-    \skug{avoid having mixed ranks, when the memory size of the lower task is not taken into account}
+%    \skug{avoid having mixed ranks, when the memory size of the lower task is not taken into account}
 
-    HEFTM-MM orders tasks as dictated by MinMem.
+\item   Finally, HEFTM-MM orders tasks  in the order returned by %as dictated by MinMem.
+the  \algo{memDag} algorithm~\cite{KAYAASLAN20181}, which corresponds to a traversal 
+of the graph that minimizes the peak memory usage.
 
-    \subsubsection{Task assignment}
+\end{itemize}
+
+\smallskip
+  \noindent  {\bf Task assignment. }
 
@@
     and greedily assign it to a processor, by trying all possible options
-    and keeping the most promising one.
+    and keeping the most promising one. We first detail how a task
+    is tentatively assigned on a processor, by carefully accounting for the memory usage. 
+    Next, we explain the steps to be taken to effectively assign a task on a given processor.
 
@@
 
-    \subsection{Approximation}
-    \hmey{Rough notes:}
-    Let's use a fork to see how the algorithm behaves and if it provides some approximation.
+%    \subsection{Approximation}
+%    \hmey{Rough notes:}
+%    Let's use a fork to see how the algorithm behaves and if it provides some approximation.
 Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).
@@
 
-    \bibliographystyle{apalike}
+Further variability commented out from the paper: changing DAG structure; changing platform...
+
+    \bibliographystyle{IEEEtran}
     \bibliography{references}

### 2024-10-05 main.tex
mention simexpal as manager of experiments

--- 2024-10-03 main.tex
@@
     All algorithms are implemented in C++ and compiled with g++ (v.11.2.0).
-    The experiments are executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
+    The experiments are managed by simexpal~\cite{DBLP:journals/algorithms/AngrimanGLMNPT19} and executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
     and CentOS 8 as OS.
@@
 
-    \bibliographystyle{apalike}
+    \bibliographystyle{IEEEtran}
     \bibliography{references}

### 2024-10-08 main.tex
Chapter V.A - Experimentsal Evaluation - Setup; still WIP

--- 2024-10-05 main.tex
@@
 
+\newcommand{\heft}{\algo{HEFT}\xspace}
 \newcommand{\heftmm}{\algo{HEFTM\_MM}\xspace}
@@
 
-        %Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
-        Scientific workflows are often represented as directed acyclic graphs (DAGs),
+        Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
+        They are often represented as directed acyclic graphs (DAGs),
         where vertices correspond to tasks and edges represent the dependencies between them.
+       % Since these graphs are often large in both the number of tasks and their resource requirements, it is important
+       % to schedule them efficiently on parallel or distributed compute systems.
         Typically, each task requires a
         certain amount of memory to be executed and needs to communicate data to its successor tasks.
-        The goal is generally to execute the workflow  as fast as possible (i.e., to minimize its makespan), 
-        while satisfying the memory constraints.  
-        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows on 
-        heterogeneous platforms, where each processor can have a different speed and a different memory size.
-        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that (in contrast to the original) accounts for memory and 
-        includes eviction strategies for cases when it might be beneficial to remove some data from memory
-	in order to have enough memory to execute other tasks.
-%
-	Furthermore, while HEFT assumes perfect knowledge of the execution time and memory usage
-	of each task, the actual values might differ upon execution. Thus, we propose an adaptive
-	scheduling strategy, where a schedule is recomputed when there has been a significant variation in terms
-	of execution time or memory. 
-%	
-	The scheduler has been closely integrated with a runtime system, allowing us to perform a thorough
-	experimental evaluation on real-world workflows. The runtime system warns the scheduler when 
-	the task parameters have changed, and a schedule can be recomputed on the fly. The memory-aware
-	strategy allows us to schedule task graphs that would run out of memory with a state-of-the-art
-	scheduler, and the adaptive setting allows us to significantly reduce the makespan.
+        The goal is thus to execute the workflow  as fast as possible (i.e., to minimize its makespan) while satisfying the memory constraints.  %and using parallelism to speed up the execution.
 
-%                when tentatively assigning tasks, in order to 
-%        Its first step is to compute the weights of the tasks.
-%        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
-%        and weights along the optimal memory traversal.
-%        In the second step, we try assigning each task to each processors and execute the assignment that
-%        is feasible with regard to memory size and gives the earliest finishing time to the task.
-%        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
-%        assign a task to a processor.
-%        We suggest two eviction strategies - largest files first and smallest files first.
+        Hence, we investigate adaptive memory-aware scheduling  of DAG-shaped workflows onto heterogeneous platforms where
+        each processor can have a different speed and a different memory size.
+        We propose a two-step HEFT-inspired heuristic.
+        Its first step is to compute the weights of the tasks.
+        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
+        and weights along the optimal memory traversal.
+        In the second step, we try assigning each task to each processors and execute the assignment that
+        is feasible with regard to memory size and gives the earliest finishing time to the task.
+        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
+        assign a task to a processor.
+        We suggest two eviction strategies - largest files first and smallest files first.
 
-%        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
-%        ones with real task and edge weights with up to 30,000 tasks shows that
-%        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
-%        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
-%        on small one.
-%        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
-%        average memory utlization by $t\%$.
+        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
+        ones with real task and edge weights with up to 30,000 tasks shows that
+        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
+        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
+        on small one.
+        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
+        average memory utlization by $t\%$.
 
@@
 
-%    \begin{IEEEkeywords}
-%        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
-%    \end{IEEEkeywords}
+    \begin{IEEEkeywords}
+        Scheduling, Adaptive Scheduling, DAG
+    \end{IEEEkeywords}
 
-    \section{Introduction} %: \skug{Full: 0, Polished: 0}}
+    \section{Introduction: \skug{Full: 0, Polished: 0}}
 
-%    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
-%
-%    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
-%    }
-%    TODO: Insert abstract
+    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
+
+    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
+    }
+    TODO: Insert abstract
 
@@
 
-    All algorithms are implemented in C++ and compiled with g++ (v.11.2.0).
-    The experiments are managed by simexpal~\cite{DBLP:journals/algorithms/AngrimanGLMNPT19} and executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
+    All algorithms are implemented in C++ and compiled with g++ (v.8.5.0).
+    The experiments are executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
     and CentOS 8 as OS.
-    Code, input data, and experiment scripts are available for review under \url{TODO:insert link}.
+    Code, input data, and experiment scripts are available for review under \skug{\url{insert link}}.
 
@@
     \skug{Todo: describe the workflows the same asin the abstract}
-    The input data for the experiments consists of two sets of workflows: real-world workflows
-    obtained from~\cite{ewels2020nf} and workflows obtained by simulating real-world workflows
-    with the WFGen generator~\cite{COLEMAN202216}.
-    First, we discuss how the graph topology is generated,
-    then we focus on the weights associated to tasks and edges.
+    We run our experiments on a dataqset comprised of real-world workflows: ones used in~\cite{lotaru} (atacseq, bacass, chipseq,
+    eager and methylseq), and ones obtained by increasing their size using the WFGen generator~\cite{COLEMAN202216}.
 
     \paragraph{Workflow graphs}
-    For real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
+    For the five real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
     respective repository and transformed into .dot format using the nextflow option ``-with-dag''.
@@
 
-    For the simulated workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
+    For the size-increased workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
     the desired number of tasks.
-    The model workflows are described on the WFGen website, and we used the following ones:
-    1000Genome, BLAST, BWA, Epigenomics, Montage, Seismology, and SoyKB.
-    Other models could not be generated without errors.
+    We used the real-world workflows as models, except for bacass that lead to errors in the generator.
 %
@@
     ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.
-    Note that for some workflow models, such as SoyKB or Montage, only a subset of the workflow sizes could be generated,
-    because the workflow generator either took a disproportionally long time or yielded errors.
 
-    Overall, this yields four workflow types, denoted by real, small, mid, and big.
+    \skug{Im still considerin adding category "tiny" with original workflows and those with 200 tasks. this needs to be double-checked after}
+    Overall, this yields four workflow types, denoted by tiny, small, mid, and big.
 
-    \paragraph{Generation of task and edge weights}
+    \paragraph{Task and edge weights}
     For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
@@
     size of files that the task sends to all its children.
+
     For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
-    To avoid underestimation, we take the maximum among all values from different runs on the same cluster node.
+    The same workflow can require different memory capacity and take different time to execute
+    depending on the size of its input.
+    We simulate these various runs by obtaining values corresponding to each input size.
+    For each of the four families, there are five input sizes, so we run each workflow in five variants corresponding to these inputs
+
     Not all tasks have historical runtime data stored in the tables.
@@
     For two more, around 40\% of tasks have no historical runtime data stored.
-    Hence, in the absence of historical data about a task, we give it a weight of~1.
+    Hence, in the absence of historical data about a task, we give it fixed weights.
+    We give it an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
+    These values align with the findings of~\cite{lotaru} about small tasks.
 %
-    Because the historical data contains absolute measured values and the cluster node information is a relative value,
-    we normalize all values extracted from the historical data by the smallest one.
-    This way, task memory weights fit into the memory of the cluster nodes.
-    Additionally, this way, the tasks without historical data receive less insignificant values compared to tasks with historical data.
-
-    For the simulated workflows, we generate random execution times and memory weights for each task as well as
-    edge weights for each precedence constraint.
-    We generate uniformly distributed values between $1$ and $10$ for edge weights,
-    $1$ and $1000$ for the workloads,
-    and $1$ and $192$ for memory weights.
-    When doing so, we try to mimic the weights observed in the historical data, hence \eg the low lower bounds for the
-    workloads.
+    \skug{insert more abt how the runtime system hanfles data reads}
 
@@
     \skug{Check that the description of speeds and memories is plausible and consistent with experimental setup. esp check for 19200 -> 192 GB.}
+
     To fully benefit from the historical data, the  {\em default} experimental environment
@@
     kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
-    We set the number of each kind of node to six, thus having 36 processors in total. % the whole cluster.
+    We set the number of each kind of node to \skug{$24$ - check if this still holds }, thus having 144 processors in total. % the whole cluster.
 
@@
     (see Table~\ref{tab:procs} in Appendix~\ref{sub:appx-cluster-config}~\cite{daghetpart_full_version}).
-    Note that, by nature,
-    memory sizes are normalized values, relative to each other, too.
-    Therefore, we additionally normalize memory weights of real-world workflows
-    to the maximum size of 192 to make sure they fit.
-    For simulated workflows, we increase memory sizes proportionally until the task
-    with the biggest memory requirement still
-    has a processor it could be executed on.
 
@@
     variants of the default cluster presented above:\\
-    $\bullet$ {\em Small} and {\em large} clusters. While the default cluster consists of 36 nodes (6 of each kind), we
-    consider a small cluster with three processors of each kind ($18$ processors in total), and a
-    large cluster with ten processors of each kind ($60$ processors in total). \skug{leaving the big cluster inside for
-    the case we do want to include it}\\
-    $\bullet$ {\em Fat-and-thin} cluster. In this cluster, the fastest processors have the smallest amount of memory (thin),
-    while the slowest processors have the largest memories (fat).
-    The overall amount of memory in the cluster, as well as the overall speed of all processors remain the same as in the
-            {\em normal} cluster.
-%\end{itemize}
+    $\bullet$ {\em tiny}, {\em small} and {\em large} clusters. While the default cluster consists of 144 nodes (24 of each kind), we
+    consider a small cluster with $18$ processors of each kind ($108$ processors in total), a tiny cluster with $72$ nodes and a
+    large cluster with $36$ processors of each kind ($216$ processors in total). \skug{check for cluster sizes and naming}\\
 
@@
     \subsubsection{Static experiments}
+    \paragraph{Small cluster}
+
+   On the small cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
+    The successful schedules were achieved almost exclusively on the tiny workflows and one one 200-task workflow.
+    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm could schedule
+    $94.9\%$ of workflows, thus having the smallest failure rate of $5\%$.
+
     \paragraph{Normal cluster}

### 2024-10-08 main.tex
started writing the dynamic experiments, WIP

@@
     \subsubsection{Static experiments}
-    \paragraph{Small cluster}
+    \paragraph{Tiny cluster}
 
-   On the small cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
+    On the tiny cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
     The successful schedules were achieved almost exclusively on the tiny workflows and one one 200-task workflow.
@@
     \paragraph{Normal cluster}
+    \skug{OLD DATA here, rewrite}
 
@@
     \heftmm takes $13.9$ minutes on average.
+    \subsubsection{Dynamic experiments}
+    \paragraph{Tiny cluster}
+    The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
+    situation - that is, if the scheduler assumed its memory to be smaller than it actually was and assigned it to a processor
+    with not enough memory capacity.
+    Due to an extremely constrained memory in the tiny cluster, only 20 experiments out of 290 ended with a valid makespan
+    in case of \heftbl, 33 in case of \heftblc and 29 in case of \heftmm.
+    In case of \heftmm, the makespan without recomputation would have been, on average, $5.77$ times longer than with the recomputation.
+
     \subsubsection{Summary}

### 2024-10-08 main.tex

--- 2024-10-05 main.tex
@@
 
-    All algorithms are implemented in C++ and compiled with g++ (v.11.2.0).
+    All algorithms are implemented in C++ and compiled with g++ (v.8.5.0).
     The experiments are managed by simexpal~\cite{DBLP:journals/algorithms/AngrimanGLMNPT19} and executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
     and CentOS 8 as OS.
-    Code, input data, and experiment scripts are available for review under \url{TODO:insert link}.
+    Code, input data, and experiment scripts are available for review under \skug{\url{insert link}}.
 
@@
     \subsubsection{Static experiments}
+    \paragraph{Tiny cluster}
+
+    On the tiny cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
+    The successful schedules were achieved almost exclusively on the tiny workflows and one one 200-task workflow.
+    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm could schedule
+    $94.9\%$ of workflows, thus having the smallest failure rate of $5\%$.
+
     \paragraph{Normal cluster}
+    \skug{OLD DATA here, rewrite}
 

### 2024-10-09 main.tex
picture and some text for the dynamic case

@@
 \usepackage{todonotes}
+\usepackage{graphicx}
 
@@
 \newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
+\newcommand{\willchange}[1]{{\color{orange}[AB: #1]}}
 
@@
     \subsection{Results \skug{Full: 3, Polished: 2}}
+
+    \subsubsection{Preliminary experiments}
+
+    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.
+
     \subsubsection{Static experiments}
@@
     On the tiny cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
-    The successful schedules were achieved almost exclusively on the tiny workflows and one one 200-task workflow.
+    The successful schedules were achieved exclusively on the tiny workflows (with only one 200-task size-increased
+    workflow among them).
     \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm could schedule
-    $94.9\%$ of workflows, thus having the smallest failure rate of $5\%$.
+    $94.9\%$ of workflows, thus having the smallest failure rate of $5.1\%$.
 
@@
 
-    \subsubsection{Impact of cluster composition}
-    On the {\em fat-and-thin} cluster, the baseline produces $60$\% invalid results on small workflows, $83.3$\% on
-    middle-sized and $88.9$\% on big workflows, and therefore performs worse than on the normal cluster.
-
-    The relative makespans are $14-29$\% worse for small workflows, $22-40$\% worse on middle-sized workflows and $16-35$\%
-    on big workflows.
-    \wrt failures, all heuristics again manage to schedule all workflows.
-    \subsubsection{Behavior of workflow types when scaling workflow size}
-    \subsubsection{Running times}
-    The geometric means of runtimes of all heuristics on small workflows lie between $0.9$ and $2.2$ seconds.
-    For middle-sized workflows, the runtimes for \heftbl and \heftblc are close to $2$ minutes, while \heftmm
-    takes $5.1$ minutes.
-    For the big workflows, again, \heftbl and \heftblc show a better runtime ($5$ and $5.4$ minutes, respectively).
-    \heftmm takes $13.9$ minutes on average.
     \subsubsection{Dynamic experiments}
@@
     with not enough memory capacity.
-    Due to an extremely constrained memory in the tiny cluster, only 20 experiments out of 290 ended with a valid makespan
-    in case of \heftbl, 33 in case of \heftblc and 29 in case of \heftmm.
-    In case of \heftmm, the makespan without recomputation would have been, on average, $5.77$ times longer than with the recomputation.
+    Due to an extremely constrained memory in the tiny cluster, only 134 experiments out of 1160 for all algorithm variants
+    ended with a valid makespan withjout recomputation. $13$ experiments ended with a valid makespan after recomputation in case of \heft,
+   $16$ in case of \heftmm, $50$ in case of \heftblc and $55$ in case of \heftbl.
+    Figure~\ref{fig:updates-ms} shows the increase in makespan in case of no recomputation for these experiments.
+    There were valid final makespans with no recomputation only for the smallest workflows - $56$ original ones, $47$ 200-task ones,
+    $25$ 1000-task ones and $6$ 2000-task ones.
+
+    With growing size of the workflow, the excess makespan of not recomputing grows - from $13.9\%$ to $20\%$ for \heftbl,
+    from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the 2000-task workflows in this case), $12.1\%$ to $23.5\%$
+    for \heftmm.
+    The larger variations for \heft can be explained by the small amount of data - for example, there were only 2 200-task
+    workflows for this case.
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=0.495\columnwidth] {images/UpdatesMss2}
+        \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm for different . Smaller is better.}
+        \label{fig:updates-ms}
+        \vspace{-0.3cm}
+    \end{figure}
 

### 2024-10-09 main.tex
storyline of intro, joint with SK

@@
     %%% CONTEXT %%%
-    Only mapping tasks to processors is not enough.
-    Reusing processors is important to achieve good makespans.
+    \begin{itemize}
+      \item What is a scientific workflow?
+      \item Scientific workflows are all over the place $\rightarrow$ important!
+      \item Scheduling (order, maybe even concrete starting times) is important for efficient workflow executions
+      \item Only mapping tasks to processors is not enough.
+      \item This also allows to reuse processors, which is important to achieve good HW utilization / makespans (explain!).
+    \end{itemize}
 
     %%% MOTIVATION %%%
-    HEFT is one of the most popular heuristics (cite a review).
-    However, existing HEFT variations do not take memory sizes into consideration (\skug{check in related work if true!}).
+    \begin{itemize}
+      \item Heterogeneous means: different CPU speeds, different mem sizes
+      \item With different mem sizes: more difficult to adhere to mem constraints
+      \item Adhering to memory constraints is very important to avoid execution failures.
+      \item In this paper, we formulate the problem that takes memories into consideration (using (nearly the) model from ICPP paper)
+      \item HEFT (see book chapter) is one of the most popular heuristics (cite a review).
+      \item However, existing HEFT variations do not take memory sizes into consideration (\skug{check in related work if true!}).
+        Note: 2 papers that deal with memory sizes, but model is very different!
+      \item Many scheduling strategies assume that task running times provided to them are accurate. This is not really the case
+        in practice. Thus, one needs to adapt the schedule when major deviations occur (which is the case in practice).
+        List-based schedulers such as HEFT are not designed for adaptation and would compute a completely schedule from scratch.
+    \end{itemize}
 
     %%% CONTRIBUTION %%%
-    In this paper, we formulate the problem that takes memories into consideration.
-    We propose three HEFT-based heuristics that take memory size into account: \heftbl, \heftblc, and \heftmm.
-    The difference is the way they order tasks for scheduling.
-    We compare our heuristics to a baseline HEFT variant that does not take memory sizes into account.
-
-    We find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
+    \paragraph*{Contribution}
+    \begin{itemize}
+      \item We propose three HEFT-based heuristics that take memory sizes into account: \heftbl, \heftblc, and \heftmm.
+      \item The difference between the new heuristics is the way they rank tasks (for processor assignment).
+      \item Static: we compare our heuristics to a baseline HEFT variant that does not take memory sizes into account.
+      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
+      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
+      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
+    \end{itemize}
 

### 2024-10-09 main.tex
about the runtime system

@@
 
+    \subsubsection{Runtime System}
+
+    To simulate the execution of a workflow, we implemented a runtime system.
+    It read the historical data and builds weights for tasks.
+    These weights are then treated as the ``truth'' about the execution during the simulation.
+    In the case without errors, these values are being sent to the scheduler.
+    However, if errors have to be simulated, the runtime system applies a deviation function to the values.
+    The deviation function computed a normally distributed deviated value, where the correct value is the mean and the deviation
+    is $10\%$.
+    This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and\skug{todo cite someone Jonathan cites}.
+
+    So, the scheduler receives deviated values and makes decision based on them.
+    This leads to several types of possible issues:
+    \begin{itemize}
+        \item Processor is blocked by another task. If the scheduler underestimated the execution time of a task, it will block another
+        one from starting.
+        \item The parent has not finished yet. A similar situation, but a task could start on its processor, but one or more of its parents a
+        are not ready yet.
+        \item Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able
+        to execute on a chosen processor.
+        \item A task took lees than expected. We only consider this case if a task took more than 10\% less than expected. In this
+        case we want to exploit the newly acquired free time by possibly starting other tasks earlier.
+    \end{itemize}
+
     \subsection{Results \skug{Full: 3, Polished: 2}}

### 2024-10-09 main.tex
rough draft for first part of intro

@@
     %%% CONTEXT %%%
+    Data analysis on massive data sets (stemming from areas such as genomics, 
+    remote sensing, or biomedical imaging -- to name just a few) has become ubiquitous in science;
+    it often happens in the form of workflows, \iec separate software components chained together
+    in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
+    Such workflows are usually represented as directed acyclic graphs (DAGs).
+    The DAG vertices represent the software components (more general: the workflow \emph{tasks}),
+    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liuPacitti}.
+    Large workflows with resource-hungry tasks can easily exhaust the capabilities of a 
+    single computer and are thus executed on a parallel or distributed platform.
+    An efficient execution of the workflows on such platforms requires not only to map the tasks
+    to a particular processor, but also a good task schedule, \iec an execution order
+    and possibly also starting times for the tasks.
+
     \begin{itemize}
-      \item What is a scientific workflow?
-      \item Scientific workflows are all over the place $\rightarrow$ important!
-      \item Scheduling (order, maybe even concrete starting times) is important for efficient workflow executions
-      \item Only mapping tasks to processors is not enough.
       \item This also allows to reuse processors, which is important to achieve good HW utilization / makespans (explain!).

### 2024-10-09 main.tex
rough draft for first part of intro

@@
     single computer and are thus executed on a parallel or distributed platform.
-    An efficient execution of the workflows on such platforms requires not only to map the tasks
-    to a particular processor, but also a good task schedule, \iec an execution order
+    An efficient execution of the workflows on such platforms requires to map the tasks
+    to a particular processor; to increase utilization by reusing finished processors,
+    one also needs a task schedule, \iec a valid execution order that respects the dependencies
     and possibly also starting times for the tasks.
-
-    \begin{itemize}
-      \item This also allows to reuse processors, which is important to achieve good HW utilization / makespans (explain!).
-    \end{itemize}
 
@@
       \item In this paper, we formulate the problem that takes memories into consideration (using (nearly the) model from ICPP paper)
+      \item Objective: makespan (proxy for total execution time)
       \item HEFT (see book chapter) is one of the most popular heuristics (cite a review).

### 2024-10-09 main.tex
some fine-tuning (adds one line, though)

--- 2024-10-09 main.tex
@@
     %%% CONTEXT %%%
-    \begin{itemize}
-      \item What is a scientific workflow?
-      \item Scientific workflows are all over the place $\rightarrow$ important!
-      \item Scheduling (order, maybe even concrete starting times) is important for efficient workflow executions
-      \item Only mapping tasks to processors is not enough.
-      \item This also allows to reuse processors, which is important to achieve good HW utilization / makespans (explain!).
-    \end{itemize}
+    The analysis of massive datasets, originating from fields such as genomics, 
+    remote sensing, or biomedical imaging -- to name just a few -- has become ubiquitous in science;
+    this often takes the form of workflows, \iec separate software components chained together
+    in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
+    These workflows are usually represented as directed acyclic graphs (DAGs).
+    The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
+    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liuPacitti}.
+    Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
+    single computer and are therefore executed on a parallel or distributed platform.
+    An efficient execution of the workflows on such platforms requires mapping tasks
+    to specific processors; to increase utilization by reusing finished processors,
+    one also needs a task schedule (\iec a valid execution order that respects the dependencies)
+    and possibly also starting times for the tasks.
 
     %%% MOTIVATION %%%
+    Today's computing platforms are often heterogeneous,
+
     \begin{itemize}

### 2024-10-09 main.tex
rough draft for second part of intro (WIP)

@@
     %%% MOTIVATION %%%
-    Today's computing platforms are often heterogeneous,
+    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
+    and memory sizes. In general, varying memory sizes make it more challenging to compute
+    a schedule that respects all memory constraints -- meaning that no task is executed on a 
+    processor with less memory than needed for the task. This is, however, very important to
+    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
+    Hence, building on previous related work~\cite{TODO}, we consider a scheduling problem 
+    formulation that takes memory sizes as explicit constraints into account. Its objective is
+    the very common \emph{makespan}, which acts as proxy for the total execution time of a
+    workflow.
 
     \begin{itemize}
-      \item Heterogeneous means: different CPU speeds, different mem sizes
-      \item With different mem sizes: more difficult to adhere to mem constraints
-      \item Adhering to memory constraints is very important to avoid execution failures.
-      \item In this paper, we formulate the problem that takes memories into consideration (using (nearly the) model from ICPP paper)
-      \item Objective: makespan (proxy for total execution time)
       \item HEFT (see book chapter) is one of the most popular heuristics (cite a review).

### 2024-10-09 main.tex
rough draft for second part of intro (WIP)

@@
     formulation that takes memory sizes as explicit constraints into account. Its objective is
-    the very common \emph{makespan}, which acts as proxy for the total execution time of a
-    workflow.
+    the very common \emph{makespan}~\cite{liu2018survey}, 
+    which acts as proxy for the total execution time of a workflow.
+
+    A seminal list-based scheduling heuristic for workflows is HEFT 
+    (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
+    It has two phases: (i) each task is assigned a priority and (ii) the tasks are assigned
+    to processors, where the ``ready'' task with the highest priority is scheduled next.
+    HEFT has been extended (e.g.\ by Shi and Dongarra~\cite{SHI2006665}) and adjusted for a variety of different scheduling problem formulations. 
+    Yet, none of them adheres to memory constraints -- also see our discussion of related work
+    in Section~\ref{sec:related-work}. (\skug{check in related work if true!}).
+        \skug{Note: 2 papers that deal with memory sizes, but model is very different!}
 
     \begin{itemize}
-      \item HEFT (see book chapter) is one of the most popular heuristics (cite a review).
-      \item However, existing HEFT variations do not take memory sizes into consideration (\skug{check in related work if true!}).
-        Note: 2 papers that deal with memory sizes, but model is very different!
       \item Many scheduling strategies assume that task running times provided to them are accurate. This is not really the case
@@
     \section{Related work: \skug{Full: 5, Polished: 4}}
+    \label{sec:related-work}
+
+    \hmey{Question: Does it make sense to place related work before the model (which is more common)?}
 

### 2024-10-09 main.tex
rough draft for second part of intro

@@
 
-    A seminal list-based scheduling heuristic for workflows is HEFT 
+    A seminal list scheduling heuristic for workflows is HEFT 
     (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
-    It has two phases: (i) each task is assigned a priority and (ii) the tasks are assigned
+    It has two phases: (i) each task is assigned a priority and (ii) the tasks in a priority-ordered list are assigned
     to processors, where the ``ready'' task with the highest priority is scheduled next.
@@
 
-    \begin{itemize}
-      \item Many scheduling strategies assume that task running times provided to them are accurate. This is not really the case
-        in practice. Thus, one needs to adapt the schedule when major deviations occur (which is the case in practice).
-        List-based schedulers such as HEFT are not designed for adaptation and would compute a completely schedule from scratch.
-    \end{itemize}
+    Another limitation in practice of HEFT (and many other scheduling strategies) is their 
+    assumption that the task running times provided to them are accurate. In practice this is 
+    not the case and deviations from user estimates or historical measurements are 
+    very common~\cite{TODO}. As a consequence, one should adapt the schedule when \emph{major}
+    deviations occur. List-based schedulers such as HEFT are, however, not designed for 
+    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement?}
+%    and would compute a completely new schedule from scratch.
 

### 2024-10-09 main.tex
minor

--- 2024-10-09 main.tex
@@
     %%% CONTEXT %%%
-    The analysis of massive datasets, originating from fields such as genomics, 
-    remote sensing, or biomedical imaging -- to name just a few -- has become ubiquitous in science;
-    this often takes the form of workflows, \iec separate software components chained together
-    in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
-    These workflows are usually represented as directed acyclic graphs (DAGs).
-    The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
-    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liuPacitti}.
-    Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
-    single computer and are therefore executed on a parallel or distributed platform.
-    An efficient execution of the workflows on such platforms requires mapping tasks
-    to specific processors; to increase utilization by reusing finished processors,
-    one also needs a task schedule (\iec a valid execution order that respects the dependencies)
-    and possibly also starting times for the tasks.
+    \begin{itemize}
+      \item What is a scientific workflow?
+      \item Scientific workflows are all over the place $\rightarrow$ important!
+      \item Scheduling (order, maybe even concrete starting times) is important for efficient workflow executions
+      \item Only mapping tasks to processors is not enough.
+      \item This also allows to reuse processors, which is important to achieve good HW utilization / makespans (explain!).
+    \end{itemize}
 
     %%% MOTIVATION %%%
-    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
-    and memory sizes. In general, varying memory sizes make it more challenging to compute
-    a schedule that respects all memory constraints -- meaning that no task is executed on a 
-    processor with less memory than needed for the task. This is, however, very important to
-    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
-    Hence, building on previous related work~\cite{TODO}, we consider a scheduling problem 
-    formulation that takes memory sizes as explicit constraints into account. Its objective is
-    the very common \emph{makespan}, which acts as proxy for the total execution time of a
-    workflow.
-
     \begin{itemize}
+      \item Heterogeneous means: different CPU speeds, different mem sizes
+      \item With different mem sizes: more difficult to adhere to mem constraints
+      \item Adhering to memory constraints is very important to avoid execution failures.
+      \item In this paper, we formulate the problem that takes memories into consideration (using (nearly the) model from ICPP paper)
       \item HEFT (see book chapter) is one of the most popular heuristics (cite a review).
@@
     \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
-    We divide the workflows into three groups by size: small ones with up to \numprint{8000} tasks, middle
-    ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.
-
-    \skug{Im still considerin adding category "tiny" with original workflows and those with 200 tasks. this needs to be double-checked after}
-    Overall, this yields four workflow types, denoted by tiny, small, mid, and big.
+    We divide the workflows into four groups by size: tiny ones with up to \numprint{200} tasks, small ones with \numprint{1000} to \numprint{8000} tasks,
+    middle ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.
 

### 2024-10-09 main.tex
rough draft for third part of intro

--- 2024-10-09 main.tex
@@
     %%% CONTRIBUTION %%%
-    \paragraph*{Contribution}
+    \paragraph*{Contribution} In this paper, we first propose three HEFT-based heuristics
+    that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
+    (M behind HEFT for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
+    BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
+    and MM for \underline{m}inimum \underline{m}emory traversal).
+    The difference between the new heuristics is the way they prioritize tasks (for processor assignment).
+    We compare the static schedule produced by these heuristics with a HEFT variant as
+    baseline that does not take memory sizes into account. \hmey{Need to clarify: 
+    why is this SotA?}
+
     \begin{itemize}
-      \item We propose three HEFT-based heuristics that take memory sizes into account: \heftbl, \heftblc, and \heftmm.
-      \item The difference between the new heuristics is the way they rank tasks (for processor assignment).
-      \item Static: we compare our heuristics to a baseline HEFT variant that does not take memory sizes into account.
       \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.

### 2024-10-09 main.tex
started editing Sec. 2 (to use time before meeting)

@@
 
-    We first describe the target applications, which are (large scientific) workflows for which we
-    do not have a perfect a priori knowlodge,
+    We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
     in Section~\ref{sec.mod.work}.  Next, we define the execution
@@
     in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
-    Section~\ref{sec.mod.pb}. The key notations are summarized in Table~\ref{tabnotation}.
+    Section~\ref{sec.mod.pb}. The key notation is summarized in Table~\ref{tabnotation}.
 

### 2024-10-09 main.tex
started editing Sec. 2 (to use time before meeting)

@@
 \AB{I transformed the sum into a max, correct? Or maybe we could assume $r_u=m_u$? }
+\hmey{I  recall a discussion of the sum with Svetlana, but the text says max, it seems.}
 

### 2024-10-09 main.tex
done with editing Sec. 2

@@
 
-Furthermore, we place ourself in a context where we do not have a perfect knowledge
-of the task parameters ($w_u$ and $m_u$) before they start their execution,
+Furthermore, we place ourselves in a context where we do not have perfect knowledge
+of the task parameters ($w_u$ and $m_u$) before the tasks start their execution,
 but only estimates. \AB{Add motivation: related work with variable task durations for instance...}
@@
     We assume that all processors are connected with the same bandwidth~$\beta$.
+    \hmey{Maybe mention that variable bandwidths are part of future work...?}
 
     We keep track of the current ready time of each processor and each communication
-    channel, $\rt_j$ and $\rt_{j,j'}$, for all processors~$(j,j')$.
+    channel, $\rt_j$ and $\rt_{j,j'}$, for each processor $j$ and all pairs~$(j,j')$.
     Initially, all the ready times are set to~$0$.
@@
 
-The goal is to find a schedule of the DAG~$G$ onto the $k$ processors,
+The goal is to find a schedule of the DAG~$G$ for the $k$ processors,
 so that the makespan (total execution time) is minimized while
 respecting memory constraints. If a processor runs out of memory to execute
-a task, the schedule is said to be {\em invalid}.
+a task \hmey{mapped to it}, the schedule is said to be {\em invalid}.
 
@@
 on the estimated task parameters.
+
+\hmey{Mention NP-hardness due to being more general than NP-hard problem?}
 

### 2024-10-09 main.tex
editing Sec. 3 (WIP)

@@
 
-    We discuss relevant scheduling approaches that either reuse processors, or respect memory requirements of the processors.
+    We discuss relevant scheduling approaches that reuse processors or respect the memory requirements of the processors.
 
@@
     They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
-    with the speed of 1.
-
-    The \textit{task duplication}-based approaches note that sometimes running a task twice on different machines can
+    with the speed of 1.\hmey{needs refs or at least a survey pointer}
+
+    The \textit{task duplication}-based approaches exploit that sometimes running a task twice on different machines can
     help reduce the makespan by saving communication costs.
-    The two approaches are scheduling with the partial duplication~(SPD), and withe the full duplication~(SFD).
+    The two categories are scheduling with partial duplication~(SPD), and with full duplication~(SFD).
     For a join task (a task whose incoming degree is larger than its outgoing degree), SPD finds a critical immediate
-    parent (the one that gives the largest start time to the join task) and duplicate only it.
+    parent (the one that gives the largest start time to the join task) and duplicates only it.
     SFD duplicates all parents of a join node.
-    \cite{dfrn1997} is a task duplication algorithm.
-    The authors duplicate first (create copies of all parent tasks) and then reduce (remove) the ones that can be removed without harming
-    the makespan.
-    The critical path fast duplication algorithm \texttt{CPFD}~\cite{5727760} classifies tasks into three categories: a critical
-    path task, and in-branch task, or out-branch task.
+    Ref.~\cite{dfrn1997} proposes a task duplication algorithm. \hmey{The prev. sentence sounds odd since we are in the paragraph on task duplication.}
+    It duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
+    The critical path fast duplication algorithm \texttt{CPFD}~\cite{5727760} classifies tasks into three categories: critical
+    path task, in-branch task, or out-branch task.
     It schedules critical path tasks first, then in-branch tasks.
+    \hmey{I'm missing the connection to our paper. Or vice versa, how is our contribution
+    connected to these works? (Einordnung in den Forschungskontext) For example, do we do similar things? Are there interesting limitations we overcome?}
 
@@
     path and repeats the procedure.
-    \texttt{Heaviest node first} algorithm~\cite{SHIRAZI1990222} assigns the tasks level by level, and in each level,
-    schedules the heaviest (with largest computation time)  ones first.
+    \texttt{Heaviest node first}~\cite{SHIRAZI1990222} \hmey{Really typewriter for algorithms?
+    Plus: this is not done consistently yet!} assigns the tasks level by level; 
+    in each level, it schedules the one with largest computation time first.
 
@@
     Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
-    It and all its successors consist of two phases: task ordering, and task assignment.
+    It and all its successors consist of two phases: task prioritization/ordering and task assignment.
     In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
     and then schedule tasks in the order of these priorities.
-    The modifications of HEFT revolve around the way the priorities of tasks are computed, and the logic
-    of processor assignment.
+    The modifications of HEFT revolve around the way the priorities of the tasks are computed and the logic of the processor assignment.
     All such algorithms assume a heterogeneous execution environment.
 
-    So, during task prioritization phase~\cite{sulaiman2021hybrid} compute the standard deviation of the computation cost
-    (between processors) and add it to the mean value to account for the differences between processor speeds.
+    So, during the task prioritization phase, Ref.~\cite{sulaiman2021hybrid} computes the standard deviation of the computation cost
+    (between processors) and adds it to the mean value to account for the differences between processor speeds.
     In the processor choice phase, they duplicate the entry task and the longest parent tasks during idle times on the processor.
 
-    \cite{alebrahim2017task} compute a the bottom level based on the difference of execution times on
-    the fastest and the slowest processors, divided by the speedup (speeds ratio) of these two processors.
+    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
+    the fastest and the slowest processors, divided by the speed ratio of these two processors.
     When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
-    They choose the processor with the lowest execution time, and cross over to other processors sometimes.
-    They build upon~\cite{shetti2013optimization}.
+    They choose the processor with the lowest execution time and cross over to other processors sometimes.
+    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}
 

### 2024-10-09 main.tex
removed old numbers in experiments, new numbers will be tomroow

@@
 
-    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{}.
+    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{TODO}.
     Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
@@
 
-    On the normal cluster, the baseline produces invalid results in $40$\% of small workflows and $66.6$\% of middle-sized
-    and big workflows.
-    In most cases, the failures of the baseline can be traced to choosing the processor with not enough memory when several
-    options are available.
-    Due to not taking memories into account, the baseline chooses a processor with already full memory, while there can
-    be another processor with the same speed that still has enough memory.
-    In this case, \heftbl produces the same makespan while producing a valid solution.
-    The heuristics produce valid results by default and also manage to find a solution in all cases (on all workflow sizes).
-
-    The excess makespans produced by the heuristics are shown in Figure~\ref{fig:excess-ms}.
-    On the normal cluster and among small workflows, the makespans produced by \heftbl and \heftblc are virtually indistinguishable
-    from those of the baseline (no change and $0.4$\% change), while \heftmm produces  $18$\% worse makespans.
-    For middle-sized ones \heftmm is $9$\% worse,  \heftblc $0.9$\% worse.
-    For the big workflows however, while results of \heftbl and \heftblc are still barely worse ($<1$\% worse),
-     \heftmm is only $5.8$\% worse.
-    The heuristics pay off more on larger workflows.
-
-    \begin{figure}[tb]
-        \centering
-        \includegraphics[width=0.495\columnwidth] {images/ms-relative-3groups}
-        \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm for different . Smaller is better.}
-       \label{fig:excess-ms}
-        \vspace{-0.3cm}
-    \end{figure}
-    \subsubsection{Memory usage}
-    Minimizing memory usage is an interesting measure of the quality of the solution produced by a heuristic.
-    We studied the average peak memory consumption in our cluster.
-    We looked at the maximum memory usage on each processor during a task execution, and averaged it over all processors
-    in the execution environment.
-
-    The results adhere to the intuition: \heftmm produces consistently smaller memory usage among all the heuristics.
-    So, for small workflows, the average memory occupation by \heftbl and \heftblc is $11.8$\% and $11.7$\%, respectively,
-    while \heftmm manged to occupy $10.2$\% of memories.
-    For middle-sized workflows, \heftbl and \heftblc occupied $27.8$\% and $28.2$\%, respectively, while \heftmm occupied
-    $21$\%.
-    For the big workflows, \heftmm occupied $20.9$\% against $33.7$\% by \heftbl and $27.6$\% by \heftblc.
-    Interestingly, for big workflows, \heftmm occupied on average even a little less that for the middle-sized ones.
-    This means that the bound of exploiting parallelization has been achieved for workflows around the size of middle-sized ones,
-    and for the big workflows, the makespan is not improved by ``spreading'' the workflow more.
-
-    \subsubsection{Impact of cluster size}
-    On a cluster with only $18$ processors ($3$ of each kind), the baseline makes more failures: from $43.8$\%  (small workflows)
-    to $77.8$\% (mid-sized) and $88.9$\% (big workflows).
-    At the same time, each heuristic is unable to schedule one big workflow and one middle-sized one.
-    On the workflows that they are able to schedule, \heftbl and \heftblc produce makespans that are $<1$\% bigger than
-    those of the baseline on small and middle-sized workflows.
-    On the big workflows, they produce makespans that are $3.5$\% and $2.3$\% bigger respectively.
-    The performance of \heftmm improves more with workflow size: it goes from producing $16$\% bigger makespans on small
-    workflows, to $8$\% on middle-sized ones, to $7$\% on big workflows.
-
-   Like in the normal cluster, \heftmm again consistently shows smaller peak memory usage in the cluster.
-    It utilizes on average $15$\% of memory on small workflows ($17.7$\% for \heftbl and \heftblc),
-    $31$\% on middle-sied workflows ($45.2$\% and $44.9$\% on \heftbl and \heftblc, respectively),
-    and $34$\% on big workflows ($50.1$\% and $50.2$\% for \heftbl and \heftblc).
+    Preliminary data: HEFT has 50\% failures, our algorithms have no failures.
+    memory utilization blc is 5\% better, makespan 1\% worse.
 

### 2024-10-10 main.tex
temp - there are some missing references

@@
 
+\renewcommand{\iec}{i.e., }
 \begin{document}
@@
     Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
-    and memory sizes. In general, varying memory sizes make it more challenging to compute
+    and memory sizes. In general, having different memory sizes per CPUs makes it more challenging to compute
     a schedule that respects all memory constraints -- meaning that no task is executed on a 
@@
 
-    A seminal list scheduling heuristic for workflows is HEFT 
+    A seminal list scheduling heuristic for workflows on heterogeneous platforms is HEFT 
     (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
-    It has two phases: (i) each task is assigned a priority and (ii) the tasks in a priority-ordered list are assigned
-    to processors, where the ``ready'' task with the highest priority is scheduled next.
-    HEFT has been extended (e.g.\ by Shi and Dongarra~\cite{SHI2006665}) and adjusted for a variety of different scheduling problem formulations. 
-    Yet, none of them adheres to memory constraints -- also see our discussion of related work
-    in Section~\ref{sec:related-work}. (\skug{check in related work if true!}).
-        \skug{Note: 2 papers that deal with memory sizes, but model is very different!}
+    It has two phases: (i) each task is assigned a priority; and (ii) the tasks in a priority-ordered list are assigned
+    to processors, where the ``ready'' task with the highest priority is scheduled next on the processor
+    where it would complete its execution first. 
+    HEFT has been extended (e.g., by Shi and Dongarra~\cite{SHI2006665}) and adjusted 
+    for a variety of different scheduling problem formulations. 
+    Yet, none of them adhere to memory constraints as we propose -- see discussion of related work
+    in Section~\ref{sec:related-work}. 
+%    (\skug{check in related work if true!}).
+%        \skug{Note: 2 papers that deal with memory sizes, but model is very different!}
 
     Another limitation in practice of HEFT (and many other scheduling strategies) is their 
-    assumption that the task running times provided to them are accurate. In practice this is 
+    assumption that the task running times provided to them are accurate. In practice, this is 
     not the case and deviations from user estimates or historical measurements are 
@@
     The difference between the new heuristics is the way they prioritize tasks (for processor assignment).
-    We compare the static schedule produced by these heuristics with a HEFT variant as
-    baseline that does not take memory sizes into account. \hmey{Need to clarify: 
-    why is this SotA?}
+    We compare the static schedule produced by these heuristics with the classical HEFT as
+    baseline, which however does not take memory sizes into account. 
+    \hmey{Need to clarify: why is this SotA?}
 
@@
 
-    \textit{Assignment to processor}
-    When assigning the task, we set the ready time of the processor~$j$ $rt_j$ to the finish time of the task.
+    \textit{Assignment to processor. }
+    When assigning the task, we set the ready time $rt_j$ of  processor~$j$ to be the finish time of the task.
     For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
@@
 
+Extend to heterogeneous bandwidths.
+
 Further variability commented out from the paper: changing DAG structure; changing platform...

### 2024-10-10 main.tex
some remaining edits in rel. work I forgot to commit yesterday

--- 2024-10-09 main.tex
@@
 
-    PEFT (Predict earliest finish time) algorithm~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
-    Cost Table OCT.
-    OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task, if this
+    PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
+    Cost Table (OCT).
+    The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
     processor is chosen for this task.
     Ranking is based on OCT values.
-    Processor choice stage minimizes the Optimistic EFT, which is EFT plus the the longest path to the exit node for each task.
+    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.
 

### 2024-10-10 main.tex
several missing refs added

@@
     avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
-    Hence, building on previous related work~\cite{TODO}, we consider a scheduling problem 
+    Hence, building on previous related \hmey{You may want to add other refs not from us} work~\cite{DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
     formulation that takes memory sizes as explicit constraints into account. Its objective is
@@
     not the case and deviations from user estimates or historical measurements are 
-    very common~\cite{TODO}. As a consequence, one should adapt the schedule when \emph{major}
+    very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
     deviations occur. List-based schedulers such as HEFT are, however, not designed for 
-    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement?}
+    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
 %    and would compute a completely new schedule from scratch.

### 2024-10-10 main.tex
minor changes in related work, addressed 2 of Henning's notes, added first picture to static experiments

--- 2024-10-10 main.tex
@@
     \hmey{I  recall a discussion of the sum with Svetlana, but the text says max, it seems.}
+    \skug{it used to be  a sum, but not anymore since the collaboration with B1}
 
@@
     They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
-    with the speed of 1.\hmey{needs refs or at least a survey pointer}
+    with the speed of 1~\cite{benoit2013survey}.\hmey{needs refs or at least a survey pointer}
+    \skug{actually, Anne cited them, so citing her :-)}
 
@@
     SFD duplicates all parents of a join node.
-    Ref.~\cite{dfrn1997} proposes a task duplication algorithm. \hmey{The prev. sentence sounds odd since we are in the paragraph on task duplication.}
-    It duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
-    The critical path fast duplication algorithm \texttt{CPFD}~\cite{5727760} classifies tasks into three categories: critical
+    The algorithm by~\cite{dfrn1997} duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
+    The critical path fast duplication algorithm CPFD~\cite{5727760} classifies tasks into three categories: critical
     path task, in-branch task, or out-branch task.
@@
     path and repeats the procedure.
-    \texttt{Heaviest node first}~\cite{SHIRAZI1990222} \hmey{Really typewriter for algorithms?
-    Plus: this is not done consistently yet!} assigns the tasks level by level; 
+    Heaviest node first~\cite{SHIRAZI1990222} assigns the tasks level by level;
     in each level, it schedules the one with largest computation time first.
@@
 
-    PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
-    Cost Table (OCT).
-    The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
+    PEFT (Predict earliest finish time) algorithm~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
+    Cost Table OCT.
+    OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task, if this
     processor is chosen for this task.
     Ranking is based on OCT values.
-    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.
+    Processor choice stage minimizes the Optimistic EFT, which is EFT plus the the longest path to the exit node for each task.
 
@@
 
-    The \texttt{SDC} algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
+    The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
     average execution cost in its weight.
@@
     \paragraph{Normal cluster}
-    \skug{OLD DATA here, rewrite}
 
-    Preliminary data: HEFT has 50\% failures, our algorithms have no failures.
-    memory utilization blc is 5\% better, makespan 1\% worse.
+    On a normal cluster, the baseline \heft had $24.2\%$ success rate ($75.7\%$ failure rate).
+    All the heuristics could schedule all workflows. \skug{check this tomorrow for two weird failures}.
+
+    Figure~\ref{fig:ms-relations} shows the relations of the makespans foound by the heuristics to the makespans found by the baseline.
+    On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline, the makespans of \heftblc are $8\%$
+    worse, and those found by \heftmm are $71\%$ worse.
+
+    \skug{TODO: how much worse are makespans for SUCCESSFUL hefts? presulmably almost identical}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=0.495\columnwidth] {images/MsRelations}
+        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.}
+        \label{fig:ms-relations}
+        \vspace{-0.3cm}
+    \end{figure}
 

### 2024-10-10 main.tex
temp

@@
     \]
-    \AB{I transformed the sum into a max, correct? Or maybe we could assume $r_u=m_u$? }
-    \hmey{I  recall a discussion of the sum with Svetlana, but the text says max, it seems.}
-    \skug{it used to be  a sum, but not anymore since the collaboration with B1}
 
@@
 respecting memory constraints. If a processor runs out of memory to execute
-a task \hmey{mapped to it}, the schedule is said to be {\em invalid}.
+a task mapped on it, the schedule is said to be {\em invalid}.
 
@@
     We need to evict at least $Res$ data.
-    For now, we propose a greedy approach, evicting the smallest files of $\PD_j$ until $Res$ data has been evicted,
-    in order to avoid costly communications.
-    \AB{FYI We initially discussed evicting the largest files, but this leads to
-    large communications and does not seem efficient after all... Maybe we can think of another
-    approach that would take into account both data size and bottom level...}
+    We propose a greedy approach, evicting the largest files of $\PD_j$ until $Res$ data has been evicted. 
+    A variant where the smallest files are evicted first has been tested, and it led to comparable results. 
+%    
+%    in order to avoid costly communications.
+%    \AB{FYI We initially discussed evicting the largest files, but this leads to
+%    large communications and does not seem efficient after all... Maybe we can think of another
+%    approach that would take into account both data size and bottom level...}
     While tentatively evicting files, we remove them from the list of pending memories and move them into a list
     of memories pending in the communication buffer.
-    We keep track of the available buffer size, too - each time a file gets moved into the pending in buffer, the available buffer size is reduced by its weight.
+    We keep track of the available buffer size, too -- each time a file gets moved into the pending in buffer, 
+    the available buffer size is reduced by its weight.
 
@@
     must be orchestrated from predecessor tasks $u\notin T(p_j)$.
-    The starting time is then
-    \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} } \]
-    Its finish time on $p_j$ is then
+    The starting time is therefore:\\[-.7cm]
+
+   {\footnotesize{ \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} .} \]}}
+
+  \noindent  Finally, its finish time on $p_j$ is 
     $FT(v,p_j) = ST(v, p_j) + \frac{w_v}{s_j}$.
@@
 
-    \subsection{Dynamic Scenario}
+    \section{Dynamic scenario}
 
@@
 
-    \subsection{Retracing the effects of change on an existing schedule}
+    \subsection*{Retracing the effects of change on an existing schedule}
     After the monitoring system has reported changes, we need to assess their impact on the existing schedule.

### 2024-10-10 main.tex
ms relations also with only valid hefts
temp

@@
     avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
-    Hence, building on previous related \hmey{You may want to add other refs not from us} work~\cite{DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
+    Hence, building on previous related \hmey{You may want to add other refs not from us} work~\cite{gou2020partitioning,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
     formulation that takes memory sizes as explicit constraints into account. Its objective is
@@
     such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
+    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
 %    and would compute a completely new schedule from scratch.

### 2024-10-10 main.tex
pictues on: success rates on mem-constrained cluster, memory usage by algorithm, makespan boxplots for larger cluster with invalid and only valid values. Almost zero text yet
temp, reorganizing

@@
     avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
-    Hence, building on previous related \hmey{You may want to add other refs not from us} work~\cite{gou2020partitioning,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
+    Hence, building on previous related %\hmey{You may want to add other refs not from us} 
+    work~\cite{gou2020partitioning,He21,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
     formulation that takes memory sizes as explicit constraints into account. Its objective is
@@
 
-    A seminal list scheduling heuristic for workflows on heterogeneous platforms is HEFT 
+    While previous work with memory constraints has been focusing on partitioning the graph, 
+    and not reusing processors during execution, 
+    a seminal list scheduling heuristic for workflows on heterogeneous platforms, without accounting
+    for the memory constraint, is HEFT 
     (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
@@
     \end{itemize}
+
+    \section{Related work} %: \skug{Full: 5, Polished: 4}}
+    \label{sec:related-work}
+
+   First, we focus on HEFT-like scheduling heuristics from the literature  that do not necessarily  
+   consider memory constraints. Then, we discuss memory-aware scheduling algorithms. 
+   Finally, we move to related work on dynamic or adaptive algorithms. 
+ %   \hmey{Question: Does it make sense to place related work before the model (which is more common)?}
+
+%    We discuss relevant scheduling approaches that reuse processors or respect the memory requirements of the processors.
+%
+%    \subsection{Early list schedulers with unlimited processors}
+%    An entire cluster of works on list schedulers has been carried out as early as the 90s.
+%    They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
+%    with the speed of 1~\cite{benoit2013survey}.\hmey{needs refs or at least a survey pointer}
+%    \skug{actually, Anne cited them, so citing her :-)}
+%
+%    The \textit{task duplication}-based approaches exploit that sometimes running a task twice on different machines can
+%    help reduce the makespan by saving communication costs.
+%    The two categories are scheduling with partial duplication~(SPD), and with full duplication~(SFD).
+%    For a join task (a task whose incoming degree is larger than its outgoing degree), SPD finds a critical immediate
+%    parent (the one that gives the largest start time to the join task) and duplicates only it.
+%    SFD duplicates all parents of a join node.
+%    The algorithm by~\cite{dfrn1997} duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
+%    The critical path fast duplication algorithm CPFD~\cite{5727760} classifies tasks into three categories: critical
+%    path task, in-branch task, or out-branch task.
+%    It schedules critical path tasks first, then in-branch tasks.
+%    \hmey{I'm missing the connection to our paper. Or vice versa, how is our contribution
+%    connected to these works? (Einordnung in den Forschungskontext) For example, do we do similar things? Are there interesting limitations we overcome?}
+%
+%    Linear clustering~\cite{KWOK1999381} acts on critical paths in the workflow.
+%    It assigns the current critical path to one processor, removes all these tasks from the workflow, recomputes the critical
+%    path and repeats the procedure.
+%    Heaviest node first~\cite{SHIRAZI1990222} assigns the tasks level by level;
+%    in each level, it schedules the one with largest computation time first.
+
+    \subsection{Static list schedulers, especially HEFT-based algorithms}
+
+    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
+    It and all its successors consist of two phases: task prioritization/ordering and task assignment.
+    In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
+    and then schedule tasks in the order of these priorities.
+    The modifications of HEFT revolve around the way the priorities of the tasks are computed and the logic of the processor assignment.
+    All such algorithms assume a heterogeneous execution environment.
+
+    So, during the task prioritization phase, Ref.~\cite{sulaiman2021hybrid} computes the standard deviation of the computation cost
+    (between processors) and adds it to the mean value to account for the differences between processor speeds.
+    In the processor choice phase, they duplicate the entry task and the longest parent tasks during idle times on the processor.
+
+    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
+    the fastest and the slowest processors, divided by the speed ratio of these two processors.
+    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
+    They choose the processor with the lowest execution time and cross over to other processors sometimes.
+    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}
+
+    PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
+    Cost Table (OCT).
+    The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
+    processor is chosen for this task.
+    Ranking is based on OCT values.
+    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.
+
+    The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
+    comparison to HEFT.
+    It combines the standard deviation with the communication cost weight on the tasks.
+    In the second stage, the algorithm duplicates the entry task if there is a need for it.
+
+    The TSHCS (Task Scheduling for Heterogeneous Comupting Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
+    by adding randomized decisions to the second phase.
+    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
+    produces the lowest finish time.
+
+    The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
+    average execution cost in its weight.
+    The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT) that
+    additionally notes how large the communication between current node and its children will be on the
+    average provided that it is scheduled on the current processor.
+
+    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement
+    learning techniques(\cite{yano2022cqga}).
+
+    \subsection{Memory-aware scheduling algorithms}
+    Respecting processor memories adds a constraint to a scheduling problem.
+    Therefore, only specifically memory-targeted algorithms address this issue.
+    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
+    is formulated and addressed in the algorithm.
+
+    Different models of memory available on processors and memory requirements of tasks have been presented.
+
+    Marchal~et~al.~\cite{marchal2018parallel} assume a memory model where each processor has an individual memory available.
+    Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
+    A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a DAG is provided,
+    as well as an ILP solution to the scheduling problem.
+    The memory model requires deleting all input data upon starting of the task and adding all output files there.
+
+    In an assumed dual-memory systems(\cite{herrmann2014memory}) a processor can have access to a memory of two different
+    kinds (red or blue), and each task can be executed on only one sort of memory.
+    The communications happen only between these two kinds of processors (communications withing each group are ignored).
+    The authors then formulate an ILP solution for this problem formulation.
+
+    The algorithm presented by Yao \etal~\cite{yao2022memory}, each processor has an own internal memory and all
+    processors share a common external one. The internal (local) memory is used to store the task files.
+    The external memory is used to store evicted files to make room for the execution of a task on a processor.
+    All processors, including the original one, can access these files.
+    Each edge in~\cite{yao2022memory} has two weights -- the size of the files transferred along it,
+    and the time of communication along this edge.
+    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
+
+    \cite{ding2024ils} assumes connected processors with individual limited memories.
+    The collective set of memories forms the global memory, to which each processor has access, however the access time
+    to global memory is different.
+    Each memory access in the graph is modelled as a memory access token on the task, while the edges have no weights.
+    The solved problem is how to allocate initial input data in processor memories so that the overall
+    execution is minimized and the memories are not exceeded.
+    They propose an integer linear programming model.
+    %that minimizes the length of the critical path, including a greedy initial solution.
+
+    In~\cite{rodriguez2019exploration}, the authors assume memory requirement on tasks represented as tiles.
+    Each processor has individual memories to process the task, but only the shared memories store the tiles containing
+    memory tiles occupied by memory tiles.
+
+    Cloud-oriented model can include costs associated with memory usage(\cite{liang2020memory}).
+
+    \subsection{Dynamic/adaptive algorithms}
+
+    DVR HEFT~\cite{SANDOKJI2019482} uses an almost unchanged HEFT algorithm in their static step, executing three slightly
+    varying variants of task weighting and choosing the variant that gives the best overall makespan.
+    In the dynamic phase, they receive new tasks and schedule them on either idle processors or those processors that give them
+    the earliest finish time.
+    Task failures are not covered.
+
+    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
+    by calculating the critical path in the graph dynamically at every step.
+    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
+    %on the start time of a task (differing by the slack this task has).
+    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
+    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
+    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
+    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
+    %execution time for the task is chosen.
+    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
+    The approach is evaluated on random workflows of the size up to 300 tasks.
+
+    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
+    The procedure involves building a first (static) schedule, periodic resource monitoring and rescheduling the remaining
+    tasks.
+    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
+    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
+    tasks to groups.
+    The static heuristic is HEFT with earliest start time as priority.
+    Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
+    is smaller than the previous one.
+    The experiments were conducted on a single workflow with 10 tasks.
+%
+    % The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
+    %The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
+
+    %The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
+    %THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
+    %They then go through tasks and map than to such processors that minimize the earliest start times of this task's
+    %successors.
+    %To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
+    %average communication and computation costs fir the dependent tasks.
+    %
+    %The rescheduling is triggered when either a load on a resource increases over a threshold, or if a new resource
+    %is added.
+
+    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
+    but constrained by cost.
+
+    Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
+    Particles are possible solution in the solution space.
+    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
+    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
+
+    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling algorithm
+    for clouds.
+    They solve a set of linear equations that represent the cost of an execution based on the criteria.
+    The authors test out 4 scenarios - one preferring each criteria, and a balanced one.
+    The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
+    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
+    %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
+ %   The authors use provenance data to make scheduling decisions.
+
+    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
+    on a single machine.
+    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
+    that these runtime changes are unrelated to each other).
+    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
+    performance measure $\phi$.
+    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
+
+    \subsection{Other notable works}
+
+    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
+    They utilize task duplication when creating the clusters (grains).
+    Their scheduler then maps clusters to processors.
+    They assume unlimited processors with the speed of 1.
+    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
+    to it as possible.
+    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
+    %release times.
+
+    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
+    to search for an optimal solution for mapping tasks on machines.
+    A solution is generated at each step, and the best solution is kept at the end.
+    The search terminates when a certain termination criterion is reached.
+    It generates better results than other algorithms, because it explores the whole solution space.
+
+    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
+    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
+    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
+    The heuristic first determines similar activities and groups them together.
+    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
+    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
+    The heuristic calculates a new execution path with these tasks.
+
+    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
+    (processor) speed heterogeneity.
+    Each accelerator has its own memory that can be zero.
+    Each accelerator's characteristics depend on the task it runs and are not fixed.
+
+    \cite{meng2018traffic} investigate scheduling on multi-core chips.
+    Their model is far from ours.
+
+    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
+    They prioritize tasks that have failed before or are well-predictable.
 
@@
 
-    \section{Related work: \skug{Full: 5, Polished: 4}}
-    \label{sec:related-work}
-
-    \hmey{Question: Does it make sense to place related work before the model (which is more common)?}
-
-    We discuss relevant scheduling approaches that reuse processors or respect the memory requirements of the processors.
-
-    \subsection{Early list schedulers with unlimited processors}
-    An entire cluster of works on list schedulers has been carried out as early as the 90s.
-    They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
-    with the speed of 1~\cite{benoit2013survey}.\hmey{needs refs or at least a survey pointer}
-    \skug{actually, Anne cited them, so citing her :-)}
-
-    The \textit{task duplication}-based approaches exploit that sometimes running a task twice on different machines can
-    help reduce the makespan by saving communication costs.
-    The two categories are scheduling with partial duplication~(SPD), and with full duplication~(SFD).
-    For a join task (a task whose incoming degree is larger than its outgoing degree), SPD finds a critical immediate
-    parent (the one that gives the largest start time to the join task) and duplicates only it.
-    SFD duplicates all parents of a join node.
-    The algorithm by~\cite{dfrn1997} duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
-    The critical path fast duplication algorithm CPFD~\cite{5727760} classifies tasks into three categories: critical
-    path task, in-branch task, or out-branch task.
-    It schedules critical path tasks first, then in-branch tasks.
-    \hmey{I'm missing the connection to our paper. Or vice versa, how is our contribution
-    connected to these works? (Einordnung in den Forschungskontext) For example, do we do similar things? Are there interesting limitations we overcome?}
-
-    Linear clustering~\cite{KWOK1999381} acts on critical paths in the workflow.
-    It assigns the current critical path to one processor, removes all these tasks from the workflow, recomputes the critical
-    path and repeats the procedure.
-    Heaviest node first~\cite{SHIRAZI1990222} assigns the tasks level by level;
-    in each level, it schedules the one with largest computation time first.
-
-    \subsection{Static list schedulers, especially HEFT-based algorithms}
-
-    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
-    It and all its successors consist of two phases: task prioritization/ordering and task assignment.
-    In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
-    and then schedule tasks in the order of these priorities.
-    The modifications of HEFT revolve around the way the priorities of the tasks are computed and the logic of the processor assignment.
-    All such algorithms assume a heterogeneous execution environment.
-
-    So, during the task prioritization phase, Ref.~\cite{sulaiman2021hybrid} computes the standard deviation of the computation cost
-    (between processors) and adds it to the mean value to account for the differences between processor speeds.
-    In the processor choice phase, they duplicate the entry task and the longest parent tasks during idle times on the processor.
-
-    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
-    the fastest and the slowest processors, divided by the speed ratio of these two processors.
-    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
-    They choose the processor with the lowest execution time and cross over to other processors sometimes.
-    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}
-
-    PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
-    Cost Table (OCT).
-    The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
-    processor is chosen for this task.
-    Ranking is based on OCT values.
-    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.
-
-    The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
-    comparison to HEFT.
-    It combines the standard deviation with the communication cost weight on the tasks.
-    In the second stage, the algorithm duplicates the entry task if there is a need for it.
-
-    The TSHCS (Task Scheduling for Heterogeneous Comupting Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
-    by adding randomized decisions to the second phase.
-    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
-    produces the lowest finish time.
-
-    The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
-    average execution cost in its weight.
-    The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT) that
-    additionally notes how large the communication between current node and its children will be on the
-    average provided that it is scheduled on the current processor.
-
-    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement
-    learning techniques(\cite{yano2022cqga}).
-
-    \subsection{Memory-aware scheduling algorithms}
-    Respecting processor memories adds a constraint to a scheduling problem.
-    Therefore, only specifically memory-targeted algorithms address this issue.
-    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
-    is formulated and addressed in the algorithm.
-
-    Different models of memory available on processors and memory requirements of tasks have been presented.
-
-    Marchal~et~al.~\cite{marchal2018parallel} assume a memory model where each processor has an individual memory available.
-    Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
-    A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a DAG is provided,
-    as well as an ILP solution to the scheduling problem.
-    The memory model requires deleting all input data upon starting of the task and adding all output files there.
-
-    In an assumed dual-memory systems(\cite{herrmann2014memory}) a processor can have access to a memory of two different
-    kinds (red or blue), and each task can be executed on only one sort of memory.
-    The communications happen only between these two kinds of processors (communications withing each group are ignored).
-    The authors then formulate an ILP solution for this problem formulation.
-
-    The algorithm presented by Yao \etal~\cite{yao2022memory}, each processor has an own internal memory and all
-    processors share a common external one. The internal (local) memory is used to store the task files.
-    The external memory is used to store evicted files to make room for the execution of a task on a processor.
-    All processors, including the original one, can access these files.
-    Each edge in~\cite{yao2022memory} has two weights -- the size of the files transferred along it,
-    and the time of communication along this edge.
-    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
-
-    \cite{ding2024ils} assumes connected processors with individual limited memories.
-    The collective set of memories forms the global memory, to which each processor has access, however the access time
-    to global memory is different.
-    Each memory access in the graph is modelled as a memory access token on the task, while the edges have no weights.
-    The solved problem is how to allocate initial input data in processor memories so that the overall
-    execution is minimized and the memories are not exceeded.
-    They propose an integer linear programming model.
-    %that minimizes the length of the critical path, including a greedy initial solution.
-
-    In~\cite{rodriguez2019exploration}, the authors assume memory requirement on tasks represented as tiles.
-    Each processor has individual memories to process the task, but only the shared memories store the tiles containing
-    memory tiles occupied by memory tiles.
-
-    Cloud-oriented model can include costs associated with memory usage(\cite{liang2020memory}).
-
-    \subsection{Dynamic/adaptive algorithms}
-
-    DVR HEFT~\cite{SANDOKJI2019482} uses an almost unchanged HEFT algorithm in their static step, executing three slightly
-    varying variants of task weighting and choosing the variant that gives the best overall makespan.
-    In the dynamic phase, they receive new tasks and schedule them on either idle processors or those processors that give them
-    the earliest finish time.
-    Task failures are not covered.
-
-    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
-    by calculating the critical path in the graph dynamically at every step.
-    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
-    %on the start time of a task (differing by the slack this task has).
-    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
-    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
-    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
-    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
-    %execution time for the task is chosen.
-    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
-    The approach is evaluated on random workflows of the size up to 300 tasks.
-
-    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
-    The procedure involves building a first (static) schedule, periodic resource monitoring and rescheduling the remaining
-    tasks.
-    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
-    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
-    tasks to groups.
-    The static heuristic is HEFT with earliest start time as priority.
-    Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
-    is smaller than the previous one.
-    The experiments were conducted on a single workflow with 10 tasks.
-%
-    % The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
-    %The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
-
-    %The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
-    %THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
-    %They then go through tasks and map than to such processors that minimize the earliest start times of this task's
-    %successors.
-    %To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
-    %average communication and computation costs fir the dependent tasks.
-    %
-    %The rescheduling is triggered when either a load on a resource increases over a threshold, or if a new resource
-    %is added.
-
-    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
-    but constrained by cost.
-
-    Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
-    Particles are possible solution in the solution space.
-    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
-    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
-
-    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling algorithm
-    for clouds.
-    They solve a set of linear equations that represent the cost of an execution based on the criteria.
-    The authors test out 4 scenarios - one preferring each criteria, and a balanced one.
-    The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
-    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
-    %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
- %   The authors use provenance data to make scheduling decisions.
-
-    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
-    on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
-    that these runtime changes are unrelated to each other).
-    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
-    performance measure $\phi$.
-    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
-
-    \subsection{Other notable works}
-
-    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
-    They utilize task duplication when creating the clusters (grains).
-    Their scheduler then maps clusters to processors.
-    They assume unlimited processors with the speed of 1.
-    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
-    to it as possible.
-    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
-    %release times.
-
-    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
-    to search for an optimal solution for mapping tasks on machines.
-    A solution is generated at each step, and the best solution is kept at the end.
-    The search terminates when a certain termination criterion is reached.
-    It generates better results than other algorithms, because it explores the whole solution space.
-
-    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
-    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
-    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
-    The heuristic first determines similar activities and groups them together.
-    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
-    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
-    The heuristic calculates a new execution path with these tasks.
-
-    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
-    (processor) speed heterogeneity.
-    Each accelerator has its own memory that can be zero.
-    Each accelerator's characteristics depend on the task it runs and are not fixed.
-
-    \cite{meng2018traffic} investigate scheduling on multi-core chips.
-    Their model is far from ours.
-
-    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
-    They prioritize tasks that have failed before or are well-predictable.
-
     \section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
@@
         \centering
-        \includegraphics[width=0.495\columnwidth] {images/MsRelations}
+        \includegraphics[width=0.495\columnwidth] {images/MsRelations2}
         \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.}
         \label{fig:ms-relations}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \paragraph{Memory Usage}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal}
+        \caption{Memory usage by algorithm by workflow size. Includes invalid heft schedules}
+        \label{fig:mem-usages}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
+        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules}
+        \label{fig:mem-usages-onlyvalid}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \paragraph{Memory-constrained cluster}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/success-rates-tiny}
+        \caption{Success ratey by algorithm by workflow size. Taller is better. }
+        \label{fig:success-rates-tiny}
         \vspace{-0.3cm}

### 2024-10-10 main.tex
pictues on: makespan relations by size, 2 pictures on runtimes

--- 2024-10-10 main.tex
@@
         \centering
-        \includegraphics[width=0.495\columnwidth] {images/MsRelations}
+        \includegraphics[width=1.1\columnwidth] {images/MsRelations2}
         \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.}
         \label{fig:ms-relations}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size}
+        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow size. Smaller is better.}
+        \label{fig:ms-relations-by-workflow}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \paragraph{Memory Usage}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal}
+        \caption{Memory usage by algorithm by workflow size. Includes invalid heft schedules}
+        \label{fig:mem-usages}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
+        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules}
+        \label{fig:mem-usages-onlyvalid}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \paragraph{Memory-constrained cluster}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/success-rates-tiny}
+        \caption{Success ratey by algorithm by workflow size. Taller is better. }
+        \label{fig:success-rates-tiny}
         \vspace{-0.3cm}
@@
 
+    \subsubsection{Runtimes of the algorithm}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/runtimes-logarithmic}
+        \caption{Runtimes of the heuristics and the baseline. The y axis ist logarithmic.}
+        \label{fig:runtimes-log}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/runtimes-absolute}
+        \caption{Runtimes of the heuristics and the baseline.}
+        \label{fig:runtimes-abs}
+        \vspace{-0.3cm}
+    \end{figure}
+
     \subsubsection{Summary}

### 2024-10-10 main.tex
temp

--- 2024-10-10 main.tex
@@
     The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
-    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liuPacitti}.
+    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liu2018survey}.
     Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
@@
         \centering
-        \includegraphics[width=0.495\columnwidth] {images/MsRelations2}
+        \includegraphics[width=1.1\columnwidth] {images/MsRelations2}
         \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.}
         \label{fig:ms-relations}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size}
+        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow size. Smaller is better.}
+        \label{fig:ms-relations-by-workflow}
         \vspace{-0.3cm}
@@
 
-    \bibliographystyle{IEEEtran}
+    \bibliographystyle{abbrv}%IEEEtran}
     \bibliography{references}

### 2024-10-10 main.tex
picture on usccess rates on the larger cluster

@@
     The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
-    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liu2018survey}.
+    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liuPacitti}.
     Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
@@
         \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
-        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules}
+        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules \skug{THIS FIGURE IS WRONG}}
         \label{fig:mem-usages-onlyvalid}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/success-rates-large}
+        \caption{Success rates on the \skug{normal} cluster.}
+        \label{fig:success-rates-large}
         \vspace{-0.3cm}
@@
 
-    \bibliographystyle{abbrv}%IEEEtran}
+    \bibliographystyle{IEEEtran}
     \bibliography{references}

### 2024-10-10 main.tex
temp

--- 2024-10-10 main.tex
@@
 of the task parameters ($w_u$ and $m_u$) before the tasks start their execution,
-but only estimates. \AB{Add motivation: related work with variable task durations for instance...}
+but only estimates~\cite{rahman2013,GARG2015256}.  
+%. \AB{Add motivation: related work with variable task durations for instance...}
 Hence, scheduling decisions are made on these estimated parameters, and
-may be reconsidered at runtime when a task starts its execution and we know
-its exact parameters.
+may be reconsidered at runtime when a task starts its execution and we know its exact parameters.
 
@@
     We assume that all processors are connected with the same bandwidth~$\beta$.
-    \hmey{Maybe mention that variable bandwidths are part of future work...?}
+%    \hmey{Maybe mention that variable bandwidths are part of future work...?}
 
@@
 
-    \begin{table}[h]
+    \begin{table}
         \begin{center}
@@
 
-\hmey{Mention NP-hardness due to being more general than NP-hard problem?}
+Note that the problem is already NP-hard even in the homogeneous case and 
+without memory constraints, because of the DAG structure of the application. 
+Hence, we focus on the design of efficient scheduling heuristics. 
+%\hmey{Mention NP-hardness due to being more general than NP-hard problem?}
 
@@
     first, they compute the task ranks,
-    and second, they assign tasks to processors in the order defined in the first step.
+    and second, they assign tasks to processors in the order defined in the first stage.
     We consider three variants of HEFT accounting for memory usage (HEFTM), which only
@@
     \noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
-    \[Res = availM_j - m_v - \sum_{u \in \Pi(v), u\notin T(p_j)}  \{c_{u,v}\}
-    - \sum_{w\in Succ(v)}  \{c_{v,w}\}.\]
+    $$Res = availM_j - m_v - \!\!\!\! \sum_{u \in \Pi(v), u\notin T(p_j)}  \!\!\!\!\!\!\!\!\{c_{u,v}\}
+    - \sum_{w\in Succ(v)} \!\!\!\!\!\! \{c_{v,w}\}.$$
 
@@
         processors, and increase the available memories or available buffer sizes on these processors.
-        \item We compute the correct amount of available memory for $p_j$ (for when the task is done).
-        For each predecessor that is mapped to the same processor, we remove the pending memory corresponding to the weight of
+        \item We compute the correct amount of available memory for~$p_j$ (for when the task is done).
+        Then, for each predecessor that is mapped to the same processor, 
+        we remove the pending memory corresponding to the weight of
         the incoming edge, also freeing the same amount of available memory (increasing $availM_j$).
-        For each successor, on the other hand, we add the edge weights to pending memories and reduce $availM_j$ by the corresponding
-        amount.
+        For each successor, we rather add the edge weights to pending memories and reduce $availM_j$ 
+        by the corresponding amount.
     \end{itemize}
 
-    \subsection{The fork}
-    We look at the behavior of these heuristics on a fork graph,
-    where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
-    to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
-
-    Without memory, this problem is NP-complete; this is equivalent
-    to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
-    and with two processors. Half of the tasks must be sent to the processor
-    on which $T_0$ is not executed, and the optimal makespan is
-    $w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
-
-    However, with an infinite number of identical processors, it can be
-    solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$;
-    the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
-    while the remaining $n-k$ tasks are executed locally (try all values of $k$).
-
-    With heterogeneous processors, it is probably NP-complete again
-    because we could ensure that there are only two processors fast enough
-    and get back to the 2-partition...
-
-    We also had an example where evicting large files first in step 2
-    can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
-    $f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
-    imposes that we free one unit of memory before executing one
-    of the tasks\ldots Actually the new version with BLC would start
-    considering $T_2$ and be fine in this case\ldots
-
-    \AB{Can we prove that we have (maybe) a 2-approximation,
-        at least for the fork? What worst-case can we think of? }
+%    \subsection{The fork}
+%    We look at the behavior of these heuristics on a fork graph,
+%    where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
+%    to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
+%
+%    Without memory, this problem is NP-complete; this is equivalent
+%    to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
+%    and with two processors. Half of the tasks must be sent to the processor
+%    on which $T_0$ is not executed, and the optimal makespan is
+%    $w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
+%
+%    However, with an infinite number of identical processors, it can be
+%    solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$;
+%    the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
+%    while the remaining $n-k$ tasks are executed locally (try all values of $k$).
+%
+%    With heterogeneous processors, it is probably NP-complete again
+%    because we could ensure that there are only two processors fast enough
+%    and get back to the 2-partition...
+%
+%    We also had an example where evicting large files first in step 2
+%    can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
+%    $f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
+%    imposes that we free one unit of memory before executing one
+%    of the tasks\ldots Actually the new version with BLC would start
+%    considering $T_2$ and be fine in this case\ldots
+%
+%
+%    \AB{Can we prove that we have (maybe) a 2-approximation,
+%        at least for the fork? What worst-case can we think of? }
 

### 2024-10-10 main.tex
minor
temp

@@
 
-In order to be able to easily adapt to variability of task parameters, we design
+%In order to be able to easily adapt to variability of task parameters, 
+We design 
 variants of HEFT that account for memory usage and aim at minimizing the makespan.
@@
     In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
-    This information may include, memory usage, runtime, graph structures, or the status of the underlying infrastructure.
+    This information may include memory usage, runtime, graph structures, or the status of the underlying infrastructure.
     In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
@@
     For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{TODO}.
+    \AB{I have no idea what to cite there... Svetlana?}
     Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
@@
 % @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
-    Such a dynamic execution environment necessitates for a dynamic scheduling method where the schedule can be recomputed during the workflow execution.
+    Such a dynamic execution environment requires a dynamic scheduling method where the schedule can be recomputed during the workflow execution.
 
@@
     These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
-    they can lead to a later finishing time (\eg if some tasks longer and delay other tasks), or they can have no effect (\eg if new processors
+    they can lead to a later finishing time (\eg if some tasks are longer and they delay other tasks), or they can have no effect (\eg if new processors
     joined the cluster, but the old schedule did not account for them).
@@
     We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
-    is a topologial ordering.
-    We then repeat steps similar to those we did during tentative assignment in our heuristic, except we do not choose a processor
-    anymore, but rather see if the current one still fits.
-
-    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from our heuristic.
-    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ or $c_{v,w}$ from successors $w$,
-    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements from other tasks).
-    If originally,$Res$ was positive (no files were evicted from memory into the communication buffer), then it has to stay this way -
-    otherwise evicted files can invalidate next tasks.
+    is a topological ordering.
+    We then repeat steps similar to those we did during tentative assignment in the heuristics, 
+    except that we do not choose a processor
+    anymore, but rather we check whether the current processor assigned to the task still fits.
+
+    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from the heuristic.
+    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ 
+    or $c_{v,w}$ from successors $w$,
+    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements 
+    from other tasks).
+    If originally, $Res$ was positive (no files were evicted from memory into the communication buffer), 
+    then it has to stay this way --
+    otherwise, evicted files can invalidate next tasks.
     If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
     If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
-    We update the $availM_j$ and $availMC_j$ according to the new memory constraints.
-
-    Then we can re-calculate the finish time of the task on its processor like in Step 3.
+    We update the $availM_j$ and $availMC_j$ values according to the new memory constraints.
+
+    Then, we can re-calculate the finish time of the task on its processor like in Step 3.
     The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
@@
 
-    Then, after having updated the processor's values, we move on to the next task.
+    Finally, after having updated the processor's values, we move on to the next task.
 

### 2024-10-10 main.tex
some progress... need to get some sleep now ;)

@@
     which acts as proxy for the total execution time of a workflow.
+However, to the best of our knowledge,
+    the only memory-aware heuristics that would account for memory constraints are  partitioning
+    the DAG and not reusing processors once they have processed a part of the graph, leading
+    to high values of makespan compared to a finer grain solution that reuses processors.
 
@@
     very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
-    deviations occur. List-based schedulers such as HEFT are, however, not designed for 
-    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
-    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
-%    and would compute a completely new schedule from scratch.
+    deviations occur. However, the original list-based schedulers, such as HEFT, are only defined
+    in a static setting with accurate task parameters. 
+%    
+%    List-based schedulers such as HEFT are, however, not designed for 
+%    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
+%    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
+%%    and would compute a completely new schedule from scratch.
 
     %%% CONTRIBUTION %%%
-    \paragraph*{Contribution} In this paper, we first propose three HEFT-based heuristics
+    %\paragraph*{Contribution} 
+    The main contributions of this paper are both algorithmic and experimental:
+    \begin{itemize}
+\item We formalize the problem with memory constraints, where communication buffers
+are used to evict data from memory if it will be later used by another processor. 
+\item We design three HEFT-based heuristics
     that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
@@
     The difference between the new heuristics is the way they prioritize tasks (for processor assignment).
-    We compare the static schedule produced by these heuristics with the classical HEFT as
-    baseline, which however does not take memory sizes into account. 
-    \hmey{Need to clarify: why is this SotA?}
-
-    \begin{itemize}
-      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
-      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
-      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
+
+    \item We implement a runtime system able to provide some feedback to the scheduler
+    when task requirements (in terms of execution time and memory) differ from the initial predictions, 
+    and we recompute a schedule, based on the reported deviations.
+
+    \item We perform extensive simulations, first in the static case by comparing the schedules produced 
+    by these heuristics with the classical HEFT as baseline, which however does not take memory sizes into account; 
+    while HEFT returns invalid schedules that exceed the processor memories and cannot execute correctly,
+    the new heuristics are able to successfully schedule large workflows, with reasonable makespans.
+
+    \item In the dynamic setting, we use a runtime system that allows us to simulate workflow executions,
+    introducing deviations in running times and memory requirements of tasks that are communicated
+    back to the scheduler; the scheduler can then recompute a schedule. Without these recomputations,
+    most schedules become invalid after deviations, since the memory constraint is exceeded 
+    for most workflows, hence demonstrating the necessity of a dynamic adjustment of the schedule.
+
+%    \hmey{Need to clarify: why is this SotA?}
+
+  %  \begin{itemize}
+%      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
+%      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
+%      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
     \end{itemize}
+
+We first review related work in Section~\ref{sec:related-work}. Then, we formalize the model in Section~\ref{sec:model}
+and the algorithms in Section~\ref{sec:heuristics}.
+The adaptation of the heuristics in a dynamic setting is discussed in Section~\ref{sec:dyn}, and experiment results are presented in Section~\ref{sec:expe}.
+Finally, we conclude
+and provide future working directions in Section~\ref{sec:conc}.
 
@@
 
-    So, during the task prioritization phase, Ref.~\cite{sulaiman2021hybrid} computes the standard deviation of the computation cost
-    (between processors) and adds it to the mean value to account for the differences between processor speeds.
-    In the processor choice phase, they duplicate the entry task and the longest parent tasks during idle times on the processor.
-
-    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
-    the fastest and the slowest processors, divided by the speed ratio of these two processors.
-    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
-    They choose the processor with the lowest execution time and cross over to other processors sometimes.
-    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}
+    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
+    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
+    In the processor choice phase, the entry task and the longest parent tasks are duplicated 
+    during idle times on the processor.
+
+%    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
+%    the fastest and the slowest processors, divided by the speed ratio of these two processors.
+%    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
+%    They choose the processor with the lowest execution time and cross over to other processors sometimes.
+%    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}
 
@@
 
-    The TSHCS (Task Scheduling for Heterogeneous Comupting Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
+    The TSHCS (Task Scheduling for Heterogeneous Computing Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
     by adding randomized decisions to the second phase.
@@
 
-    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement
-    learning techniques(\cite{yano2022cqga}).
+    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement learning techniques~\cite{yano2022cqga}.
 
@@
     is formulated and addressed in the algorithm.
-
+%
     Different models of memory available on processors and memory requirements of tasks have been presented.
@@
 
-    In an assumed dual-memory systems(\cite{herrmann2014memory}) a processor can have access to a memory of two different
+    In an assumed dual-memory systems~\cite{herrmann2014memory},  a processor can have access 
+    to a memory of two different
     kinds (red or blue), and each task can be executed on only one sort of memory.
-    The communications happen only between these two kinds of processors (communications withing each group are ignored).
+    The communications happen only between these two kinds of processors (communications within 
+    each group are ignored).
     The authors then formulate an ILP solution for this problem formulation.
 
-    The algorithm presented by Yao \etal~\cite{yao2022memory}, each processor has an own internal memory and all
+    %The algorithm presented by Yao et al.
+    Yao et al.~\cite{yao2022memory} consider that each processor has an own internal memory and all
     processors share a common external one. The internal (local) memory is used to store the task files.
@@
     All processors, including the original one, can access these files. 
-    Each edge in~\cite{yao2022memory} has two weights -- the size of the files transferred along it,
+    Each edge %in~\cite{yao2022memory} 
+    has two weights -- the size of the files transferred along it,
     and the time of communication along this edge.
@@
 
-    \cite{ding2024ils} assumes connected processors with individual limited memories.
+    In~\cite{ding2024ils}, there are connected processors with individual limited memories.
     The collective set of memories forms the global memory, to which each processor has access, however the access time
     to global memory is different.
-    Each memory access in the graph is modelled as a memory access token on the task, while the edges have no weights.
+    Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
     The solved problem is how to allocate initial input data in processor memories so that the overall
     execution is minimized and the memories are not exceeded.
-    They propose an integer linear programming model.
+    The authors propose an integer linear programming model.
     %that minimizes the length of the critical path, including a greedy initial solution.
@@
 
-    Cloud-oriented model can include costs associated with memory usage(\cite{liang2020memory}).
+   Finally, there are some cloud-oriented models that include costs associated with memory usage~\cite{liang2020memory}.
+
+   Overall, there are a variety of memory models, but, to the best of our knowledge, the only study on a multiprocessor
+   platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24},
+   but where a partition of the workflow is proposed, hence preventing processor reuse. 
+   Hence, in ~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need of communication buffers to store data 
+   that should be communicated
+   between two processors when tasks are ready to execute.
 
@@
 
-    DVR HEFT~\cite{SANDOKJI2019482} uses an almost unchanged HEFT algorithm in their static step, executing three slightly
+We now review related work in a dynamic setting. With no variation in task parameters, 
+    DVR HEFT~\cite{SANDOKJI2019482} rather considers that new tasks arrive in the system. 
+    They use an almost unchanged HEFT algorithm in the static step, executing three slightly
     varying variants of task weighting and choosing the variant that gives the best overall makespan.
-    In the dynamic phase, they receive new tasks and schedule them on either idle processors or those processors that give them
+    In the dynamic phase, they receive new tasks and schedule them on either idle processors or 
+    those processors that give them
     the earliest finish time.
-    Task failures are not covered.
+    %Task failures are not covered.
 
@@
     that these runtime changes are unrelated to each other).
-    The authors formulate a decision space of all permutations of n jobs, and the optimal schedule in relation to a
+    The authors formulate a decision space of all permutations of $n$ jobs, and the optimal schedule in relation to a
     performance measure $\phi$.
@@
 
-    \subsection{Other notable works}
-
-    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
-    They utilize task duplication when creating the clusters (grains).
-    Their scheduler then maps clusters to processors.
-    They assume unlimited processors with the speed of 1.
-    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
-    to it as possible.
-    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
-    %release times.
-
-    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
-    to search for an optimal solution for mapping tasks on machines.
-    A solution is generated at each step, and the best solution is kept at the end.
-    The search terminates when a certain termination criterion is reached.
-    It generates better results than other algorithms, because it explores the whole solution space.
-
-    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
-    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
-    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
-    The heuristic first determines similar activities and groups them together.
-    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
-    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
-    The heuristic calculates a new execution path with these tasks.
-
-    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
-    (processor) speed heterogeneity.
-    Each accelerator has its own memory that can be zero.
-    Each accelerator's characteristics depend on the task it runs and are not fixed.
-
-    \cite{meng2018traffic} investigate scheduling on multi-core chips.
-    Their model is far from ours.
-
-    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
-    They prioritize tasks that have failed before or are well-predictable.
+\AB{Say why our approach in dynamic setting is different and novel}
+
+%    \subsection{Other notable works}
+%
+%    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
+%    They utilize task duplication when creating the clusters (grains).
+%    Their scheduler then maps clusters to processors.
+%    They assume unlimited processors with the speed of 1.
+%    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
+%    to it as possible.
+%    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
+%    %release times.
+%
+%    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
+%    to search for an optimal solution for mapping tasks on machines.
+%    A solution is generated at each step, and the best solution is kept at the end.
+%    The search terminates when a certain termination criterion is reached.
+%    It generates better results than other algorithms, because it explores the whole solution space.
+%
+%    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
+%    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
+%    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
+%    The heuristic first determines similar activities and groups them together.
+%    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
+%    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
+%    The heuristic calculates a new execution path with these tasks.
+%
+%
+%    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
+%    (processor) speed heterogeneity.
+%    Each accelerator has its own memory that can be zero.
+%    Each accelerator's characteristics depend on the task it runs and are not fixed.
+%
+%
+%
+%
+%    \cite{meng2018traffic} investigate scheduling on multi-core chips.
+%    Their model is far from ours.
+%
+%
+%    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
+%    They prioritize tasks that have failed before or are well-predictable.
+%
+%
 
 \section{Model} %: \skug{Full: 4, polished: 3}}
+\label{sec:model}
 
@@
 \section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
-\label{sec.heuristics}
+\label{sec:heuristics}
 
@@
 \section{Dynamic scenario}
+\label{sec:dyn}
 
@@
 \section{Experimental evaluation: \skug{Full: 3, Polished: 3}}
+\label{sec:expe}
+
 \subsection{Setup: \skug{Full: 5, Polished: 4}}
@@
         \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
-        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules \skug{THIS FIGURE IS WRONG}}
+        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules}
         \label{fig:mem-usages-onlyvalid}
@@
 \section{Conclusion  \skug{Full: 0, Polished: 0}}
+\label{sec:conc}
 

### 2024-10-11 main.tex
rewrote the description of clusters (incl memory-constrained)

@@
     \subsubsection{Target computing systems}
-    \skug{Check that the description of speeds and memories is plausible and consistent with experimental setup. esp check for 19200 -> 192 GB.}
+    \skug{Because of the normalized nature of the task/edge weights, I will now be saying that our larger cluster has 192 GB and it
+    is represented as 192 000 (with 3 zeroes). For the memory-constrained cluster (192 00 with 2 zeroes) I will say that it has 10 times
+    less memory - 19.2 GB. Other processors will be dealt with similarly.}
 
@@
     kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
-    We set the number of each kind of node to \skug{$24$ - check if this still holds }, thus having 144 processors in total. % the whole cluster.
+    We set the number of each kind of node to $12$, thus having 72 processors in total. % the whole cluster.
 
     Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
-    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines; ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and large memory
-    (see Table~\ref{tab:procs} in Appendix~\ref{sub:appx-cluster-config}~\cite{daghetpart_full_version}).
-
-    To test various settings, we also vary the cluster configuration and consider
-    variants of the default cluster presented above:\\
-    $\bullet$ {\em tiny}, {\em small} and {\em large} clusters. While the default cluster consists of 144 nodes (24 of each kind), we
-    consider a small cluster with $18$ processors of each kind ($108$ processors in total), a tiny cluster with $72$ nodes and a
-    large cluster with $36$ processors of each kind ($216$ processors in total). \skug{check for cluster sizes and naming}\\
+    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
+    ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and
+    large memory (see Table~\ref{tab:procs}).
+
+    \begin{table}[htb]
+        \begin{center}
+            \begin{tabular}{rcc}
+                \toprule
+                Processor name %& Amount
+                &  CPU speed (GHz)  &Memory size (GB) \\
+                \midrule
+                local                    & 4                    & 16     \\
+                A1                      & 32                   & 32     \\
+                A2                      & 6                    & 64     \\
+                N1                      & 12                   & 16     \\
+                N2                      & 8                    & 8      \\
+                C2                      & 32                   & 192    \\
+                \bottomrule
+            \end{tabular}
+        \end{center}
+        \caption{Cluster configuration (default).}
+        \label{tab:procs}
+    \end{table}
+
+    To more constrained settings, we also vary the cluster configuration and consider
+    a memory-constrained cluster.
+    Like the default cluster, it consists of 72 nodes (12 of each kind), but we
+    give each node $10$ times less memory.
+    So, the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, $A2$ has $6.4$~\GB instead
+    of $64$~GB etc.
+    The processor speeds and their relations stay unchanged.
 
@@
     \subsubsection{Static experiments}
-    \paragraph{Tiny cluster}
+    \paragraph{Memory-constrained cluster}
 

### 2024-10-11 main.tex
wrote nmore about makespan and memory usage on normal cluster

@@
     \subsubsection{Static experiments}
-    \paragraph{Memory-constrained cluster}
-
-    On the tiny cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
-    The successful schedules were achieved exclusively on the tiny workflows (with only one 200-task size-increased
-    workflow among them).
-    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm could schedule
-    $94.9\%$ of workflows, thus having the smallest failure rate of $5.1\%$.
-
+
+    \skug{TODO for all pictures: substitute 0 in sizes through <100. Will be done in one commit for all.}
     \paragraph{Normal cluster}
 
-    On a normal cluster, the baseline \heft had $24.2\%$ success rate ($75.7\%$ failure rate).
-    All the heuristics could schedule all workflows. \skug{check this tomorrow for two weird failures}.
-
-    Figure~\ref{fig:ms-relations} shows the relations of the makespans foound by the heuristics to the makespans found by the baseline.
+    On a normal cluster,  all three of our heuristics were able to schedule all workflows(Figure~\ref{fig:success-rates-large}).
+    The baseline \heft had $24.2\%$ success rate ($75.7\%$ failure rate).
+    It was able to schedule preliminarily small workflows, with no workflows over the size of $4000$ tasks being scheduled
+    correctly.
+    Small workflows are less constrained by the memory requirement, so in the situation of abundant processing resources, even
+    a not memory-aware approach is able to produce valid schedules.
+    However, this advantage quickly vanishes as workflows grow larger.
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=0.9\columnwidth] {images/success-rates-large}
+        \caption{Success rates on the \skug{normal} cluster.}
+        \label{fig:success-rates-large}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    Figure~\ref{fig:ms-relations-by-workflow} shows the relations of the makespans found by the heuristics to the
+    makespans found by the baseline (including invalid ones), by workflow size.
     On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline, the makespans of \heftblc are $8\%$
-    worse, and those found by \heftmm are $71\%$ worse.
+    worse, and those found by \heftmm are $82.6\%$ worse.
 
@@
         \includegraphics[width=1.1\columnwidth] {images/MsRelations2}
-        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.}
+        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.
+        \skug{SUGESTION: remove this picture, as it duplicates Fig 3}
+        }
+
         \label{fig:ms-relations}
@@
         \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size}
-        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow size. Smaller is better.}
+        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow
+        size. Smaller is better. \skug{CHECK this pictue. It might have a wrong mean function}}
         \label{fig:ms-relations-by-workflow}
@@
     \paragraph{Memory Usage}
+
+    The percentage of memory occupied by the schedule is another its important characteristic.
+    Fig.~\ref{fig:mem-usages-normal} shows the percentage of memory occupied on average by the schedule produced by different
+    algorithms, by workflow size.
+
+    \heftmm continuously outperforms other heuristics in memory usage, using from $46\%$ less memory on the smallest
+    workflows to $4$ times less memory on the largest \numprint{30000}-task workflows.
+    If we consider the invalid \heft schedules, too, we see that they would require more and more memory on average.
+    This is because some assignments require more that $100\%$ of memory (which makes them invalid).
+    We can assess the degree of invalidity by comparing \heft memory usage with the memory usage of \heftbl.
+    \heftbl differs from the baseline only in the sense that it respects available memory on the processors.
+    For the largest workflows, \heft schedules require almost double the memory of \heftbl.
 
@@
         \caption{Memory usage by algorithm by workflow size. Includes invalid heft schedules}
-        \label{fig:mem-usages}
+        \label{fig:mem-usages-normal}
         \vspace{-0.3cm}
     \end{figure}
+
+    Figure~\ref{fig:mem-usages-onlyvalid} shows memory usages if only valid heft schedules are presented.
+    There are no valid heft scheduler from middle-sized and large workflows, so no red bars are present.
 
@@
 
-    \begin{figure}[tb]
-        \centering
-        \includegraphics[width=1.1\columnwidth] {images/success-rates-large}
-        \caption{Success rates on the \skug{normal} cluster.}
-        \label{fig:success-rates-large}
-        \vspace{-0.3cm}
-    \end{figure}
+    \paragraph{Memory-constrained cluster}
+
+    On the tiny cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
+    The successful schedules were achieved exclusively on the tiny workflows (with only one 200-task size-increased
+    workflow among them).
+    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm could schedule
+    $94.9\%$ of workflows, thus having the smallest failure rate of $5.1\%$.
 

### 2024-10-11 main.tex
about successes on the memory-constrained cluster

@@
 
-    On the tiny cluster, \heft could produce valid assignments in $14$ experiments out of $290$ ($4.8\%$).
-    The successful schedules were achieved exclusively on the tiny workflows (with only one 200-task size-increased
+    On the memory-constrained cluster, \heft could produce valid assignments in $14$ experiments out of $290$~($4.8\%$
+    success rate).
+    The successful schedules were achieved exclusively on the tiny workflows (with only two 200-task size-increased
     workflow among them).
-    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm could schedule
-    $94.9\%$ of workflows, thus having the smallest failure rate of $5.1\%$.
-
-    \paragraph{Memory-constrained cluster}
+    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm
+    could still schedule all of the workflows, including even the largest ones.
+    As also observed on the normal cluster, \heftmm seems to be less affested by the size of the workflow when scheduling
+    it.
+    The makespans it produces are larger, but it seems to order the vertices in a similar way on both the small and the large workflows,
+    rather than resorting to more memory-consuming ordering that other heuristics and the baseline produces.
+    This makes it especially valuable for very large workflows in situations of very constrained memory in the cluster.
 

### 2024-10-11 main.tex
wrote about the runtimes of the algorithms

@@
     could still schedule all of the workflows, including even the largest ones.
-    As also observed on the normal cluster, \heftmm seems to be less affested by the size of the workflow when scheduling
+    As also observed on the normal cluster, \heftmm seems to be less affected by the size of the workflow when scheduling
     it.
@@
     \subsubsection{Runtimes of the algorithm}
+    To be able to answer the runtime system without holding it up for too long, the scheduler needs to provide a fast answer.
+    Our bottom-level-based heuristics \heftbl and \heftblc provide smaller runtimes and also scale better with growing workflow
+    sizes~(see Fig.~\ref{fig:runtimes-abs} and Fig.~\ref{fig:runtimes-abs}).
+    Their runtimes are similar and grow from tens of miliseconds for the smallest workflows to $25-27$ seconds for the largest workflows.
+    \heftmm, however, needs to compute a memory-optimal traversal of the entire workflow to compute the ranks of the tasks,
+    so its runtime scales from also rens of miliseconds for the smalest workflows to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
+    $2994.9$ seconds for the largest, \numprint{30000}-task workflows.
+    This increased runtime is, though, offset by the unique $100\%$ success rate this algorithm shows when scheduling large
+    workflows in difficult (memory-constrained) setups.
 

### 2024-10-11 main.tex
fixed compile errors and added some conclusion (WIP)

@@
 \usepackage{graphicx}
+\usepackage{booktabs}
 
@@
     give each node $10$ times less memory.
-    So, the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, $A2$ has $6.4$~\GB instead
+    So, the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, $A2$ has $6.4$~GB instead
     of $64$~GB etc.
@@
 
+ \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately a half of them in a memory-constrained
+    one.
+    They provide a runtime that scales well with the growing size of a workflow.
+
+    \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
+    It is the only heuristic that had $100\%$ success rate.
+    However, the makespans it produces are larger than those of the the other two
+
 Extend to heterogeneous bandwidths.

### 2024-10-11 main.tex
conclusions slightly edited

@@
 
- \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately a half of them in a memory-constrained
+ \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
     one.
-    They provide a runtime that scales well with the growing size of a workflow.
+    They provide a running time that scales well with the growing size of a workflow.
 
     \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
-    It is the only heuristic that had $100\%$ success rate.
-    However, the makespans it produces are larger than those of the the other two
-
-Extend to heterogeneous bandwidths.
-
-Further variability commented out from the paper: changing DAG structure; changing platform...
+    It is the only heuristic that has $100\%$ success rate in our experiments.
+    However, the makespans it produces are larger than those of the the other two.
+
+\begin{itemize}
+  \item Extend to heterogeneous bandwidths. 
+  \item Further variability commented out from the paper: changing DAG structure; changing platform...
+\end{itemize}
 

### 2024-10-11 main.tex
runtimes -> running times

--- 2024-10-11 main.tex
@@
 
-    \subsubsection{Runtimes of the algorithm}
+    \subsubsection{Running times of the algorithm}
     To be able to answer the runtime system without holding it up for too long, the scheduler needs to provide a fast answer.
-    Our bottom-level-based heuristics \heftbl and \heftblc provide smaller runtimes and also scale better with growing workflow
+    Our bottom-level-based heuristics \heftbl and \heftblc provide smaller running times and also scale better with growing workflow
     sizes~(see Fig.~\ref{fig:runtimes-abs} and Fig.~\ref{fig:runtimes-abs}).
-    Their runtimes are similar and grow from tens of miliseconds for the smallest workflows to $25-27$ seconds for the largest workflows.
+    Their running times are similar and grow from tens of miliseconds for the smallest workflows to $25-27$ seconds for the largest workflows.
     \heftmm, however, needs to compute a memory-optimal traversal of the entire workflow to compute the ranks of the tasks,
-    so its runtime scales from also rens of miliseconds for the smalest workflows to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
+    so its runnning time scales from also rens of miliseconds for the smalest workflows to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
     $2994.9$ seconds for the largest, \numprint{30000}-task workflows.
-    This increased runtime is, though, offset by the unique $100\%$ success rate this algorithm shows when scheduling large
+    This increased running time is, though, offset by the unique $100\%$ success rate this algorithm shows when scheduling large
     workflows in difficult (memory-constrained) setups.
@@
         \includegraphics[width=1.1\columnwidth] {images/runtimes-logarithmic}
-        \caption{Runtimes of the heuristics and the baseline. The y axis ist logarithmic.}
+        \caption{Running times of the heuristics and the baseline. The y axis ist logarithmic.}
         \label{fig:runtimes-log}
@@
         \includegraphics[width=1.1\columnwidth] {images/runtimes-absolute}
-        \caption{Runtimes of the heuristics and the baseline.}
+        \caption{Running times of the heuristics and the baseline.}
         \label{fig:runtimes-abs}
@@
     one.
-    They provide a runtime that scales well with the growing size of a workflow.
+    They provide a running time that scales well with the growing size of a workflow.
 

### 2024-10-11 main.tex

@@
 
- \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately a half of them in a memory-constrained
+ \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
     one.
@@
     \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
-    It is the only heuristic that had $100\%$ success rate.
-    However, the makespans it produces are larger than those of the the other two
-
-Extend to heterogeneous bandwidths.
-
-Further variability commented out from the paper: changing DAG structure; changing platform...
+    It is the only heuristic that has $100\%$ success rate in our experiments.
+    However, the makespans it produces are larger than those of the the other two.
+
+\begin{itemize}
+  \item Extend to heterogeneous bandwidths. 
+  \item Further variability commented out from the paper: changing DAG structure; changing platform...
+\end{itemize}
 

### 2024-10-11 main.tex

@@
 
-\section{Experimental evaluation: \skug{Full: 3, Polished: 3}}
+\section{Experimental evaluation} %: \skug{Full: 3, Polished: 3}}
 \label{sec:expe}
 
-\subsection{Setup: \skug{Full: 5, Polished: 4}}
+We first describe the experimental setup in Section~\ref{sec:setup}. 
+Then, we report results on static experiments to assess the performance
+of the memory-aware heuristics in Section~\ref{sec.expe.static}, before
+discussing the heuristics behavior in a dynamic setting in Section~\ref{sec.expe.dyn}. 
+Finally, we report running times of the heuristics in Section~\ref{sec.expe.t}.
+
+\subsection{Experimental setup} %: \skug{Full: 5, Polished: 4}}
     \label{sec:setup}
@@
 
-    \subsection{Results \skug{Full: 3, Polished: 2}}
+\subsection{Results in a static setting} % \skug{Full: 3, Polished: 2}}
+\label{sec.expe.static}
 
@@
 
-    \subsubsection{Dynamic experiments}
+\subsection{Dynamic experiments}
+\label{sec.expe.dyn}
+
     \paragraph{Tiny cluster}
@@
 
-    \subsubsection{Running times of the algorithm}
+\subsection{Running times of the heuristics}
+\label{sec.expe.t}
+
     To be able to answer the runtime system without holding it up for too long, the scheduler needs to provide a fast answer.
@@
 
-    \subsubsection{Summary}
+%    \subsubsection{Summary}
 

### 2024-10-11 main.tex
temp

@@
     and CentOS 8 as OS.
-    Code, input data, and experiment scripts are available for review under \skug{\url{insert link}}.
-
-    Next, we describe the set of workflows used in the evaluation and then the clusters on which the
+    Code, input data, and experiment scripts are available to allow reproducibility of results, at~\skug{\url{insert link}}.
+
+    We first describe the set of workflows used for the evaluation, and then the clusters on which the
     workflows are scheduled.
@@
     \subsubsection{Workflow instances}
-    \skug{Todo: describe the workflows the same asin the abstract}
-    We run our experiments on a dataqset comprised of real-world workflows: ones used in~\cite{lotaru} (atacseq, bacass, chipseq,
-    eager and methylseq), and ones obtained by increasing their size using the WFGen generator~\cite{COLEMAN202216}.
+  %  \skug{Todo: describe the workflows the same asin the abstract}
+    We run  experiments on a dataset that consists of real-world workflows: 
+    we use workflows coming from~\cite{lotaru} (atacseq, bacass, chipseq,
+    eager and methylseq), and also we increase their size using the WFGen generator~\cite{COLEMAN202216}.
 
@@
 %
-    \skug{insert more abt how the runtime system hanfles data reads}
+ %   \skug{insert more abt how the runtime system hanfles data reads}
 
     \subsubsection{Target computing systems}
-    \skug{Because of the normalized nature of the task/edge weights, I will now be saying that our larger cluster has 192 GB and it
-    is represented as 192 000 (with 3 zeroes). For the memory-constrained cluster (192 00 with 2 zeroes) I will say that it has 10 times
-    less memory - 19.2 GB. Other processors will be dealt with similarly.}
+%    \skug{Because of the normalized nature of the task/edge weights, I will now be saying that our larger cluster has 192 GB and it
+%    is represented as 192 000 (with 3 zeroes). For the memory-constrained cluster (192 00 with 2 zeroes) I will say that it has 10 times
+%    less memory - 19.2 GB. Other processors will be dealt with similarly.}
 
@@
     We set the number of each kind of node to $12$, thus having 72 processors in total.  % the whole cluster.
-
+%
     Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
@@
         \begin{center}
-            \begin{tabular}{rcc}
+            \begin{tabular}{c|c|cc}
                 \toprule
-                Processor name %& Amount
-                &  CPU speed (GHz)  &Memory size (GB) \\
+                Processor  %& Amount
+                &  CPU speed   & \multicolumn{2}{c}{Memory size (GB)} \\
+                name & (GHz) & {\em default} & {\em mem-constrained} \\
                 \midrule
-                local                    & 4                    & 16     \\
-                A1                      & 32                   & 32     \\
-                A2                      & 6                    & 64     \\
-                N1                      & 12                   & 16     \\
-                N2                      & 8                    & 8      \\
-                C2                      & 32                   & 192    \\
+                local                    & 4                    & 16     & 1.6 \\
+                A1                      & 32                   & 32     & 3.2 \\
+                A2                      & 6                    & 64    & 6.4 \\
+                N1                      & 12                   & 16     & 1.6 \\
+                N2                      & 8                    & 8      & 0.8\\
+                C2                      & 32                   & 192   &  19.2\\
                 \bottomrule
@@
         \end{center}
-        \caption{Cluster configuration (default).}
+        \caption{Cluster configuration.}
         \label{tab:procs}
@@
 
-    To more constrained settings, we also vary the cluster configuration and consider
-    a memory-constrained cluster.
-    Like the default cluster, it consists of 72 nodes (12 of each kind), but we
-    give each node $10$ times less memory.
-    So, the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, $A2$ has $6.4$~GB instead
-    of $64$~GB etc.
-    The processor speeds and their relations stay unchanged.
-
-    \subsubsection{Runtime System}
+    We also consider a more constrained setting, by varying the cluster configuration. The
+     {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
+     but each node has 10 times less memory. Hence,  
+     the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, 
+     $A2$ has $6.4$~GB instead, of $64$~GB etc.
+    The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).
+
+    \subsubsection{Runtime system}
 
     To simulate the execution of a workflow, we implemented a runtime system.
-    It read the historical data and builds weights for tasks.
+    It reads the historical data and builds weights for tasks.
     These weights are then treated as the ``truth'' about the execution during the simulation.

### 2024-10-11 main.tex
added memory usages on the constrained cluster

@@
         \includegraphics[width=1.1\columnwidth] {images/success-rates-tiny}
-        \caption{Success ratey by algorithm by workflow size. Taller is better. }
+        \caption{Success rates by algorithm by workflow size. Taller is better. }
         \label{fig:success-rates-tiny}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained}.
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/mem-usage-constrained-onlyvalid}
+        \caption{Memory usage by different heuristics by workflow size.}
+        \label{fig:memory-usage-constrained}
         \vspace{-0.3cm}

### 2024-10-11 main.tex
relative makespans for constrained cluster
temp

@@
     To simulate the execution of a workflow, we implemented a runtime system.
-    It reads the historical data and builds weights for tasks.
-    These weights are then treated as the ``truth'' about the execution during the simulation.
-    In the case without errors, these values are being sent to the scheduler.
-    However, if errors have to be simulated, the runtime system applies a deviation function to the values.
-    The deviation function computed a normally distributed deviated value, where the correct value is the mean and the deviation
-    is $10\%$.
+    It reads the historical data and builds weights for tasks, as explained above. 
+%    These weights are then treated as the ``truth'' about the execution during the simulation.
+    In the static case, these values are being sent to the scheduler, which builds a schedule
+    according to these weights. 
+    However, in the dynamic setting, the runtime system applies a deviation function to the values.
+    The deviation function computes a normally distributed deviated value, where the initial value 
+    is the mean and the deviation is~$10\%$.
     This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and\skug{todo cite someone Jonathan cites}.
 
-    So, the scheduler receives deviated values and makes decision based on them.
+    Hence, the scheduler receives deviated values and makes decision based on them.
     This leads to several types of possible issues:
     \begin{itemize}
-        \item Processor is blocked by another task. If the scheduler underestimated the execution time of a task, it will block another
-        one from starting.
-        \item The parent has not finished yet. A similar situation, but a task could start on its processor, but one or more of its parents a
-        are not ready yet.
-        \item Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able
-        to execute on a chosen processor.
-        \item A task took lees than expected. We only consider this case if a task took more than 10\% less than expected. In this
-        case we want to exploit the newly acquired free time by possibly starting other tasks earlier.
+        \item A processor is blocked by another task. If the scheduler underestimated the execution time of a task, 
+        it will block another one from starting.
+        \item A predecessor has not yet finished. 
+        The scheduler may request a task to start on its processor, while
+        some of the predecessors of the tasks have in fact not yet completed their execution, and the
+        task is therefore not yet ready. 
+%        A similar situation, but a task could start on its processor, but one or more of its parents a
+%        are not ready yet.
+        \item Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able to execute on a chosen processor.
+        \item A task took less time than expected. We only consider this case if a task took more than 10\% less 
+        time than expected. In this
+        case, we want to exploit the newly acquired free time by possibly starting other tasks earlier.
     \end{itemize}
@@
 
-    \subsubsection{Preliminary experiments}
-
-    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.
-
-    \subsubsection{Static experiments}
+   % \subsubsection{Preliminary experiments}
+We first study the heuristics behavior when the weights do not change upon runtime, 
+hence the scheduler has a perfect knowledge of task memory requirements and execution times. 
+We have compared two eviction strategies, starting with large files first or small files first, 
+and did not report any significant changes in terms of validity of schedule or makespan. 
+Hence, we only present results with the eviction of largest files first. 
+%    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.
+
+    \subsubsection{Scheduling on the default cluster}
 
     \skug{TODO for all pictures: substitute 0 in sizes through <100. Will be done in one commit for all.}
-    \paragraph{Normal cluster}
-
-    On a normal cluster,  all three of our heuristics were able to schedule all workflows(Figure~\ref{fig:success-rates-large}).
-    The baseline \heft had $24.2\%$ success rate ($75.7\%$ failure rate).
-    It was able to schedule preliminarily small workflows, with no workflows over the size of $4000$ tasks being scheduled
-    correctly.
-    Small workflows are less constrained by the memory requirement, so in the situation of abundant processing resources, even
-    a not memory-aware approach is able to produce valid schedules.
-    However, this advantage quickly vanishes as workflows grow larger.
+%    \paragraph{Normal cluster}
+
+    On the default cluster,  the three memory-aware heuristics are able to schedule all workflows 
+    (see Figure~\ref{fig:success-rates-large}), while 
+    the baseline \heft has $24.2\%$ success rate ($75.7\%$ failure rate). 
+    Indeed, \heft is only able to schedule small workflows, but no workflow over the size of $4000$ tasks     
+    can be scheduled correctly; some tasks run out of memory. As soon as we are not in a setting
+    with abundant processing resources for small workflows, it is hence necessary to adopt
+    a memory-aware strategy in order to produce valid schedules. 
+%    Small workflows are less constrained by the memory requirement, so in the situation of abundant processing resources, even
+%    a not memory-aware approach is able to produce valid schedules.
+%    However, this advantage quickly vanishes as workflows grow larger.
 
@@
 
-    Figure~\ref{fig:ms-relations-by-workflow} shows the relations of the makespans found by the heuristics to the
-    makespans found by the baseline (including invalid ones), by workflow size.
-    On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline, the makespans of \heftblc are $8\%$
-    worse, and those found by \heftmm are $82.6\%$ worse.
-
-    \skug{TODO: how much worse are makespans for SUCCESSFUL hefts? presulmably almost identical}
-
-    \begin{figure}[tb]
-        \centering
-        \includegraphics[width=1.1\columnwidth] {images/MsRelations2}
-        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.
-        \skug{SUGESTION: remove this picture, as it duplicates Fig 3}
-        }
-
-        \label{fig:ms-relations}
-        \vspace{-0.3cm}
-    \end{figure}
+   We also report in  Figure~\ref{fig:ms-relations-by-workflow}  the relative makespan found 
+   by the memory-aware heuristics, normalized to the
+   makespan achieved by \heft, often through an invalid over-optimistic schedule that exceeds
+   the bound on memory. 
+%    Figure~\ref{fig:ms-relations-by-workflow} shows the relations of the makespans found by the heuristics to the
+%    makespans found by the baseline (including invalid ones), by workflow size.
+    On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline, 
+    the makespans of \heftblc are $8\%$
+    worse, and those found by \heftmm are $82.6\%$ worse. These are still very encouraging results,
+    in particular for \heftbl and \heftblc, since the makespans of \heft correspond to invalid schedules.
+
+ %   \skug{TODO: how much worse are makespans for SUCCESSFUL hefts? presulmably almost identical}
+
+%    \begin{figure}[tb]
+%        \centering
+%        \includegraphics[width=1.1\columnwidth] {images/MsRelations2}
+%        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.
+%        \skug{SUGESTION: remove this picture, as it duplicates Fig 3}
+%        }
+%
+%        \label{fig:ms-relations}
+%        \vspace{-0.3cm}
+%    \end{figure}
 
@@
 
-    \paragraph{Memory Usage}
-
-    The percentage of memory occupied by the schedule is another its important characteristic.
-    Fig.~\ref{fig:mem-usages-normal} shows the percentage of memory occupied on average by the schedule produced by different
-    algorithms, by workflow size.
-
-    \heftmm continuously outperforms other heuristics in memory usage, using from $46\%$ less memory on the smallest
+%    \paragraph{Memory Usage}
+
+    Finally, we study the percentage of memory occupied by the schedule, which is another good indicator
+    of the memory usage of the heuristic and its ability to produce valid schedules.  %its important characteristic.
+    Fig.~\ref{fig:mem-usages-normal} and~\ref{fig:mem-usages-onlyvalid} show the percentage of memory 
+    occupied on average by the schedule 
+    produced by the different heuristics, for different workflow sizes, first on all schedules (hence including
+    \heft schedules that were not valid), and then only on valid schedules (hence, no results for \heft
+    on large workflow sizes). 
+%    
+%   Figure shows memory usages if only valid heft schedules are presented.
+%    There are no valid heft scheduler from middle-sized and large workflows, so no red bars are present.
+
+    \heftmm continuously outperforms other heuristics in terms of memory usage, 
+    using from $46\%$ less memory on the smallest
     workflows to $4$ times less memory on the largest \numprint{30000}-task workflows.
-    If we consider the invalid \heft schedules, too, we see that they would require more and more memory on average.
+    If we consider the invalid \heft schedules, too, we see that they would require more and more memory on average,
+    which explains why these schedules rapidly become invalid.
     This is because some assignments require more that $100\%$ of memory (which makes them invalid).
@@
     \heftbl differs from the baseline only in the sense that it respects available memory on the processors.
-    For the largest workflows, \heft schedules require almost double the memory of \heftbl.
+    For the largest workflows, \heft schedules require almost twice the memory of \heftbl.
 
@@
         \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal}
-        \caption{Memory usage by algorithm by workflow size. Includes invalid heft schedules}
+        \caption{Memory usage by algorithm by workflow size. Includes invalid schedules for \heft. }
         \label{fig:mem-usages-normal}
@@
 
-    Figure~\ref{fig:mem-usages-onlyvalid} shows memory usages if only valid heft schedules are presented.
-    There are no valid heft scheduler from middle-sized and large workflows, so no red bars are present.
-
     \begin{figure}[tb]
@@
         \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
-        \caption{Memory usage by algorithm by workflow size. Includes only valid heft schedules}
+        \caption{Memory usage by algorithm by workflow size. Includes only valid schedules for \heft. }
         \label{fig:mem-usages-onlyvalid}
@@
 
-    \paragraph{Memory-constrained cluster}
+\subsubsection{Scheduling on the memory-constrained cluster}
 

### 2024-10-11 main.tex
merged

@@
         \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow
-        size. Smaller is better. \skug{CHECK this pictue. It might have a wrong mean function}}
+        size. Smaller is better.}
         \label{fig:ms-relations-by-workflow}
@@
 
+    Relative makespans to the baseline (including the invalid values of the baseline) are shown in Fig.~\ref{fig:ms-relations-by-workflow-constrained}.
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
+        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow
+        size. Smaller is better.}
+        \label{fig:ms-relations-by-workflow-constrained}
+        \vspace{-0.3cm}
+    \end{figure}
+
 \subsection{Dynamic experiments}
@@
 
-    \begin{figure}[tb]
-        \centering
-        \includegraphics[width=1.1\columnwidth] {images/runtimes-absolute}
-        \caption{Running times of the heuristics and the baseline.}
-        \label{fig:runtimes-abs}
-        \vspace{-0.3cm}
-    \end{figure}
+    %\begin{figure}[tb]
+    %    \centering
+    %    \includegraphics[width=1.1\columnwidth] {images/runtimes-absolute}
+    %    \caption{Running times of the heuristics and the baseline.}
+    %    \label{fig:runtimes-abs}
+    %    \vspace{-0.3cm}
+    %\end{figure}
 

### 2024-10-11 main.tex
temp

@@
         \includegraphics[width=0.9\columnwidth] {images/success-rates-large}
-        \caption{Success rates on the \skug{normal} cluster.}
+        \caption{Success rates on the default cluster. Higher is better.}
         \label{fig:success-rates-large}
@@
         \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size}
-        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow
-        size. Smaller is better.}
+        \caption{Relative makespans of heuristics normalized by \heft makespan, by workflow
+        size, on default cluster. Smaller is better.}
         \label{fig:ms-relations-by-workflow}
@@
     of the memory usage of the heuristic and its ability to produce valid schedules.  %its important characteristic.
-    Fig.~\ref{fig:mem-usages-normal} and~\ref{fig:mem-usages-onlyvalid} show the percentage of memory
+    Figures \ref{fig:mem-usages-normal} and~\ref{fig:mem-usages-onlyvalid} show the percentage of memory
     occupied on average by the schedule
@@
         \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal}
-        \caption{Memory usage by algorithm by workflow size. Includes invalid schedules for \heft. }
+        \caption{Memory usage on default cluster, including invalid \heft schedules. }
+ %by algorithm by workflow size
          \label{fig:mem-usages-normal}
@@
         \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
-        \caption{Memory usage by algorithm by workflow size. Includes only valid schedules for \heft. }
+        \caption{Memory usage on default cluster, including only valid \heft schedules. }
         \label{fig:mem-usages-onlyvalid}
@@
 
-    On the memory-constrained cluster, \heft could produce valid assignments in $14$ experiments out of $290$~($4.8\%$
-    success rate).
+    On the memory-constrained cluster, \heft could produce valid assignments in only $14$ experiments 
+    out of $290$~($4.8\%$ success rate).
     The successful schedules were achieved exclusively on the tiny workflows (with only two 200-task size-increased
     workflow among them).
-    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$, while \heftmm
-    could still schedule all of the workflows, including even the largest ones.
-    As also observed on the normal cluster, \heftmm seems to be less affected by the size of the workflow when scheduling
-    it.
-    The makespans it produces are larger, but it seems to order the vertices in a similar way on both the small and the large workflows,
-    rather than resorting to more memory-consuming ordering that other heuristics and the baseline produces.
-    This makes it especially valuable for very large workflows in situations of very constrained memory in the cluster.
+    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$ of them, while \heftmm
+    could still schedule all the workflows, including even the largest ones, see Figure~\ref{fig:success-rates-tiny}. 
+    As also observed on the normal cluster, \heftmm seems to be less affected by the size of the workflow 
+    when scheduling it than the other heuristics.
+
+    Similarly to the default cluster, we observe that relative makespan of \heftmm is usually greater
+    than those of \heft (see Fig.~\ref{fig:ms-relations-by-workflow-constrained}), but \heft schedules are
+    almost all invalid. It is therefore very interesting to resort to \heftmm for large workflows 
+    in a constrained cluster, since tasks are processed in an order that minimizes
+    the memory usage of schedule. 
+%
+%    The makespans it produces are larger, but it seems to order the vertices in a similar way on both the small and the large workflows,
+%    rather than resorting to more memory-consuming ordering that other heuristics and the baseline produces.
+%    This makes it especially valuable for very large workflows in situations of very constrained memory in the cluster.
+%
+%
+
+    Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained},
+    and we observe that the memory footprint of \heftmm remains constant with workflow size.
 
@@
         \includegraphics[width=1.1\columnwidth] {images/success-rates-tiny}
-        \caption{Success rates by algorithm by workflow size. Taller is better. }
+        \caption{Success rates on the memory-constrained cluster. Higher is better. }
+%        by algorithm by workflow size. 
         \label{fig:success-rates-tiny}
@@
 
-    Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained}.
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
+        \caption{Relative makespans on the memory-constrained cluster. 
+%        produced by the heuristics to the makespan produced by the baseline, by workflow size. 
+        Smaller is better.}
+        \label{fig:ms-relations-by-workflow-constrained}
+        \vspace{-0.3cm}
+    \end{figure}
 
@@
         \includegraphics[width=1.1\columnwidth] {images/mem-usage-constrained-onlyvalid}
-        \caption{Memory usage by different heuristics by workflow size.}
+        \caption{Memory usage on the memory-constrained  cluster. } % by different heuristics by workflow size.}
         \label{fig:memory-usage-constrained}
-        \vspace{-0.3cm}
-    \end{figure}
-
-    Relative makespans to the baseline (including the invalid values of the baseline) are shown in Fig.~\ref{fig:ms-relations-by-workflow-constrained}.
-
-    \begin{figure}[tb]
-        \centering
-        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
-        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline, by workflow
-        size. Smaller is better.}
-        \label{fig:ms-relations-by-workflow-constrained}
         \vspace{-0.3cm}

### 2024-10-11 main.tex
added refs from JOnathan, removed 2 minor comments by me

--- 2024-10-11 main.tex
@@
     is the mean and the deviation is~$10\%$.
-    This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and\skug{todo cite someone Jonathan cites}.
+    This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and
+    other works dedicated to predicting runtimes of tasks~\cite{da2015online,da2013toward}.
 
@@
 
-    \skug{TODO for all pictures: substitute 0 in sizes through <100. Will be done in one commit for all.}
 %    \paragraph{Normal cluster}
@@
         \includegraphics[width=0.9\columnwidth] {images/success-rates-large}
-        \caption{Success rates on the \skug{normal} cluster.}
+        \caption{Success rates on the normal cluster.}
         \label{fig:success-rates-large}

### 2024-10-11 main.tex
temp - shall we discuss briefly for dynamic expe?

--- 2024-10-11 main.tex
@@
 
+\AB{Any info on success rate of adaptive heuristic compared to static ones? 
+For which cluster do we have some results? }
+
     \paragraph{Tiny cluster}
@@
     Due to an extremely constrained memory in the tiny cluster, only 134 experiments out of 1160 for all algorithm variants
-    ended with a valid makespan withjout recomputation. $13$ experiments ended with a valid makespan after recomputation in case of \heft,
+    ended with a valid makespan without recomputation. $13$ experiments ended with a valid makespan after recomputation in case of \heft,
    $16$ in case of \heftmm, $50$ in case of \heftblc and $55$ in case of \heftbl.
@@
 
-    To be able to answer the runtime system without holding it up for too long, the scheduler needs to provide a fast answer.
-    Our bottom-level-based heuristics \heftbl and \heftblc provide smaller running times and also scale better with growing workflow
-    sizes~(see Fig.~\ref{fig:runtimes-abs} and Fig.~\ref{fig:runtimes-abs}).
-    Their running times are similar and grow from tens of miliseconds for the smallest workflows to $25-27$ seconds for the largest workflows.
-    \heftmm, however, needs to compute a memory-optimal traversal of the entire workflow to compute the ranks of the tasks,
-    so its runnning time scales from also rens of miliseconds for the smalest workflows to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
+    To be able to answer the runtime system without holding it up for too long, the scheduler needs to 
+    be able to compute a schedule rapidly. %provide a fast answer.
+    The bottom-level-based heuristics \heftbl and \heftblc provide smaller running times than \heftmm, 
+    and also scale better with growing workflow
+    sizes~(see Fig.~\ref{fig:runtimes-log}).
+    Their running times are similar and grow from tens of milliseconds for the smallest workflows to $25-27$ seconds 
+    for the largest workflows.
+    \heftmm, however, needs to compute a memory-optimal traversal of the entire workflow to compute the ranks 
+    of the tasks,
+    so its running time scales from also tens of milliseconds for the smallest workflows, 
+    to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
     $2994.9$ seconds for the largest, \numprint{30000}-task workflows.
-    This increased running time is, though, offset by the unique $100\%$ success rate this algorithm shows when scheduling large
+    This increased running time is, though, offset by the unique $100\%$ success rate this algorithm 
+    obtained when scheduling large
     workflows in difficult (memory-constrained) setups.
@@
         \includegraphics[width=1.1\columnwidth] {images/runtimes-logarithmic}
-        \caption{Running times of the heuristics and the baseline. The y axis ist logarithmic.}
+        \caption{Running times of the heuristics. The y axis is logarithmic.}
         \label{fig:runtimes-log}
@@
 
-\section{Conclusion  \skug{Full: 0, Polished: 0}}
+\section{Conclusion} %  \skug{Full: 0, Polished: 0}}
 \label{sec:conc}

### 2024-10-11 main.tex
some text about failures of no recomputation.

@@
 
-    \paragraph{Tiny cluster}
+    \paragraph{Memory-constrained cluster}
     The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
@@
     Due to an extremely constrained memory in the tiny cluster, only 134 experiments out of 1160 for all algorithm variants
-    ended with a valid makespan without recomputation. $13$ experiments ended with a valid makespan after recomputation in case of \heft,
-   $16$ in case of \heftmm, $50$ in case of \heftblc and $55$ in case of \heftbl.
+    ended with a valid makespan without recomputation.
+    $13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
+    rate. However, there were only $14$ successful \heft experiments overall in this setup, so for the successful experiments
+    only, the success rate of not recomputing is $92\%$.
+    In case of \heftmm, $16$ experiments ended with a successfully computed makespan without recomputation.
+    Because \heftmm managed to schedule all experiments, the rate is $5.5\%$ success without recomputation out of all successful
+    schedules with recomputation.
+    $50$ experiments were successful in case of \heftblc ($35\%$ out of all successful schedules) and $55$ in case of
+    \heftbl (roughly $50\%$ of all successful schedules).
+
+    \heftbl and \heftblc were successful on smaller workflows and failed on larger ones, the success of the strategy without
+    recomputation was limited to even the smallest of these smaller workflows.
+    So, the strategy without recomputation dfelivered valid makespans (independently of the algorithm) on $56$ original
+    workflows with $<100$ tasks, $47$ 200-task ones, $25$ 1000-task ones and $6$ 2000-task ones.
+
     Figure~\ref{fig:updates-ms} shows the increase in makespan in case of no recomputation for these experiments.
-    There were valid final makespans with no recomputation only for the smallest workflows - $56$ original ones, $47$ 200-task ones,
-    $25$ 1000-task ones and $6$ 2000-task ones.
-
     With growing size of the workflow, the excess makespan of not recomputing grows - from $13.9\%$ to $20\%$ for \heftbl,

### 2024-10-11 main.tex
conclusion

--- 2024-10-11 main.tex
@@
 \usepackage{booktabs}
+\usepackage{balance}
 
@@
 
+Note that in both clusters, we set the size of the communication buffer to be equal
+to the memory size. \todo[inline]{Is this correct? }
+
     \subsubsection{Runtime system}
@@
 
- \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
-    one.
-    They provide a running time that scales well with the growing size of a workflow.
-
-    \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
-    It is the only heuristic that has $100\%$ success rate in our experiments.
-    However, the makespans it produces are larger than those of the the other two.
-
-\begin{itemize}
-  \item Extend to heterogeneous bandwidths. 
-  \item Further variability commented out from the paper: changing DAG structure; changing platform...
-\end{itemize}
-
+We have formalized a scheduling problem in memory-constrained environments, where tasks
+may exceed the memory available on a processor and resort to communication buffers
+to store and communicate data between processors. In this context, we have 
+designed three memory-aware HEFT-based heuristics, which account for memory
+constraints while scheduling tasks.
+
+Two heuristics rely on an ordering of tasks based on the bottom level of tasks, with the objective
+of minimizing the makespan and hence scheduling critical tasks first. The third one, \heftmm,
+goes one step further and handles tasks in an order dictated by an efficient traversal
+of the graph in terms of memory requirements, hence reducing the memory used
+by the schedule. Experimental results on a large set of workflows coming from real-life applications
+demonstrate that the memory-aware heuristics are  successfully producing valid schedules. 
+In the most memory-constrained setting, \heftmm succeeds to schedule even the largest workflows,
+while other heuristics return invalid schedules exceeding the memory capacity of processors. 
+It comes at a price of a makespan that is not quite as good as the one obtained by \heftbl
+and \heftblc, which focus on makespan by ordering tasks by bottom level. 
+As expected, the baseline \heft that is not memory aware returns invalid schedules in almost
+all settings, except for very small workflows.
+
+%Hence, \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
+%    one.
+%    They provide a running time that scales well with the growing size of a workflow.
+%
+%    \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
+%    It is the only heuristic that has $100\%$ success rate in our experiments.
+%    However, the makespans it produces are larger than those of the the other two.
+
+Another key contribution is that we have adapted these heuristics for a dynamic setting, where exact
+task parameters (execution times and memory requirements) are not know in advance. 
+We have implemented a runtime system that interacts with the scheduler, returning exact
+parameter values once a task arrives in the system, while only estimates are known for future tasks. 
+Some preliminary experiments have been conducted in this setting, and demonstrated that
+it is necessary to adapt the schedule on the fly in order to avoid an execution failure because
+of a shortage in memory.
+
+This work could be extended in several directions. First, the model could be refined to include
+heterogeneous bandwidths, while we consider a homogeneous communication network. 
+More importantly, it would be interesting to consider other types of variability, for instance
+if new tasks appear in the graph (or disappear), or if there is variability in the platform, 
+with processors arriving and departing. We believe that we could adapt the current approach
+by recomputing schedules on the fly, and we plan to perform a new set of experiments
+to further assess the impact of dynamic scheduling.
+
+%
+%\begin{itemize}
+%  \item Extend to heterogeneous bandwidths. 
+%  \item Further variability commented out from the paper: changing DAG structure; changing platform...
+%\end{itemize}
+
+\balance
     \bibliographystyle{abbrv}%IEEEtran}

### 2024-10-11 main.tex
related work

@@
 
-\AB{Say why our approach in dynamic setting is different and novel}
+While several related work consider building a new schedule once some variation has been observed,
+we are not aware of work implementing a real runtime system that interacts with the scheduler,
+and tested on workflows with thousands of tasks, as we propose in this paper. Furthermore, 
+we are not aware of any previous work discussing dynamic algorithms combined with memory constraints. 
+%\AB{Say why our approach in dynamic setting is different and novel}
 
@@
 it is necessary to adapt the schedule on the fly in order to avoid an execution failure because
-of a shortage in memory.
+of a shortage in memory. To the best of our knowledge, this is the first study of adaptive
+algorithms accounting for memory constraints.
 

### 2024-10-11 main.tex
rewrote about final successful schedules

--- 2024-10-11 main.tex
@@
 \usepackage{booktabs}
-\usepackage{balance}
 
@@
 
-Note that in both clusters, we set the size of the communication buffer to be equal
-to the memory size. \todo[inline]{Is this correct? }
-
     \subsubsection{Runtime system}
@@
     ended with a valid makespan without recomputation.
-    $13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
-    rate. However, there were only $14$ successful \heft experiments overall in this setup, so for the successful experiments
-    only, the success rate of not recomputing is $92\%$.
-    In case of \heftmm, $16$ experiments ended with a successfully computed makespan without recomputation.
-    Because \heftmm managed to schedule all experiments, the rate is $5.5\%$ success without recomputation out of all successful
-    schedules with recomputation.
-    $50$ experiments were successful in case of \heftblc ($35\%$ out of all successful schedules) and $55$ in case of
-    \heftbl (roughly $50\%$ of all successful schedules).
+    In case of \heft, $14$ valid initial makespans were computed.
+    Out of them, $13$ managed to stay valid until the end, after all the update requests.
+    The same $13$ experiments ended with a valid makespan in case of no recomputation.
+    So, these workflows required so few resources that there was no point in re-scheduling them.
+    %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
+    %rate. However, there were only $14$ successful \heft experiments overall in this setup, so for the successful experiments
+    %only, the success rate of not recomputing is $92\%$.
+    In case of \heftmm, all $290$ workflows could be scheduled initially and all the schedules remained valid until the end.
+    $16$ experiments ended with a successful makespan without recomputation.
+    The rate of successful schedules without recomputation is therefore $5.5\%$
+    \heftblc could produce $142$ valid initial schedules and kept $141$ of them valid until the end.
+    $50$ experiments were successful without recomputation, too, making it $35\%$ out of all successful final schedules.
+    \heftbl kept $105$ schedules valid until the end out of $110$ initial valid schedules.
+    $55$ were able to be kept valid until the end even without recomputation, roughly $50\%$ of all succeses.
 
@@
 
-We have formalized a scheduling problem in memory-constrained environments, where tasks
-may exceed the memory available on a processor and resort to communication buffers
-to store and communicate data between processors. In this context, we have 
-designed three memory-aware HEFT-based heuristics, which account for memory
-constraints while scheduling tasks.
-
-Two heuristics rely on an ordering of tasks based on the bottom level of tasks, with the objective
-of minimizing the makespan and hence scheduling critical tasks first. The third one, \heftmm,
-goes one step further and handles tasks in an order dictated by an efficient traversal
-of the graph in terms of memory requirements, hence reducing the memory used
-by the schedule. Experimental results on a large set of workflows coming from real-life applications
-demonstrate that the memory-aware heuristics are  successfully producing valid schedules. 
-In the most memory-constrained setting, \heftmm succeeds to schedule even the largest workflows,
-while other heuristics return invalid schedules exceeding the memory capacity of processors. 
-It comes at a price of a makespan that is not quite as good as the one obtained by \heftbl
-and \heftblc, which focus on makespan by ordering tasks by bottom level. 
-As expected, the baseline \heft that is not memory aware returns invalid schedules in almost
-all settings, except for very small workflows.
-
-%Hence, \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
-%    one.
-%    They provide a running time that scales well with the growing size of a workflow.
-%
-%    \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
-%    It is the only heuristic that has $100\%$ success rate in our experiments.
-%    However, the makespans it produces are larger than those of the the other two.
-
-Another key contribution is that we have adapted these heuristics for a dynamic setting, where exact
-task parameters (execution times and memory requirements) are not know in advance. 
-We have implemented a runtime system that interacts with the scheduler, returning exact
-parameter values once a task arrives in the system, while only estimates are known for future tasks. 
-Some preliminary experiments have been conducted in this setting, and demonstrated that
-it is necessary to adapt the schedule on the fly in order to avoid an execution failure because
-of a shortage in memory.
-
-This work could be extended in several directions. First, the model could be refined to include
-heterogeneous bandwidths, while we consider a homogeneous communication network. 
-More importantly, it would be interesting to consider other types of variability, for instance
-if new tasks appear in the graph (or disappear), or if there is variability in the platform, 
-with processors arriving and departing. We believe that we could adapt the current approach
-by recomputing schedules on the fly, and we plan to perform a new set of experiments
-to further assess the impact of dynamic scheduling.
-
-%
-%\begin{itemize}
-%  \item Extend to heterogeneous bandwidths. 
-%  \item Further variability commented out from the paper: changing DAG structure; changing platform...
-%\end{itemize}
-
-\balance
+ \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
+    one.
+    They provide a running time that scales well with the growing size of a workflow.
+
+    \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
+    It is the only heuristic that has $100\%$ success rate in our experiments.
+    However, the makespans it produces are larger than those of the the other two.
+
+\begin{itemize}
+  \item Extend to heterogeneous bandwidths. 
+  \item Further variability commented out from the paper: changing DAG structure; changing platform...
+\end{itemize}
+
     \bibliographystyle{abbrv}%IEEEtran}

### 2024-10-11 main.tex
temp, have to go to class

@@
         \centering
-        \includegraphics[width=1.1\columnwidth] {images/success-rates-tiny}
+        \includegraphics[width=1\columnwidth] {images/success-rates-tiny}
         \caption{Success rates on the memory-constrained cluster. Higher is better. }
@@
         \centering
-        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
+        \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
         \caption{Relative makespans on the memory-constrained cluster.
@@
 
-\AB{Any info on success rate of adaptive heuristic compared to static ones? 
-For which cluster do we have some results? }
+%\AB{Any info on success rate of adaptive heuristic compared to static ones? 
+%For which cluster do we have some results? }
 
@@
     with not enough memory capacity.
-    Due to an extremely constrained memory in the tiny cluster, only 134 experiments out of 1160 for all algorithm variants
+    Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160 for all algorithm variants
     ended with a valid makespan without recomputation.
@@
     $16$ experiments ended with a successful makespan without recomputation.
-    The rate of successful schedules without recomputation is therefore $5.5\%$
+    The rate of successful schedules without recomputation is therefore~$5.5\%$.
     \heftblc could produce $142$ valid initial schedules and kept $141$ of them valid until the end.
@@
     \heftbl kept $105$ schedules valid until the end out of $110$ initial valid schedules.
-    $55$ were able to be kept valid until the end even without recomputation, roughly $50\%$ of all succeses.
+    $55$ were able to be kept valid until the end even without recomputation, roughly $50\%$ of all successes.
 
@@
     recomputation was limited to even the smallest of these smaller workflows.
-    So, the strategy without recomputation dfelivered valid makespans (independently of the algorithm) on $56$ original
+    So, the strategy without recomputation delivered valid makespans (independently of the algorithm) on $56$ original
     workflows with $<100$ tasks, $47$ 200-task ones, $25$ 1000-task ones and $6$ 2000-task ones.
@@
     With growing size of the workflow, the excess makespan of not recomputing grows - from $13.9\%$ to $20\%$ for \heftbl,
-    from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the 2000-task workflows in this case), $12.1\%$ to $23.5\%$
-    for \heftmm.
-    The larger variations for \heft can be explained by the small amount of data - for example, there were only 2 200-task
+    from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the 2000-task workflows in this case), 
+    $12.1\%$ to $23.5\%$ for \heftmm.
+    The larger variations for \heft can be explained by the small amount of data -- for instance, there were only 2 200-task
     workflows for this case.
@@
         \centering
-        \includegraphics[width=0.495\columnwidth] {images/UpdatesMss2}
-        \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm for different . Smaller is better.}
+        \includegraphics[width=0.8\columnwidth] {images/UpdatesMss2}
+        \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm. Smaller is better.}
         \label{fig:updates-ms}

### 2024-10-15 main.tex
last changes before submission (done on Friday)

--- 2024-10-11 main.tex
@@
 \newcommand{\heft}{\algo{HEFT}\xspace}
-\newcommand{\heftmm}{\algo{HEFTM\_MM}\xspace}
-\newcommand{\heftbl}{\algo{HEFTM\_BL}\xspace}
-\newcommand{\heftblc}{\algo{HEFTM\_BLC}\xspace}
+\newcommand{\heftmm}{\algo{HEFTM-MM}\xspace}
+\newcommand{\heftbl}{\algo{HEFTM-BL}\xspace}
+\newcommand{\heftblc}{\algo{HEFTM-BLC}\xspace}
 
@@
 
-    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures of approximately 15\% is shown~\cite{TODO}.
-    \AB{I have no idea what to cite there... Svetlana?}
+    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures
+    of approximately 15\% is shown~\cite{malik2013execution}.
     Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
@@
     and CentOS 8 as OS.
-    Code, input data, and experiment scripts are available to allow reproducibility of results, at~\skug{\url{insert link}}.
+    Code, input data, and experiment scripts are available to allow reproducibility of results, at~\url{https://zenodo.org/records/13919214}~and~\url{https://zenodo.org/records/13919302}.
 
@@
 Note that in both clusters, we set the size of the communication buffer to be equal
-to the memory size. \todo[inline]{Is this correct? }
+to ten times the the memory size.
 
@@
 \label{sec.expe.dyn}
-
-\AB{Any info on success rate of adaptive heuristic compared to static ones? 
-For which cluster do we have some results? }
 

### 2024-10-15 submission_IPDPS/main.tex
minor

--- 2024-10-11 main.tex
@@
 \label{sec.expe.dyn}
-
-%\AB{Any info on success rate of adaptive heuristic compared to static ones? 
-%For which cluster do we have some results? }
 

### 2024-10-28 main-disk-model.tex
new main file for the new model, first changes in Chapter "Model"

--- 2024-10-15 submission_IPDPS/main.tex
@@
 
+\skug{
 \section{Model} %: \skug{Full: 4, polished: 3}}
@@
 
+CHAGES IN THIS CHAPTER
+}
     We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
@@
 consists of $k$ processors $p_1, \dots, p_k$.
-    Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$, a communication
-    buffer of size $MC_j$ and a speed~$s_j$.
-    We can decide to evict some data from the main memory if we are sending the data
-    to another processor; it then stays in the communication buffer until it has been sent.
+Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
+All processors have access to a common disk space of unlimited size.
+All communications happen over the disk - after one processor has finished writing data there, another one can read
+it into its individual memory.
+We can decide to evict some data from the individual memory onto the disk to free space in the individuaö memory.
+    This data can be read back into memory at any time later.
+
 The execution time of a single task~$u\in V$ on a processor~$p_j$ is expressed as $\frac{w_u}{s_j}$.

### 2024-10-28 main-disk-model.tex
new text in chapter "Model"

@@
 Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
-All processors have access to a common disk space of unlimited size.
+All processors have access to a common disk space of unlimited size, but of much slower speed.
 All communications happen over the disk - after one processor has finished writing data there, another one can read
 it into its individual memory.
-We can decide to evict some data from the individual memory onto the disk to free space in the individuaö memory.
-    This data can be read back into memory at any time later.
-
-The execution time of a single task~$u\in V$ on a processor~$p_j$ is expressed as $\frac{w_u}{s_j}$.
-We assume that all processors are connected with the same bandwidth~$\beta$.
+We can decide to evict some data from the individual memory onto the disk to free space in the individual memory.
+This data can later be read back into memory at any time, or transferred to another processor.
+
+If the entire memory required by the task ~$u\in V$ fits in the free memory of a processor~$p_j$,
+    then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
+If the memory requirement of the task exceeds the available memory on the processor, the task can still be executed there,
+    albeit slower.
+The part of memory requirement that exceeds the available memory will be offloaded to the disk, slowing down the execution.
+
+    \[
+        w_{real} = \frac{w_u}{s_j} * (1 + (m_i - M_j)/m_i ) * q_s.
+
+    \]
+where $q_s$ sit he slowdown coefficient.
+%We assume that all processors are connected with the same bandwidth~$\beta$.
 %    \hmey{Maybe mention that variable bandwidths are part of future work...?}
 
-We keep track of the current ready time of each processor and each communication
-channel, $\rt_j$ and $\rt_{j,j'}$, for each processor $j$ and all pairs~$(j,j')$.
+Multiple processors can communicate with the disk in parallel, but the communication of each processor  with the disk is sequential.
+The processor can read and write at the same time, but only one file can be read and only one written at any point in time.
+We keep track of the current ready times of each processor $j$ , ready time for computation $\rt_j^c$,
+ready time for writing $\rt_j^w$, and ready time for reading $\rt_j^r$.
 Initially, all the ready times are set to~$0$.
-We also keep track of the currently available memory, $availM_j$ and $availC_j$,
-on the processor memory and communication buffer, respectively.
+We also keep track of the currently available memory, $availM_j$ on the processor memory.
+
 Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
-that are in the memory of size $\MM_j$ but may be evicted to be communicated if
-more memory is needed on~$p_j$. They are ordered by non-decreasing size and
-correspond to some $c_{u,v}$.
+that are in the memory of size $\MM_j$ but may be evicted to the disk
+more memory is needed on~$p_j$.
+They are ordered by non-decreasing size and correspond to some $c_{u,v}$.
+When scheduling a tass, the decision has to be made to execute this task on a processor with enough memory
+    without slowdown, but with possibly more time to read the input files, or to execute the same task
+    on a processor with less memory (and hence a slowdown), but potentially with a higher processor speed or with less times
+    spent reading files.
 
@@
 
+\subsection{Scheduling approach - general description}
+
+In the first step, we order the tasks by non-increasing ranks (e.g. their bottom levels).
+In the next step, we repeatedly schedule ready tasks.
+For each task, we first choose the best processor to put it on - the one that minimizes the expected finishing time.
+For each processor, we compute this earliest finishing time by first calculating the earliest starting time on this processor.
+    It is computed as the maximum of the ready time of computation on this processor, and the time to read input files.
+
 \subsection{Optimization problem}
@@
 so that the makespan (total execution time) is minimized while
-respecting memory constraints. If a processor runs out of memory to execute
-a task mapped on it, the schedule is said to be {\em invalid}.
+respecting memory constraints. \skug{do we still formulate it like that?}
+    %If a processor runs out of memory to execute
+%a task mapped on it, the schedule is said to be {\em invalid}.
 

### 2024-10-28 main-disk-model.tex
minor- changed typo
minor

@@
 
+\newcommand{\bw}{bw}
+\newcommand{\br}{br}
+
 \newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
@@
 may be reconsidered at runtime when a task starts its execution and we know its exact parameters.
+\AB{Variability only on $w_u$ for now}
 
@@
 consists of $k$ processors $p_1, \dots, p_k$.
-Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
-All processors have access to a common disk space of unlimited size, but of much slower speed.
+Each processor $p_u$ ($1 \leq u \leq k$) has an individual memory of size $M_u$ and a speed~$s_u$.
+All processors have access to a shared disk of unlimited size, where they can write
+and read data: Processor~$p_u$ has a bandwidth $\bw_u$ (resp. $\br_u$) to write  (resp. read),
+and hence it takes a time $\frac{c_{u,v}}{bw_u}$ to write the data produced for~$p_v$, 
+if this data is evicted from memory.
+
+%, but of much slower speed.
 All communications happen over the disk - after one processor has finished writing data there, another one can read
@@
     \[
-        w_{real} = \frac{w_u}{s_j} * (1 + (m_i - M_j)/m_i ) * q_s.
-
+        w_{real} = \frac{w_u}{s_j} \* (1 + (m_i - M_j)/m_i ) \* q_s.
     \]
@@
 The goal is to find a schedule of the DAG~$G$ for the $k$ processors,
-so that the makespan (total execution time) is minimized while
-respecting memory constraints. \skug{do we still formulate it like that?}
+so that the makespan (total execution time) is minimized. 
+\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
+I'll probably add a few sentences here...}
+% while respecting memory constraints. \skug{do we still formulate it like that?}
     %If a processor runs out of memory to execute
@@
 on the estimated task parameters.
+\AB{Probably need to clarify offline vs online problem at some point...}
 

### 2024-10-29 main-disk-model.tex
improved the description of the static case

--- 2024-10-28 main-disk-model.tex
@@
     Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.
+   \skug{here?: For each edge, we also keep information on where it is currently stored - in memory of a processor (with processor id),
+    on the disk or nowhere - if this file has not yet been generated or has already been consumed by the successor task.}
     Note that $m_u$ is the total memory usage
@@
     \label{sec.mod.plat}
-
+\skug{changes here}
 The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
@@
 We can decide to evict some data from the individual memory onto the disk to free space in the individual memory.
-This data can later be read back into memory at any time, or transferred to another processor.
+This data can later be read back into memory of the same or any other processpr at any time.
+\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).}
 
@@
     \[
-        w_{real} = \frac{w_u}{s_j} * (1 + (m_i - M_j)/m_i ) * q_s.
-
+        w_{real} = \frac{w_u}{s_j} \* (1 + (m_i - M_j)/m_i ) \* q_s.
     \]
-where $q_s$ sit he slowdown coefficient.
+where $q_s$ sit he slowdown coefficient and $(m_i - M_j)/m_i$ is the part of the task memory that needs to be offloaded.
+    \skug{check this equation}
 %We assume that all processors are connected with the same bandwidth~$\beta$.
@@
 The processor can read and write at the same time, but only one file can be read and only one written at any point in time.
-We keep track of the current ready times of each processor $j$ , ready time for computation $\rt_j^c$,
-ready time for writing $\rt_j^w$, and ready time for reading $\rt_j^r$.
+We keep track of the current ready times of each processor $j$, ready time for computation $\rt_j^c$ (when one task has finished
+computing and another one can start),
+ready time for writing $\rt_j^w$ (when a file has been written to disk and another file can be written), and ready time for
+ reading+$\rt_j^r$.
 Initially, all the ready times are set to~$0$.

### 2024-10-29 main-disk-model.tex
more description of the approach, also abt lazy writes

@@
 and then finds the traversal that leads to the minimum memory consumption.
+
+\skug{Todo: describe deviations}
 
@@
 It is computed as the maximum of the ready time of computation on this processor, and the time to read input files.
+For the currently reviewed processor, we preliminarily schedule necessary file writes on other processors holding input files,
+and file reads on the current processor.
+Doing so, we respect the ready times on writes on other processors and reads on the current processor.
+We possibly plan ``into the past'' with ready and writes.
+
+The earliest finish time is calculated by adding the execution time (possibly with delay due to memory overflow) to the start time.
+Further scheduling decisions happen when another task finishes.
+\skug{check this: when a task finishes earlier or later than planned, we review our earlier scheduling decisions and may
+move the task execution to another processor if this proves more beneficial.}
+
+When a task finishes, we look at all of its sucessors, rather than only the ready ones.
+For tasks that lie further in the future, we try to plan rather early.
+Therefore, we employ lazy writes.
+We try to write the files for these tasks to the disk early.
+So, we preliminarily schedule these file writes. However, they are not high priority and can be cancelled.
+We keep two ready times on writes, one soft, with lazy writes, and another hard onw, with scheduled writes.
+If a task requires a write during the lazy write, we cancel it and move it to later point in time.
 

### 2024-10-29 main-disk-model.tex
minor, equation

@@
     Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.
-   \skug{here?: For each edge, we also keep information on where it is currently stored - in memory of a processor (with processor id),
-    on the disk or nowhere - if this file has not yet been generated or has already been consumed by the successor task.}
+
     Note that $m_u$ is the total memory usage
@@
     \[
-        w_{real} = \frac{w_u}{s_j} \* (1 + (m_i - M_j)/m_i ) \* q_s.
+        w_{real} = \frac{w_u}{s_j} +    \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
     \]
@@
 
+\skug{For each edge, we also keep information on where it is currently stored - in memory of a processor (with processor id),
+        on the disk or nowhere - if this file has not yet been generated or has already been consumed by the successor task.}
+
 \subsection{Optimization problem}

### 2024-10-30 main-disk-model.tex
wrote about events, part 1

@@
 
-\subsection{Scheduling approach - general description}
+\subsection{Optimization problem}
+\label{sec.mod.pb}
+
+The goal is to find a schedule of the DAG~$G$ for the $k$ processors,
+so that the makespan (total execution time) is minimized. 
+\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
+I'll probably add a few sentences here...}
+% while respecting memory constraints. \skug{do we still formulate it like that?}
+    %If a processor runs out of memory to execute
+%a task mapped on it, the schedule is said to be {\em invalid}.
+
+Since tasks are subject to variability, we aim at minimizing the actual makespan
+achieved at the end of the execution, while decisions may be taken building
+on the estimated task parameters.
+\AB{Probably need to clarify offline vs online problem at some point...}
+
+Note that the problem is already NP-hard even in the homogeneous case and 
+without memory constraints, because of the DAG structure of the application. 
+Hence, we focus on the design of efficient scheduling heuristics. 
+%\hmey{Mention NP-hardness due to being more general than NP-hard problem?}
+
+%    \paragraph{Workflow-related changes}
+%
+%    \begin{itemize}
+%        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
+%        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
+%
+%    \end{itemize}
+%
+%    The following changes are not a part of this article's scope:
+%
+%    \begin{itemize}
+%        \item The workflow structure changes: edges or tasks come in or leave.
+%    \end{itemize}
+%
+%    \paragraph{Execution environment-related changes }
+%
+%
+%    \begin{itemize}
+%        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
+%        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
+%
+%    \end{itemize}
+%
+%    The following changes are not a part of this article's scope:
+%
+%    \begin{itemize}
+%        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
+%    \end{itemize}
+%
+%    \subsection{Time of changes }
+%
+%    We consider discrete time in seconds.
+%    The time point(s) at which the changes happen is unambiguously defined.
+%
+%    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
+%    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
+%    The start time of the source task in the workflow is zero.
+%    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
+%
+%    \subsection{Changes and knowledge horizon - important questions TBA}
+%
+%    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
+%
+%    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
+%    However, how do we model the following:
+%    \begin{itemize}
+%        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
+This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
+%        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
+%        \item How many times of change do we model - one per workflow run, or multiple?
+%        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
+%    \end{itemize}
+
+\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
+\label{sec:heuristics}
+\subsection{\skug{New scheduling approach - general description}}
 
@@
 
-\subsection{Optimization problem}
-\label{sec.mod.pb}
-
-The goal is to find a schedule of the DAG~$G$ for the $k$ processors,
-so that the makespan (total execution time) is minimized. 
-\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
-I'll probably add a few sentences here...}
-% while respecting memory constraints. \skug{do we still formulate it like that?}
-    %If a processor runs out of memory to execute
-%a task mapped on it, the schedule is said to be {\em invalid}.
-
-Since tasks are subject to variability, we aim at minimizing the actual makespan
-achieved at the end of the execution, while decisions may be taken building
-on the estimated task parameters.
-\AB{Probably need to clarify offline vs online problem at some point...}
-
-Note that the problem is already NP-hard even in the homogeneous case and 
-without memory constraints, because of the DAG structure of the application. 
-Hence, we focus on the design of efficient scheduling heuristics. 
-%\hmey{Mention NP-hardness due to being more general than NP-hard problem?}
-
-%    \paragraph{Workflow-related changes}
-%
-%    \begin{itemize}
-%        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
-%        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
-%
-%    \end{itemize}
-%
-%    The following changes are not a part of this article's scope:
-%
-%    \begin{itemize}
-%        \item The workflow structure changes: edges or tasks come in or leave.
-%    \end{itemize}
-%
-%    \paragraph{Execution environment-related changes }
-%
-%
-%    \begin{itemize}
-%        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
-%        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
-%
-%    \end{itemize}
-%
-%    The following changes are not a part of this article's scope:
-%
-%    \begin{itemize}
-%        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
-%    \end{itemize}
-%
-%    \subsection{Time of changes }
-%
-%    We consider discrete time in seconds.
-%    The time point(s) at which the changes happen is unambiguously defined.
-%
-%    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
-%    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
-%    The start time of the source task in the workflow is zero.
-%    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
-%
-%    \subsection{Changes and knowledge horizon - important questions TBA}
-%
-%    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
-%
-%    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
-%    However, how do we model the following:
-%    \begin{itemize}
-%        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
-This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
-%        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
-%        \item How many times of change do we model - one per workflow run, or multiple?
-%        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
-%    \end{itemize}
-
-\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
-\label{sec:heuristics}
+\subsection{Data structures }
+    To optimally represent the timeline of the execution of a workflow, we represent the execution as a series of events
+    with attached unique timestamps.
+    The events are kept in the priority queue ordered by timestamp.
+    Events are created when other events are triggered and are being inserted into the queue.
+    When their timestamp is the smallest, they are then triggered, possibly creating new events.
+    The execution starts with placing the event of starting the starting node into the queue with the timestamp 0.
+    The execution ends when there are no events in the queue.
+
+    An event can happen on finish or start of a task execution or an io action (file read or write).
+    Each event has several parameters:
+    \begin{itemize}
+        \item Event type: OnTaskStart, OnTaskFinish, OnIoStart, OnIoFinish
+        \item The action associated with the event: the task to be executed or an edge whose file needs to be read or written
+        \item The processor id on which the action should take place
+        \item Expected and actual time of firing.
+        \item A set of predecessor events. The event is called ready if all its predecessors have been finished.
+        \item A set of successor events to keep track of.
+    \end{itemize}
+
+    Per processor, we keep the currently latest finishing task, read and write, and their estimated finishing times (ready times
+    for computation, reads and writes).
+
+    \paragraph{On Task Finish}
+    When a task finishes, it first cleans after itself:
+    \begin{itemize}
+        \item It frees its memory: it increases the available memory on the processor by the amount it occupied
+        \item If there had been a memory overflow onto disk, then this memory is being freed implicitly
+        \item it sets its status to finished in the workflow.
+    \end{itemize}
+    It the goes over its successor events (Tasks and IO actions) and removes itself from their predecessor lists.
+    It also updates their actual trigger times with its actual finish time.
+    Successor tasks can be tasks scheduled on the same processor after this one.
+    Successor io events can be reads that could not start until the task finishes and frees its memory.
+    \skug{can writes be successor events of a task?}
+
+    Then task goes over its successor tasks in the workflow and schedules ready ones.
+    Ready tasks are those whose all predecessors have the finished status.
+
+    For each ready task, the scheduler tries to tentatively put it on each processor in the cluster and chooses the processor
+    that promises the earliest finish time.
+    The finish time on each processor is calculated as follows:
+
+    \begin{itemize}
+        \item For each incoming edge, its location is determined.
+        \begin{itemize}
+                 \item If the edge is in the memory of the target processor, nothing needs to be done
+                  \item If the edge is on the disk, it first needs to be read into the memory of the target processor (the computation is delayed by at least this read time).
+                 If there is not enough memory to hold the file, the task cannot be scheduled on this processor
+                  \item If the edge is in another processor's memory, it needs to be first written to disk, then read by the target processor. Same holds with memory amount estimation.
+        \end{itemize}
+        \item The finish time is the earliest start time (max of incoming edge times and the ready time of computation of the processor) plus compuation time of the task.
+        \item The computation time of the task is calculated with the memory overflow formula.
+    \end{itemize}
+
+    When a processor is chosen, the scheduler inserts new events into the queue:
+    \begin{itemize}
+        \item For the task, OnTaskStart and OnTaskFinish events with the estimated trigger times, the OnTaskFinish event being dependent on the OnTaskStart event.
+        \item For each io operation necessary for the execution: OnIoStart event of the corresponding type and on corresponding processor
+        \item It connects all the event with predecessor-successor relations.
+    \end{itemize}
+    THe scheduler also updates the ready times of the processors and updates the new currently last events there.
+
+    \paragraph{On Task Start}
+    When a task start event is triggered, it determines the deviation of the runtime of the task.
+    It applies the deviation function on its estimated runtime (estimated finish minus estimated start time).
+    It then adds this deviated runtime to its actual start time and receives an updated actual finish time.
+    With this time, it updates the OnTaskFinish event of this task.
 

### 2024-10-30 main-disk-model.tex
minor
more written on events

@@
         \item A set of successor events to keep track of.
+        \item For a write event, a boolean if this file should be removed from memory after being written.
     \end{itemize}
@@
     \end{itemize}
-    THe scheduler also updates the ready times of the processors and updates the new currently last events there.
+    The scheduler also updates the ready times of the processors and updates the new currently last events there.
 
     \paragraph{On Task Start}
-    When a task start event is triggered, it determines the deviation of the runtime of the task.
+    When a task start event is triggered, it first determines if it can start.
+    If some predecessor tasks have not yet finished (the predecessor set is not empty), then the actual time of this event
+    is set to the maximum of the estimated end times of predecessor events and this event is reinserted in the queue
+    with this new time.
+
+    If the task can start, the scheduler determines the deviation of the runtime of the task.
     It applies the deviation function on its estimated runtime (estimated finish minus estimated start time).
@@
     With this time, it updates the OnTaskFinish event of this task.
+
+    \paragraph{OnIoFinish}
+    When the io operation finishes, it removes itself from the predecessor sets of all its successor events.
+    It also sets their actual time to its own actual finish time.
+
+    \paragraph{OnIOStart}
+    When the IO operation starts, it updates the corresponding IoFinish event
+
+    After being triggered and completed, each event is deleted from the queue.
+
+--------------------------------------
 

### 2024-10-30 main-disk-model.tex
corrected about reads and evictions

@@
         \item Event type: OnTaskStart, OnTaskFinish, OnIoStart, OnIoFinish
-        \item The action associated with the event: the task to be executed or an edge whose file needs to be read or written
+        \item The workflow entity associated with the event: the task to be executed or an edge whose file needs to be read or written
         \item The processor id on which the action should take place
@@
 
-    Per processor, we keep the currently latest finishing task, read and write, and their estimated finishing times (ready times
-    for computation, reads and writes).
-
-    \paragraph{On Task Finish}
+    Per processor,  in addition to keeping the fixed values (memory, speed), we keep track also of the current state in its memory.
+    Because during and after execution of each even the state of the processor memory changes, we need to keep snapshots of what files
+    were in memory and how much of it was available.
+    Because keeping such snapshots for each timestap is too complicated, we only keep two timestamps.
+    First, we keep the information of available memory and tasks pending in memory during the execution of the last
+    task scheduled to this processor.
+    Then, we keep the same information on what it will be immediately after this task finishes.
+
+    \subsubsection{On Task Finish}
     When a task finishes, it first cleans after itself:
@@
 
+    \paragraph{Earliest start and finish time}
+    We find out when the last task on this processor finishes.
+    If there is enough memory to start our task after that, then the end of previous task is the earliest start time for our task.
+    If files need to be evicted first, we need to find a balance between evicting tasks (and waiting for their writes to finish),
+    then running our task, and starting the task immediately and letting it run longer due to memory overflow.
+    We compare the following cases:
     \begin{itemize}
-        \item For each incoming edge, its location is determined.
-        \begin{itemize}
-                 \item If the edge is in the memory of the target processor, nothing needs to be done
-                  \item If the edge is on the disk, it first needs to be read into the memory of the target processor (the computation is delayed by at least this read time).
-                 If there is not enough memory to hold the file, the task cannot be scheduled on this processor
-                  \item If the edge is in another processor's memory, it needs to be first written to disk, then read by the target processor. Same holds with memory amount estimation.
+        \item Evict everything, then start task
+        \item Greedily evict the largest file, starting the task
+        \item Not evicting anything, starting the task
     \end{itemize}
-        \item The finish time is the earliest start time (max of incoming edge times and the ready time of computation of the processor) plus compuation time of the task.
-        \item The computation time of the task is calculated with the memory overflow formula.
-    \end{itemize}
-
-    When a processor is chosen, the scheduler inserts new events into the queue:
+    Whatever combination gives the best estimated finishing time for the task, we note the estimated starting time of the task in this case.
+     The finish time is the earliest start time  plus computation time of the task.
+
+    \paragraph{Incoming edges} For each incoming edge, its location is determined.
+    If the edge is in the memory of the target processor, nothing needs to be done.
+    The estimated start and finish times for the task are unaffected.
+
+    If the edge is on the disk, it first needs to be read into the memory of the target processor.
+    We delay the reads to avoid holding unnecessary files in the memory.
+    When we preliminarily schedule a read, we first assess its length and other actions that happen on this processor in the meantime.
+    So, the first estimated start of the read is the estimated start time of the task minus the estimated length of the read.
+    We then see if this start of the read is happening during the runtime of the previous task or already after it (during the writes).
+    If the starting time of the read is during the execution of the previous task and there is enough memory to read this file, then we move the read
+    forward to the beginning of this previous task.
+    If there is not enough memory, then the entire read is rescheduled for after the task finishes.
+    The read will then be bottleneck and affect the eraliest start time of the desired task.
+
+     If the edge is in another processor's memory, it needs to be first written to disk, then read by the target processor.
+    So we look at the latest write on that processor, preliminarily schedule the write to disk and then proceed like in the
+    previous case, reading the file from disk.
+
+   \paragraph{When processor is chosen} When a processor is chosen, the scheduler inserts new events into the queue:
     \begin{itemize}
@@
         \item It connects all the event with predecessor-successor relations.
+        \item If files needed to be evicted
     \end{itemize}
@@
 
-    \paragraph{On Task Start}
+    \subsubsection{On Task Start}
     When a task start event is triggered, it first determines if it can start.
@@
 
-    \paragraph{OnIoFinish}
+    \subsubsection{OnIoFinish}
     When the io operation finishes, it removes itself from the predecessor sets of all its successor events.
@@
 
-    \paragraph{OnIOStart}
+    \subsubsection{OnIOStart}
     When the IO operation starts, it updates the corresponding IoFinish event

### 2024-10-30 main-disk-model.tex
text about fork instance, example not quite accurate yet

@@
 
-%    \subsection{The fork}
-%    We look at the behavior of these heuristics on a fork graph,
-%    where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
-%    to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
-%
-%    Without memory, this problem is NP-complete; this is equivalent
-%    to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
-%    and with two processors. Half of the tasks must be sent to the processor
-%    on which $T_0$ is not executed, and the optimal makespan is
-%    $w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
-%
-%    However, with an infinite number of identical processors, it can be
-%    solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$;
-%    the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
-%    while the remaining $n-k$ tasks are executed locally (try all values of $k$).
-%
-%    With heterogeneous processors, it is probably NP-complete again
-%    because we could ensure that there are only two processors fast enough
-%    and get back to the 2-partition...
-%
-%    We also had an example where evicting large files first in step 2
-%    can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
-%    $f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
-%    imposes that we free one unit of memory before executing one
-%    of the tasks\ldots Actually the new version with BLC would start
-%    considering $T_2$ and be fine in this case\ldots
-%
-%
-%    \AB{Can we prove that we have (maybe) a 2-approximation,
-%        at least for the fork? What worst-case can we think of? }
+\hmey{BEGIN FORK}
+    \subsection{The fork}
+    We look at the behavior of these heuristics on a fork graph,
+    where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
+    to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).
+
+    Without memory, this problem is NP-complete; this is equivalent
+    to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
+    and with two processors. Half of the tasks must be sent to the processor
+    on which $T_0$ is not executed, and the optimal makespan is
+    $w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.
+
+    However, with an infinite number of identical processors, it can be
+    solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$;
+    the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
+    while the remaining $n-k$ tasks are executed locally (try all values of $k$).
+
+    With heterogeneous processors, it is probably NP-complete again
+    because we could ensure that there are only two processors fast enough
+    and get back to the 2-partition...
+
+    We also had an example where evicting large files first in step 2
+    can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
+    $f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
+    imposes that we free one unit of memory before executing one
+    of the tasks\ldots Actually the new version with BLC would start
+    considering $T_2$ and be fine in this case\ldots
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{images/bad-blc-instance.png} \\[1ex]
+  \includegraphics[width=0.999\columnwidth]{images/blc-bad-cost.png} \\[1ex]
+  \includegraphics[width=0.58\columnwidth]{images/opt-cost.png}
+  \caption{Top: Example instance $I$, middle: BLC applied to $I$ leads to makespan $w_0 + nk/2 + k$, bottom: optimal solution has makespan $w_0 + n/2$.
+  \hmey{TODO: adapt resolution of figure and fix $\epsilon$ problem}}
+  \label{fig:fork-bad-blc-instance}
+\end{figure}
+
+Figure~\ref{fig:fork-bad-blc-instance} shows an instance where \heftblc\ performs very poorly.
+Since the sorting criterion $w_i + c_{0i}$ yields an arbitrary ordering of the tasks $T_i$ (with $1 \leq i \leq n$), we may
+assume that the algorithm processes the nodes in level $1$ of instance $I$ from left to right. 
+As shown in the middle part of the figure, $T_1$ is processed...
+\hmey{TODO: adapt with $\epsilon$ to enforce even tasks to be ``sent away''}
+This leads to a makespan
+of $w_0 + nk/2 + k$. The optimal solution, in turn, as shown in the bottom,
+requires only $w_0 + n/2$ timesteps. Since $k$ can be chosen arbitrarily large if the value $10$ is scaled accordingly,
+this instance shows that \heftblc\ may yield solutions with an arbitrarily large gap to optimality.
+
+\AB{Can we prove that we have (maybe) a 2-approximation,
+        at least for the fork? What worst-case can we think of? }
+\hmey{END FORK}
 

### 2024-10-31 main-disk-model.tex

@@
     (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
-    In the processor choice phase, the entry task and the longest parent tasks are duplicated 
+    In the processor choice phase, the entry task and the longest predecessor tasks are duplicated 
     during idle times on the processor.
@@
     %execution time for the task is chosen.
-    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
+    The heuristic also uses the same processor to schedule predecessor and successor tasks, as to avoid data transfer between processors.
     The approach is evaluated on random workflows of the size up to 300 tasks.
@@
 
-\skug{
 \section{Model} %: \skug{Full: 4, polished: 3}}
@@
 
-CHAGES IN THIS CHAPTER
-}
+%\skug{
+%CHANGES IN THIS CHAPTER
+%}
     We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
@@
     A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
-    $E$ is a set of directed edges of the form $e=(u,v)$, with $u,v\in V$, expressing precedence constraints between tasks.
+    $E$ is a set of directed edges of the form $e=(u,v)$, with \mbox{$u,v\in V$}, expressing precedence constraints between tasks.
     Each task~$u \in V$  is performing $w_u$ operations, and it also
@@
     of a task during its execution, including input and output files currently being read and written,
-    and hence the total memory requirement  for the execution of task~$u$ consists of the maximum
-    between the input files
-    (total size of the files to be received from the parents),
-    the output files (total size of the files to be sent to the children),
-    and the total memory size~$m_u$ (usually achieving the maximum):
-    \[
-        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
-    \]
-
-    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
-    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without parents is called a {\it source task}.
-    The children tasks of~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
-    $ \children{u} = \{v \in V: (u,v) \in E\}$. A task without children is called a {\it target task}.
-    Each task may have multiple parents and children.
+    and hence it is greater than the total size of input files
+    (received from the predecessor tasks) and than 
+    the total size of  output files (sent to successor tasks): 
+    $$   m_u \geq \max \left\{ \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v} \right\} . $$
+
+%     the total memory requirement  for the execution of task~$u$ consists of the maximum
+%    between the input files
+%    (total size of the files to be received from the parents),
+%    the output files (total size of the files to be sent to the children),
+%    and the total memory size~$m_u$ (usually achieving the maximum):
+%    \[
+%        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
+%    \]
+
+    The predecessors of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of predecessors is
+    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without predecessors is called a {\it source task}.
+    The successors of a task~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
+    $ \children{u} = \{v \in V: (u,v) \in E\}$. A task without successors is called a {\it target task}.
+    Each task may have multiple predecessors and successors.
 
 Furthermore, we place ourselves in a context where we do not have perfect knowledge
-of the task parameters ($w_u$ and $m_u$) before the tasks start their execution,
+of the length of the tasks, i.e., $w_u$, % ($w_u$ and $m_u$) 
+before the tasks complete their execution,
 but only estimates~\cite{rahman2013,GARG2015256}.  
@@
 Hence, scheduling decisions are made on these estimated parameters, and
-may be reconsidered at runtime when a task starts its execution and we know its exact parameters.
-\AB{Variability only on $w_u$ for now}
+may be reconsidered at runtime when a task completes its execution. 
+%when a task starts its execution and we know its exact parameters.
+%\AB{Variability only on $w_u$ for now}
 
@@
     \label{sec.mod.plat}
-\skug{changes here}
+
+%\skug{changes here}
 The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
 consists of $k$ processors $p_1, \dots, p_k$.
-Each processor $p_u$ ($1 \leq u \leq k$) has an individual memory of size $M_u$ and a speed~$s_u$.
+Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
 All processors have access to a shared disk of unlimited size, where they can write
-and read data: Processor~$p_u$ has a bandwidth $\bw_u$ (resp. $\br_u$) to write  (resp. read),
-and hence it takes a time $\frac{c_{u,v}}{bw_u}$ to write the data produced for~$p_v$, 
+and read data: Processor~$p_j$ has a bandwidth $\bw_j$ (resp.~$\br_j$) to write  (resp. read),
+and hence it takes a time $\frac{c_{u,v}}{bw_j}$ to write the data produced by task~$u$
+on~$p_j$ for task~$v$, 
 if this data is evicted from memory.
@@
 %, but of much slower speed.
-All communications happen over the disk - after one processor has finished writing data there, another one can read
-it into its individual memory.
-We can decide to evict some data from the individual memory onto the disk to free space in the individual memory.
-This data can later be read back into memory of the same or any other processpr at any time.
-\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).}
-
-If the entire memory required by the task ~$u\in V$ fits in the free memory of a processor~$p_j$,
+All communications happen over the disk -- after one processor has finished writing data there, another one can read
+it to load the data on its individual memory. 
+Hence, we can decide to evict some data from the individual processor memories if it has been 
+written onto the disk, 
+in order to free some space in the individual memories.
+This data can later be read back into memory of the same or any other processor at any time.
+%\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).} \AB{We can keep two copies: evict if we need to }
+
+If the entire memory required by  task ~$u\in V$ fits in the available memory of processor~$p_j$,
     then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
-If the memory requirement of the task exceeds the available memory on the processor, the task can still be executed there,
-    albeit slower.
-The part of memory requirement that exceeds the available memory will be offloaded to the disk, slowing down the execution.
+However, if the memory requirement of the task exceeds the available memory on the processor, 
+then task can still be executed there, but it will be slowed down. Indeed, 
+%    albeit slower.
+the part of memory requirement that exceeds the available memory on processor~$j$, $availM_j$, 
+will be offloaded to the disk, slowing down the execution:
 
     \[
-        w_{real} = \frac{w_u}{s_j} +    \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
+        w_{real} = \frac{w_u}{s_j} +  q_s \times   (m_u - availM_j) , %   \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
     \]
-where $q_s$ sit he slowdown coefficient and $(m_i - M_j)/m_i$ is the part of the task memory that needs to be offloaded.
-    \skug{check this equation}
+where $q_s$ is the slowdown coefficient, and $(m_u - availM_j)$ is the part 
+of the task memory that needs to be offloaded.
+\AB{Svetlana please check, I think we said in the end that we use absolute value 
+of the amount of memory not fitting locally...}
+%    \skug{check this equation}
 %We assume that all processors are connected with the same bandwidth~$\beta$.
@@
 ready time for writing $\rt_j^w$ (when a file has been written to disk and another file can be written), and ready time for
- reading+$\rt_j^r$.
+ reading~$\rt_j^r$.
 Initially, all the ready times are set to~$0$.
@@
 Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
-that are in the memory of size $\MM_j$ but may be evicted to the disk
-more memory is needed on~$p_j$.
-They are ordered by non-decreasing size and correspond to some $c_{u,v}$.
-When scheduling a tass, the decision has to be made to execute this task on a processor with enough memory
+that are in the memory of processor~$p_j$,  but that may be evicted to the disk
+if more memory is needed on~$p_j$.
+They are ordered by non-decreasing size and correspond to some files~$c_{u,v}$.
+When scheduling a task, the decision has to be made to execute this task on a processor with enough memory
 without slowdown, but with possibly more time to read the input files, or to execute the same task
-    on a processor with less memory (and hence a slowdown), but potentially with a higher processor speed or with less times
-    spent reading files.
+on a processor with less memory (and hence a slowdown), but potentially with a higher processor speed 
+or with less times spent reading files.
 
@@
 and then finds the traversal that leads to the minimum memory consumption.
-
-\skug{Todo: describe deviations}
+\AB{When do we need this? Probably more in Algorithms, right? }
+
+%\skug{Todo: describe deviations}
 
@@
                 $G = (V, E)$                          & Workflow graph, set of vertices (tasks) and edges        \\
-                $\parents{u}$, $\children{u}$         & Parents of a task $u$, children of a task $u$            \\
+                $\parents{u}$, $\children{u}$         & Predecessors of a task $u$, successors of a task $u$            \\
                 $m_u$                                 & Memory weight of task $u$                                \\

### 2024-10-31 main-disk-model.tex
temp

@@
 \newcommand{\br}{br}
+\newcommand{\Moff}[1]{m^{\text{off}}_{#1}}
 
@@
 %    albeit slower.
-the part of memory requirement that exceeds the available memory on processor~$j$, $availM_j$, 
+the part of memory requirement that exceeds the available memory on processor~$j$ for task~$u$, 
+denoted by $\Moff{u}$, %M\text{off}_u$, 
 will be offloaded to the disk, slowing down the execution:
@@
     \[
-        w_{real} = \frac{w_u}{s_j} +  q_s \times   (m_u - availM_j) , %   \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
+        w_{real} = \frac{w_u}{s_j} +  q_j \times  \Moff{u}, %  (m_u - Mav_u) , %   \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
     \]
-where $q_s$ is the slowdown coefficient, and $(m_u - availM_j)$ is the part 
-of the task memory that needs to be offloaded.
+where $q_j$ is the slowdown coefficient for $p_j$. %, and $(m_u - Mav_u)$ is the part 
+%of the task memory that needs to be offloaded.
 \AB{Svetlana please check, I think we said in the end that we use absolute value 
@@
 
-The goal is to find a schedule of the DAG~$G$ for the $k$ processors,
-so that the makespan (total execution time) is minimized. 
-\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
-I'll probably add a few sentences here...}
+In the {\bf offline setting}, the goal is to find a {\em schedule} of the DAG~$G$ for the $k$ processors,
+so that the makespan (total execution time) is minimized. Formally, a schedule contains:
+\begin{itemize}
+\item for each task, its processor allocation and starting time, 
+as well as the amount of data offloaded for the execution of the task; 
+\item for each read and write operation to disk, the starting time of the operation; 
+\item for each data and processor, the time intervals during which the file is available
+on the local memory of the processor (once it has been generated or read, and before 
+it is used or evicted). 
+\end{itemize}
+
+\AB{I'm confused again about the offloading model, we need to talk}
+
+%\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
+%I'll probably add a few sentences here...}
 % while respecting memory constraints. \skug{do we still formulate it like that?}
@@
 
-Since tasks are subject to variability, we aim at minimizing the actual makespan
-achieved at the end of the execution, while decisions may be taken building
+\medskip
+We also consider the {\bf online setting} where tasks are subject to variability, 
+and then we know the exact time required to complete a task only when 
+it ends its execution. Hence, we aim at minimizing the actual makespan
+achieved at the end of the execution, while scheduling decisions 
+have to be taken building
 on the estimated task parameters.
-\AB{Probably need to clarify offline vs online problem at some point...}
-
-Note that the problem is already NP-hard even in the homogeneous case and 
-without memory constraints, because of the DAG structure of the application. 
+In this case, we do not build the whole schedule offline, but we build
+it on the fly, as tasks become ready (their predecessor tasks have been completed). 
+%\AB{Probably need to clarify offline vs online problem at some point...}
+
+Note that the offline problem is already NP-hard even in the homogeneous case and 
+without memory constraints (i.e., no need to offload data onto disk), and even for
+a graph without precedence constraints (independent tasks). 
+%because of the DAG structure of the application. 
 Hence, we focus on the design of efficient scheduling heuristics. 

### 2024-10-31 main-disk-model.tex
power point figures updated

--- 2024-10-31 main-disk-model.tex
@@
   \centering
-  \includegraphics[width=0.5\columnwidth]{images/bad-blc-instance.png} \\[1ex]
+  \includegraphics[width=0.6\columnwidth]{images/bad-blc-instance.png} \\[1ex]
   \includegraphics[width=0.999\columnwidth]{images/blc-bad-cost.png} \\[1ex]
-  \includegraphics[width=0.58\columnwidth]{images/opt-cost.png}
+  \includegraphics[width=0.67\columnwidth]{images/opt-cost.png}
   \caption{Top: Example instance $I$, middle: BLC applied to $I$ leads to makespan $w_0 + nk/2 + k$, bottom: optimal solution has makespan $w_0 + n/2$.
-  \hmey{TODO: adapt resolution of figure and fix $\epsilon$ problem}}
+  \hmey{TODO: adapt resolution of figure}}
   \label{fig:fork-bad-blc-instance}
@@
 assume that the algorithm processes the nodes in level $1$ of instance $I$ from left to right. 
-As shown in the middle part of the figure, $T_1$ is processed...
-\hmey{TODO: adapt with $\epsilon$ to enforce even tasks to be ``sent away''}
+\hmey{As shown in the middle part of the figure, $T_1$ is processed...}
 This leads to a makespan
-of $w_0 + nk/2 + k$. The optimal solution, in turn, as shown in the bottom,
-requires only $w_0 + n/2$ timesteps. Since $k$ can be chosen arbitrarily large if the value $10$ is scaled accordingly,
-this instance shows that \heftblc\ may yield solutions with an arbitrarily large gap to optimality.
+of $w_0 + nk/2 + 1 + \epsilon$. The optimal solution, in turn, as shown in the bottom,
+requires only $w_0 + n/2 + k$ timesteps. This means that \heftblc\ may yield solutions whose 
+gap of optimality $\approx k$.
 

### 2024-10-31 main-disk-model.tex
text on fork instance updated

@@
   \includegraphics[width=0.67\columnwidth]{images/opt-cost.png}
-  \caption{Top: Example instance $I$, middle: BLC applied to $I$ leads to makespan $w_0 + nk/2 + k$, bottom: optimal solution has makespan $w_0 + n/2$.
+  \caption{Top: Example instance $I$, middle: BLC applied to $I$ leads to makespan $w_0 + nk/2 + 1+\epsilon$, bottom: optimal solution has makespan $w_0 + n/2 + k$.
   \hmey{TODO: adapt resolution of figure}}
@@
 assume that the algorithm processes the nodes in level $1$ of instance $I$ from left to right. 
-\hmey{As shown in the middle part of the figure, $T_1$ is processed...}
-This leads to a makespan
-of $w_0 + nk/2 + 1 + \epsilon$. The optimal solution, in turn, as shown in the bottom,
-requires only $w_0 + n/2 + k$ timesteps. This means that \heftblc\ may yield solutions whose 
-gap of optimality $\approx k$.
+As shown in the middle part of the figure, $T_1$ is processed on $P_1$ due the earliest finishing time on this processor.
+$T_2$ is then sent to $P_2$ because it finishes there $\epsilon$ timesteps earlier than on $P_1$. The same rationale applies
+to the other odd/even tasks. This leads to a makespan of $w_0 + nk/2 + 1 + \epsilon$. The optimal solution, in turn, as shown in the bottom,
+requires only $w_0 + n/2 + k$ timesteps (assuming $k > n\epsilon/2$). This means that \heftblc\ may yield solutions whose 
+gap of optimality is $\Omega(k)$.
 

### 2024-10-31 main-disk-model.tex
new fork example

@@
 
+Here is the new fork example, that would lead to both \heftbl and \heftblc to get arbitrarily bad result. 
+Consider a fork with $n=2$, $w_0=0$, $w_1=1$, $w_2=2$. Furthermore, 
+$f_1=1$ and $f_2=K$, where $K$ is an arbitrarily large number. 
+The memory requirements of tasks are $m_0=K+1$, $m_1=1$, and $m_2=2K$, 
+while the memory size on each processor is~$2K$.
+
+Both heuristics would first try to schedule task~$T_2$. At the end of $T_0$ on processor~1,
+the memory usage is $1+K$; to execute~$T_2$ on $p_1$, we would go up to $1+2K = f_1+m_2$, which
+exceeds the available memory on~$p_1$. If the offloading penalty $q_1$ is very high,
+we would rather schedule $T_2$ on $p_2$ at time $2K$ (write $f_2$ on disk, read it back from disk,
+assuming write/read bandwidths to be~$1$), hence completing execution at time~$2K+2$. 
+Then, $T_1$ is scheduled on $p_1$ and completes at time~$1$. 
+\AB{Actually, we can consider having a slow bandwidth as well, but offloading is even worth,
+hence that would be enough to take the decision to execute $T_2$ on~$p_2$. }
+
+However, if we schedule~$T_1$ first, we execute it on~$p_1$, free the memory of $f_1$, 
+and then we can execute~$T_2$ on $p_1$ as well without offloading, 
+hence an optimal makespan of~$3$.
+
+However, \heftmm would behave nicely here, since the traversal to minimize peak memory
+would first consider~$T_1$ and lead to the optimal solution.
+
 \section{Dynamic scenario}

### 2024-12-04 rebuttalIPDPS.tex
first version of the rebuttal, very WIP! Reviews are not answered after section 4.1. Everything is still to be improved.

\documentclass{article}
\usepackage{xcolor}
\newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
\newcommand{\hmey}[1]{{\color{red}[HM: #1]}}
\newcommand{\AB}[1]{{\color{purple}[AB: #1]}}

\newcommand{\review}[1]{{\color{orange}[Reviewer: #1]}}

\begin{document}
    \title{Rebuttal}

    \maketitle

    \section{Writing and chapters 1-4}

   Reviewer 2 found that the paper has similarities with paper [20].
   We will rewrite the paragraphs that show unacceptable formulations.
   However, the cited sentences are so similar because they refer to standard scheduling aspects.
   As the reviewer notes, the approach of our paper and paper [20] are different.

    We will also make the related work section more concise and address all minor comments related to the text.

    \section{Model}

    We designed our model to be realistic enough to represent the actual executions of the workflow, but still concise
    enough to be able to formulate algorithms for it.

    If running the heuristics is in the order of thousands of seconds,
    is possible that a reassignment occurs that have already completed? What
    happens in this case? It seems like a very important discussion to
    claim  the advantage of dynamicity.

    \review{I dont fully understand the argument of reusing finished processors during execution.
    Is this referring to task reuse maybe? }
    The reuse of processors refers to scheduling tasks on processors, rather than only mapping them there.
    When in the authors of DagHetPart~[20] map a block of tasks to a processor, then this processor remains idle after
    all tasks in this block have finished.
    In our approach, this does not happen, because tasks are being put on processors under consideration of their ready times.
    When a processor becomes free, it is likely that a new task will be assigned there soon.
    This allows us to achieve better makespans than in the approaches with mapping.

    \review{I disagree with the assumption that input files need to be loaded entirely; why not consider partial reads?
    Is this a limitation introduced by the experimental data traces from Sec. VI-A(b)?}
    In our approach, we assume that all input files need to be held in memory.
    Attempting partial reads would unnecessarily complicate the algorithmic setup.
    However, it does not bring a special knowledge, because sooner or later all the files will need to be in the memory
    and at that point in time there should be enough of this memory.
    This would not solve the problems with invalid schedules if there is not enough memory on processors to hold the input files.

    \review{Is the notation in $r_u$ definition correct? It is confusing because  it is looking into all pairs as
    bidirectional (from u to v and from v to  u).}
    The definition of $r_u$ is not bidirectional, but rather contains the sums of weights of both incoming and outgoing edges.
    In the second term, the edges (v,u) are incoming edges into our task u.
    In the thrid term, the edges (u,v) are outgoing edges of the same task u.
    v refers to different set of tasks in these terms.
    We will explain this formula better, though.

    \review{Section 4.a.3: How does the runtime handle nodes that are down?}
    In our runtime model, processors do not fail.
    The task can take much longer to execute on a processor than expected, but the execution finishes sooner or later.
    Adding processor failures is a good use case for the future work.

    \section{Description of the algorithms}

    \review{Page 5, the greedy eviction approach: So the algorithm doesn't take into account whether the to-be-evicted
    data could be required for some future next tasks on that processor?}
    We evict files greedily, without taking into consideration when they will be required for the execution of a task,
    because this would require a look into the future.
    This look, however, is difficult, because the present (i.e. the assignment of the current task) is not defined yet
    and the fitire depends on it.
    It is possible that we evict a file that will be needed soon, but the fact that we are required to evist it means that
    there was no (good) way of assigning the task without doing so.

    \subsection{Description of the dynamic algorithm}
    \review{Section V is quite confusing. Please provide an algorithm to describe the dynamic scenario.

    Description of the method reads unclear.
    }

    We will improve the description of the dynamic scenario in Section V and describe all algorithms, especially the dynamic one
    in more detail.

    \section{Experimental evaluation}

    \subsection{Description of experiments}

    \review{Page 7, right column, 2nd paragraph: "set the size of the communication buffer to be equal to 10x the memory size" -- I thought communication buffer is allocated from a part of device memory, so its size can't be larger than the memory size.
    Could you please explain why the communication buffer can be larger than the memory size?}
    We model the communication buffer as some sort of larger, slower memory, similar to a hard drive disc.
    Therefore, the size of the buffer is much larger than that of the memory, and we can store more files there.
    These files then can wait to be consumed by their target task without unnecessarily filling the memory of their processor.

    \review{The runtime system does not execute real workflows. It just simulates the execution, applying a deviation function.}
    \skug{I have trouble formulating rebuttal for this. How do I say that we don't want to run actual workflow in actual cluster,
        because this is crazy complicated to organize?}
    Our simulation is based on the measured historical data.
    The deviation function is also in accordance to values observed in practice.

    \review{No comparison with other dynamic scheduling baselines, e.g., with reference [20] in the paper.}

    TODO for Svetlana: say that comparison is unfair

    Page 8, charts: How do you aggregate the numbers across many different workflow graphs to report in the charts? Average, median, etc.?

    \subsection{Execution of experiments}

    \review{Why is the memory footprint of HEFTM-MM decreasing in some cases in Fig. 4?}

    For some larger workflows, HEFT-MM was able to achieve better memory usage.
    For these size categories of workflows, HEFT-MM can achieve a better memory usage. On a sufficiently large cluster, a larger workflow is not necessarily always bad for the memory usage.
    A large workflow brings more chances at parallel execution, and more variants of rearranging the tasks in relation to each other.
    These freedoms can lead to a better memory usage in comparison to small workflows that offer less flexibility and require tasks to be executed in a more fixed order.

    This is especially true on less fanned-out workflows like bacass.

    Fig. 9: For workflow size < 100, why is the runtime of HEFT much higher than others'?

    A better discussion on the experimental results could help deriving
    insights on the suitability of the different heuristics in different
    scenarios. First, it is unclear whether all workflow traces were mixed
    in the evaluation and what are their computational requirements. Second,
    in the static scenario, HEFTM-BL leads to a consistently higher memory
    usage (Fig. 4), but minimises slowdown. Also, HEFTM-MM is consistently
    more robust when it comes to deliver valid scheduling plans, even under
    memory constraints. Are there any characteristics of the tested
    workflows (besides task number) that can be aligned with this behaviour,
    and to what extent can this be generalised? More comments below.

    \section{For us to discuss}

    \begin{itemize}
        \item Page 4, 1st paragraph, ``mu is the best total memory usage of a task during its execution, including input and output files currently being read and written''
        Doesn't this mean mu will always be at least as large as the size of the input files and the size of the output files combined? If so, ru should just be mu

        The reviewer is right, the sentence they refer to is a bit confusing. We need to formulate this differently.
        But I don't think we should include it explicitly, it is covered by ``addressing minor review points''.

        \item  Page 7, left column, 2nd to last paragraph: Please report nonzero statistics for the real, historical data too.

        don't understand this one
        \item Reviewer 2: The experimental comparison is insufficient

        Don't know if a reaction to this is necessary?
        \item  Section 4.a.1.b: Please summarize execution time, memory requirement, and I/O size statistics for existing historical data as well.

        should I address this separately?
        \item
        \item

    \end{itemize}

\end{document}

### 2024-12-22 main.tex
first part of abstract (slightly) rewritten to get rid of duplicate text

--- 2024-10-11 main.tex
@@
 \begin{abstract}
-
-        %Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
-        Scientific workflows are often represented as directed acyclic graphs (DAGs),
-        where vertices correspond to tasks and edges represent the dependencies between them.
-        Typically, each task requires a
-        certain amount of memory to be executed and needs to communicate data to its successor tasks.
-        The goal is generally to execute the workflow  as fast as possible (i.e., to minimize its makespan),
-        while satisfying the memory constraints.
-        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows on
+The analysis of massive scientific data often happens in the form of workflows with
+interdependent tasks. When such a scientific workflow needs to be scheduled on a parallel or distributed
+system, one usually represents the workflow as a directed acyclic graph (DAG). 
+The vertices of the DAG represent the tasks, while 
+its edges model the dependencies between the tasks (usually data to be communicated to
+successor tasks). When executed, each task requires a certain amount of memory and if that
+exceeds the available memory, the execution fails.
+The typical goal is to execute the workflow without failures (i.e., satisfying the memory
+constraints) and with the shortest possible execution time (i.e., to minimize its makespan).
+
+To address this problem, we investigate in this paper the memory-aware scheduling of DAG-shaped workflows on
 heterogeneous platforms, where each processor can have a different speed and a different memory size.
@@
 Note that in both clusters, we set the size of the communication buffer to be equal
-to the memory size. \todo[inline]{Is this correct? }
+to ten times the memory size.
 
@@
 
-%\AB{Any info on success rate of adaptive heuristic compared to static ones? 
-%For which cluster do we have some results? }
+\AB{Need to add results on the default cluster!}
 

### 2024-12-22 main.tex
first part of intro slightly edited

@@
 
-%    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
-%
-%    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
-%    }
-%    TODO: Insert abstract
-
 %%% CONTEXT %%%
@@
 single computer and are therefore executed on a parallel or distributed platform.
-    An efficient execution of the workflows on such platforms requires mapping tasks
+An efficient execution of the workflows on such platforms requires the mapping of tasks
 to specific processors; to increase utilization by reusing finished processors,
@@
 %%% MOTIVATION %%%
-    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
-    and memory sizes. In general, having different memory sizes per CPUs makes it more challenging to compute
-    a schedule that respects all memory constraints -- meaning that no task is executed on a 
-    processor with less memory than needed for the task. This is, however, very important to
-    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
+Modern parallel and in particular distributed computing platforms are often heterogeneous,
+meaning they feature varying CPU speeds and memory sizes. 
+In general, having different memory sizes per CPUs makes it more challenging for an algorithm 
+to compute a schedule that respects all memory constraints -- meaning that no task is executed on a 
+processor with less memory than needed for the task. Violating a memory constraint is, however, 
+very important to avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
 Hence, building on previous related %\hmey{You may want to add other refs not from us} 
 work~\cite{gou2020partitioning,He21,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
-    formulation that takes memory sizes as explicit constraints into account. Its objective is
-    the very common \emph{makespan}~\cite{liu2018survey}, 
+formulation that takes memory sizes as explicit constraints into account. 
+Its objective is the very common \emph{makespan}~\cite{liu2018survey}, 
 which acts as proxy for the total execution time of a workflow.
-However, to the best of our knowledge,
-    the only memory-aware heuristics that would account for memory constraints are  partitioning
-    the DAG and not reusing processors once they have processed a part of the graph, leading
-    to high values of makespan compared to a finer grain solution that reuses processors.
-
-    While previous work with memory constraints has been focusing on partitioning the graph, 
-    and not reusing processors during execution, 
-    a seminal list scheduling heuristic for workflows on heterogeneous platforms, without accounting
-    for the memory constraint, is HEFT 
+However, to the best of our knowledge, the only memory-aware heuristics that would account for 
+memory constraints partition the DAG and do not reuse processors once they have processed a part
+of the graph. This approach leads to high values of makespan compared to a more fine-grained solution
+with processor reuse.
+
+While previous work with memory constraints has focused on partitioning the DAG and not on 
+reusing processors during execution, a seminal list scheduling heuristic for workflows on 
+heterogeneous platforms, without accounting for the memory constraint, is HEFT 
 (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
-    It has two phases: (i) each task is assigned a priority; and (ii) the tasks in a priority-ordered list are assigned
+It has two phases: (i) each task is assigned a priority and (ii) the tasks in a priority-ordered list are assigned
 to processors, where the ``ready'' task with the highest priority is scheduled next on the processor
@@
 for a variety of different scheduling problem formulations. 
-    Yet, none of them adhere to memory constraints as we propose -- see discussion of related work
-    in Section~\ref{sec:related-work}. 
+Yet, none of them adhere to memory constraints as addressed in this paper -- 
+see the discussion of related work in Section~\ref{sec:related-work}. 
 %    (\skug{check in related work if true!}).
 %        \skug{Note: 2 papers that deal with memory sizes, but model is very different!}
-
-    Another limitation in practice of HEFT (and many other scheduling strategies) is their 
+%    
+Another limitation of HEFT (and many other scheduling strategies) in practice is their 
 assumption that the task running times provided to them are accurate. In practice, this is 
 not the case and deviations from user estimates or historical measurements are 
-    very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
-    deviations occur. However, the original list-based schedulers, such as HEFT, are only defined
-    in a static setting with accurate task parameters. 
+very common~\cite{hirales2012multiple}. As a consequence, it is advisable to adapt the schedule when \emph{major}
+deviations occur. However, the original list-based schedulers, such as HEFT, are designed
+for a static setting with accurate task parameters. 
 %    

### 2024-12-22 main.tex
small edit in the beginning of Model section (addresses 2 reviewer´s complaints)

@@
 
-    We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
+    We first describe the model for our target applications, large scientific workflows for which we do not have exact a priori knowledge,
     in Section~\ref{sec.mod.work}.  Next, we define the execution
-    environment, a heterogeneous system (in terms of processor speed and memory size),
+    environment, a cluster that is heterogeneous in terms of processor speed and memory size,
    in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in

### 2024-12-22 main.tex
small change in the beginning of experimental evaluation

@@
 
-    We first describe the set of workflows used for the evaluation, and then the clusters on which the
-    workflows are scheduled.
+    Before presenting the results, we first describe the set of workflows we used for the evaluation, and then the clusters on which the
+    workflows were scheduled.
+We also explain the way the runtime system works and interacts with our scheduler.
 

### 2024-12-22 main.tex
intro edited

@@
 \item We implement a runtime system able to provide some feedback to the scheduler
-when task requirements (in terms of execution time and memory) differ from the initial predictions, 
+when task requirements (in terms of execution time and/or memory) differ from the initial predictions, 
 and we recompute a schedule, based on the reported deviations.
@@
 \item We perform extensive simulations, first in the static case by comparing the schedules produced 
-by these heuristics with the classical HEFT as baseline, which however does not take memory sizes into account; 
+by these heuristics with the classical HEFT as baseline (the latter does not take memory sizes into account); 
 while HEFT returns invalid schedules that exceed the processor memories and cannot execute correctly,
-the new heuristics are able to successfully schedule large workflows, with reasonable makespans.
-
-\item In the dynamic setting, we use a runtime system that allows us to simulate workflow executions,
-introducing deviations in running times and memory requirements of tasks that are communicated
-back to the scheduler; the scheduler can then recompute a schedule. Without these recomputations,
+the new heuristics are able to successfully schedule large workflows and do so with reasonable makespans.
+
+\item In the dynamic setting, we use a runtime system that allows us to simulate workflow executions.
+The runtime system introduces deviations in running times and task memory requirements and communicates
+them to the scheduler; the scheduler can then recompute a schedule. Without these recomputations,
 most schedules become invalid after deviations, since the memory constraint is exceeded 
-for most workflows, hence demonstrating the necessity of a dynamic adjustment of the schedule.
+for most workflows, which demonstrates the necessity of a dynamic adjustment of the schedule.
 
@@
 We first review related work in Section~\ref{sec:related-work}. Then, we formalize the model in Section~\ref{sec:model}
-and the algorithms in Section~\ref{sec:heuristics}.
-The adaptation of the heuristics in a dynamic setting is discussed in Section~\ref{sec:dyn}, and experiment results are presented in Section~\ref{sec:expe}.
+and present our algorithms in Section~\ref{sec:heuristics}.
+The adaptation of the heuristics in a dynamic setting is discussed in Section~\ref{sec:dyn}, and the results of our experiments are presented in Section~\ref{sec:expe}.
 Finally, we conclude

### 2024-12-22 main.tex
editing rel. work section (II-A done)

@@
 Finally, we move to related work on dynamic or adaptive algorithms.
- %   \hmey{Question: Does it make sense to place related work before the model (which is more common)?}
 
@@
 \subsection{Static list schedulers, especially HEFT-based algorithms}
-
+\label{sub:static-list-schedulers}
+%
 Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
-    It and all its successors consist of two phases: task prioritization/ordering and task assignment.
+HEFT and its successors consist of two phases: task prio\-ri\-tization/ordering and task assignment.
 In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
@@
 
-    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
-    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
+During the task prioritization phase in~\cite{sulaiman2021hybrid}, the standard deviation of the computation cost % (between processors) 
+is computed and added to the mean value to account for differences between processor speeds.
+% TODO: Why ``between processors''? And what exactly is ``computation cost'' in the first place?
 In the processor choice phase, the entry task and the longest parent tasks are duplicated 
 during idle times on the processor.
+% TODO: Context of previous sentence unclear. Why is this done? What is the difference to others / our work in this respect?
 
@@
 processor is chosen for this task.
+% TODO: is ``target task'' defined/clear?
 Ranking is based on OCT values.
-    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.
+The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task. % TODO: is ``exit node'' == ``target task''?
 
@@
 In the second stage, the algorithm duplicates the entry task if there is a need for it.
+% TODO: Sounds like Ref. [32] above. What is the difference? (Why) Is it an improvement? Do we need to describe both?
 
@@
 
-    The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
+The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to a task’s
 average execution cost in its weight.
-    The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT) that
-    additionally notes how large the communication between current node and its children will be on the
-    average provided that it is scheduled on the current processor.
+The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT),
+which additionally notes how large the communication between the current node and its children will be on 
+average when scheduled on the current processor.
 

### 2024-12-22 main.tex
editing rel. work section (II-B done)

@@
 \subsection{Memory-aware scheduling algorithms}
-    Respecting processor memories adds a constraint to a scheduling problem.
-    Therefore, only specifically memory-targeted algorithms address this issue.
+\label{sub:mem-aware-algs}
+%
+Only memory-aware scheduling algorithms are designed to respect memory constraints.
+%Respecting processor memories adds a constraint to a scheduling problem.
+%Therefore, only specifically memory-targeted algorithms address this issue.
 Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
@@
 Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
-    A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a DAG is provided,
-    as well as an ILP solution to the scheduling problem.
-    The memory model requires deleting all input data upon starting of the task and adding all output files there.
-
-    In an assumed dual-memory systems~\cite{herrmann2014memory},  a processor can have access 
-    to a memory of two different
-    kinds (red or blue), and each task can be executed on only one sort of memory.
-    The communications happen only between these two kinds of processors (communications within 
-    each group are ignored).
-    The authors then formulate an ILP solution for this problem formulation.
+A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a workflow DAG is provided,
+as well as an integer linear programming (ILP) solution to the scheduling problem.
+The memory model requires the deletion of all input files when starting the task.
+% and adding all output files there.
+% TODO: This sounded strange. I changed ``input data'' to ``input files''.
+% How can I process input data when they are deleted? 
+% TODO: And what is ``there''?
+
+In an assumed dual-memory system~\cite{herrmann2014memory},  a processor can have access 
+to two different kinds of memory, and each task can be executed on only one sort of memory.
+Communication happens only between these two kinds of processors (communication within 
+each group is ignored).
+The authors then present an ILP-based solution for this problem formulation.
 
@@
 
-    In~\cite{ding2024ils}, there are connected processors with individual limited memories.
-    The collective set of memories forms the global memory, to which each processor has access, however the access time
-    to global memory is different.
+In~\cite{ding2024ils}, the authors consider connected processors with individual limited memories.
+The collective set of memories forms the global memory to which each processor has access;
+however, the access time to global memory is different.
 Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
-    The solved problem is how to allocate initial input data in processor memories so that the overall
+The solved problem is how to allocate the initial input data in processor memories so that the overall
 execution is minimized and the memories are not exceeded.
-    The authors propose an integer linear programming model.
+To this end, the authors propose an ILP model.
 %that minimizes the length of the critical path, including a greedy initial solution.
 
-    In~\cite{rodriguez2019exploration}, the authors assume memory requirement on tasks represented as tiles.
+In~\cite{rodriguez2019exploration}, the authors assume memory requirements on tasks represented as tiles.
 Each processor has individual memories to process the task, but only the shared memories store the tiles containing
@@
 Overall, there are a variety of memory models, but, to the best of our knowledge, the only study on a multiprocessor
-   platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24},
-   but where a partition of the workflow is proposed, hence preventing processor reuse. 
-   Hence, in ~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need of communication buffers to store data 
-   that should be communicated
-   between two processors when tasks are ready to execute.
+platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24}.
+Yet, they propose only a partitioning-based mapping of the workflow without processor reuse.
+Hence, in~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need for communication buffers to store data 
+that should be communicated between two processors when tasks are ready to execute. 
+% TODO: Why is the last sentence necessary? More specifically, why is it a consequence of the previous
+% sentence (as implied by ``Hence''?)?
 

### 2024-12-22 main.tex
shortened related work to fit in 10 pages with ref

@@
 
-To address this problem, we investigate in this paper the memory-aware scheduling of DAG-shaped workflows on
+To address this problem, we investigate the memory-aware scheduling of DAG-shaped workflows on
 heterogeneous platforms, where each processor can have a different speed and a different memory size.
@@
 
-First, we focus on HEFT-like scheduling heuristics from the literature  that do not necessarily  
-consider memory constraints. Then, we discuss memory-aware scheduling algorithms. 
-Finally, we move to related work on dynamic or adaptive algorithms.
+%First, we focus on HEFT-like scheduling heuristics from the literature  that do not necessarily  
+%consider memory constraints. Then, we discuss memory-aware scheduling algorithms. 
+%Finally, we move to related work on dynamic or adaptive algorithms.
 
@@
 
-\subsection{Static list schedulers, especially HEFT-based algorithms}
-\label{sub:static-list-schedulers}
-%
-Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
-HEFT and its successors consist of two phases: task prio\-ri\-tization/ordering and task assignment.
+\noindent{\bf HEFT-based algorithms. }
+%\subsection{Static list schedulers, especially HEFT-based algorithms}
+%\label{sub:static-list-schedulers}
+%
+Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic, consisting 
+%HEFT and its successors consist 
+of two phases: task prio\-ri\-tization/ordering and task assignment.
 In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
@@
 
-During the task prioritization phase in~\cite{sulaiman2021hybrid}, the standard deviation of the computation cost % (between processors) 
-is computed and added to the mean value to account for differences between processor speeds.
-% TODO: Why ``between processors''? And what exactly is ``computation cost'' in the first place?
-In the processor choice phase, the entry task and the longest parent tasks are duplicated 
-during idle times on the processor.
-% TODO: Context of previous sentence unclear. Why is this done? What is the difference to others / our work in this respect?
+%During the task prioritization phase in~\cite{sulaiman2021hybrid}, the standard deviation of the computation cost % (between processors) 
+%is computed and added to the mean value to account for differences between processor speeds.
+%% TODO: Why ``between processors''? And what exactly is ``computation cost'' in the first place?
+%In the processor choice phase, the entry task and the longest parent tasks are duplicated 
+%during idle times on the processor.
+%% TODO: Context of previous sentence unclear. Why is this done? What is the difference to others / our work in this respect?
 
@@
 
-PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
-Cost Table (OCT).
-The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
-processor is chosen for this task.
-% TODO: is ``target task'' defined/clear?
-Ranking is based on OCT values.
-The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task. % TODO: is ``exit node'' == ``target task''?
-
-The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
-comparison to HEFT.
-It combines the standard deviation with the communication cost weight on the tasks.
-In the second stage, the algorithm duplicates the entry task if there is a need for it.
-% TODO: Sounds like Ref. [32] above. What is the difference? (Why) Is it an improvement? Do we need to describe both?
-
-The TSHCS (Task Scheduling for Heterogeneous Computing Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
-by adding randomized decisions to the second phase.
-The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
-produces the lowest finish time.
-
+Some variants have been designed with various ways of ordering tasks, for instance based
+on an optimistic cost table in 
+PEFT (Predict earliest finish time)~\cite{arabnejad2014list}, or by combining the standard deviation
+with the communication cost weight on the tasks in HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip}. 
+% is a HEFT variant that computes an Optimistic
+%Cost Table (OCT).
+%The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
+%processor is chosen for this task.
+%% TODO: is ``target task'' defined/clear?
+%Ranking is based on OCT values.
+%The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task. % TODO: is ``exit node'' == ``target task''? 
+%
+%The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
+%comparison to HEFT.
+%It combines the standard deviation with the communication cost weight on the tasks.
+%In the second stage, the algorithm duplicates the entry task if there is a need for it.
+%% TODO: Sounds like Ref. [32] above. What is the difference? (Why) Is it an improvement? Do we need to describe both?
+
+The TSHCS (Task Scheduling for Heterogeneous Computing Systems) algorithm~\cite{alebrahim2017task} 
+improves on HEFT by adding randomized decisions to the second phase.
+%The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
+%produces the lowest finish time.
 The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to a task’s
 average execution cost in its weight.
-The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT),
-which additionally notes how large the communication between the current node and its children will be on 
-average when scheduled on the current processor.
-
-HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement learning techniques~\cite{yano2022cqga}.
-
-\subsection{Memory-aware scheduling algorithms}
-\label{sub:mem-aware-algs}
-%
-Only memory-aware scheduling algorithms are designed to respect memory constraints.
-%Respecting processor memories adds a constraint to a scheduling problem.
-%Therefore, only specifically memory-targeted algorithms address this issue.
-Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
+%The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT),
+%which additionally notes how large the communication between the current node and its children will be on 
+%average when scheduled on the current processor.
+HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement learning techniques~\cite{yano2022cqga}, but none of these variants of HEFT consider memory constraints,
+to the best of our knowledge.
+
+\medskip
+%\subsection
+\noindent{\bf Memory-aware scheduling algorithms. }
+%\label{sub:mem-aware-algs}
+%
+%Only memory-aware scheduling algorithms are designed to respect memory constraints.
+%%Respecting processor memories adds a constraint to a scheduling problem.
+%%Therefore, only specifically memory-targeted algorithms address this issue.
+%Moreover, 
+The way processor memories are represented in the model has a decisive impact on the way the constraint
 is formulated and addressed in the algorithm.
@@
 
-In an assumed dual-memory system~\cite{herrmann2014memory},  a processor can have access 
+Other models consider for instance a 
+%In an assumed 
+dual-memory system~\cite{herrmann2014memory} where a processor can have access 
 to two different kinds of memory, and each task can be executed on only one sort of memory.
-Communication happens only between these two kinds of processors (communication within 
-each group is ignored).
+%Communication happens only between these two kinds of processors (communication within 
+%each group is ignored).
 The authors then present an ILP-based solution for this problem formulation.
-
+%
 %The algorithm presented by Yao et al.
-Yao et al.~\cite{yao2022memory} consider that each processor has an own internal memory and all
+Yao et al.~\cite{yao2022memory} consider that each processor has its own internal memory, and all
 processors share a common external one. The internal (local) memory is used to store the task files.
 The external memory is used to store evicted files to make room for the execution of a task on a processor.
-All processors, including the original one, can access these files. 
-Each edge %in~\cite{yao2022memory} 
-has two weights -- the size of the files transferred along it,
-and the time of communication along this edge.
-The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
-
-In~\cite{ding2024ils}, the authors consider connected processors with individual limited memories.
-The collective set of memories forms the global memory to which each processor has access;
-however, the access time to global memory is different.
-Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
-The solved problem is how to allocate the initial input data in processor memories so that the overall
-execution is minimized and the memories are not exceeded.
-To this end, the authors propose an ILP model.
-%that minimizes the length of the critical path, including a greedy initial solution.
-
-In~\cite{rodriguez2019exploration}, the authors assume memory requirements on tasks represented as tiles.
-Each processor has individual memories to process the task, but only the shared memories store the tiles containing
-memory tiles occupied by memory tiles.
-
-Finally, there are some cloud-oriented models that include costs associated with memory usage~\cite{liang2020memory}.
+%All processors, including the original one, can access these files. 
+%Each edge %in~\cite{yao2022memory} 
+%has two weights -- the size of the files transferred along it,
+%and the time of communication along this edge.
+%The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
+%
+In~\cite{ding2024ils}, the authors consider connected processors with individual limited memories
+forming a global memory, with different access times to memory, and no weights on edges. An ILP
+is proposed in this setting. 
+%The collective set of memories forms the global memory to which each processor has access;
+%however, the access time to global memory is different.
+%Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
+%The solved problem is how to allocate the initial input data in processor memories so that the overall
+%execution is minimized and the memories are not exceeded.
+%To this end, the authors propose an ILP model.
+%%that minimizes the length of the critical path, including a greedy initial solution.
+%
+%In~\cite{rodriguez2019exploration}, the authors assume memory requirements on tasks represented as tiles.
+%Each processor has individual memories to process the task, but only the shared memories store the tiles containing
+%memory tiles occupied by memory tiles.
+%
+There are also some cloud-oriented models that include costs associated with memory usage~\cite{liang2020memory}.
 
@@
 platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24}.
-Yet, they propose only a partitioning-based mapping of the workflow without processor reuse.
+Yet, they propose only a partitioning-based mapping of the workflow, without processor reuse.
 Hence, in~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need for communication buffers to store data 
@@
 
-    \subsection{Dynamic/adaptive algorithms}
-
-We now review related work in a dynamic setting. With no variation in task parameters, 
+%    \subsection
+\smallskip
+\noindent{\bf Dynamic/adaptive algorithms. }
+We finally review related work in a dynamic setting. With no variation in task parameters, 
     DVR HEFT~\cite{SANDOKJI2019482} rather considers that new tasks arrive in the system. 
@@
     varying variants of task weighting and choosing the variant that gives the best overall makespan.
-    In the dynamic phase, they receive new tasks and schedule them on either idle processors or 
-    those processors that give them
-    the earliest finish time.
-    %Task failures are not covered.
-
-    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
-    by calculating the critical path in the graph dynamically at every step.
+%    In the dynamic phase, they receive new tasks and schedule them on either idle processors or 
+%    those processors that give them
+%    the earliest finish time.
+%    %Task failures are not covered.
+%
+%    Rahman~\etal 
+    The dynamic critical path (DCP) algorithm for grids maps tasks to machines
+    by calculating the critical path in the graph dynamically at every step~\cite{rahman2013}.
     %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
@@
     The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
-    The approach is evaluated on random workflows of the size up to 300 tasks.
+%    The approach is evaluated on random workflows of the size up to 300 tasks.
 
     Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
-    The procedure involves building a first (static) schedule, periodic resource monitoring and rescheduling the remaining
-    tasks.
-    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
-    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
-    tasks to groups.
-    The static heuristic is HEFT with earliest start time as priority.
+    The procedure involves building a first (static) schedule with HEFT, periodic resource monitoring, 
+    and rescheduling the remaining tasks. 
+%    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
+%    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
+%    tasks to groups.
+%    The static heuristic is HEFT with earliest start time as priority.
     Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
     is smaller than the previous one.
-    The experiments were conducted on a single workflow with 10 tasks.
+%    The experiments were conducted on a single workflow with 10 tasks.
 %
@@
     %is added.
-
-    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
-    but constrained by cost.
-
-    Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
-    Particles are possible solution in the solution space.
-    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
-    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
+%
+%
+%    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
+%    but constrained by cost.
+%
+% In a cloud framework,  Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm.
+% % to schedule workflows in a cloud.
+%%    Particles are possible solution in the solution space.
+%    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
+%    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
 
@@
     They solve a set of linear equations that represent the cost of an execution based on the criteria.
-    The authors test out 4 scenarios - one preferring each criteria, and a balanced one.
+    The authors test out four scenarios - one preferring each criteria, and a balanced one.
     The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
-    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
+    The authors used workflows with less than ten tasks, but repeated them so that the execution had up to 200 tasks.
     %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
@@
 
-\subsection{Dynamic experiments}
+\subsection{Dynamic experiments on the memory-constrained cluster}
 \label{sec.expe.dyn}
 
-\AB{Need to add results on the default cluster!}
-
-    \paragraph{Memory-constrained cluster}
+%\AB{Need to add results on the default cluster!}
+%
+%    \paragraph{Memory-constrained cluster}
     The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size

### 2024-12-22 main.tex
Pass on section 3, lots of rephrasing and fixing table

@@
 \newcommand{\criticalpath}{\mathcal{P}}
-\newcommand{\parents}[1]{\,\Pi_{#1}}
-\newcommand{\children}[1]{\,C_{#1}}
+\newcommand{\parents}[1]{\,parent_{#1}}
+\newcommand{\children}[1]{\,child_{#1}}
 \newcommand{\cluster}{\,\mathcal{S}}
@@
 
-    We first describe the model for our target applications, large scientific workflows for which we do not have exact a priori knowledge,
-    in Section~\ref{sec.mod.work}.  Next, we define the execution
-    environment, a cluster that is heterogeneous in terms of processor speed and memory size,
-   in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
-    Section~\ref{sec.mod.pb}. The key notation is summarized in Table~\ref{tabnotation}.
-
-    \subsection{Workflow}
+The target applications, large scientific workflows for which we do not have exact a priori knowledge,
+are described in Section~\ref{sec.mod.work}. Then, the heterogeneous system on which
+the applications are to be executed is presented in Section~\ref{sec.mod.plat}. The optimization problem
+is defined in 
+%
+%    We first describe the model for our target applications, large scientific workflows for which we do not have exact a priori knowledge,
+%    in Section~\ref{sec.mod.work}.  Next, we define the execution
+%    environment, a cluster that is heterogeneous in terms of processor speed and memory size,
+%   in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
+    Section~\ref{sec.mod.pb}, and the key notation is summarized in Table~\ref{tabnotation}.
+
+    \subsection{Applications}
     \label{sec.mod.work}
-    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
-    $E$ is a set of directed edges of the form $e=(u,v)$, with $u,v\in V$, expressing precedence constraints between tasks.
-    Each task~$u \in V$  is performing $w_u$ operations, and it also
-    requires some amount of memory to be executed, denoted as~$m_u$.
-    Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.
+    The target applications, corresponding to large scientific workflows, are represented with a DAG (Directed
+    Acyclic Graph) $G=(V,E)$. The set of vertices~$V$ corresponds to tasks, while edges express the
+    precedence constraints. Hence, a directed edge $e=(u,v)\in E$ means that task $u\in V$ must be executed 
+    before task~$v\in V$. A cost~$c_{u,v}$ is associated to each edge, representing the size of the 
+    output of task~$u$, to be used by task~$v$. 
+    Furthermore, $w_u$ is the number of operations performed by task~$u\in V$, 
+    and $m_u$ is the amount of memory required by task~$u$ to be executed. 
+%    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
+%    $E$ is a set of directed edges of the form $e=(u,v)$, with $u,v\in V$, expressing precedence constraints between tasks.
+%    Each task~$u \in V$  is performing $w_u$ operations, and it also
+%    requires some amount of memory to be executed, denoted as~$m_u$.
+%    For an edge $(u,v) \in E$, the weight~$c_{u,v}$  corresponds to the size of the output file 
+%    written by task~$u$ and used as input by task~$v$.
    Note that $m_u$ is the total memory usage
@@
 
-    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
-    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without parents is called a {\it source task}.
-    The children tasks of~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
-    $ \children{u} = \{v \in V: (u,v) \in E\}$. A task without children is called a {\it target task}.
-    Each task may have multiple parents and children.
+    We denote by $\parents{u}$ the tasks preceding task~$u\in V$, that must be completed before $u$ can be started:   
+%    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
+    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A {\em source} task is a task without parents. % is called a {\it source task}.
+ Similarly, the children of task $u\in V$ are %   The children tasks of~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
+    $ \children{u} = \{v \in V: (u,v) \in E\}$, and a {\em target} task has no children. 
+    %A task without children is called a {\it target task}.
+    Note that each task may have multiple parents and children.
 
@@
 
-    \subsection{Execution environment}
+    \subsection{Heterogeneous system}
     \label{sec.mod.plat}
 
-    The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
-    consists of $k$ processors $p_1, \dots, p_k$.
-    Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$, a communication
+The target platform is a heterogeneous system $\cluster$ with $k$ processors 
+%    The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
+%    consists of $k$ processors 
+    $p_1, \dots, p_k$.
+    For $1 \leq j \leq k$, each processor $p_j$  has an individual memory of size $M_j$, a communication
     buffer of size $MC_j$ and a speed~$s_j$.
@@
     to another processor; it then stays in the communication buffer until it has been sent.
-    The execution time of a single task~$u\in V$ on a processor~$p_j$ is expressed as $\frac{w_u}{s_j}$.
-    We assume that all processors are connected with the same bandwidth~$\beta$.
+    The execution time of a single task~$u\in V$ on a processor~$p_j$ is~$\frac{w_u}{s_j}$,
+    and all 
+    %We assume that all 
+    processors are connected with an identical bandwidth~$\beta$.
 %    \hmey{Maybe mention that variable bandwidths are part of future work...?}
@@
 
-    We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
-    the memory requirement; it transforms the workflow into a series-parallel graph
+In oder to compute the memory requirement of a DAG, we use \algo{memDag}~\cite{KAYAASLAN20181},
+an algorithm that 
+%    We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
+%    the memory requirement; it 
+    transforms the workflow into a series-parallel graph
     and then finds the traversal that leads to the minimum memory consumption.
@@
                 \hline
-                $G = (V, E)$                          & Workflow graph, set of vertices (tasks) and edges        \\
-                $\parents{u}$, $\children{u}$         & Parents of a task $u$, children of a task $u$            \\
-                $m_u$                                 & Memory weight of task $u$                                \\
-                $w_u$                                 & Workload of task $u$  (normalized execution time)          \\
-                $c_{u,v}$                             & Communication volume along the edge $(u,v)\in E$         \\
-                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
-                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
-                $\cluster$, $k$                    & Computing system and its number of processors           \\
-                $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
+                $G = (V, E)$                          & Application DAG (tasks and edges)       \\
+ %               $\parents{u}$, $\children{u}$         & Parents of task $u$, children of task $u$            \\
+                $m_u$                                 & Amount of memory required by task $u$                                \\
+                $w_u$                                 & Number of operations of task $u$         \\
+                $c_{u,v}$                             & Size of output file for edge $(u,v)\in E$         \\
+%                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
+%                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
+                $\cluster$, $k$                    & Computing platform, total number of processors           \\
+%                $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
                 $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer size, and speed of proc.\ $p_j$                          \\
                 $\beta$                     & Bandwidth in the compute system                                \\
-                $\bottomlevel{u}$                      & Bottom weight of task $u$ \\
-                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
-                $\Gamma = (\mathcal{V}, \mathcal{E})$                      & Quotient graph, its vertices and its edges        \\
-                $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
+                $bl{u}$                      & Bottom level of task $u$ \\
+%                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
+%                $\Gamma = (\mathcal{V}, \mathcal{E})$                      & Quotient graph, its vertices and its edges        \\
+%                $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
                 \hline

### 2024-12-23 main.tex
minor edits in Sec. II-C

@@
     %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
-    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
+    The authors schedule the first task on the critical path to the best suitable processor and recompute the critical path.
     %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
@@
     They solve a set of linear equations that represent the cost of an execution based on the criteria.
-    The authors test out four scenarios - one preferring each criteria, and a balanced one.
+    The authors test four scenarios -- one preferring each criterion and a balanced one.
     The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
@@
 
-    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
+    Daniels \etal~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
     on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
-    that these runtime changes are unrelated to each other).
-    The authors formulate a decision space of all permutations of $n$ jobs, and the optimal schedule in relation to a
-    performance measure $\phi$.
+    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (and thus unrelated to each other).
+    The authors search for an optimal schedule
+%    formulate a decision space of all permutations of $n$ jobs, and the optimal schedule 
+    in relation to a performance measure. % $\phi$.
     Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
 
-While several related work consider building a new schedule once some variation has been observed,
-we are not aware of work implementing a real runtime system that interacts with the scheduler,
-and tested on workflows with thousands of tasks, as we propose in this paper. Furthermore, 
+While several related works consider building a new schedule once some variation has been observed,
+we are not aware of work implementing a real runtime system that interacts with the scheduler
+and has been tested on workflows with thousands of tasks, as we propose in this paper. Furthermore, 
 we are not aware of any previous work discussing dynamic algorithms combined with memory constraints. 

### 2024-12-23 main.tex
minor edits in Sec. III-A

@@
 \label{sec:model}
-
-The target applications, large scientific workflows for which we do not have exact a priori knowledge,
-are described in Section~\ref{sec.mod.work}. Then, the heterogeneous system on which
+%
+The applications we target, large scientific workflows for which we do not have exact a priori knowledge,
+are described in Section~\ref{sec.mod.work}. Then, the type of heterogeneous system on which
 the applications are to be executed is presented in Section~\ref{sec.mod.plat}. The optimization problem
-is defined in 
-%
-%    We first describe the model for our target applications, large scientific workflows for which we do not have exact a priori knowledge,
-%    in Section~\ref{sec.mod.work}.  Next, we define the execution
-%    environment, a cluster that is heterogeneous in terms of processor speed and memory size,
-%   in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
-    Section~\ref{sec.mod.pb}, and the key notation is summarized in Table~\ref{tabnotation}.
-
-    \subsection{Applications}
+is defined in Section~\ref{sec.mod.pb}, and the key notation is summarized in Table~\ref{tabnotation}.
+
+\subsection{Applications: Large Scientific Workflows}
 \label{sec.mod.work}
-    The target applications, corresponding to large scientific workflows, are represented with a DAG (Directed
-    Acyclic Graph) $G=(V,E)$. The set of vertices~$V$ corresponds to tasks, while edges express the
+%
+Following common practice, we represent a workflow by a DAG (Directed Acyclic Graph) 
+% The target applications, corresponding to large scientific workflows, are represented with a DAG (Directed Acyclic Graph) 
+$G=(V,E)$. The set of vertices~$V$ corresponds to tasks, while edges express the
 precedence constraints. Hence, a directed edge $e=(u,v)\in E$ means that task $u\in V$ must be executed 
-    before task~$v\in V$. A cost~$c_{u,v}$ is associated to each edge, representing the size of the 
+before task~$v\in V$. A cost~$c_{u,v}$ is associated with each edge, representing the size of the 
 output of task~$u$, to be used by task~$v$. 
@@
 %    written by task~$u$ and used as input by task~$v$.
-   Note that $m_u$ is the total memory usage
-    of a task during its execution, including input and output files currently being read and written,
-    and hence the total memory requirement  for the execution of task~$u$ consists of the maximum
-    between the input files
-    (total size of the files to be received from the parents),
-    the output files (total size of the files to be sent to the children),
-    and the total memory size~$m_u$ (usually achieving the maximum):
-    \[
-        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
-    \]
-
-    We denote by $\parents{u}$ the tasks preceding task~$u\in V$, that must be completed before $u$ can be started:   
+We denote by $\parents{u}$ the tasks preceding task~$u\in V$, which must be completed before $u$ can be started:   
 %    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
@@
 %A task without children is called a {\it target task}.
-    Note that each task may have multiple parents and children.
-
-Furthermore, we place ourselves in a context where we do not have perfect knowledge
+Each task may have multiple parents and children.
+
+Note that $m_u$ is the total memory usage
+of a task during its execution, including input and output files currently being read and written,
+and hence the total memory requirement for executing task~$u$ is the maximum of the following:
+(i) the total size of the files to be received from the parents, (ii) the total size of the files
+to be sent to the children, and (iii) the total memory size~$m_u$ (which often reaches the maximum):
+\[
+    r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
+\]
+
+Furthermore, we operate in a context where we do not have perfect knowledge
 of the task parameters ($w_u$ and $m_u$) before the tasks start their execution,
@@
 \label{sec.mod.plat}
-
+%
 The target platform is a heterogeneous system $\cluster$ with $k$ processors 
@@
         \end{center}
-        \caption{Notation} \label{tabnotation}
+        \caption{Key Notation} \label{tabnotation}
     \end{table}

### 2024-12-23 main.tex
minor edits in Sec. III-B to IV-A, indentation in IV-B

@@
 
-\subsection{Applications: Large Scientific Workflows}
+\subsection{Applications: Large scientific workflows}
 \label{sec.mod.work}
@@
     buffer of size $MC_j$ and a speed~$s_j$.
-    We can decide to evict some data from the main memory if we are sending the data
+    We can decide to evict some data from the main memory when we are sending the data
     to another processor; it then stays in the communication buffer until it has been sent.
@@
 Since tasks are subject to variability, we aim at minimizing the actual makespan
-achieved at the end of the execution, while decisions may be taken building
-on the estimated task parameters.
+achieved at the end of the execution, while decisions may be taken \wrt 
+the estimated task parameters.
 
@@
 
-\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
+\section{Scheduling heuristics}
 \label{sec:heuristics}
-
-%    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
-%    but favor processor reuse as is done in HEFT.
-%    Furthermore, this would allow us to handle variability on the fly, by updating
-%    the bottom levels if some parameters vary, and computing the schedule
-%    only for the near future...
-
-%In order to be able to easily adapt to variability of task parameters, 
-We design 
-variants of HEFT that account for memory usage and aim at minimizing the makespan.
-First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic that does not account for the memory
-(and hence, may return invalid schedules that will not be able to run on the platform
-by running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
+%
+We design variants of HEFT that account for memory usage and aim at minimizing the makespan.
+First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic; as it does not account for the memory,
+it may return invalid schedules that will not be able to run successfully on the platform (due to 
+running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
 heuristics, including eviction strategies to move some data in communication buffers
@@
 \label{sec.heft}
-
-    Original HEFT does not consider memory sizes.
-    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
-    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.
+%
+Original HEFT does not account for memory sizes.
+Its schedules can be invalid if tasks are assigned to processors without enough memory.
+These solutions can be viewed, however, as a ``lower bound'' for a valid solution that 
+respects memory constraints.
 
@@
 The bottom level of a task is defined as
-    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
-    (the max is 0 if there is no outgoing edge).
+\[
+bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}
+\]
+(with the max yielding $0$ if there is no outgoing edge).
 The tasks are sorted by non-decreasing ranks.
@@
 We tentatively assign each task~$v$ to each processor~$j$.
-    The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$,
+The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$
 and all communications that
@@
 \label{sec.heftm}
+%
 Like the original HEFT, the memory-aware versions of HEFT consist of two stages:
@@
 \begin{itemize}
-\item    HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
+\item    
+  HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
   level is defined as
@@
 
-\item    HEFTM-BLC %: from the study of the fork (see below), it seems important
+\item    
+  HEFTM-BLC %: from the study of the fork (see below), it seems important
   %  to also account for the size of the data as input of a task,
@@
 
-\item   Finally, HEFTM-MM orders tasks  in the order returned by %as dictated by MinMem.
+\item   
+  Finally, HEFTM-MM orders tasks  in the order returned by %as dictated by MinMem.
   the  \algo{memDag} algorithm~\cite{KAYAASLAN20181}, which corresponds to a traversal
   of the graph that minimizes the peak memory usage.
-
 \end{itemize}

### 2024-12-23 main.tex
editing IV-B, enlarged some of the figures in the experiments section (still at 10 pages)

@@
 
-We design three variants of memory-aware HEFT:
+Our three variants of memory-aware HEFT work as follows:
 \begin{itemize}
@@
   $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
-  (the max is 0 if there is no outgoing edge).
+  (max yields $0$ if there is no outgoing edge).
 
@@
   %  to also account for the size of the data as input of a task,
-  is  giving more priority at tasks with potential large incoming communications,
+  gives more priority to tasks with potential large incoming communications,
   hence aiming at clearing the memory used by files as soon as possible,
@@
   the  \algo{memDag} algorithm~\cite{KAYAASLAN20181}, which corresponds to a traversal
-  of the graph that minimizes the peak memory usage.
+  of the graph that minimizes peak memory usage.
 \end{itemize}
@@
 and keeping the most promising one. We first detail how a task
-    is tentatively assigned on a processor, by carefully accounting for the memory usage.
-    Next, we explain the steps to be taken to effectively assign a task on a given processor.
+is tentatively assigned to a processor, by carefully accounting for the memory usage.
+Next, we explain the steps to be taken to effectively assign a task to a given processor.
 
@@
         \centering
-        \includegraphics[width=0.9\columnwidth] {images/success-rates-large}
+        \includegraphics[width=0.99\columnwidth] {images/success-rates-large}
         \caption{Success rates on the default cluster. Higher is better.}
@@
         \centering
-        \includegraphics[width=0.8\columnwidth] {images/UpdatesMss2}
+        \includegraphics[width=0.95\columnwidth] {images/UpdatesMss2}
         \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm. Smaller is better.}

### 2024-12-23 main.tex
editing IV-B, finished Step 2

@@
 
-$T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
+$T(p_j)$ is the set of tasks already scheduled on $p_j$; by Step 1, their files are
 already in the memory of~$p_j$. However, the files from the
-other predecessor tasks must be loaded in memory before executing task~$v$,
-as well as $m_v$ and the data generated for all successor tasks.
-$Res$ is then checking whether there was enough memory; if it is negative,
-it means that we have exceeded the memory of~$p_j$ with this tentative
-assignment.
-
-In this case ($Res <0$), we try evicting
-some data from memory so that we have enough memory to execute task~$v$.
-We need to evict at least $Res$ data.
-We propose a greedy approach, evicting the largest files of $\PD_j$ until $Res$ data has been evicted. 
-A variant where the smallest files are evicted first has been tested, and it led to comparable results. 
+other predecessor tasks must be loaded into memory before executing task~$v$,
+as well as the task data of size $m_v$, and the data generated for all successor tasks.
+$Res$ then checks whether there is enough memory; if it is negative,
+we have exceeded the memory of~$p_j$ with this tentative assignment.
+%
+In this case ($Res <0$), we try to evict some data from memory so that there is enough memory to execute task~$v$.
+Clearly, we need to evict at least $Res$ data.
+To this end, we propose a greedy approach, which evicts the largest files of $\PD_j$ until data of size $Res$ have been evicted. 
+A variant where the smallest files are evicted first has been tested; it led to comparable results. 
 %    
@@
 of memories pending in the communication buffer.
-We keep track of the available buffer size, too -- each time a file gets moved into the pending in buffer, 
-the available buffer size is reduced by its weight.
-
-If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
-or if while doing so we exceeded the size of the available buffer,
-we set the finish time to~$+\infty$ (invalid choice).
+We keep track of the available buffer size, too -- every time a file is moved into the pending buffer, 
+the available buffer size is reduced by the file size.
+
+If, after tentatively evicting all files from $\PD_j$, we still do not have enough memory, or if we exceed the size of the available buffer during this process, we set the finish time to $+\infty$ (indicating an invalid choice).
 

### 2024-12-23 main.tex
pagestyle plain
Sec. IV edited

@@
 \renewcommand{\iec}{i.e., }
+
+\pagestyle{plain}
+
 \begin{document}
@@
 \noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
-Its starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
+Its starting time $st_v$ is dictated by the maximum between $rt_j$ and all communications that
 must be orchestrated from predecessor tasks $u\notin T(p_j)$.
@@
 Once we have computed all finish times for task~$v$,
-    we keep the minimum $FT(v,p_j)$ and assign task~$v$
+we keep the minimum $FT(v,p_j)$ and assign~$v$
 to processor~$p_j$. 
-    In detail, we:
+In detail:
 \begin{itemize}
-        \item  Evict the file memories that correspond to edge weights that need to be evicted to free the memory.
+\item 
+  We evict the files corresponding to edge weights that need to be evicted to free the memory.
   We remove these files from pending memories
   $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
-        \item    Calculate the new $availM_j$ on the processor.
-        We subtract the weights of all incoming files from predecessors assigned to the same processor,
+\item 
+  We calculate the new $availM_j$ on the processor.
+  Next, we subtract the weights of all incoming files from predecessors assigned to the same processor
   and add the weights of outgoing files generated by the currently assigned task.
-        \item  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
-        communication buffers $rt_{j', j}$ for the processor~$j'$that the predecessor $u$ has been assigned to: we increase them by the
+
+\item  
+  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
+  communication buffers $rt_{j', j}$ for the processor~$j'$ that the predecessor $u$ has been assigned to: we increase them by the
   communication time $c( u,v) / \beta$.
   We also remove the incoming files from either the pending memories or pending data in buffers of these other
-        processors, and increase the available memories or available buffer sizes on these processors.
-        \item We compute the correct amount of available memory for~$p_j$ (for when the task is done).
+  processors, and increase the available memory or buffer sizes on these processors.
+
+\item 
+  We compute the correct amount of available memory for~$p_j$ (for when the task is done).
   Then, for each predecessor that is mapped to the same processor, 

### 2024-12-23 main.tex
Sec. V edited

@@
 \label{sec:dyn}
-
+%
 In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
-    This information may include memory usage, runtime, graph structures, or the status of the underlying infrastructure.
+This information may include memory usage, running time, graph structures, or the status of the underlying infrastructure.
 In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
@@
 Examples are processor failures, node recoveries, or acquisition of new nodes.
-    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
-    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.
+Even if the hardware infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
+As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during workflow runtime, the previous schedule may become invalid, so that a new one must be calculated.
 
 For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures
-    of approximately 15\% is shown~\cite{malik2013execution}.
-    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
+of approximately 15\% has been observed~\cite{malik2013execution}.
+Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
 %For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
@@
 % @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
-    Such a dynamic execution environment requires a dynamic scheduling method where the schedule can be recomputed during the workflow execution.
+Such a dynamic execution environment requires a dynamic scheduling method where the schedule can be recomputed during workflow execution.
 
@@
 
-    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
+First, we find out if at least one processor that had assigned tasks has terminated operation -- this instantly invalidates the
 entire schedule.
-
-    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
-    is a topological ordering.
+%
+We then iterate over all tasks of the workflow in a topological order -- 
+any of the orderings given by rankings BL, BLC or MM is a topological ordering.
 We then repeat steps similar to those we did during tentative assignment in the heuristics, 
@@
 from other tasks).
-    If originally, $Res$ was positive (no files were evicted from memory into the communication buffer), 
-    then it has to stay this way --
-    otherwise, evicted files can invalidate next tasks.
-    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
-    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
+If $Res$ was originally positive (no files evicted from memory into the communication buffer), 
+then it has to stay this way -- otherwise, evicted files can invalidate subsequent tasks.
+If $Res$ was originally negative, then we need to make sure that evicted files still fit into the communication buffer.
+If either $Res$ is newly negative or the communication buffer is not large enough,
+then this invalidates the schedule.
 We update the $availM_j$ and $availMC_j$ values according to the new memory constraints.

### 2024-12-23 main.tex
Sec. V-A-1-a edited

@@
 
-\section{Experimental evaluation} %: \skug{Full: 3, Polished: 3}}
+\section{Experimental evaluation}
 \label{sec:expe}
-
+%
 We first describe the experimental setup in Section~\ref{sec:setup}. 
@@
 of the memory-aware heuristics in Section~\ref{sec.expe.static}, before
-discussing the heuristics behavior in a dynamic setting in Section~\ref{sec.expe.dyn}. 
+discussing the heuristics' behavior in a dynamic setting in Section~\ref{sec.expe.dyn}. 
 Finally, we report running times of the heuristics in Section~\ref{sec.expe.t}.
 
-\subsection{Experimental setup} %: \skug{Full: 5, Polished: 4}}
+\subsection{Experimental setup}
 \label{sec:setup}
-
+%
 All algorithms are implemented in C++ and compiled with g++ (v.8.5.0).
@@
 and CentOS 8 as OS.
-    Code, input data, and experiment scripts are available to allow reproducibility of results, at~\url{https://zenodo.org/records/13919214}~and~\url{https://zenodo.org/records/13919302}.
-
-    Before presenting the results, we first describe the set of workflows we used for the evaluation, and then the clusters on which the
-    workflows were scheduled.
+Code, input data, and experiment scripts are available to allow reproducibility of the results at~\url{https://zenodo.org/records/13919214}~and~\url{https://zenodo.org/records/13919302}.
+
+Before presenting the results, we first describe the set of workflows we used for the evaluation and
+then the clusters on which the workflows were scheduled.
 We also explain the way the runtime system works and interacts with our scheduler.
@@
 \subsubsection{Workflow instances}
-  %  \skug{Todo: describe the workflows the same asin the abstract}
-    We run  experiments on a dataset that consists of real-world workflows: 
-    we use workflows coming from~\cite{lotaru} (atacseq, bacass, chipseq,
-    eager and methylseq), and also we increase their size using the WFGen generator~\cite{COLEMAN202216}.
+We run experiments on a set of several real-world workflows from Ref.~\cite{lotaru}
+(atacseq, bacass, chipseq, eager, and methylseq). We use the WFGen generator~\cite{COLEMAN202216} to
+create larger variants of the workflows above and add the instances created this way to the set.
 
@@
 the desired number of tasks.
-    We used the real-world workflows as models, except for bacass that lead to errors in the generator.
+We used the real-world workflows as models, except for bacass since it leads to errors in the generator.
 %
 As number of tasks, we use: 200, \numprint{1000}, \numprint{2000}, \numprint{4000}, \numprint{8000}, \numprint{10000},
-    \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
+\numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, and \numprint{30000}.
 We divide the workflows into four groups by size: tiny ones with up to \numprint{200} tasks, small ones with \numprint{1000} to \numprint{8000} tasks,

### 2024-12-23 main.tex
Sec. V-A-2 edited

@@
 to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
-    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
+In a similar manner, the historical data provided by Ref.~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
 size of files that the task sends to all its children.
@@
 For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
-    The same workflow can require different memory capacity and take different time to execute
+The same workflow can require different memory capacities and take different times to execute
 depending on the size of its input.
 We simulate these various runs by obtaining values corresponding to each input size.
-    For each of the four families, there are five input sizes, so we run each workflow in five variants corresponding to these inputs
+For each of the four families, there are five input sizes; we thus run each workflow in five variants corresponding to these inputs.
 
@@
 In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
-    For two more, around 40\% of tasks have no historical runtime data stored.
-    Hence, in the absence of historical data about a task, we give it fixed weights.
-    We give it an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
-    These values align with the findings of~\cite{lotaru} about small tasks.
+For two more, around 40\% of the tasks have no historical runtime data stored.
+Hence, in the absence of historical data about a task, we give it fixed weights:
+an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
+These values align with the findings of Ref.~\cite{lotaru} about small tasks.
 %
@@
 that we consider is a cluster based on the same six
-    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
-    We set the number of each kind of node to $12$, thus having 72 processors in total.  % the whole cluster.
-%
-    Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
+kinds of real-world machines that were used in the experimental evaluation in Ref.~\cite{lotaru}.
+We set the number of each kind of node to $12$, yielding 72 processors in total.  % the whole cluster.
+%
+Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as tuples (name, speed, memory):
 ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
@@
     \end{center}
-        \caption{Cluster configuration.}
+    \caption{Cluster configurations.}
     \label{tab:procs}
@@
 
-    We also consider a more constrained setting, by varying the cluster configuration. The
-     {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
+We also consider a more constrained setting by varying the cluster configuration. 
+The {\em memory-constrained cluster} consists of 72 nodes (12 of each kind) as the default cluster,
 but each node has 10 times less memory. Hence,  

### 2024-12-23 main.tex
done editing Sec. V-A

@@
 \subsubsection{Runtime system}
-
+\label{ss:runtime-sys}
+%
 To simulate the execution of a workflow, we implemented a runtime system.
-    It reads the historical data and builds weights for tasks, as explained above.
+It reads the historical data and builds weights for tasks as explained above.
 %    These weights are then treated as the ``truth'' about the execution during the simulation.
@@
 according to these weights.
-    However, in the dynamic setting, the runtime system applies a deviation function to the values.
-    The deviation function computes a normally distributed deviated value, where the initial value
+In the dynamic setting, in turn, the runtime system applies a deviation function to the values.
+This function computes a normally distributed random deviation value, where the initial value
 is the mean and the deviation is~$10\%$.
-    This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and
-    other works dedicated to predicting runtimes of tasks~\cite{da2015online,da2013toward}.
-
-    Hence, the scheduler receives deviated values and makes decision based on them.
+This scenario corresponds to the real-life scenarios identified in Ref.~\cite{lotaru} and
+other works dedicated to predicting task running times~\cite{da2015online,da2013toward}.
+
+Thus, the scheduler receives deviation values and makes decision based on them.
 This leads to several types of possible issues:
+%
 \begin{itemize}
-        \item A processor is blocked by another task. If the scheduler underestimated the execution time of a task,
+\item 
+  A processor is blocked by another task. If the scheduler underestimated the execution time of a task,
   it will block another one from starting.
-        \item A predecessor has not yet finished.
+\item 
+  A predecessor has not finished yet.
   The scheduler may request a task to start on its processor, while
-        some of the predecessors of the tasks have in fact not yet completed their execution, and the
-        task is therefore not yet ready.
+  some of the predecessors of the task have in fact not yet completed their execution -- 
+  the task is therefore not yet ready.
 %        A similar situation, but a task could start on its processor, but one or more of its parents a
 %        are not ready yet.
-        \item Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able to execute on a chosen processor.
-        \item A task took less time than expected. We only consider this case if a task took more than 10\% less
-        time than expected. In this
-        case, we want to exploit the newly acquired free time by possibly starting other tasks earlier.
+\item 
+  Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able to execute on a chosen processor.
+\item 
+  A task took less time than expected. We only consider this case if a task took more than 10\% less
+  time than expected. In this case, we want to exploit the newly acquired free time by possibly starting other tasks earlier.
 \end{itemize}

### 2024-12-23 main.tex
done editing Sec. V-B

@@
 
-\subsection{Results in a static setting} % \skug{Full: 3, Polished: 2}}
+\subsection{Results in a static setting}
 \label{sec.expe.static}
-
-   % \subsubsection{Preliminary experiments}
-We first study the heuristics behavior when the weights do not change upon runtime,
-hence the scheduler has a perfect knowledge of task memory requirements and execution times.
+%
+We first study the heuristics' behavior when the weights do not change during runtime,
+hence the scheduler has perfect knowledge of the task memory requirements and execution times.
 We have compared two eviction strategies, starting with large files first or small files first,
-and did not report any significant changes in terms of validity of schedule or makespan.
-Hence, we only present results with the eviction of largest files first.
+and did not observe any significant changes in terms of schedule validity or makespan.
+Thus, we only present results with the eviction of largest files. % first. TODO: Why ``first''?
 %    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.
@@
 \subsubsection{Scheduling on the default cluster}
-
+%
 %    \paragraph{Normal cluster}
-
+%
 On the default cluster,  the three memory-aware heuristics are able to schedule all workflows
 (see Figure~\ref{fig:success-rates-large}), while
-    the baseline \heft has $24.2\%$ success rate ($75.7\%$ failure rate).
-    Indeed, \heft is only able to schedule small workflows, but no workflow over the size of $4000$ tasks
-    can be scheduled correctly; some tasks run out of memory. As soon as we are not in a setting
+the baseline \heft has a success rate of only $24.2\%$ ($75.7\%$ failure rate).
+Indeed, \heft is only able to schedule small workflows; it cannot schedule any workflow 
+with more than $4000$ tasks correctly as some tasks run out of memory. As soon as we are not in a setting
 with abundant processing resources for small workflows, it is hence necessary to adopt
@@
   \centering
-        \includegraphics[width=0.99\columnwidth] {images/success-rates-large}
+  \includegraphics[width=1\columnwidth] {images/success-rates-large}
   \caption{Success rates on the default cluster. Higher is better.}
@@
 by the memory-aware heuristics, normalized to the
-   makespan achieved by \heft, often through an invalid over-optimistic schedule that exceeds
+makespan achieved by \heft, often with an invalid over-optimistic schedule that exceeds
 the bound on memory.
@@
 On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline,
-    the makespans of \heftblc are $8\%$
+the makespans of \heftblc are $8.0\%$
 worse, and those found by \heftmm are $82.6\%$ worse. These are still very encouraging results,
@@
 Finally, we study the percentage of memory occupied by the schedule, which is another good indicator
-    of the memory usage of the heuristic and its ability to produce valid schedules.  %its important characteristic.
+of the memory usage of the heuristics and their ability to produce valid schedules.  %its important characteristic.
 Figures \ref{fig:mem-usages-normal} and~\ref{fig:mem-usages-onlyvalid} show the percentage of memory
 occupied on average by the schedule
-    produced by the different heuristics, for different workflow sizes, first on all schedules (hence including
-    \heft schedules that were not valid), and then only on valid schedules (hence, no results for \heft
-    on large workflow sizes).
+produced by the different heuristics for different workflow sizes;
+first on all schedules (including invalid \heft schedules) and then only on 
+valid schedules (hence, no results for \heft on large workflow sizes).
 %
@@
 
-    \heftmm continuously outperforms other heuristics in terms of memory usage,
+\heftmm continuously outperforms the other heuristics in terms of memory usage,
 using from $46\%$ less memory on the smallest
-    workflows to $4$ times less memory on the largest \numprint{30000}-task workflows.
+workflows to $4$ times less memory on the largest ones with \numprint{30000} tasks.
 If we consider the invalid \heft schedules, too, we see that they would require more and more memory on average,
 which explains why these schedules rapidly become invalid.
-    This is because some assignments require more that $100\%$ of memory (which makes them invalid).
+This is because some assignments require more than $100\%$ of the memory (which makes them invalid).
 We can assess the degree of invalidity by comparing \heft memory usage with the memory usage of \heftbl.
-    \heftbl differs from the baseline only in the sense that it respects available memory on the processors.
-    For the largest workflows, \heft schedules require almost twice the memory of \heftbl.
+\heftbl differs from the baseline only in the sense that it respects the available memory on the processors.
+For the largest workflows, \heft schedules require almost twice as much memory as \heftbl.
 
@@
 \subsubsection{Scheduling on the memory-constrained cluster}
-
-    On the memory-constrained cluster, \heft could produce valid assignments in only $14$ experiments
-    out of $290$~($4.8\%$ success rate).
-    The successful schedules were achieved exclusively on the tiny workflows (with only two 200-task size-increased
+\label{ss:mem-constrained-cluster}
+%
+On the memory-constrained cluster, \heft produces valid assignments in only $14$ out of $290$ experiments
+$290$~($4.8\%$ success rate).
+The successful schedules are achieved exclusively on the tiny workflows (with only two 200-task size-increased
 workflow among them).
-    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$ of them, while \heftmm
-    could still schedule all the workflows, including even the largest ones, see Figure~\ref{fig:success-rates-tiny}.
-    As also observed on the normal cluster, \heftmm seems to be less affected by the size of the workflow
-    when scheduling it than the other heuristics.
-
-    Similarly to the default cluster, we observe that relative makespan of \heftmm is usually greater
-    than those of \heft (see Fig.~\ref{fig:ms-relations-by-workflow-constrained}), but \heft schedules are
-    almost all invalid. It is therefore very interesting to resort to \heftmm for large workflows
+\heftbl successfully schedules $38\%$ of the workflows, \heftblc is able to schedule $49\%$ of them, while \heftmm
+still schedules all of them, including even the largest ones, see Figure~\ref{fig:success-rates-tiny}.
+As also observed on the default cluster, \heftmm seems to be less affected by the workflow size
+than the other heuristics.
+
+Similarly to the default cluster, we observe that the makespan of \heftmm is usually higher
+than with \heft (see Fig.~\ref{fig:ms-relations-by-workflow-constrained}), but the \heft schedules are
+almost all invalid. It is thus worthwhile to resort to \heftmm for large workflows
 in a constrained cluster, since tasks are processed in an order that minimizes

### 2024-12-23 main.tex
done editing Sec. VI

@@
 The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
-    situation - that is, if the scheduler assumed its memory to be smaller than it actually was and assigned it to a processor
-    with not enough memory capacity.
-    Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160 for all algorithm variants
-    ended with a valid makespan without recomputation.
+situation -- that is, if the scheduler assumed the task's memory to be smaller than actually needed and
+assigned the task to a processor with not enough memory capacity.
+Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160 
+for all algorithm variants end with a valid makespan without recomputation.
 In case of \heft, $14$ valid initial makespans were computed.
-    Out of them, $13$ managed to stay valid until the end, after all the update requests.
-    The same $13$ experiments ended with a valid makespan in case of no recomputation.
-    So, these workflows required so few resources that there was no point in re-scheduling them.
+Out of them, $13$ remain valid until the end, after all the update requests.
+The same $13$ experiments end with a valid makespan in case of no recomputation.
+So, these workflows require so few resources that there is no point in re-scheduling them.
 %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
@@
 %only, the success rate of not recomputing is $92\%$.
-    In case of \heftmm, all $290$ workflows could be scheduled initially and all the schedules remained valid until the end.
-    $16$ experiments ended with a successful makespan without recomputation.
-    The rate of successful schedules without recomputation is therefore~$5.5\%$.
-    \heftblc could produce $142$ valid initial schedules and kept $141$ of them valid until the end.
-    $50$ experiments were successful without recomputation, too, making it $35\%$ out of all successful final schedules.
-    \heftbl kept $105$ schedules valid until the end out of $110$ initial valid schedules.
-    $55$ were able to be kept valid until the end even without recomputation, roughly $50\%$ of all successes.
-
-    \heftbl and \heftblc were successful on smaller workflows and failed on larger ones, the success of the strategy without
-    recomputation was limited to even the smallest of these smaller workflows.
-    So, the strategy without recomputation delivered valid makespans (independently of the algorithm) on $56$ original
-    workflows with $<100$ tasks, $47$ 200-task ones, $25$ 1000-task ones and $6$ 2000-task ones.
+In case of \heftmm, all $290$ workflows can be scheduled initially and all the schedules remain valid until the end.
+$16$ experiments end with a successful schedule without recomputation, a rate of~$5.5\%$.
+\heftblc produces $142$ valid initial schedules and keeps $141$ of them valid until the end.
+$50$ experiments are successful without recomputation, a rate of $35\%$ out of all successful final schedules.
+\heftbl keeps $105$ schedules valid until the end out of $110$ initial valid schedules.
+$55$ remain valid until the end even without recomputation, roughly $50\%$ of all successes.
+
+\heftbl and \heftblc are successful on smaller workflows and fail on larger ones, 
+the success of the strategy without recomputation is limited to even the smallest of these smaller workflows.
+Thus, the strategy without recomputation delivers valid makespans (independently of the algorithm) 
+on $56$ original workflows with $<100$ tasks, $47$ $200$-task ones, $25$ $1000$-task ones, and $6$ $2000$-task ones.
 
 Figure~\ref{fig:updates-ms} shows the increase in makespan in case of no recomputation for these experiments.
-    With growing size of the workflow, the excess makespan of not recomputing grows - from $13.9\%$ to $20\%$ for \heftbl,
-    from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the 2000-task workflows in this case),
+With growing size of the workflow, the excess makespan of not recomputing grows -- from $13.9\%$ to $20\%$ for \heftbl,
+from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the $2000$-task workflows in this case), % TODO: why not?
 $12.1\%$ to $23.5\%$ for \heftmm.
-    The larger variations for \heft can be explained by the small amount of data -- for instance, there were only 2 200-task
-    workflows for this case.
+The larger variations for \heft can be explained by the small amount of data -- for instance, there are
+only $2$ $200$-task workflows for this case.
 
@@
 \label{sec.expe.t}
-
-    To be able to answer the runtime system without holding it up for too long, the scheduler needs to 
-    be able to compute a schedule rapidly. %provide a fast answer.
-    The bottom-level-based heuristics \heftbl and \heftblc provide smaller running times than \heftmm, 
-    and also scale better with growing workflow
-    sizes~(see Fig.~\ref{fig:runtimes-log}).
-    Their running times are similar and grow from tens of milliseconds for the smallest workflows to $25-27$ seconds 
-    for the largest workflows.
-    \heftmm, however, needs to compute a memory-optimal traversal of the entire workflow to compute the ranks 
-    of the tasks,
-    so its running time scales from also tens of milliseconds for the smallest workflows, 
-    to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
-    $2994.9$ seconds for the largest, \numprint{30000}-task workflows.
-    This increased running time is, though, offset by the unique $100\%$ success rate this algorithm 
-    obtained when scheduling large
-    workflows in difficult (memory-constrained) setups.
+%
+In order to answer the runtime system without holding it up for too long, the scheduler needs to 
+compute a schedule reasonably quickly. %provide a fast answer.
+The bottom-level-based heuristics \heftbl and \heftblc yield smaller running times than \heftmm, 
+and also scale better with growing workflow sizes~(see Fig.~\ref{fig:runtimes-log}).
+Their running times are similar and grow from tens of milliseconds for the smallest workflows to 
+$25$-$27$ seconds for the largest workflows.
+\heftmm, in turn, computes a memory-optimal traversal of the entire workflow to compute the task ranks. 
+It also takes only tens of milliseconds for the smallest inputs, but requires
+thousands of seconds for the large(st) workflows ($1172.7$s for \numprint{20000}-task workflows and
+$2994.9$s for \numprint{30000}-task ones).
+This increased running time is, however, offset by the unique $100\%$ success rate this algorithm 
+obtained in our experiments when scheduling large workflows in difficult (memory-constrained) setups.
 
@@
 
-\section{Conclusion} %  \skug{Full: 0, Polished: 0}}
+\section{Conclusion}
 \label{sec:conc}

### 2024-12-23 main.tex
done editing Sec. VII

@@
 by the schedule. Experimental results on a large set of workflows coming from real-life applications
-demonstrate that the memory-aware heuristics are  successfully producing valid schedules. 
+demonstrate that the memory-aware heuristics produce valid schedules successfully. 
 In the most memory-constrained setting, \heftmm succeeds to schedule even the largest workflows,
-while other heuristics return invalid schedules exceeding the memory capacity of processors. 
-It comes at a price of a makespan that is not quite as good as the one obtained by \heftbl
-and \heftblc, which focus on makespan by ordering tasks by bottom level. 
-As expected, the baseline \heft that is not memory aware returns invalid schedules in almost
+while the other heuristics return invalid schedules that exceed the memory capacity. 
+This advantage of \heftmm comes at the price of worse makespans than with \heftbl
+and \heftblc. % which focus on makespan by ordering tasks by bottom level. 
+As expected, the baseline \heft, which is not memory-aware, returns invalid schedules in almost
 all settings, except for very small workflows.
@@
 Another key contribution is that we have adapted these heuristics for a dynamic setting, where exact
-task parameters (execution times and memory requirements) are not know in advance. 
-We have implemented a runtime system that interacts with the scheduler, returning exact
+task parameters (execution times and memory requirements) are not known in advance. 
+We have implemented a runtime system that interacts with the scheduler, providing exact
 parameter values once a task arrives in the system, while only estimates are known for future tasks. 
-Some preliminary experiments have been conducted in this setting, and demonstrated that
+Some preliminary experiments have been conducted in this setting and demonstrated that
 it is necessary to adapt the schedule on the fly in order to avoid an execution failure because
-of a shortage in memory. To the best of our knowledge, this is the first study of adaptive
+of a memory shortage. To the best of our knowledge, this is the first study of adaptive
 algorithms accounting for memory constraints.
@@
 More importantly, it would be interesting to consider other types of variability, for instance
-if new tasks appear in the graph (or disappear), or if there is variability in the platform, 
+if new tasks appear in (or disappear from) the workflow graph, or if there is variability in the platform, 
 with processors arriving and departing. We believe that we could adapt the current approach

### 2024-12-23 main.tex
figures are quite small, but hard to change in a 2-column layout

@@
 
-\begin{table}[htb]
+\begin{table}[b]
     \begin{center}
@@
     \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
-    \caption{Memory usage on default cluster, including only valid \heft schedules. }
+    \caption{Memory usage on default cluster, considering only valid \heft schedules. }
     \label{fig:mem-usages-onlyvalid}
@@
     \centering
-    \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
+    \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
     \caption{Relative makespans on the memory-constrained cluster.

### 2024-12-23 main.tex
minor things to improve layout

@@
   \label{fig:success-rates-large}
-  \vspace{-0.3cm}
+  \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:ms-relations-by-workflow}
-    \vspace{-0.3cm}
+    \vspace{-0.15cm}
 \end{figure}
@@
      \label{fig:mem-usages-normal}
-    \vspace{-0.3cm}
+    \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:mem-usages-onlyvalid}
-    \vspace{-0.3cm}
+    \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:success-rates-tiny}
-    \vspace{-0.3cm}
+    \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:ms-relations-by-workflow-constrained}
-    \vspace{-0.3cm}
+    \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:memory-usage-constrained}
-    \vspace{-0.3cm}
+    \vspace{-0.15cm}
 \end{figure}
@@
     \includegraphics[width=1.1\columnwidth] {images/runtimes-logarithmic}
-    \caption{Running times of the heuristics. The y axis is logarithmic.}
+    \caption{Average running times (in seconds) of the heuristics \wrt workflow size. The $y$-axis is logarithmic.}
     \label{fig:runtimes-log}
@@
 
-We have formalized a scheduling problem in memory-constrained environments, where tasks
+We have formalized a scheduling problem in memory-constrained environments where tasks
 may exceed the memory available on a processor and resort to communication buffers
 to store and communicate data between processors. In this context, we have 
-designed three memory-aware HEFT-based heuristics, which account for memory
-constraints while scheduling tasks.
-
-Two heuristics rely on an ordering of tasks based on the bottom level of tasks, with the objective
+designed three memory-aware HEFT-based heuristics which account for memory
+constraints when scheduling workflow tasks.
+
+Two of these heuristics rely on a task ordering \wrt the bottom level of tasks, with the objective
 of minimizing the makespan and hence scheduling critical tasks first. The third one, \heftmm,
-goes one step further and handles tasks in an order dictated by an efficient traversal
-of the graph in terms of memory requirements, hence reducing the memory used
-by the schedule. Experimental results on a large set of workflows coming from real-life applications
+% goes one step further and 
+handles tasks in an order dictated by an efficient traversal
+of the workflow in terms of memory requirements, hence reducing the memory used
+by the schedule. Experimental results on a large set of workflows from real-world applications
 demonstrate that the memory-aware heuristics produce valid schedules successfully. 
@@
 while the other heuristics return invalid schedules that exceed the memory capacity. 
-This advantage of \heftmm comes at the price of worse makespans than with \heftbl
+This advantage of \heftmm comes at the price of worse makespans than those of \heftbl
 and \heftblc. % which focus on makespan by ordering tasks by bottom level. 
 As expected, the baseline \heft, which is not memory-aware, returns invalid schedules in almost
-all settings, except for very small workflows.
+all cases, except for very small workflows.
 
@@
 
-This work could be extended in several directions. First, the model could be refined to include
-heterogeneous bandwidths, while we consider a homogeneous communication network. 
-More importantly, it would be interesting to consider other types of variability, for instance
-if new tasks appear in (or disappear from) the workflow graph, or if there is variability in the platform, 
+This work can be extended in several directions. First, the model could be refined to include
+heterogeneous bandwidths. %, while we consider a homogeneous communication network. 
+More importantly, it would be interesting to consider other types of variability, \egc
+if new tasks (dis)appear in the workflow graph, or if there is variability in the platform, 
 with processors arriving and departing. We believe that we could adapt the current approach
-by recomputing schedules on the fly, and we plan to perform a new set of experiments
-to further assess the impact of dynamic scheduling.
-
-%
-%\begin{itemize}
-%  \item Extend to heterogeneous bandwidths. 
-%  \item Further variability commented out from the paper: changing DAG structure; changing platform...
-%\end{itemize}
-
+based on recomputing schedules on the fly, and we plan to perform a new set of experiments
+to further assess the impact of dynamic scheduling techniques.
+
+%
 \balance

### 2024-12-23 main.tex
funding ack placeholder (anonymity preserved)
Still working, added paragraph about lack of comparison with icpp paper (from rebuttal)

@@
 \newcommand{\criticalpath}{\mathcal{P}}
-\newcommand{\parents}[1]{\,parent_{#1}}
+\newcommand{\parents}[1]{\,\Pi_{#1}}
 \newcommand{\children}[1]{\,child_{#1}}
@@
 %    written by task~$u$ and used as input by task~$v$.
-We denote by $\parents{u}$ the tasks preceding task~$u\in V$, which must be completed before $u$ can be started:   
+We denote by $\parents{u}$ the tasks preceding task~$u\in V$, also called {\em parents}, 
+which must be completed before $u$ can be started:   
 %    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
@@
 of a task during its execution, including input and output files currently being read and written,
-and hence the total memory requirement for executing task~$u$ is the maximum of the following:
+and hence the total memory requirement for executing task~$u$, denoted by~$r_u$,  is the maximum of the following:
 (i) the total size of the files to be received from the parents, (ii) the total size of the files
@@
 
-In oder to compute the memory requirement of a DAG, we use \algo{memDag}~\cite{KAYAASLAN20181},
+In order to compute the memory requirement of a DAG, we use \algo{memDag}~\cite{KAYAASLAN20181},
 an algorithm that 
@@
                 $G = (V, E)$                          & Application DAG (tasks and edges)       \\
- %               $\parents{u}$, $\children{u}$         & Parents of task $u$, children of task $u$            \\
-                $m_u$                                 & Amount of memory required by task $u$                                \\
+                $\parents{u}$, $\children{u}$         & Parents of task $u$, children of task $u$            \\
                 $w_u$                                 & Number of operations of task $u$         \\
                 $c_{u,v}$                             & Size of output file for edge $(u,v)\in E$         \\
+                $m_u$                                 & Amount of memory required by task $u$                                \\
+                $r_u$ 				& Total memory requirement for task~$u$ \\
 %                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
@@
                 $\beta$                     & Bandwidth in the compute system                                \\
-                $bl{u}$                      & Bottom level of task $u$ \\
+                $bl(u)$                      & Bottom level of task $u$ \\
+                $ST(u,p_j)$	 & Start time of task~$u$ on~$p_j$ \\
+               $FT(u,p_j)$	 & Finish time of task~$u$ on~$p_j$ \\
 %                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
@@
 has the earliest finish time.
-We tentatively assign each task~$v$ to each processor~$j$.
-The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$
+We tentatively assign each task~$v$ to each processor~$p_j$.
+The task's starting time $ST(v,p_j)$ on processor~$p_j$ is dictated by the maximum between the ready time of the processor~$rt_j$
 and all communications that
@@
 
-{\footnotesize{   \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]}}
+{\footnotesize{   \[ST(v, p_j) = \max{ \Large\{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ \frac{c_{u,v}}{\beta} , 
+rt_{proc(u), p_j} + \frac{c_{u,v}}{\beta}  \} \Large\} } . \]}}
 
@@
 When assigning the task, we set the ready time $rt_j$ of  processor~$j$ to be the finish time of the task.
-For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
-communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
+For every predecessor~$u$ of task~$v$ that has been assigned to another processor~$j'$, we adjust ready times on
+communication buffers $rt_{j', j}$: % for every predecessor $u$'s processor $j'$: 
+we increase them by the
 communication time $c( u,v) / \beta$.
@@
 \noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
-Its starting time $st_v$ is dictated by the maximum between $rt_j$ and all communications that
+Its starting time $ST(v, p_j)$ on $p_j$ is dictated by the maximum between $rt_j$ and all communications that
 must be orchestrated from predecessor tasks $u\notin T(p_j)$.
@@
 
-{\footnotesize{ \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} .} \]}}
+{\small{ \begin{align*}
+ST(v, p_j) & = \max  \large\{rt_j,  \\
+& \max_{ u \in \parent(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + \frac{c_{u,v}}{\beta} \Large\} . 
+\end{align*}
+}}
 
@@
 Finally, we report running times of the heuristics in Section~\ref{sec.expe.t}.
+
+Note that we consider HEFT as the baseline; experiments to compare the novel heuristics
+with the closest competitor~\cite{DBLP:conf/icpp/KulaginaMB24} obtained much better results 
+(at least 10x smaller makespans). 
+Indeed, we have much more flexibility in the current setting where processors are reused, since tasks
+are not just partitioned. Hence, the comparison with~\cite{DBLP:conf/icpp/KulaginaMB24} seemed not fair
+and we believe the natural baseline is HEFT in this setting. 
+We are not aware of other dynamic scheduling algorithms that account for memory usage 
+and hence could not find other competitors from the literature apart from HEFT.
 
@@
 
-%
+\subsubsection*{Acknowledgements}
+Funding acknowledgement suppressed to preserve anonymity.
+%This work is partially supported by Collaborative Research Center (CRC) 1404 FONDA
+%–- Foundations of Workflows for Large-Scale Scientific Data Analysis, which is funded
+%by German Research Foundation (DFG).
+% TODO: include FONDA ack in final version
+
 \balance

### 2024-12-23 main.tex
Some figs larger

@@
   \centering
-  \includegraphics[width=1\columnwidth] {images/success-rates-large}
+  \includegraphics[width=1.08\columnwidth] {images/success-rates-large2}
+  \vspace{-0.8cm}
   \caption{Success rates on the default cluster. Higher is better.}
@@
     \centering
-    \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size}
+  \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size2}
+    \vspace{-0.8cm}
     \caption{Relative makespans of heuristics normalized by \heft makespan, by workflow
@@
     \centering
-    \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal}
+    \includegraphics[width=1\columnwidth] {images/mem-usage-normal2}
     \caption{Memory usage on default cluster, including invalid \heft schedules. }
@@
     \centering
-    \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
+    \includegraphics[width=1\columnwidth] {images/mem-usage-normal-onlyvalid2}
     \caption{Memory usage on default cluster, considering only valid \heft schedules. }
@@
     \centering
-    \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
+    \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size-constrained2}
     \caption{Relative makespans on the memory-constrained cluster.
@@
     \centering
-    \includegraphics[width=1.1\columnwidth] {images/mem-usage-constrained-onlyvalid}
+    \includegraphics[width=1\columnwidth] {images/mem-usage-constrained-onlyvalid2}
     \caption{Memory usage on the memory-constrained  cluster. } % by different heuristics by workflow size.}
@@
     \centering
-    \includegraphics[width=1.1\columnwidth] {images/runtimes-logarithmic}
+    \includegraphics[width=1\columnwidth] {images/runtimes-logarithmic2}
     \caption{Average running times (in seconds) of the heuristics \wrt workflow size. The $y$-axis is logarithmic.}

### 2025-01-14 main-disk-model.tex
put first figures into the text, described the setups. Everything WIP.

@@
     \subsubsection{Target computing systems}
-%    \skug{Because of the normalized nature of the task/edge weights, I will now be saying that our larger cluster has 192 GB and it
-%    is represented as 192 000 (with 3 zeroes). For the memory-constrained cluster (192 00 with 2 zeroes) I will say that it has 10 times
-%    less memory - 19.2 GB. Other processors will be dealt with similarly.}
 
@@
     kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
-    We set the number of each kind of node to $12$, thus having 72 processors in total.  % the whole cluster.
-%
-    Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
+    We set the number of each kind of node to $6$, thus having 36 processors in total.  % the whole cluster.
+%
+    Each machine has a normalized CPU speed and a normalized memory size, and we list them as (name, speed, memory):
     ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
@@
                 Processor  %& Amount
-                &  CPU speed   & \multicolumn{2}{c}{Memory size (GB)} \\
-                name & (GHz) & {\em default} & {\em mem-constrained} \\
+                &  CPU speed   & \multicolumn{2}{c}{Memory size (relative)} \\
+                name & (relative) & {\em default} & {\em mem-constrained} \\
                 \midrule
@@
 
-    We also consider a more constrained setting, by varying the cluster configuration. The
-     {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
-     but each node has 10 times less memory. Hence,  
-     the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, 
-     $A2$ has $6.4$~GB instead, of $64$~GB etc.
-    The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).
-
-Note that in both clusters, we set the size of the communication buffer to be equal
-to ten times the the memory size.
+    When we consider a default execution environment, we multiply the relative memory size with $10^6$.
+    We multiply the relative speeds with $100$, and let read-write operations unpenalized (multiple the size of files with 1).
+    Effectively, we thus reward the computations with $100$ in comparison  with IO operations.
+    We further penalize offloading of memory by multiplying its execution time by $1000$.
+
+    For a memory-constrained scenario, we give the processors ten time less memory: we multiply the relative memory size
+    with $10^5$.
+    We keep all the settings related to speeds the same.
+    For a more memory-loo scenario, we give the processors ten times more memory.
+
+    For the IO-based scenarios, we keep the amount of memory the same as in the default scenario and change only the penalties for the read-write
+    operations and offloading.
+    For the ``IO slower'' case, we penalize both the read-write operations and the offloading ten times more: we multiply the length of
+    read-write operations by $10$, and the offloading by $10^4$.
+    In the ``IO faster'' case, we do the opposite and reward read-write operations by dividing their time by to, while penalize the offloading only
+    by $100$ instead of $1000$.
+    Finally, we examine the effect of the offloading penalty reducing and increasing only it ten times for the experimental setups ``offload faster''
+    and ``offload slower''.
+
+    TODO: describe with more and less nodes.
+   % We also consider a more constrained setting, by varying the cluster configuration. The
+   %  {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
+   %  but each node has 10 times less memory. Hence,
+   %  the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB,
+   %%  $A2$ has $6.4$~GB instead, of $64$~GB etc.
+   % The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).
 
@@
 
-    To simulate the execution of a workflow, we implemented a runtime system.
-    It reads the historical data and builds weights for tasks, as explained above.
-%    These weights are then treated as the ``truth'' about the execution during the simulation.
-    In the static case, these values are being sent to the scheduler, which builds a schedule
-    according to these weights.
-    However, in the dynamic setting, the runtime system applies a deviation function to the values.
-    The deviation function computes a normally distributed deviated value, where the initial value
-    is the mean and the deviation is~$10\%$.
-    This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and
-    other works dedicated to predicting runtimes of tasks~\cite{da2015online,da2013toward}.
-
-    Hence, the scheduler receives deviated values and makes decision based on them.
-    This leads to several types of possible issues:
-    \begin{itemize}
-        \item A processor is blocked by another task. If the scheduler underestimated the execution time of a task,
-        it will block another one from starting.
-        \item A predecessor has not yet finished.
-        The scheduler may request a task to start on its processor, while
-        some of the predecessors of the tasks have in fact not yet completed their execution, and the
-        task is therefore not yet ready.
-%        A similar situation, but a task could start on its processor, but one or more of its parents a
-%        are not ready yet.
-        \item Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able to execute on a chosen processor.
-        \item A task took less time than expected. We only consider this case if a task took more than 10\% less
-        time than expected. In this
-        case, we want to exploit the newly acquired free time by possibly starting other tasks earlier.
-    \end{itemize}
+    No ``runtime system'' as such in the static case.
+    Todo: either remove entirely or explain how simulation of the runtime system works in the dynamic case.
 
@@
 
-   % \subsubsection{Preliminary experiments}
     We first study the heuristics behavior when the weights do not change upon runtime,
@@
     Hence, we only present results with the eviction of largest files first.
-%    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.
+
+    \subsubsection{Relative makespan}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-normal-bar}
+        \caption{Relative makespans on the default cluster setup. Lower is better.}
+        \label{fig:rel-ms-normal-bare}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-iofast-bar}
+        \caption{Relative makespans on the cluster setup with faster IO. Lower is better.}
+        \label{fig:rel-ms-iofast-bar}
+        \vspace{-0.3cm}
+    \end{figure}
+
+    \begin{figure}[tb]
+        \centering
+        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-ioslow-bar}
+        \caption{Relative makespans on the cluster setup with slower IO. Lower is better.}
+        \label{fig:rel-ms-ioslow-bar}
+        \vspace{-0.3cm}
+    \end{figure}
 
     \subsubsection{Scheduling on the default cluster}
-
+    AFTER HERE IS OLD.
 %    \paragraph{Normal cluster}

### 2025-01-27 main-disk-model.tex
minor changes from programming

@@
     When we preliminarily schedule a read, we first assess its length and other actions that happen on this processor in the meantime.
-    So, the first estimated start of the read is the estimated start time of the task minus the estimated length of the read.
+    So, first estimated start of the read is \skug{the maximum between the desired start time (estimated start time of the task minus the estimated length of the read),
+    and the feasible one (when the lastest read event on this processor finishes)}.
+
     We then see if this start of the read is happening during the runtime of the previous task or already after it (during the writes).
@@
     with this new time.
+    \skug{We  also update the actual times of all its successor events.}
 

### 2025-02-18 main.tex
last submitted version
author names

@@
 
-%\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\author{\IEEEauthorblockN{Svetlana Kulagina}
+\IEEEauthorblockA{\textit{Humboldt-Universität zu Berlin, Germany} \\
+ORCID: 0000-0002-2108-9425}
+\and
+\IEEEauthorblockN{Anne Benoit}
+\IEEEauthorblockA{\textit{ENS Lyon, France} \\
+ORCID: 0000-0003-2910-3540
+}
 %\textit{name of organization (of Aff.)}\\
 %City, Country \\
-%email address or ORCID}
-%\and
-%\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%\textit{name of organization (of Aff.)}\\
-%City, Country \\
-%email address or ORCID} }
+%email address or ORCID
+\and
+\IEEEauthorblockN{Henning Meyerhenke}
+\IEEEauthorblockA{\textit{Humboldt-Universität zu Berlin, Germany} \\
+ORCID: 0000-0002-7769-726X}
+}
 
@@
     \includegraphics[width=1\columnwidth] {images/mem-usage-normal2}
+    \vspace{-0.6cm}
     \caption{Memory usage on default cluster, including invalid \heft schedules. }
@@
     \includegraphics[width=1\columnwidth] {images/mem-usage-normal-onlyvalid2}
+    \vspace{-0.6cm}
     \caption{Memory usage on default cluster, considering only valid \heft schedules. }
@@
 
+\bigskip
 \section{Conclusion}
@@
 to store and communicate data between processors. In this context, we have 
-designed three memory-aware HEFT-based heuristics which account for memory
+designed three memory-aware HEFT-based heuristics that account for memory
 constraints when scheduling workflow tasks.

### 2025-02-18 main.tex
FONDA ack

@@
 \subsubsection*{Acknowledgements}
-Funding acknowledgement suppressed to preserve anonymity.
-%This work is partially supported by Collaborative Research Center (CRC) 1404 FONDA
-%–- Foundations of Workflows for Large-Scale Scientific Data Analysis, which is funded
-%by German Research Foundation (DFG).
-% TODO: include FONDA ack in final version
+This work is partially supported by Collaborative Research Center (CRC) 1404 FONDA
+–- Foundations of Workflows for Large-Scale Scientific Data Analysis, which is funded
+by German Research Foundation (DFG).
 

### 2025-02-24 main.tex
removed some parentheses

@@
 
+\newcommand{\new}[1]{{\color{blue}#1}}
+
 \newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
@@
 heterogeneous platforms, where each processor can have a different speed and a different memory size.
-We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that (in contrast to the original) accounts for memory and
+We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that, in contrast to the original, accounts for memory and
 includes eviction strategies for cases when it might be beneficial to remove some data from memory
@@
 An efficient execution of the workflows on such platforms requires the mapping of tasks
-to specific processors; to increase utilization by reusing finished processors,
-one also needs a task schedule (\iec a valid execution order that respects the dependencies)
-and possibly also starting times for the tasks.
+to specific processors.
+\new{A task schedule - \iec a valid execution order that respects the dependencies,  and possibly also starting times for
+the tasks are needed to increase utilization by reusing finished processors.}
 
@@
 processor with less memory than needed for the task. Violating a memory constraint is, however, 
-very important to avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
+very important to avoid possibly expensive runtime failures and to provide a satisfactory user experience.
 Hence, building on previous related %\hmey{You may want to add other refs not from us} 
@@
 \item We design three HEFT-based heuristics
-that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
-(M behind HEFT for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
+that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm.
+M behind HEFT \new{stands for} for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
 BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
-and MM for \underline{m}inimum \underline{m}emory traversal).
-The difference between the new heuristics is the way they prioritize tasks (for processor assignment).
+and MM for \underline{m}inimum \underline{m}emory traversal.
+The difference between the new heuristics is the way they prioritize tasks for processor assignment.
 
@@
     on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (and thus unrelated to each other).
+    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related
+    and thus unrelated to each other.
     The authors search for an optimal schedule

### 2025-02-25 main.tex
improving figures v2. WIP
last figure
some small explanations of experiments added

@@
   \vspace{-0.8cm}
-  \caption{Success rates on the default cluster. Higher is better.}
+  \caption{Success rates \new{by workflow size and algorithm} on the default cluster. Higher is better.}
   \label{fig:success-rates-large}
@@
 the memory usage of schedule.
+\new{Due to a varying nature of the workflow generator, some larger workflows can have smaller makespans than their
+smaller counterparts due to more parallel execution due to higher fanout or other factors of the workflow's
+internal composition.
+}
+
 %
@@
 
-Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained},
+Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained}.
+\new{We plot the average memory usage in the cluster per algorithm and workflow size},
 and we observe that the memory footprint of \heftmm remains constant with workflow size.
@@
     \includegraphics[width=0.95\columnwidth] {images/UpdatesMss2}
-    \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm. Smaller is better.}
+    \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm \new{in case of no recomputation}. Smaller is better.}
     \label{fig:updates-ms}

### 2025-03-04 main.tex
implemented some of Henning's todos

@@
 as well as an integer linear programming (ILP) solution to the scheduling problem.
-The memory model requires the deletion of all input files when starting the task.
-% and adding all output files there.
-% TODO: This sounded strange. I changed ``input data'' to ``input files''.
-% How can I process input data when they are deleted? 
-% TODO: And what is ``there''?
+The memory model \new{requires deleting all input files when the task starts and adding all output files to memory.}
 
@@
 Yet, they propose only a partitioning-based mapping of the workflow, without processor reuse.
-Hence, in~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need for communication buffers to store data 
-that should be communicated between two processors when tasks are ready to execute. 
-% TODO: Why is the last sentence necessary? More specifically, why is it a consequence of the previous
-% sentence (as implied by ``Hence''?)?
+\new{Because a processor is idle after a partition of a workflow has finished executing}, in~\cite{DBLP:conf/icpp/KulaginaMB24},
+there is no need for communication buffers to store data
+that should be \new{exchanged} between two processors when tasks are ready to execute.
 
@@
 and did not observe any significant changes in terms of schedule validity or makespan.
-Thus, we only present results with the eviction of largest files. % first. TODO: Why ``first''?
+Thus, we only present results with the eviction of largest files. % first.
+%TODO: Why ``first''? SK: don't see a problem here. We first kick the largest file, then the second largest etc
+%as opposed to first the smallest, then second smallest ...
+
 %    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.
@@
 With growing size of the workflow, the excess makespan of not recomputing grows -- from $13.9\%$ to $20\%$ for \heftbl,
-from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the $2000$-task workflows in this case), % TODO: why not?
+from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the $2000$-task workflows in this case),
+% TODO: why not? SK: no (or not enough) valid solutions
 $12.1\%$ to $23.5\%$ for \heftmm.

### 2025-03-07 main.tex
new figures 2 and 6 (WIP)

@@
     \centering
-  \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size2}
+  \includegraphics[width=1\columnwidth] {images/ms_relations_by_wf_size_barplot}
     \vspace{-0.8cm}
@@
     \centering
-    \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size-constrained2}
+    \includegraphics[width=1\columnwidth] {images/ms_relations_by_wf_size-constrained-barplot}
     \caption{Relative makespans on the memory-constrained cluster.

### 2025-03-07 main.tex
rewrote the "on average" with more concrete numbers

@@
 %    makespans found by the baseline (including invalid ones), by workflow size.
-On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline,
-the makespans of \heftblc are $8.0\%$
-worse, and those found by \heftmm are $82.6\%$ worse. These are still very encouraging results,
+\new{The makespans found by \heftbl are $<13\%$ worse than those found by the baseline for workflows with less than \numprint{2000}
+tasks, and up to $27\%$ worse for the largest workflows.
+The makespans of \heftblc are very similar,being $<17\%$ worse on smallest workflows, $19$-$28\%$ worse on middle-sized,
+and up to $30\%$ worse on the largest workflows.
+The makespans found by \heftmm are $25$-$30\%$ worse on the smallest workflows, $40$-$70\%$ on middle-sized and up to
+$210\%$ on the largest ones. These are still very encouraging results,
 in particular for \heftbl and \heftblc, since the makespans of \heft correspond to invalid schedules.
-
+}
  %   \skug{TODO: how much worse are makespans for SUCCESSFUL hefts? presulmably almost identical}

### 2025-03-07 main-disk-model.tex
main-disk-model.tex cleaned until II.C

@@
 
-        %Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
-        Scientific workflows are often represented as directed acyclic graphs (DAGs),
-        where vertices correspond to tasks and edges represent the dependencies between them.
-        Typically, each task requires a
-        certain amount of memory to be executed and needs to communicate data to its successor tasks.
-        The goal is generally to execute the workflow  as fast as possible (i.e., to minimize its makespan),
-        while satisfying the memory constraints.
-        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows on
-        heterogeneous platforms, where each processor can have a different speed and a different memory size.
-        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that (in contrast to the original) accounts for memory and
-        includes eviction strategies for cases when it might be beneficial to remove some data from memory
-	in order to have enough memory to execute other tasks.
-%
-	Furthermore, while HEFT assumes perfect knowledge of the execution time and memory usage
-	of each task, the actual values might differ upon execution. Thus, we propose an adaptive
-	scheduling strategy, where a schedule is recomputed when there has been a significant variation in terms
-	of execution time or memory.
-%
-	The scheduler has been closely integrated with a runtime system, allowing us to perform a thorough
-	experimental evaluation on real-world workflows. The runtime system warns the scheduler when
-	the task parameters have changed, and a schedule can be recomputed on the fly. The memory-aware
-	strategy allows us to schedule task graphs that would run out of memory with a state-of-the-art
-	scheduler, and the adaptive setting allows us to significantly reduce the makespan.
-
-%                when tentatively assigning tasks, in order to
-%        Its first step is to compute the weights of the tasks.
-%        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
-%        and weights along the optimal memory traversal.
-%        In the second step, we try assigning each task to each processors and execute the assignment that
-%        is feasible with regard to memory size and gives the earliest finishing time to the task.
-%        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
-%        assign a task to a processor.
-%        We suggest two eviction strategies - largest files first and smallest files first.
-
-%        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
-%        ones with real task and edge weights with up to 30,000 tasks shows that
-%        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
-%        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
-%        on small one.
-%        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
-%        average memory utlization by $t\%$.
+      TODO
 
@@
 
-\section{Introduction} %: \skug{Full: 0, Polished: 0}}
-
-%    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
-%
-%    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
-%    }
-%    TODO: Insert abstract
+\section{Introduction}
 
     %%% CONTEXT %%%
-    The analysis of massive datasets, originating from fields such as genomics, 
-    remote sensing, or biomedical imaging -- to name just a few -- has become ubiquitous in science;
-    this often takes the form of workflows, \iec separate software components chained together
-    in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
-    These workflows are usually represented as directed acyclic graphs (DAGs).
-    The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
-    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liu2018survey}.
-    Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
-    single computer and are therefore executed on a parallel or distributed platform.
-    An efficient execution of the workflows on such platforms requires mapping tasks
-    to specific processors; to increase utilization by reusing finished processors,
-    one also needs a task schedule (\iec a valid execution order that respects the dependencies)
-    and possibly also starting times for the tasks.
 
     %%% MOTIVATION %%%
-    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
-    and memory sizes. In general, having different memory sizes per CPUs makes it more challenging to compute
-    a schedule that respects all memory constraints -- meaning that no task is executed on a 
-    processor with less memory than needed for the task. This is, however, very important to
-    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
-    Hence, building on previous related %\hmey{You may want to add other refs not from us} 
-    work~\cite{gou2020partitioning,He21,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
-    formulation that takes memory sizes as explicit constraints into account. Its objective is
-    the very common \emph{makespan}~\cite{liu2018survey}, 
-    which acts as proxy for the total execution time of a workflow.
-However, to the best of our knowledge,
-    the only memory-aware heuristics that would account for memory constraints are  partitioning
-    the DAG and not reusing processors once they have processed a part of the graph, leading
-    to high values of makespan compared to a finer grain solution that reuses processors.
-
-    While previous work with memory constraints has been focusing on partitioning the graph, 
-    and not reusing processors during execution, 
-    a seminal list scheduling heuristic for workflows on heterogeneous platforms, without accounting
-    for the memory constraint, is HEFT 
-    (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
-    It has two phases: (i) each task is assigned a priority; and (ii) the tasks in a priority-ordered list are assigned
-    to processors, where the ``ready'' task with the highest priority is scheduled next on the processor
-    where it would complete its execution first. 
-    HEFT has been extended (e.g., by Shi and Dongarra~\cite{SHI2006665}) and adjusted 
-    for a variety of different scheduling problem formulations. 
-    Yet, none of them adhere to memory constraints as we propose -- see discussion of related work
-    in Section~\ref{sec:related-work}. 
-%    (\skug{check in related work if true!}).
-%        \skug{Note: 2 papers that deal with memory sizes, but model is very different!}
-
-    Another limitation in practice of HEFT (and many other scheduling strategies) is their 
-    assumption that the task running times provided to them are accurate. In practice, this is 
-    not the case and deviations from user estimates or historical measurements are 
-    very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
-    deviations occur. However, the original list-based schedulers, such as HEFT, are only defined
-    in a static setting with accurate task parameters. 
-%    
-%    List-based schedulers such as HEFT are, however, not designed for 
-%    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
-%    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
-%%    and would compute a completely new schedule from scratch.
 
@@
     \begin{itemize}
-\item We formalize the problem with memory constraints, where communication buffers
-are used to evict data from memory if it will be later used by another processor. 
-\item We design three HEFT-based heuristics
-    that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
-    (M behind HEFT for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
-    BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
-    and MM for \underline{m}inimum \underline{m}emory traversal).
-    The difference between the new heuristics is the way they prioritize tasks (for processor assignment).
-
-    \item We implement a runtime system able to provide some feedback to the scheduler
-    when task requirements (in terms of execution time and memory) differ from the initial predictions, 
-    and we recompute a schedule, based on the reported deviations.
-
-    \item We perform extensive simulations, first in the static case by comparing the schedules produced 
-    by these heuristics with the classical HEFT as baseline, which however does not take memory sizes into account; 
-    while HEFT returns invalid schedules that exceed the processor memories and cannot execute correctly,
-    the new heuristics are able to successfully schedule large workflows, with reasonable makespans.
-
-    \item In the dynamic setting, we use a runtime system that allows us to simulate workflow executions,
-    introducing deviations in running times and memory requirements of tasks that are communicated
-    back to the scheduler; the scheduler can then recompute a schedule. Without these recomputations,
-    most schedules become invalid after deviations, since the memory constraint is exceeded 
-    for most workflows, hence demonstrating the necessity of a dynamic adjustment of the schedule.
-
-%    \hmey{Need to clarify: why is this SotA?}
-
-  %  \begin{itemize}
-%      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
-%      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
-%      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
+    \item We formalize the problem with memory constraints on heterogeneous processors, where a disk space
+   stores files evicted from memory that will be later used by another processor.
+    \item We design static and dynamic (adaptive?) heuristics
+    \item We implement ...
+    \item We perform extensive simulations,
+    \item In the dynamic setting, we
     \end{itemize}
 
-We first review related work in Section~\ref{sec:related-work}. Then, we formalize the model in Section~\ref{sec:model}
-and the algorithms in Section~\ref{sec:heuristics}.
-The adaptation of the heuristics in a dynamic setting is discussed in Section~\ref{sec:dyn}, and experiment results are presented in Section~\ref{sec:expe}.
-Finally, we conclude
-and provide future working directions in Section~\ref{sec:conc}.
-
-\section{Related work} %: \skug{Full: 5, Polished: 4}}
+Todo: list what we do in each chapter
+
+\section{Related work}
     \label{sec:related-work}
 
-   First, we focus on HEFT-like scheduling heuristics from the literature  that do not necessarily  
-   consider memory constraints. Then, we discuss memory-aware scheduling algorithms. 
-   Finally, we move to related work on dynamic or adaptive algorithms. 
- %   \hmey{Question: Does it make sense to place related work before the model (which is more common)?}
-
-%    We discuss relevant scheduling approaches that reuse processors or respect the memory requirements of the processors.
-%
-%    \subsection{Early list schedulers with unlimited processors}
-%    An entire cluster of works on list schedulers has been carried out as early as the 90s.
-%    They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
-%    with the speed of 1~\cite{benoit2013survey}.\hmey{needs refs or at least a survey pointer}
-%    \skug{actually, Anne cited them, so citing her :-)}
-%
-%    The \textit{task duplication}-based approaches exploit that sometimes running a task twice on different machines can
-%    help reduce the makespan by saving communication costs.
-%    The two categories are scheduling with partial duplication~(SPD), and with full duplication~(SFD).
-%    For a join task (a task whose incoming degree is larger than its outgoing degree), SPD finds a critical immediate
-%    parent (the one that gives the largest start time to the join task) and duplicates only it.
-%    SFD duplicates all parents of a join node.
-%    The algorithm by~\cite{dfrn1997} duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
-%    The critical path fast duplication algorithm CPFD~\cite{5727760} classifies tasks into three categories: critical
-%    path task, in-branch task, or out-branch task.
-%    It schedules critical path tasks first, then in-branch tasks.
-%    \hmey{I'm missing the connection to our paper. Or vice versa, how is our contribution
-%    connected to these works? (Einordnung in den Forschungskontext) For example, do we do similar things? Are there interesting limitations we overcome?}
-%
-%    Linear clustering~\cite{KWOK1999381} acts on critical paths in the workflow.
-%    It assigns the current critical path to one processor, removes all these tasks from the workflow, recomputes the critical
-%    path and repeats the procedure.
-%    Heaviest node first~\cite{SHIRAZI1990222} assigns the tasks level by level;
-%    in each level, it schedules the one with largest computation time first.
-
-    \subsection{Static list schedulers, especially HEFT-based algorithms}
-
-    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
-    It and all its successors consist of two phases: task prioritization/ordering and task assignment.
-    In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
-    and then schedule tasks in the order of these priorities.
-    The modifications of HEFT revolve around the way the priorities of the tasks are computed and the logic of the processor assignment.
-    All such algorithms assume a heterogeneous execution environment.
-
-    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
-    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
-    In the processor choice phase, the entry task and the longest predecessor tasks are duplicated 
-    during idle times on the processor.
-
-%    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
-%    the fastest and the slowest processors, divided by the speed ratio of these two processors.
-%    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
-%    They choose the processor with the lowest execution time and cross over to other processors sometimes.
-%    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}
-
-    PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
-    Cost Table (OCT).
-    The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
-    processor is chosen for this task.
-    Ranking is based on OCT values.
-    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.
-
-    The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
-    comparison to HEFT.
-    It combines the standard deviation with the communication cost weight on the tasks.
-    In the second stage, the algorithm duplicates the entry task if there is a need for it.
-
-    The TSHCS (Task Scheduling for Heterogeneous Computing Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
-    by adding randomized decisions to the second phase.
-    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
-    produces the lowest finish time.
-
-    The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
-    average execution cost in its weight.
-    The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT) that
-    additionally notes how large the communication between current node and its children will be on the
-    average provided that it is scheduled on the current processor.
-
-    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement learning techniques~\cite{yano2022cqga}.
-
-    \subsection{Memory-aware scheduling algorithms}
-    Respecting processor memories adds a constraint to a scheduling problem.
-    Therefore, only specifically memory-targeted algorithms address this issue.
-    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
-    is formulated and addressed in the algorithm.
-%
-    Different models of memory available on processors and memory requirements of tasks have been presented.
-
-    Marchal~et~al.~\cite{marchal2018parallel} assume a memory model where each processor has an individual memory available.
-    Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
-    A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a DAG is provided,
-    as well as an ILP solution to the scheduling problem.
-    The memory model requires deleting all input data upon starting of the task and adding all output files there.
-
-    In an assumed dual-memory systems~\cite{herrmann2014memory},  a processor can have access 
-    to a memory of two different
-    kinds (red or blue), and each task can be executed on only one sort of memory.
-    The communications happen only between these two kinds of processors (communications within 
-    each group are ignored).
-    The authors then formulate an ILP solution for this problem formulation.
-
-    %The algorithm presented by Yao et al.
-    Yao et al.~\cite{yao2022memory} consider that each processor has an own internal memory and all
-    processors share a common external one. The internal (local) memory is used to store the task files.
-    The external memory is used to store evicted files to make room for the execution of a task on a processor.
-    All processors, including the original one, can access these files. 
-    Each edge %in~\cite{yao2022memory} 
-    has two weights -- the size of the files transferred along it,
-    and the time of communication along this edge.
-    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.
-
-    In~\cite{ding2024ils}, there are connected processors with individual limited memories.
-    The collective set of memories forms the global memory, to which each processor has access, however the access time
-    to global memory is different.
-    Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
-    The solved problem is how to allocate initial input data in processor memories so that the overall
-    execution is minimized and the memories are not exceeded.
-    The authors propose an integer linear programming model.
-    %that minimizes the length of the critical path, including a greedy initial solution.
-
-    In~\cite{rodriguez2019exploration}, the authors assume memory requirement on tasks represented as tiles.
-    Each processor has individual memories to process the task, but only the shared memories store the tiles containing
-    memory tiles occupied by memory tiles.
-
-   Finally, there are some cloud-oriented models that include costs associated with memory usage~\cite{liang2020memory}.
-
-   Overall, there are a variety of memory models, but, to the best of our knowledge, the only study on a multiprocessor
-   platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24},
-   but where a partition of the workflow is proposed, hence preventing processor reuse. 
-   Hence, in ~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need of communication buffers to store data 
-   that should be communicated
-   between two processors when tasks are ready to execute.
-
-    \subsection{Dynamic/adaptive algorithms}
-
-We now review related work in a dynamic setting. With no variation in task parameters, 
-    DVR HEFT~\cite{SANDOKJI2019482} rather considers that new tasks arrive in the system. 
-    They use an almost unchanged HEFT algorithm in the static step, executing three slightly
-    varying variants of task weighting and choosing the variant that gives the best overall makespan.
-    In the dynamic phase, they receive new tasks and schedule them on either idle processors or 
-    those processors that give them
-    the earliest finish time.
-    %Task failures are not covered.
-
-    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
-    by calculating the critical path in the graph dynamically at every step.
-    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
-    %on the start time of a task (differing by the slack this task has).
-    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
-    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
-    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
-    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
-    %execution time for the task is chosen.
-    The heuristic also uses the same processor to schedule predecessor and successor tasks, as to avoid data transfer between processors.
-    The approach is evaluated on random workflows of the size up to 300 tasks.
-
-    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
-    The procedure involves building a first (static) schedule, periodic resource monitoring and rescheduling the remaining
-    tasks.
-    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
-    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
-    tasks to groups.
-    The static heuristic is HEFT with earliest start time as priority.
-    Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
-    is smaller than the previous one.
-    The experiments were conducted on a single workflow with 10 tasks.
-%
-    % The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
-    %The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.
-
-    %The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
-    %THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
-    %They then go through tasks and map than to such processors that minimize the earliest start times of this task's
-    %successors.
-    %To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
-    %average communication and computation costs fir the dependent tasks.
-    %
-    %The rescheduling is triggered when either a load on a resource increases over a threshold, or if a new resource
-    %is added.
-
-    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
-    but constrained by cost.
-
-    Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
-    Particles are possible solution in the solution space.
-    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
-    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.
-
-    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling algorithm
-    for clouds.
-    They solve a set of linear equations that represent the cost of an execution based on the criteria.
-    The authors test out 4 scenarios - one preferring each criteria, and a balanced one.
-    The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
-    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
-    %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
- %   The authors use provenance data to make scheduling decisions.
-
-    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
-    on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
-    that these runtime changes are unrelated to each other).
-    The authors formulate a decision space of all permutations of $n$ jobs, and the optimal schedule in relation to a
-    performance measure $\phi$.
-    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
-
-While several related work consider building a new schedule once some variation has been observed,
-we are not aware of work implementing a real runtime system that interacts with the scheduler,
-and tested on workflows with thousands of tasks, as we propose in this paper. Furthermore, 
-we are not aware of any previous work discussing dynamic algorithms combined with memory constraints. 
-%\AB{Say why our approach in dynamic setting is different and novel}
-
-%    \subsection{Other notable works}
-%
-%    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
-%    They utilize task duplication when creating the clusters (grains).
-%    Their scheduler then maps clusters to processors.
-%    They assume unlimited processors with the speed of 1.
-%    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
-%    to it as possible.
-%    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
-%    %release times.
-%
-%    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
-%    to search for an optimal solution for mapping tasks on machines.
-%    A solution is generated at each step, and the best solution is kept at the end.
-%    The search terminates when a certain termination criterion is reached.
-%    It generates better results than other algorithms, because it explores the whole solution space.
-%
-%    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
-%    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
-%    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
-%    The heuristic first determines similar activities and groups them together.
-%    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
-%    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
-%    The heuristic calculates a new execution path with these tasks.
-%
-%
-%    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
-%    (processor) speed heterogeneity.
-%    Each accelerator has its own memory that can be zero.
-%    Each accelerator's characteristics depend on the task it runs and are not fixed.
-%
-%
-%
-%
-%    \cite{meng2018traffic} investigate scheduling on multi-core chips.
-%    Their model is far from ours.
-%
-%
-%    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
-%    They prioritize tasks that have failed before or are well-predictable.
-%
-%
-
-\section{Model} %: \skug{Full: 4, polished: 3}}
+   Todo
+
+\section{Model}
 \label{sec:model}
-
-%\skug{
-%CHANGES IN THIS CHAPTER
-%}
-    We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
-    in Section~\ref{sec.mod.work}.  Next, we define the execution
-    environment, a heterogeneous system (in terms of processor speed and memory size),
-    in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
-    Section~\ref{sec.mod.pb}. The key notation is summarized in Table~\ref{tabnotation}.
 
@@
 \label{sec.mod.work}
-    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
+
+    Todo: this section holds. Needs rewriting to avoid plagiarism accusatons. Also insert other references from related work
+
+Our workflow is a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
 $E$ is a set of directed edges of the form $e=(u,v)$, with \mbox{$u,v\in V$}, expressing precedence constraints between tasks.
@@
 
-%     the total memory requirement  for the execution of task~$u$ consists of the maximum
-%    between the input files
-%    (total size of the files to be received from the parents),
-%    the output files (total size of the files to be sent to the children),
-%    and the total memory size~$m_u$ (usually achieving the maximum):
-%    \[
-%        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
-%    \]
-
 The predecessors of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of predecessors is
@@
 Furthermore, we place ourselves in a context where we do not have perfect knowledge
-of the length of the tasks, i.e., $w_u$, % ($w_u$ and $m_u$) 
+of the length of the tasks, i.e., $w_u$,
 before the tasks complete their execution,
 but only estimates~\cite{rahman2013,GARG2015256}.
-%. \AB{Add motivation: related work with variable task durations for instance...}
+
 Hence, scheduling decisions are made on these estimated parameters, and
 may be reconsidered at runtime when a task completes its execution.
-%when a task starts its execution and we know its exact parameters.
-%\AB{Variability only on $w_u$ for now}
 
@@
 
-%\skug{changes here}
+    Todo: this section holds and is new (in parts).
+
 The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
@@
 
-%, but of much slower speed.
 All communications happen over the disk -- after one processor has finished writing data there, another one can read
@@
 This data can later be read back into memory of the same or any other processor at any time.
-%\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).} \AB{We can keep two copies: evict if we need to }
 
@@
 then task can still be executed there, but it will be slowed down. Indeed,
-%    albeit slower.
+
 the part of memory requirement that exceeds the available memory on processor~$j$ for task~$u$, 
-denoted by $\Moff{u}$, %M\text{off}_u$, 
+denoted by $\Moff{u}$,
 will be offloaded to the disk, slowing down the execution:
@@
     \[
-        w_{real} = \frac{w_u}{s_j} +  q_j \times  \Moff{u}, %  (m_u - Mav_u) , %   \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
+        w_{real} = \frac{w_u}{s_j} +  q_j \times  \Moff{u},
     \]
-where $q_j$ is the slowdown coefficient for $p_j$. %, and $(m_u - Mav_u)$ is the part 
-%of the task memory that needs to be offloaded.
-\AB{Svetlana please check, I think we said in the end that we use absolute value 
-of the amount of memory not fitting locally...}
-%    \skug{check this equation}
-%We assume that all processors are connected with the same bandwidth~$\beta$.
-%    \hmey{Maybe mention that variable bandwidths are part of future work...?}
+where $q_j$ is the slowdown coefficient for $p_j$.
 
@@
 
-We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
+\textcolor{orange}{ TODO: Move the next paragraph to Algorithms.} We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
 the memory requirement; it transforms the workflow into a series-parallel graph
@@
 \AB{When do we need this? Probably more in Algorithms, right? }
-
-%\skug{Todo: describe deviations}
 

### 2025-03-10 main.tex
minor: commented keywords back in

@@
 
-%    \begin{IEEEkeywords}
-%        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
-%    \end{IEEEkeywords}
+    \begin{IEEEkeywords}
+        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
+    \end{IEEEkeywords}
 
@@
 to specific processors.
-\new{A task schedule - \iec a valid execution order that respects the dependencies,  and possibly also starting times for
-the tasks are needed to increase utilization by reusing finished processors.}
+\new{Moreover, a task schedule -- \iec a valid execution order that respects the dependencies --
+and possibly also starting times for the tasks are needed to increase utilization by reusing 
+processors that have become idle after finishing a task.}
 
@@
 that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm.
-M behind HEFT \new{stands for} for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
+M behind HEFT stands for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
 BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
@@
 Marchal~et~al.~\cite{marchal2018parallel} assume a memory model where each processor has an individual memory available.
-Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
+\new{Workflow tasks themselves have no memory requirements for their computations,
+but they have input and output files that need to be stored in the memory.}
 A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a workflow DAG is provided,
@@
 %
-In~\cite{ding2024ils}, the authors consider connected processors with individual limited memories
-forming a global memory, with different access times to memory, and no weights on edges. An ILP
+Ding \etal~\cite{ding2024ils}, in turn, consider connected processors with individual limited memories
+forming a global memory, with different access times to memory and no weights on edges. An ILP
 is proposed in this setting. 
@@
 Overall, there are a variety of memory models, but, to the best of our knowledge, the only study on a multiprocessor
-platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24}.
-Yet, they propose only a partitioning-based mapping of the workflow, without processor reuse.
-\new{Because a processor is idle after a partition of a workflow has finished executing}, in~\cite{DBLP:conf/icpp/KulaginaMB24},
-there is no need for communication buffers to store data
-that should be \new{exchanged} between two processors when tasks are ready to execute.
+platform that is fully heterogeneous, with individual memories, \new{is our own previous work~\cite{DBLP:conf/icpp/KulaginaMB24}.}
+\new{Yet, in that paper we only propose} a partitioning-based mapping of the workflow, without processor reuse.
+\new{Since a processor is idle after a workflow partition has finished its execution,
+there is no need in Ref.~\cite{DBLP:conf/icpp/KulaginaMB24} for communication buffers to store data
+that should be \new{exchanged} between processors when tasks are ready to execute.}
 
@@
     is smaller than the previous one.
+    \todo[inline]{Maybe mention: from scratch computation because of tiny workflows??}
 %    The experiments were conducted on a single workflow with 10 tasks.
@@
 for clouds.
-    They solve a set of linear equations that represent the cost of an execution based on the criteria.
+They solve a set of linear equations\todo{Mis\-lea\-ding, TBD!} that represent the cost of an execution based on the criteria.
 The authors test four scenarios -- one preferring each criterion and a balanced one.
 The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
-    The authors used workflows with less than ten tasks, but repeated them so that the execution had up to 200 tasks.
+The authors use workflows with less than ten tasks, but repeat them so that the execution has up to 200 tasks.
 %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
@@
 on a single machine.
-    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related
+The changes in task running times are not due to changing machine properties, but are rather task-related
 and thus unrelated to each other.
@@
 in relation to a performance measure. % $\phi$.
-    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.
+Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem \new{with} a set of linear constraints.
 
@@
 we are not aware of any previous work discussing dynamic algorithms combined with memory constraints.
-%\AB{Say why our approach in dynamic setting is different and novel}
 
@@
 
-\begin{table}[b]
+\begin{table}[tb]
     \begin{center}
@@
 is the mean and the deviation is~$10\%$.
-This scenario corresponds to the real-life scenarios identified in Ref.~\cite{lotaru} and
+This scenario corresponds to the real-world scenarios identified in Ref.~\cite{lotaru} and
 other works dedicated to predicting task running times~\cite{da2015online,da2013toward}.
@@
   \centering
-  \includegraphics[width=1.08\columnwidth] {images/success-rates-large2}
-  \vspace{-0.8cm}
-  \caption{Success rates \new{by workflow size and algorithm} on the default cluster. Higher is better.}
+  \includegraphics[width=1.0\columnwidth] {images/success-rates-large2}
+%  \vspace{-0.8cm}
+  \caption{Success rates by workflow size and algorithm on the default cluster. Higher is better.}
   \label{fig:success-rates-large}
-  \vspace{-0.15cm}
+%  \vspace{-0.15cm}
 \end{figure}
@@
 by the memory-aware heuristics, normalized to the
-makespan achieved by \heft, often with an invalid over-optimistic schedule that exceeds
-the bound on memory.
+makespan achieved by \heft, \new{the latter} often with an invalid over-optimistic schedule that exceeds
+the memory bound.
 %    Figure~\ref{fig:ms-relations-by-workflow} shows the relations of the makespans found by the heuristics to the
@@
 tasks, and up to $27\%$ worse for the largest workflows.
-The makespans of \heftblc are very similar,being $<17\%$ worse on smallest workflows, $19$-$28\%$ worse on middle-sized,
+The makespans of \heftblc are very similar, being $<17\%$ worse on smallest workflows, $19$-$28\%$ worse on middle-sized,
 and up to $30\%$ worse on the largest workflows.
 The makespans found by \heftmm are $25$-$30\%$ worse on the smallest workflows, $40$-$70\%$ on middle-sized and up to
-$210\%$ on the largest ones. These are still very encouraging results,
-in particular for \heftbl and \heftblc, since the makespans of \heft correspond to invalid schedules.
-}
+$210\%$ on the largest ones.} These are still very encouraging results,
+in particular for \heftbl and \heftblc, since the makespans of \heft\ \new{often} correspond to invalid schedules.
+
  %   \skug{TODO: how much worse are makespans for SUCCESSFUL hefts? presulmably almost identical}
@@
   \includegraphics[width=1\columnwidth] {images/ms_relations_by_wf_size_barplot}
-    \vspace{-0.8cm}
+%    \vspace{-0.8cm}
     \caption{Relative makespans of heuristics normalized by \heft makespan, by workflow
@@
     \label{fig:ms-relations-by-workflow}
-    \vspace{-0.15cm}
+%    \vspace{-0.15cm}
 \end{figure}
@@
     \includegraphics[width=1\columnwidth] {images/mem-usage-normal2}
-    \vspace{-0.6cm}
+%    \vspace{-0.6cm}
     \caption{Memory usage on default cluster, including invalid \heft schedules. }
@@
      \label{fig:mem-usages-normal}
-    \vspace{-0.15cm}
+%    \vspace{-0.15cm}
 \end{figure}
@@
     \includegraphics[width=1\columnwidth] {images/mem-usage-normal-onlyvalid2}
-    \vspace{-0.6cm}
+%    \vspace{-0.6cm}
     \caption{Memory usage on default cluster, considering only valid \heft schedules. }
     \label{fig:mem-usages-onlyvalid}
-    \vspace{-0.15cm}
+%    \vspace{-0.15cm}
 \end{figure}
@@
 %
-On the memory-constrained cluster, \heft produces valid assignments in only $14$ out of $290$ experiments
-$290$~($4.8\%$ success rate).
+On the memory-constrained cluster, \heft produces valid assignments in only $14$ out of $290$ experiments ($4.8\%$ success rate).
 The successful schedules are achieved exclusively on the tiny workflows (with only two 200-task size-increased
@@
     \label{fig:success-rates-tiny}
-    \vspace{-0.15cm}
+%    \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:ms-relations-by-workflow-constrained}
-    \vspace{-0.15cm}
+%    \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:memory-usage-constrained}
-    \vspace{-0.15cm}
+%    \vspace{-0.15cm}
 \end{figure}
@@
     \label{fig:updates-ms}
-    \vspace{-0.3cm}
+%    \vspace{-0.3cm}
 \end{figure}
@@
     \label{fig:runtimes-log}
-    \vspace{-0.3cm}
+%    \vspace{-0.3cm}
 \end{figure}
@@
 
+\medskip
+
 \subsubsection*{Acknowledgements}

### 2025-03-10 main-disk-model.tex
further cleaned main-disk-model

@@
 on the local memory of the processor (once it has been generated or read, and before 
-it is used or evicted). 
+it is used or evicted). \skug{not sure if we need this interval?}
 \end{itemize}
@@
 \AB{I'm confused again about the offloading model, we need to talk}
-
-%\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
-%I'll probably add a few sentences here...}
-% while respecting memory constraints. \skug{do we still formulate it like that?}
-    %If a processor runs out of memory to execute
-%a task mapped on it, the schedule is said to be {\em invalid}.
 
@@
 Hence, we focus on the design of efficient scheduling heuristics.
-%\hmey{Mention NP-hardness due to being more general than NP-hard problem?}
-
-%    \paragraph{Workflow-related changes}
-%
-%    \begin{itemize}
-%        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
-%        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
-%
-%    \end{itemize}
-%
-%    The following changes are not a part of this article's scope:
-%
-%    \begin{itemize}
-%        \item The workflow structure changes: edges or tasks come in or leave.
-%    \end{itemize}
-%
-%    \paragraph{Execution environment-related changes }
-%
-%
-%    \begin{itemize}
-%        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
-%        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
-%
-%    \end{itemize}
-%
-%    The following changes are not a part of this article's scope:
-%
-%    \begin{itemize}
-%        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
-%    \end{itemize}
-%
-%    \subsection{Time of changes }
-%
-%    We consider discrete time in seconds.
-%    The time point(s) at which the changes happen is unambiguously defined.
-%
-%    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
-%    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
-%    The start time of the source task in the workflow is zero.
-%    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
-%
-%    \subsection{Changes and knowledge horizon - important questions TBA}
-%
-%    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
-%
-%    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
-%    However, how do we model the following:
-%    \begin{itemize}
-%        \item Do we know the new weights of currently running tasks and tasks that have not yet started?
-This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
-%        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
-%        \item How many times of change do we model - one per workflow run, or multiple?
-%        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
-%    \end{itemize}
-
-\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
+
+\section{Scheduling heuristics}
 \label{sec:heuristics}
-\subsection{\skug{New scheduling approach - general description}}
+
+\subsection{Offline setting}
 
@@
 
-\subsection{Data structures }
-    To optimally represent the timeline of the execution of a workflow, we represent the execution as a series of events
-    with attached unique timestamps.
-    The events are kept in the priority queue ordered by timestamp.
-    Events are created when other events are triggered and are being inserted into the queue.
-    When their timestamp is the smallest, they are then triggered, possibly creating new events.
-    The execution starts with placing the event of starting the starting node into the queue with the timestamp 0.
-    The execution ends when there are no events in the queue.
-
-    An event can happen on finish or start of a task execution or an io action (file read or write).
-    Each event has several parameters:
-    \begin{itemize}
-        \item Event type: OnTaskStart, OnTaskFinish, OnIoStart, OnIoFinish
-        \item The workflow entity associated with the event: the task to be executed or an edge whose file needs to be read or written
-        \item The processor id on which the action should take place
-        \item Expected and actual time of firing.
-        \item A set of predecessor events. The event is called ready if all its predecessors have been finished.
-        \item A set of successor events to keep track of.
-        \item For a write event, a boolean if this file should be removed from memory after being written.
-    \end{itemize}
-
-    Per processor,  in addition to keeping the fixed values (memory, speed), we keep track also of the current state in its memory.
-    Because during and after execution of each even the state of the processor memory changes, we need to keep snapshots of what files
-    were in memory and how much of it was available.
-    Because keeping such snapshots for each timestap is too complicated, we only keep two timestamps.
-    First, we keep the information of available memory and tasks pending in memory during the execution of the last
-    task scheduled to this processor.
-    Then, we keep the same information on what it will be immediately after this task finishes.
-
-    \subsubsection{On Task Finish}
-    When a task finishes, it first cleans after itself:
-    \begin{itemize}
-        \item It frees its memory: it increases the available memory on the processor by the amount it occupied
-        \item If there had been a memory overflow onto disk, then this memory is being freed implicitly
-        \item it sets its status to finished in the workflow.
-    \end{itemize}
-    It the goes over its successor events (Tasks and IO actions) and removes itself from their predecessor lists.
-    It also updates their actual trigger times with its actual finish time.
-    Successor tasks can be tasks scheduled on the same processor after this one.
-    Successor io events can be reads that could not start until the task finishes and frees its memory.
-    \skug{can writes be successor events of a task?}
-
-    Then task goes over its successor tasks in the workflow and schedules ready ones.
-    Ready tasks are those whose all predecessors have the finished status.
-
-    For each ready task, the scheduler tries to tentatively put it on each processor in the cluster and chooses the processor
-    that promises the earliest finish time.
-    The finish time on each processor is calculated as follows:
-
-    \paragraph{Earliest start and finish time}
-    We find out when the last task on this processor finishes.
-    If there is enough memory to start our task after that, then the end of previous task is the earliest start time for our task.
-    If files need to be evicted first, we need to find a balance between evicting tasks (and waiting for their writes to finish),
-    then running our task, and starting the task immediately and letting it run longer due to memory overflow.
-    We compare the following cases:
-    \begin{itemize}
-        \item Evict everything, then start task
-        \item Greedily evict the largest file, starting the task
-        \item Not evicting anything, starting the task
-    \end{itemize}
-    Whatever combination gives the best estimated finishing time for the task, we note the estimated starting time of the task in this case.
-     The finish time is the earliest start time  plus computation time of the task.
-
-    \paragraph{Incoming edges} For each incoming edge, its location is determined.
-    If the edge is in the memory of the target processor, nothing needs to be done.
-    The estimated start and finish times for the task are unaffected.
-
-    If the edge is on the disk, it first needs to be read into the memory of the target processor.
-    We delay the reads to avoid holding unnecessary files in the memory.
-    When we preliminarily schedule a read, we first assess its length and other actions that happen on this processor in the meantime.
-    So, first estimated start of the read is \skug{the maximum between the desired start time (estimated start time of the task minus the estimated length of the read),
-    and the feasible one (when the lastest read event on this processor finishes)}.
-
-    We then see if this start of the read is happening during the runtime of the previous task or already after it (during the writes).
-    If the starting time of the read is during the execution of the previous task and there is enough memory to read this file, then we move the read
-    forward to the beginning of this previous task.
-    If there is not enough memory, then the entire read is rescheduled for after the task finishes.
-    The read will then be bottleneck and affect the eraliest start time of the desired task.
-
-     If the edge is in another processor's memory, it needs to be first written to disk, then read by the target processor.
-    So we look at the latest write on that processor, preliminarily schedule the write to disk and then proceed like in the
-    previous case, reading the file from disk.
-
-   \paragraph{When processor is chosen} When a processor is chosen, the scheduler inserts new events into the queue:
-    \begin{itemize}
-        \item For the task, OnTaskStart and OnTaskFinish events with the estimated trigger times, the OnTaskFinish event being dependent on the OnTaskStart event.
-        \item For each io operation necessary for the execution: OnIoStart event of the corresponding type and on corresponding processor
-        \item It connects all the event with predecessor-successor relations.
-        \item If files needed to be evicted
-    \end{itemize}
-    The scheduler also updates the ready times of the processors and updates the new currently last events there.
-
-    \subsubsection{On Task Start}
-    When a task start event is triggered, it first determines if it can start.
-    If some predecessor tasks have not yet finished (the predecessor set is not empty), then the actual time of this event
-    is set to the maximum of the estimated end times of predecessor events and this event is reinserted in the queue
-    with this new time.
-    \skug{We  also update the actual times of all its successor events.}
-
-    If the task can start, the scheduler determines the deviation of the runtime of the task.
-    It applies the deviation function on its estimated runtime (estimated finish minus estimated start time).
-    It then adds this deviated runtime to its actual start time and receives an updated actual finish time.
-    With this time, it updates the OnTaskFinish event of this task.
-
-    \subsubsection{OnIoFinish}
-    When the io operation finishes, it removes itself from the predecessor sets of all its successor events.
-    It also sets their actual time to its own actual finish time.
-
-    \subsubsection{OnIOStart}
-    When the IO operation starts, it updates the corresponding IoFinish event
-
-    After being triggered and completed, each event is deleted from the queue.
-
---------------------------------------
-
-%    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
-%    but favor processor reuse as is done in HEFT.
-%    Furthermore, this would allow us to handle variability on the fly, by updating
-%    the bottom levels if some parameters vary, and computing the schedule
-%    only for the near future...
-
-%In order to be able to easily adapt to variability of task parameters, 
 We design 
@@
 
+    \subsection{Online setting }
+    To optimally represent the timeline of the execution of a workflow, we represent the execution as a series of events
+    with attached unique timestamps.
+    The events are kept in the priority queue ordered by timestamp.
+    Events are created when other events are triggered and are being inserted into the queue.
+    When their timestamp is the smallest, they are then triggered, possibly creating new events.
+    The execution starts with placing the event of starting the starting node into the queue with the timestamp 0.
+    The execution ends when there are no events in the queue.
+
+    An event can happen on finish or start of a task execution or an io action (file read or write).
+    Each event has several parameters:
+    \begin{itemize}
+        \item Event type: OnTaskStart, OnTaskFinish, OnIoStart, OnIoFinish
+        \item The workflow entity associated with the event: the task to be executed or an edge whose file needs to be read or written
+        \item The processor id on which the action should take place
+        \item Expected and actual time of firing.
+        \item A set of predecessor events. The event is called ready if all its predecessors have been finished.
+        \item A set of successor events to keep track of.
+        \item For a write event, a boolean if this file should be removed from memory after being written.
+    \end{itemize}
+
+    Per processor,  in addition to keeping the fixed values (memory, speed), we keep track also of the current state in its memory.
+    Because during and after execution of each even the state of the processor memory changes, we need to keep snapshots of what files
+    were in memory and how much of it was available.
+    Because keeping such snapshots for each timestap is too complicated, we only keep two timestamps.
+    First, we keep the information of available memory and tasks pending in memory during the execution of the last
+    task scheduled to this processor.
+    Then, we keep the same information on what it will be immediately after this task finishes.
+
+    \subsubsection{On Task Finish}
+    When a task finishes, it first cleans after itself:
+    \begin{itemize}
+        \item It frees its memory: it increases the available memory on the processor by the amount it occupied
+        \item If there had been a memory overflow onto disk, then this memory is being freed implicitly
+        \item it sets its status to finished in the workflow.
+    \end{itemize}
+    It the goes over its successor events (Tasks and IO actions) and removes itself from their predecessor lists.
+    It also updates their actual trigger times with its actual finish time.
+    Successor tasks can be tasks scheduled on the same processor after this one.
+    Successor io events can be reads that could not start until the task finishes and frees its memory.
+    \skug{can writes be successor events of a task?}
+
+    Then task goes over its successor tasks in the workflow and schedules ready ones.
+    Ready tasks are those whose all predecessors have the finished status.
+
+    For each ready task, the scheduler tries to tentatively put it on each processor in the cluster and chooses the processor
+    that promises the earliest finish time.
+    The finish time on each processor is calculated as follows:
+
+    \paragraph{Earliest start and finish time}
+    We find out when the last task on this processor finishes.
+    If there is enough memory to start our task after that, then the end of previous task is the earliest start time for our task.
+    If files need to be evicted first, we need to find a balance between evicting tasks (and waiting for their writes to finish),
+    then running our task, and starting the task immediately and letting it run longer due to memory overflow.
+    We compare the following cases:
+    \begin{itemize}
+        \item Evict everything, then start task
+        \item Greedily evict the largest file, starting the task
+        \item Not evicting anything, starting the task
+    \end{itemize}
+    Whatever combination gives the best estimated finishing time for the task, we note the estimated starting time of the task in this case.
+    The finish time is the earliest start time  plus computation time of the task.
+
+    \paragraph{Incoming edges} For each incoming edge, its location is determined.
+    If the edge is in the memory of the target processor, nothing needs to be done.
+    The estimated start and finish times for the task are unaffected.
+
+    If the edge is on the disk, it first needs to be read into the memory of the target processor.
+    We delay the reads to avoid holding unnecessary files in the memory.
+    When we preliminarily schedule a read, we first assess its length and other actions that happen on this processor in the meantime.
+    So, first estimated start of the read is \skug{the maximum between the desired start time (estimated start time of the task minus the estimated length of the read),
+        and the feasible one (when the lastest read event on this processor finishes)}.
+
+    We then see if this start of the read is happening during the runtime of the previous task or already after it (during the writes).
+    If the starting time of the read is during the execution of the previous task and there is enough memory to read this file, then we move the read
+    forward to the beginning of this previous task.
+    If there is not enough memory, then the entire read is rescheduled for after the task finishes.
+    The read will then be bottleneck and affect the eraliest start time of the desired task.
+
+    If the edge is in another processor's memory, it needs to be first written to disk, then read by the target processor.
+    So we look at the latest write on that processor, preliminarily schedule the write to disk and then proceed like in the
+    previous case, reading the file from disk.
+
+    \paragraph{When processor is chosen} When a processor is chosen, the scheduler inserts new events into the queue:
+    \begin{itemize}
+        \item For the task, OnTaskStart and OnTaskFinish events with the estimated trigger times, the OnTaskFinish event being dependent on the OnTaskStart event.
+        \item For each io operation necessary for the execution: OnIoStart event of the corresponding type and on corresponding processor
+        \item It connects all the event with predecessor-successor relations.
+        \item If files needed to be evicted
+    \end{itemize}
+    The scheduler also updates the ready times of the processors and updates the new currently last events there.
+
+    \subsubsection{On Task Start}
+    When a task start event is triggered, it first determines if it can start.
+    If some predecessor tasks have not yet finished (the predecessor set is not empty), then the actual time of this event
+    is set to the maximum of the estimated end times of predecessor events and this event is reinserted in the queue
+    with this new time.
+    \skug{We  also update the actual times of all its successor events.}
+
+    If the task can start, the scheduler determines the deviation of the runtime of the task.
+    It applies the deviation function on its estimated runtime (estimated finish minus estimated start time).
+    It then adds this deviated runtime to its actual start time and receives an updated actual finish time.
+    With this time, it updates the OnTaskFinish event of this task.
+
+    \subsubsection{OnIoFinish}
+    When the io operation finishes, it removes itself from the predecessor sets of all its successor events.
+    It also sets their actual time to its own actual finish time.
+
+    \subsubsection{OnIOStart}
+    When the IO operation starts, it updates the corresponding IoFinish event
+
+    After being triggered and completed, each event is deleted from the queue.
+
+    --------------------------------------
+
     In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.

### 2025-03-10 main.tex
minor: implemented one Henning's todo
temp

@@
 \newcommand{\hmey}[1]{{\color{red}[HM: #1]}}
-\newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
-\newcommand{\willchange}[1]{{\color{orange}[AB: #1]}}
+\newcommand{\AB}[1]{\todo[inline]{{\color{purple}[AB: #1]}}}
 
@@
 \IEEEauthorblockN{Anne Benoit}
-\IEEEauthorblockA{\textit{ENS Lyon, France} \\
+\IEEEauthorblockA{\textit{ENS Lyon \& IUF, France} \\
 ORCID: 0000-0003-2910-3540
@@
 The memory model \new{requires deleting all input files when the task starts and adding all output files to memory.}
+\AB{Is it not when task completes? We need input files while processing the task, correct?}
 
@@
 \new{Yet, in that paper we only propose} a partitioning-based mapping of the workflow, without processor reuse.
-\new{Since a processor is idle after a workflow partition has finished its execution,
-there is no need in Ref.~\cite{DBLP:conf/icpp/KulaginaMB24} for communication buffers to store data
+\new{Since a processor is idle after a workflow partition has finished its execution 
+in~\cite{DBLP:conf/icpp/KulaginaMB24},
+there is no need for communication buffers to store data
 that should be \new{exchanged} between processors when tasks are ready to execute.}
@@
 %    The static heuristic is HEFT with earliest start time as priority.
-    Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
+    Upon rescheduling, a new mapping is computed from scratch, 
+    and this mapping is accepted if the resulting makespan
     is smaller than the previous one.
-    \todo[inline]{Maybe mention: from scratch computation because of tiny workflows??}
-%    The experiments were conducted on a single workflow with 10 tasks.
+    \new{Only small workflows with less than 100 tasks are considered, hence it is possible to do
+    these repeated computations from scratch in a reasonable time. }
+%    The repeated recomputation from scratch is feasible, because the largest workflow size tested is 100 task big.}
+%    \todo[inline]{Maybe mention: from scratch computation because of tiny workflows??}
+%%    The experiments were conducted on a single workflow with 10 tasks.
 %

### 2025-03-10 main.tex
vspaces

--- 2025-03-10 main.tex
@@
     \begin{IEEEkeywords}
-        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
+        DAG, Heterogeneous platform, Adaptive \new{workflow} scheduling, Memory constraint.
     \end{IEEEkeywords}
@@
   \centering
+  \vspace{-0.4cm}
   \includegraphics[width=1.0\columnwidth] {images/success-rates-large2}
-%  \vspace{-0.8cm}
   \caption{Success rates by workflow size and algorithm on the default cluster. Higher is better.}
@@
     \centering
+  \vspace{-0.34cm}
   \includegraphics[width=1\columnwidth] {images/ms_relations_by_wf_size_barplot}
-%    \vspace{-0.8cm}
     \caption{Relative makespans of heuristics normalized by \heft makespan, by workflow
@@
 \new{Due to a varying nature of the workflow generator, some larger workflows can have smaller makespans than their
-smaller counterparts due to more parallel execution due to higher fanout or other factors of the workflow's
-internal composition.
-}
+smaller counterparts due to more parallelism at nodes with higher out-degree or other factors of the workflow's
+internal composition.\todo{TBD: Why here?}}
 
@@
 
-Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained}.
+Memory usage on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained}.
 \new{We plot the average memory usage in the cluster per algorithm and workflow size},
@@
     \centering
+    \vspace{-0.3cm}
     \includegraphics[width=1\columnwidth] {images/ms_relations_by_wf_size-constrained-barplot}
@@
     \label{fig:ms-relations-by-workflow-constrained}
-%    \vspace{-0.15cm}
 \end{figure}
@@
 %    \paragraph{Memory-constrained cluster}
-The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
+The makespan in case of no recomputation\todo{TBD: clear?} becomes invalid as soon as at least one task finds itself in an invalid memory size
 situation -- that is, if the scheduler assumed the task's memory to be smaller than actually needed and
@@
     \includegraphics[width=0.95\columnwidth] {images/UpdatesMss2}
-    \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm \new{in case of no recomputation}. Smaller is better.}
+    \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm\ \new{in case of no recomputation}. Smaller is better.}
     \label{fig:updates-ms}

### 2025-03-10 main.tex
temp

--- 2025-03-10 main.tex
@@
 Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160 
-for all algorithm variants end with a valid makespan without recomputation.
-In case of \heft, $14$ valid initial makespans were computed.
-Out of them, $13$ remain valid until the end, after all the update requests.
+for all algorithm variants \new{obtain a valid makespan when no recomputation is done}.
+
+In the case of \heft, $14$ valid initial makespans were computed.
+Out of them, $13$ remain valid until the end, after all the update requests,
+\new{i.e., when accounting for the final execution time of each task and their memory requirement.}
 The same $13$ experiments end with a valid makespan in case of no recomputation.
-So, these workflows require so few resources that there is no point in re-scheduling them.
+So, these workflows require so few resources that there is \new{nothing to be gained by re-scheduling them.}
 %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
@@
 %only, the success rate of not recomputing is $92\%$.
-In case of \heftmm, all $290$ workflows can be scheduled initially and all the schedules remain valid until the end.
+
+In case of \heftmm, all $290$ workflows can be scheduled initially and all the schedules remain valid until the end,
+\new{when new schedules are computed in a dynamic way. Only}  
 $16$ experiments end with a successful schedule without recomputation, a rate of~$5.5\%$.
+
 \heftblc produces $142$ valid initial schedules and keeps $141$ of them valid until the end.
 $50$ experiments are successful without recomputation, a rate of $35\%$ out of all successful final schedules.
+
 \heftbl keeps $105$ schedules valid until the end out of $110$ initial valid schedules.
@@
 
-\heftbl and \heftblc are successful on smaller workflows and fail on larger ones, 
-the success of the strategy without recomputation is limited to even the smallest of these smaller workflows.
+\heftbl and \heftblc are successful on smaller workflows and fail on larger ones. 
+\new{Note that the success of the strategy without recomputation is limited to  the smallest  workflows,
+while a dynamic approach is mandatory when dealing with larger workflows.}
 Thus, the strategy without recomputation delivers valid makespans (independently of the algorithm) 
-on $56$ original workflows with $<100$ tasks, $47$ $200$-task ones, $25$ $1000$-task ones, and $6$ $2000$-task ones.
+on $56$ original workflows with $<100$ tasks, \new{$47$ workflows with  $200$ tasks, 
+$25$ workflows with $1000$ tasks, and only $6$ workflows with $2000$ tasks.
+The most representative case for the dynamic experiments is hence \heftmm, since it succeeds
+to achieve valid schedules in all cases thanks to recomputation, while most schedules
+cannot be successful when no recomputation is done. 
+}
 
@@
 The larger variations for \heft can be explained by the small amount of data -- for instance, there are
-only $2$ $200$-task workflows for this case.
+only \new{two workflows with  $200$ tasks in this case.}
 
@@
     \includegraphics[width=0.95\columnwidth] {images/UpdatesMss2}
-    \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm \new{in case of no recomputation}. Smaller is better.}
+    \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm  %\new{ 
+    when there is no recomputation. Smaller is better.}
     \label{fig:updates-ms}

### 2025-03-10 main.tex
minor fix

--- 2025-03-10 main.tex
@@
 $210\%$ on the largest ones.} These are still very encouraging results,
-in particular for \heftbl and \heftblc, since the makespans of \heft\ \new{often} correspond to invalid schedules.
+in particular for \heftbl and \heftblc, since the makespans of \heft\ \new{nearly always} correspond to invalid schedules.
 

### 2025-03-10 main.tex
implemented todos until (not including) section 6.C

@@
 \usepackage{todonotes}
-\usepackage{graphicx}
 \usepackage{booktabs}
@@
 as well as an integer linear programming (ILP) solution to the scheduling problem.
-The memory model \new{requires deleting all input files when the task starts and adding all output files to memory.}
-\AB{Is it not when task completes? We need input files while processing the task, correct?}
+The memory model \new{ includes no memory requirement of the task itself, only the weights of incoming and outgoing edges.
+When the task starts, all input files are deleted  and all output files are added to memory.}
 
@@
 
-De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling algorithm
-for clouds.
-They solve a set of linear equations\todo{Mis\-lea\-ding, TBD!} that represent the cost of an execution based on the criteria.
-The authors test four scenarios -- one preferring each criterion and a balanced one.
-The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
+De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) \new{cost model and an adaptive
+ scheduling algorithm for clouds (scheduling virtual machines).
+The greedy algorithm based on this cost model  works in 4 steps, choosing the best resources type to execute the virtual
+machine on, producing new cloud activities, setting up the granularity factor, and finally adapting the amount of resources
+to fit the budget.}
+The authors test four scenarios -- one preferring each criterion in the cost model and a balanced one.
 The authors use workflows with less than ten tasks, but repeat them so that the execution has up to 200 tasks.
@@
 These values align with the findings of Ref.~\cite{lotaru} about small tasks.
-%
- %   \skug{insert more abt how the runtime system hanfles data reads}
+
+\new{Due to a varying nature of the workflow generator, some workflows can have smaller makespans than their
+smaller counterparts due to more parallelism at nodes with higher out-degree or other factors of the workflow's
+internal composition.}
 
@@
 the memory usage of schedule.
-\new{Due to a varying nature of the workflow generator, some larger workflows can have smaller makespans than their
-smaller counterparts due to more parallelism at nodes with higher out-degree or other factors of the workflow's
-internal composition.\todo{TBD: Why here?}}
+\new{Note how the relative makespan does not grow steadily with the size of the workflows.
+Rather, it grows at the beginning, but then falls for the sizes of workflows where scheduling becomes problematic.
+This indicates that \heft schedules become relatively worse for larger workflows, in addition to being unfeasible memory-wise.
+Finally, the relative makespan for the largest workflows, the ones for which only \heftmm can produce valid schedules,
+is growing larger again.
+This is due to the highest degree of memory adaptation necessary to produce a valid schedule.
+}
 
@@
     \includegraphics[width=0.95\columnwidth] {images/UpdatesMss2}
-    \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm  %\new{ 
-    when there is no recomputation. Smaller is better.}
+    \caption{
+    \new{Relative makespan improvement of \heftbl, \heftblc, \heftmm and baseline \heft against
+    the execution with the same heuristic, but without recomputation. Higher is better. } }
     \label{fig:updates-ms}

### 2025-03-11 main.tex
described execution with and without recomputation. Minor edits in Section 6.C

@@
 
+\new{We call this execution mode execution \textit{with recomputation}.
+Alternatively, the runtime system can execute \textit{without recomputation}.
+It then follows the original schedule without sending the deviation values to the scheduler.
+
+In this case, the runtime system waits for a blocked processor to become available if another task scheduled to it takes longer than expected.
+If the predecessor task takes longer to execute, the runtime system leaves the processor of the current task
+temporarily idle.
+If any tasks take less time then expected, the runtime system ignores the opportunity to start other tasks earlier.
+Finally, if there is not enough memory, the schedule is declared invalid and the execution stops.
+
+The comparison to the makespan of the execution without recomputation shows how much improvement the exploration of dynamic
+changes brings.
+Moreover, the number of schedules that manage to stay valid despite the changes is an indicator of their robustness and the quality of the scheduler.
+}
+
 \subsection{Results in a static setting}
@@
 \label{sec.expe.dyn}
-
-%\AB{Need to add results on the default cluster!}
-%
-%    \paragraph{Memory-constrained cluster}
-The makespan in case of no recomputation\todo{TBD: clear?} becomes invalid as soon as at least one task finds itself in an invalid memory size
+\new{In this section, we compare the makespans of executions with and without recomputation.}
+The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
 situation -- that is, if the scheduler assumed the task's memory to be smaller than actually needed and
@@
 The same $13$ experiments end with a valid makespan in case of no recomputation.
-So, these workflows require so few resources that there is \new{nothing to be gained by re-scheduling them.}
+So, these workflows require so few resources that
+\new{the devations do not invalidate their scheduler. However, there is also nothing to be gained by re-scheduling them.}
 %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success

### 2025-03-11 main.tex
temp

@@
 
-    \title{Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\
+    \title{Memory-aware Adaptive Scheduling of Scientific Workflows on Heterogeneous Architectures\\
 %{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 platform that is fully heterogeneous, with individual memories, \new{is our own previous work~\cite{DBLP:conf/icpp/KulaginaMB24}.}
-\new{Yet, in that paper we only propose} a partitioning-based mapping of the workflow, without processor reuse.
+\new{Yet, in that paper, we only propose} a partitioning-based mapping of the workflow, without processor reuse.
 \new{Since a processor is idle after a workflow partition has finished its execution 
@@
  scheduling algorithm for clouds (scheduling virtual machines).
-The greedy algorithm based on this cost model  works in 4 steps, choosing the best resources type to execute the virtual
+The greedy algorithm based on this cost model  works in four steps, choosing the best resource types to execute the virtual
 machine on, producing new cloud activities, setting up the granularity factor, and finally adapting the amount of resources
@@
 in relation to a performance measure. % $\phi$.
-Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem \new{with} a set of linear constraints.
+Then, they proceed to formulate the Absolute Deviation Robust Scheduling Problem \new{with} a set of linear constraints.
 
@@
 
-Thus, the scheduler receives deviation values and makes decision based on them.
+Thus, the scheduler receives deviation values and makes decision based on them\new{, hence
+leading to an execution mode  \textit{with recomputation}.}
 This leads to several types of possible issues:
@@
 
-\new{We call this execution mode execution \textit{with recomputation}.
-Alternatively, the runtime system can execute \textit{without recomputation}.
+\new{Alternatively, the runtime system can execute \textit{without recomputation}.
 It then follows the original schedule without sending the deviation values to the scheduler.
@@
 temporarily idle.
-If any tasks take less time then expected, the runtime system ignores the opportunity to start other tasks earlier.
+If any task takes less time than expected, the runtime system ignores the opportunity to start other tasks earlier.
 Finally, if there is not enough memory, the schedule is declared invalid and the execution stops.
@@
 changes brings.
-Moreover, the number of schedules that manage to stay valid despite the changes is an indicator of their robustness and the quality of the scheduler.
-}
+Moreover, the number of schedules that manage to stay valid despite the changes is an indicator of their robustness and of the quality of the scheduler.}
 

### 2025-03-11 main.tex
moved paragraph in Sec. IV-A, some minor fixes

@@
 
+\new{Note that, due to a varying nature of the workflow generator, some workflows can have smaller makespans than their
+smaller counterparts due to more parallelism at nodes with higher out-degree or other factors of the workflow's
+internal composition.}
+
 \paragraph{Task and edge weights}
@@
 These values align with the findings of Ref.~\cite{lotaru} about small tasks.
-
-\new{Due to a varying nature of the workflow generator, some workflows can have smaller makespans than their
-smaller counterparts due to more parallelism at nodes with higher out-degree or other factors of the workflow's
-internal composition.}
 
@@
 
-Thus, the scheduler receives deviation values and makes decision based on them\new{, hence
-leading to an execution mode  \textit{with recomputation}.}
-This leads to several types of possible issues:
+Thus, the scheduler receives deviation values and makes decisions based on them\new{, which
+leads to an execution mode  \textit{with recomputation}.}
+This mode leads to several types of possible issues:
 %

### 2025-03-11 main.tex
end of VI-A

@@
 
-In this case, the runtime system waits for a blocked processor to become available if another task scheduled to it takes longer than expected.
-If the predecessor task takes longer to execute, the runtime system leaves the processor of the current task
-temporarily idle.
-If any task takes less time than expected, the runtime system ignores the opportunity to start other tasks earlier.
+In this scenario, the runtime system waits with task execution for a busy processor to become available if another task scheduled on it takes longer than expected.
+If, however, the predecessor task is faster than expected, the runtime system leaves the processor of the current task temporarily idle.
 Finally, if there is not enough memory, the schedule is declared invalid and the execution stops.

### 2025-03-11 main.tex
VI-C

@@
 \new{Note how the relative makespan does not grow steadily with the size of the workflows.
-Rather, it grows at the beginning, but then falls for the sizes of workflows where scheduling becomes problematic.
+Rather, it grows at the beginning, but then falls for the workflow sizes where scheduling becomes very challenging.
 This indicates that \heft schedules become relatively worse for larger workflows, in addition to being unfeasible memory-wise.
@@
 \label{sec.expe.dyn}
-\new{In this section, we compare the makespans of executions with and without recomputation.}
+\new{In this section, we try to assess the advantage of the dynamic approach.
+To this end, we compare the makespans of executions with and without recomputation.}
 The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
@@
 The same $13$ experiments end with a valid makespan in case of no recomputation.
-So, these workflows require so few resources that
-\new{the devations do not invalidate their scheduler. However, there is also nothing to be gained by re-scheduling them.}
+Hence, these workflows require so few resources that
+\new{the devations do not invalidate their scheduler. However, there is also nothing to be gained by rescheduling them.}
 %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
@@
 
-In case of \heftmm, all $290$ workflows can be scheduled initially and all the schedules remain valid until the end,
+In case of \heftmm, all $290$\todo{SK: check!} workflows can be scheduled initially and all the schedules remain valid until the end,
 \new{when new schedules are computed in a dynamic way. Only}  
@@
 
-\heftbl keeps $105$ schedules valid until the end out of $110$ initial valid schedules.
+\heftbl keeps $105$ schedules valid until the end out of $110$ initially valid schedules.
 $55$ remain valid until the end even without recomputation, roughly $50\%$ of all successes.
@@
 
-Figure~\ref{fig:updates-ms} shows the increase in makespan in case of no recomputation for these experiments.
+Figure~\ref{fig:updates-ms} shows \new{the self-relative improvement in makespan when doing recomputation 
+compared to no recomputation taking place} for these experiments.
 With growing size of the workflow, the excess makespan of not recomputing grows -- from $13.9\%$ to $20\%$ for \heftbl,
@@
 
+\new{
+These results indicate that a meaningful comparison proves difficult due to
+many invalid schedules. However, they still show that doing recomputation 
+is beneficial for both makespan and the number of valid solutions.
+}
+
 \begin{figure}[tb]
@@
     \caption{
-    \new{Relative makespan improvement of \heftbl, \heftblc, \heftmm and baseline \heft against
+    \new{Self-relative makespan improvement of \heftbl, \heftblc, \heftmm and baseline \heft against
     the execution with the same heuristic, but without recomputation. Higher is better. } }

### 2025-03-11 main.tex
rewrote the text about successes of with/without recomputation in VI.C

@@
 Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160
-for all algorithm variants \new{obtain a valid makespan when no recomputation is done}.
+(290 experiments for each of the 3 heuristics and the baseline \heft)
+ \new{obtain a valid makespan when no recomputation is done}.
 
 In the case of \heft, $14$ valid initial makespans were computed.
-Out of them, $13$ remain valid until the end, after all the update requests,
-\new{i.e., when accounting for the final execution time of each task and their memory requirement.}
-The same $13$ experiments end with a valid makespan in case of no recomputation.
+Out of them, $13$ remain valid \new{until the end of an execution with recomputation},
+\new{i.e., the schedule is still valid when accounting for the final execution time of each task and their memory requirement.}
+The same $13$ experiments end \new{valid in case of no recomputation, i.e. when following the initial schedule without ever recomputing it}.
 Hence, these workflows require so few resources that
-\new{the devations do not invalidate their scheduler. However, there is also nothing to be gained by rescheduling them.}
+\new{the devations do not invalidate their schedules. However, there is also nothing to be gained by rescheduling them.}
 %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
@@
 
-In case of \heftmm, all $290$\todo{SK: check!} workflows can be scheduled initially and all the schedules remain valid until the end,
-\new{when new schedules are computed in a dynamic way. Only}  
+In case of \heftmm, all $290$ workflows can be scheduled initially \new{(static case)} and all these schedules remain valid until the end,
+\new{when new schedules are computed in a dynamic way. However, only}
 $16$ experiments end with a successful schedule without recomputation, a rate of~$5.5\%$.
 
-\heftblc produces $142$ valid initial schedules and keeps $141$ of them valid until the end.
-$50$ experiments are successful without recomputation, a rate of $35\%$ out of all successful final schedules.
-
-\heftbl keeps $105$ schedules valid until the end out of $110$ initially valid schedules.
+\heftblc produces $142$ valid initial schedules and keeps $141$ of them valid until the end \new{when new schedules are computed in a dynamic way.}
+$50$ experiments \new{remain} successful \new{when following the original schedule} without recomputation, a rate of $35\%$ out of all successful final schedules.
+
+\heftbl keeps $105$ schedules valid until the end out of $110$ initial \new{(static)} valid schedules.
 $55$ remain valid until the end even without recomputation, roughly $50\%$ of all successes.

### 2025-03-11 main.tex
temp

@@
 
-\noindent{\bf HEFT-based algorithms. }
+%\noindent{\bf 
+\subsection{HEFT-based algorithms}
 %\subsection{Static list schedulers, especially HEFT-based algorithms}
@@
 
-\medskip
-%\subsection
-\noindent{\bf Memory-aware scheduling algorithms. }
+%\medskip
+\subsection{
+%\noindent{\bf 
+Memory-aware scheduling algorithms}
 %\label{sub:mem-aware-algs}
@@
 
-%    \subsection
-\smallskip
-\noindent{\bf Dynamic/adaptive algorithms. }
+ \subsection{%
+%\bigskip
+%\noindent{\bf 
+Dynamic/adaptive algorithms}
 We finally review related work in a dynamic setting. With no variation in task parameters, 
@@
 Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160
-(290 experiments for each of the 3 heuristics and the baseline \heft)
- \new{obtain a valid makespan when no recomputation is done}.
-
-In the case of \heft, $14$ valid initial makespans were computed.
-Out of them, $13$ remain valid \new{until the end of an execution with recomputation},
+(290 experiments for each of the three heuristics and the baseline \heft)
+ \new{lead to a valid makespan when no recomputation is done}.
+
+In the case of \heft, $14$ valid initial schedules were computed.
+Out of them, $13$ remained valid \new{until the end of an execution with recomputation},
 \new{i.e., the schedule is still valid when accounting for the final execution time of each task and their memory requirement.}
-The same $13$ experiments end \new{valid in case of no recomputation, i.e. when following the initial schedule without ever recomputing it}.
-Hence, these workflows require so few resources that
-\new{the devations do not invalidate their schedules. However, there is also nothing to be gained by rescheduling them.}
+The same $13$ experiments also led to \new{valid schedules in case of no recomputation, i.e., when following the initial schedule without ever recomputing it}. 
+Indeed, these workflows required so few resources that
+\new{the deviations did not invalidate their schedules. Also, there is also nothing to be gained by rescheduling them.}
 %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
@@
 In case of \heftmm, all $290$ workflows can be scheduled initially \new{(static case)} and all these schedules remain valid until the end,
-\new{when new schedules are computed in a dynamic way. However, only}
+\new{when new schedules are computed in a dynamic way (with recomputation). However, only}
 $16$ experiments end with a successful schedule without recomputation, a rate of~$5.5\%$.
@@
 many invalid schedules. However, they still show that doing recomputation 
-is beneficial for both makespan and the number of valid solutions.
+is beneficial for both the makespan and the number of valid solutions.
 }

### 2025-05-15 CCGRID Presentation/main.tex
added presentation slides

\documentclass[xcolor=svgnames,titlepage,english,presentation]{beamer}
\usetheme{HUB}

\makeatletter
\let\HUB@realinsertinstitute\insertinstitute
\makeatother

\setbeamercovered{invisible}
\usepackage[noborder,colbullets]{beamerMACSYdefs}
\usepackage[normalem]{ulem}
% -----------------------------------------------------------------------------
% -- packages
\usepackage{framed}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{bbding}
\usepackage{booktabs}
\usepackage{adjustbox}
\usefonttheme[onlymath]{serif}
\usetikzlibrary{patterns,positioning}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{positioning}
\usepackage[absolute,overlay]{textpos}

\HUBfaculty{HU Berlin}

\graphicspath{{./images/}}

\setbeamertemplate{title page}[mydefault]

% -----------------------------------------------------------------------------

\title{Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures}
\subtitle{}

\author[S. Kulagina]{\underline{Svetlana Kulagina$^1$}, Anne Benoit$^2$ and Henning Meyerhenke$^{1,3}$}

%\institute{
%   \begin{minipage}{\linewidth}
%        \centering
%%        $^1$: Department of Computer Science, Humboldt-Universit\"at zu Berlin, Germany\\
%       $^2$: LIP, ENS Lyon, France\\
%        $^3$: Institute of Informatics, Humboldt-Universit\"at zu Berlin, Germany
%   \end{minipage}
%}

%\parbox{\linewidth}{
%    $^1$: Department of Computer Science, Humboldt-Universit\"at zu Berlin, Germany  \\
%%    $^2$: Department of Computer Science, Humboldt-Universit\"at zu Berlin, Germany \\
%    $^3$: Department of Computer Science, Humboldt-Universit\"at zu Berlin, Germany
%}

%\TitleImage[width=0.3\textwidth]{diagrams/cropped-1024px-FONDA_LogoLogo.jpg}

\begin{document}

\setlength\textheight{7cm} %required for correct vertical alignment, if [t] is not used as documentclass parameter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\makeatletter
\begin{frame}[plain]
   % \setbeamertemplate{frametitle}{}
    \setbeamercolor{background canvas}{bg=white}

    % ——— recreate HUB frame & logos ———
    \begin{tikzpicture}[remember picture,overlay]
        \fill[white] (current page.north west) rectangle (current page.south east);
        \draw[HUBrframe,line width=\HUB@RFramewd]
        (current page.north west) rectangle (current page.south east);

        % Top-right HUB logo
        \node[anchor=north east,inner sep=8pt] at (current page.north east)
            {\pgfimage[width=25\K@mm]{\HUB@logo}};
        % Bottom-right MACSy logo
        \node[anchor=south east,inner sep=8pt] at (current page.south east)
            {\pgfimage[width=25\K@mm]{\MACSY@logo}};
    \end{tikzpicture}

    % ——— content ———
    \vfill
    \begin{center}
    {\usebeamerfont{title}\usebeamercolor[fg]{title}\color{HUBblue}\LARGE\bfseries
    Memory-aware Adaptive Scheduling of Scientific Workflows\\
    On Heterogeneous Architectures\par}
                \vskip1.5em
            {\usebeamerfont{author}\usebeamercolor[fg]{author}\small
        \underline{Svetlana Kulagina$^1$}, Anne Benoit$^2$ and Henning Meyerhenke$^{1,3}$\par}
        \vskip1.5em

        % Custom institute box
        \setlength{\fboxsep}{0.8em}
        \colorlet{myBG}{HUBblue!15}
        \colorlet{myFG}{HUBblue!80}
        \fcolorbox{myFG}{myBG}{%
            \parbox{0.75\textwidth}{\centering
                $^1$ Humboldt‑Universität zu Berlin, Germany\\
                $^2$ LIP, ENS Lyon and IUF, France\\
                $^3$ Karlsruhe Institute of Technology (KIT), Germany
            }%
        }
        \vskip2em

        % Title image
        \usebox{\HUB@titimg}
    \end{center}
    \vfill

\end{frame}
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Motivation}

    \begin{itemize}
        \item Large applications represented as DAG-shaped workflows
        \item Heterogeneous execution environment with limited memories
         \begin{itemize}
                  \item Exceed memory = expensive or even fatal
        \end{itemize}
        \item Execute the workflow fast
    \end{itemize}

\begin{columns}
    \column{0.59\textwidth}
    \begin{figure}
            \centering
            \includegraphics[scale=0.25]{diagrams/images/methylseq-sparse.png}

            %\caption{Caption }
        \end{figure}
    \column{0.39\textwidth}
    \only<1>{
     \begin{figure}
            \centering
            \includegraphics[scale=0.3]{diagrams/images/CCGrid-muffin.png}

            %\caption{Caption }
        \end{figure}
    }
    \only<2>{
     \begin{figure}
            \centering
            \includegraphics[scale=0.3]{diagrams/images/CCgrid-muffin-no.png}

            %\caption{Caption }
        \end{figure}
    }

\end{columns}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Contributions}

    3 variants of a HEFT-based scheduler: HEFTM-\textbf{BL}, HEFTM-\textbf{BLC}, and HEFTM-\textbf{MM}

    ~~~~~~~~~~~~~~~~~

    \begin{itemize}
        \item Respects processor memory sizes
        \begin{itemize}
            \item HEFTM-\textbf{BL} and HEFTM-\textbf{BLC} produce valid makespans that are only $10$-$35\%$ worse than invalid HEFT makespans
            \item HEFTM-\textbf{MM} able to schedule all workflows even under extreme memory constraint
        \end{itemize}

        \item + adaptive strategy for scenario with uncertainty

        \begin{itemize}
            \item Adapting the schedule leads to $>20\%$ makespan improvement even on smallest workflows
        \end{itemize}

    \end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{Model: Workflow}

    ~~~~~

      \begin{columns}

        \column{0.4\textwidth}
         \begin{itemize}
             \item Tasks have \textcolor{orange}{memory} and \textcolor{magenta}{workload} weights
             \item Edges have (file) weights 
         \end{itemize}

        \column{0.4\textwidth}

       \only<1>{   \raggedleft{
         \begin{figure}
            \centering
            \includegraphics[scale=0.4]{diagrams/images/weights-noparts-rand.png}      
        \end{figure}
        }}
        \only<2>{   
        \raggedleft{
         \begin{figure}
            \centering
            \includegraphics[scale=0.4]{diagrams/images/CCGRID-workflow-weights-for-ru.png}      
        \end{figure}
        }}

    \end{columns}

    ~~~~~~~~~~~

 \only<2>{
\begin{figure}
                \centering
                \includegraphics[width=0.5\linewidth]{diagrams/images/ru_formula_ccgrid_colors.png}                
                \label{fig:enter-label}
      \end{figure}

      }

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{Model: Execution Environment}

    ~~~~~

      \begin{columns}
      \column{0.05\textwidth}
        \column{0.4\textwidth}
         \begin{itemize}
             \item Each processor has

             \begin{itemize}
                 \item a limited memory of size $M_j$ and a communication buffer $MC_j$
                 \item speed $s_j$
             \end{itemize}

             \uncover<3,4>{
             \item Keep track of ready times on each
                  \begin{itemize}
                      \item Communication channel: $rt_{j,j'}$
                      \item Communication buffer:  $rt_j$
                  \end{itemize}  
                  }
         \end{itemize}

        \column{0.4\textwidth}
        \raggedleft{
         \begin{figure}
            \centering
            \includegraphics[scale=0.35]{diagrams/images/ClusterCCGRID.png}

            %\caption{Caption }
        \end{figure}

        \uncover<2,3,4>{
            \begin{figure}
            \centering
            \includegraphics[scale=0.4]{diagrams/images/clusterCommunications.png}

            %\caption{Caption }
        \end{figure}
        }
        }
        \column{0.1\textwidth}
    \end{columns}

    ~~~~~~~~~~~

\uncover<4>{
   Objective: makespan minimization with memory constraints
   }

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{HEFT (in our model)}

  \begin{columns}
     \column{0.05\textwidth}
        \column{0.65\textwidth}
  HEFT~[1]:

  \begin{itemize}
      \item Assign each task a rank 
      \item Schedule tasks in decreasing rank order
     \only<2>{ \begin{itemize}
          \item Tentatively assign the task to each processor
          \item Choose the one with shortest finish time
      \end{itemize}
      }
  \end{itemize}

 \column{0.34\textwidth}
  \begin{tikzpicture}
   % \draw (4,2) ellipse (2cm and 1cm) node {What about memory?};
  \end{tikzpicture}

       % \column{0.1\textwidth}
    \end{columns}

\uncover<2>{ 
  \begin{figure}
      \centering
        \includegraphics[scale=0.33]{diagrams/images/HEFT.png}
         %\caption{Caption }
  \end{figure}
  \vspace{-0.4cm}
  }

\footnotesize{[1] H. Topcuoglu, S. Hariri, M.-Y. Wu. Performance-effective and
low-complexity task scheduling for heterogeneous computing. 2002.}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{HEFT (in our model)}

  \begin{columns}

    \column{0.05\textwidth}
        \column{0.65\textwidth}
  HEFT~[1]:

  \begin{itemize}
      \item Assign each task a rank 
      \item Schedule tasks in decreasing rank order
      \begin{itemize}
          \item Tentatively assign the task to each processor
          \item Choose the one with shortest finish time
      \end{itemize}
  \end{itemize}

 \column{0.34\textwidth}
  \begin{tikzpicture}
    \draw (4,2) ellipse (2cm and 1cm) node {What about memory?};
  \end{tikzpicture}

       % \column{0.1\textwidth}
    \end{columns}

  \begin{figure}
      \centering
        \includegraphics[scale=0.33]{diagrams/images/HEFT.png}
         %\caption{Caption }
  \end{figure}

\vspace{-0.4cm}
\footnotesize{[1] H. Topcuoglu, S. Hariri, M.-Y. Wu. Performance-effective and
low-complexity task scheduling for heterogeneous computing. 2002.}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{Memory Traversal}

    Fit into limited memory $\rightarrow$ use as little memory as possible

    ~~~~~~~~~~~~~~

  Peak memory depends on the traversal

 \begin{columns}
      \column{0.05\textwidth}
        \column{0.6\textwidth}
     \begin{itemize}            
               \item Traversal $1 \rightarrow 3 \rightarrow 2$ 
             \begin{itemize}
                \item Execution of 3: $\textcolor{blue}{1000} (c_{1,2})+ \textcolor{blue}{10} (m_3) = 1011$
                \item Execution of 2: $\textcolor{blue}{1000} (m_2) + \textcolor{blue}{1} (c_{1,2}) = 1001$
            \end{itemize}
            \item Traversal $1 \rightarrow 2 \rightarrow 3$ 
            \begin{itemize}
                \item Execution of 2: $\textcolor{blue}{1000} (m_2)+ \textcolor{blue}{1} (c_{1,2}) + \textcolor{blue}{1000} (c_{1,3}) = 2001$
            \end{itemize}

            \item Memory-optimal traversal \textsc{MemDag} [1]

        \end{itemize}

      \column{0.3\textwidth}
        \raggedleft{
        %\vspace{-2.0cm}
         \begin{figure}
            \centering
            \includegraphics[scale=0.4]{diagrams/images/MemTraversal.png}
            %\caption{Caption }
        \end{figure}
        }
        \column{0.05\textwidth}
    \end{columns}

%\vfill
~~~~~~~~~~~~~

 \footnotesize {[1] Kayaaslan, E., Lambert, T., Marchal, L., U{\c{c}}ar, B. (2018). Scheduling series-parallel task graphs to minimize peak memory. Theoretical Computer Science, 707, 1-23.}

 ~~~~~~
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{Scenario and Model: Eviction}

      When a task executes, it

         \begin{itemize}
             \item consumes (deletes from memory) its incoming files,
             \item creates (inserts into memory) its outgoing files.

             \end{itemize}

  To fit a task into memory, we may need to evict pending memories into the communication buffer

          \begin{figure}
            \centering
            \includegraphics[scale=0.4]{diagrams/images/evictionCCGRID.png}    
        \end{figure}

No returning tasks back into memory!

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{HEFTM-* heuristics: Ranking}

    Similar 2 phases to HEFT: rank and assign. \\[0.5ex]
    Implementation of the phases different.

 ~~~~~
\pause

  ~~~~~

    \begin{itemize}
        \item HEFTM-\textbf{BL}: orders by non-increasing bottom levels
        \item HEFTM-\textbf{BLC}: same, but prioritizes tasks with large incoming communications \\ 
            \begin{figure}
            \centering
            \includegraphics[scale=0.2]{diagrams/images/BLC-formula.png}
            %\caption{Caption }
        \end{figure}

        \item HEFTM-\textbf{MM}: orders according to \textsc{MemDag}'s memory-optimal traversal

    \end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{HEFTM-* heuristics: Task Assignment}
    ~~~~~

 For each task $v$, \textbf{tentatively} try it on each processor $p_j$

  \begin{itemize}
      \item Check that for all predecessors assigned to $p_j$ the data is still in memory $\rightarrow$ otherwise invalid choice
      \item Address the memory constraint on the processor \\

       \begin{figure}
            \centering
            \includegraphics[width=.7\textwidth]{diagrams/images/Res-formula}
        \end{figure}
  \end{itemize}

    \begin{center}
        \uncover<2>{
            \includegraphics[scale=0.4]{diagrams/images/CCgridEvict1}
        }
    \end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{HEFTM-* heuristics: Task Assignment}
    ~~~~~

    Greedily assign each task $v$ to processor $p_j$ that minimizes finish time.

    \begin{itemize}
        \item Check that for all predecessors assigned to $p_j$ the data is still in memory $\rightarrow$ otherwise invalid choice
        \item Address the memory constraint on the processor \\

        \begin{figure}
            \centering
            \includegraphics[width=.7\textwidth]{diagrams/images/Res-formula}
        \end{figure}
    \end{itemize}

    \begin{center}
            \includegraphics[scale=0.4]{diagrams/images/CCgridEvict2}
    \end{center}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Start and Finish Time Calculation}

    % --- Put background image using TikZ ---
    \begin{tikzpicture}[remember picture, overlay]
        \node[anchor=south east, xshift=0cm, yshift=1.5cm] at (current page.south east) {
            \includegraphics[width=0.55\paperwidth]{diagrams/images/StartTime.png}
        };
    \end{tikzpicture}

     \vspace{-0.6cm}
       % --- Foreground content (text appears over image) ---
    \begin{itemize}
        \item Start after all communications have finished
        \begin{figure}
            \centering
            \hskip -5.5cm
            \includegraphics[width=.6\textwidth]{diagrams/images/formula-starttime-paperwithrts.png}
        \end{figure}

    \end{itemize}

    \begin{columns}
        \column{0.02\textwidth}
        \column{0.35\textwidth}
        \begin{itemize}
            \item  Finish time
            \begin{align*}
                FT(v, p_j) &= \\
                &ST(v, p_j) + \frac{w_v}{s_j}
            \end{align*}
        \end{itemize}
        \column{0.6\textwidth}
        \column{0.08\textwidth}
    \end{columns}

    \begin{overlayarea}{\paperwidth}{\paperheight}
            \begin{tikzpicture}[remember picture, overlay]
                \draw[red, thick]
                ([xshift=-9.8cm,yshift=7.5cm]current page.south east)
                ellipse (0.5cm and 0.5cm);
            \end{tikzpicture}
    \end{overlayarea}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Start and Finish Time Calculation}

    % --- Put background image using TikZ ---
    \begin{tikzpicture}[remember picture, overlay]
        \node[anchor=south east, xshift=0cm, yshift=1.5cm] at (current page.south east) {
            \includegraphics[width=0.55\paperwidth]{diagrams/images/StartTime.png}
        };
    \end{tikzpicture}

    \vspace{-0.6cm}
    % --- Foreground content (text appears over image) ---
    \begin{itemize}
        \item Start after all communications have finished
        \begin{figure}
            \centering
            \hskip -5.5cm
            \includegraphics[width=.6\textwidth]{diagrams/images/formula-starttime-paperwithrts.png}
        \end{figure}

    \end{itemize}

    \begin{columns}
        \column{0.02\textwidth}
        \column{0.35\textwidth}
        \begin{itemize}
            \item  Finish time
            \begin{align*}
                FT(v, p_j) &= \\
                &ST(v, p_j) + \frac{w_v}{s_j}
            \end{align*}
        \end{itemize}
        \column{0.6\textwidth}
        \column{0.08\textwidth}
    \end{columns}

    \begin{overlayarea}{\paperwidth}{\paperheight}
        \begin{tikzpicture}[remember picture, overlay]
            \draw[red, thick]
            ([xshift=-8.4cm,yshift=6.8cm]current page.south east)
            ellipse (0.5cm and 0.5cm);
        \end{tikzpicture}
    \end{overlayarea}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Start and Finish Time Calculation}

    % --- Put background image using TikZ ---
    \begin{tikzpicture}[remember picture, overlay]
        \node[anchor=south east, xshift=0cm, yshift=1.5cm] at (current page.south east) {
            \includegraphics[width=0.55\paperwidth]{diagrams/images/StartTime.png}
        };
    \end{tikzpicture}

    \vspace{-0.6cm}
    % --- Foreground content (text appears over image) ---
    \begin{itemize}
        \item Start after all communications have finished
        \begin{figure}
            \centering
            \hskip -5.5cm
            \includegraphics[width=.6\textwidth]{diagrams/images/formula-starttime-paperwithrts.png}
        \end{figure}

    \end{itemize}

    \begin{columns}
        \column{0.02\textwidth}
        \column{0.35\textwidth}
        \begin{itemize}
            \item  Finish time
            \begin{align*}
                FT(v, p_j) &= \\
                &ST(v, p_j) + \frac{w_v}{s_j}
            \end{align*}
        \end{itemize}
        \column{0.6\textwidth}
        \column{0.08\textwidth}
    \end{columns}

    \begin{overlayarea}{\paperwidth}{\paperheight}
        \begin{tikzpicture}[remember picture, overlay]
            \draw[red, thick]
            ([xshift=-7.0cm,yshift=6.8cm]current page.south east)
            ellipse (1.0cm and 0.5cm);
        \end{tikzpicture}
    \end{overlayarea}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Final Assignment}

    Final assignment to processor

    \begin{itemize}
        \item Evict files that need to be evicted from memory,
        \begin{itemize}
            \item  remove them from pending memory and add to communication buffer
            \item reduce buffers sizes accordingly
        \end{itemize}
        \item New $availM_j$: - eviction, - incoming files, + outgoing files
        \item For each predecessor on another processor $j'$: ready time on communication buffer $rt_{j, j'}$

    \end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Experimental Setup}

    ~~~~~~~~~~~~~~~~~~~~~~~

    Schedule real (scientific) workflows \\[0.35ex]

    Task and edge weights: historical data from Lotaru [6]

    ~~~~~~~~~~~~~~~~~~~

        \begin{itemize}  
        \item Workflows
        \begin{itemize}
             \item Real-world workflows
            \item Generated with WFGen [7] from 7 workflow: sizes from 1K to 30K tasks
            \item Divided in groups by size
        \end{itemize}

        \item Execution environments
         \begin{itemize}
             \item 36 processors: 6 pieces of same 6 kinds of processors as in [6]
            \item Memory-constrained: same processors, 10 times less memory
        \end{itemize} 
    \end{itemize}

~~~~~~~~~~~~~~~~~~~~~~~~

\footnotesize{[6] Bader, J., Lehmann, F., Thamsen, L., Will, J., Leser, U., and Kao, O. 2022. Lotaru: Locally Estimating Runtimes of Scientific Workflow Tasks in Heterogeneous Clusters. SSDBM.}

\footnotesize{[7] Coleman, T., Casanova, H., Pottier, L., Kaushik, M., Deelman, E., and da Silva, R. F. 2022. WfCommons: A framework for enabling scientific
workflow research and development. Future Generation Computer Systems.}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Results}
    \framesubtitle{Default cluster - Success Rates}

    \vspace{-0.8cm}
    \begin{figure}    
        \centering
         \hskip -0.5cm
        \includegraphics[scale=0.38]{diagrams/images/success-rates-large2.png}
        % \caption{Makespans of HomPart in comparison to HetPart1.}
    \end{figure}

    % Overlay rectangle on top
    \begin{tikzpicture}[remember picture, overlay]
        \node[
            fill=white,
            draw=black,
            rounded corners,
            minimum width=4cm,
            minimum height=1.5cm,
            align=center
        ] at ([xshift=-5cm,yshift=-4.5cm]current page.north east)
        {HEFT invalid over 2000 tasks \\ \\ HEFTM-* valid for all workflows };
    \end{tikzpicture}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Results}
    \framesubtitle{Default cluster - Relative Makespans}

      \begin{textblock*}{5cm}(7cm,4cm) % {block width} (coords)
        %HEFT fails, HEFTM-* valid
    \end{textblock*}

     \vspace{-0.2cm}
        \begin{figure}
            \centering
            \hskip -0.7cm
            \includegraphics[scale=0.36]{diagrams/images/ms_relations_by_wf_size_barplot.png}
          %  \caption{Runtimes of HetPart2 in comparison to HetPart 1. }
        \end{figure}

    \begin{tikzpicture}[remember picture, overlay]
            \node[
                fill=white,
                draw=black,
                rounded corners,
                minimum width=4cm,
                minimum height=1.5cm,
                align=center
            ] at ([xshift=-5cm,yshift=-3.8cm]current page.north east)
            {HEFTM-* makespans valid, HEFT not \\ HEFTM-BL and BLC : 10 - 35\% worse \\ HEFTM-MM makespans 40\% to 2x worse};
        \end{tikzpicture}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Results}
    \framesubtitle{Memory-Constrained Cluster: Success Rates}

      \vspace{-0.3cm}
        \begin{figure}
            \centering
            \hskip -0.5cm
            \includegraphics[scale=1.25]{diagrams/images/success-rates-tiny.png}
           % \caption{Makespans of HomPart in comparison to HetPart1.}
        \end{figure}

          \begin{tikzpicture}[remember picture, overlay]
        \node[
            fill=white,
            draw=black,
            rounded corners,
            minimum width=4cm,
            minimum height=1.5cm,
            align=center
        ] at ([xshift=-3.5cm,yshift=-5.5cm]current page.north east)
        {HEFT invalid over 100 tasks \\ HEFTM-BL / HEFTM-BLC invalid over  \\8000 tasks
        \\ HEFTM-* valid for all workflows };
    \end{tikzpicture}

    \begin{tikzpicture}[remember picture, overlay]
        \node[
            fill=white,
            draw=black,
            rounded corners,
            minimum width=3.0cm,
            minimum height=1.5cm,
            align=center
        ] at ([xshift=-3.7cm,yshift=-1.8cm]current page.north east)
            {\textcolor{red}{\parbox{2.9cm}{Less realistic, \\extreme strain for\\ the algorithm}}};
    \end{tikzpicture}

   \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Results}
    \framesubtitle{Memory-Constrained Cluster: Relative Makespans}
         \vspace{-0.8cm}
        \begin{figure}
            \centering
            \hskip -0.8cm
            \includegraphics[scale=0.48]{diagrams/images/ms_relations_by_wf_size-constrained-barplot.png}
          %  \caption{Runtimes of HetPart2 in comparison to HetPart 1. }
        \end{figure}

    \begin{tikzpicture}[remember picture, overlay]
        \node[
            fill=white,
            draw=black,
            rounded corners,
            minimum width=4cm,
            minimum height=1.5cm,
            align=center
        ] at ([xshift=-5.0cm,yshift=-4.5cm]current page.north east)
        {Most makespans invalid, \\
        even by HEFTM-BL and HEFTM-BLC; \\ HEFTM-MM makespans the only valid ones };
    \end{tikzpicture}
   \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[t]
    \frametitle{Results}
    \framesubtitle{Runtime}

    \makebox[0pt][l]{\hspace{-1.0cm}%
        \includegraphics[scale=0.26]{diagrams/images/runtimes-logarithmic.png}
    }

    \begin{tikzpicture}[remember picture, overlay]
        \node[
            fill=white,
            draw=black,
            rounded corners,
            minimum width=4.8cm,
            minimum height=1.7cm,
            align=center
        ] at ([xshift=-4.5cm,yshift=-4.5cm]current page.north east)
        {Runtimes from 1/100 of seconds to minutes\\
         HEFTM-MM has longest runtimes\\
         Computation of optimal traversal costly};
    \end{tikzpicture}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Adaptive scenario}
\vspace{-2.0cm}

\begin{itemize}
    \item No perfect knowledge of task runtimes and memory usage
    \item Workflow changes: more/less memory, more/less time to execute
\end{itemize}

\vspace{-0.5cm}
 \uncover<2>{ 
 \begin{columns}

    %\column{0.05\textwidth}
    \column{0.69\textwidth}
     \begin{figure}    
        \centering
         %\hskip -0.5cm
        \includegraphics[scale=0.3]{diagrams/images/schedulerAndRUntime.png}
        % \caption{Makespans of HomPart in comparison to HetPart1.}
    \end{figure}
    \column{0.29\textwidth}
   Runtime reports changes

       \begin{itemize}
        \item Can invalidate - not enough memory anymore
        \item Can lead to later finish time

    \end{itemize}

\end{columns}

 }

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
    \frametitle{Adaptive Scenario}

    No recomputation: the majority of schedules becomes invalid!

        \begin{figure}
            \centering
            \includegraphics[scale=0.3]{diagrams/images/MsImprovDynamicCCGRID.png}

        \end{figure}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{Summary}
\framesubtitle{HEFTM-*: memory-valid schedules for each taste!}

\begin{itemize}
    \item Extend an influential predecessor (HEFT)
    \item Memory-aware, communicate over communication buffers
    \item Valid makespans, only 10-35\% worse than invalid HEFT makespans
    \item HEFT-\textbf{MM} proves unique resilience in memory-constrained scenario
\end{itemize}

\pause
    \begin{table}[ht]
        \centering
        \setlength{\tabcolsep}{12pt} % adjust horizontal padding between columns
        \renewcommand{\arraystretch}{1.3} % increase row height for readability
        \begin{tabular}{|p{6.5cm}|p{2.5cm}|}
            \hline
            \textbf{Scenario} & \textbf{Winner} \\
            \hline
            Memory very constrained & HEFT-\textbf{MM} \\
            Memory somewhat constrained, need smallest makespan & HEFTM-\textbf{BL} \\
            Need balance, willing to trade (very little) makespan for less memory requirement & HEFTM-\textbf{BLC} \\
            Lots of memory, need smallest makespan & \only<1>{HEFT?} \only<2> {\sout{HEFT} ~~~~~~~ HEFTM-\textbf{BL}} \\
            \hline
        \end{tabular}
    \end{table}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
    \centering
    \Large{Thank you!}

    \vspace{2cm}

    \small{This work is partially supported by Collaborative Research Center (CRC) 1404 FONDA – Foundations of Workﬂows for Large-Scale Scientiﬁc Data
Analysis, which is funded by German Research Foundation (DFG).}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}% [t]
    \frametitle{Future Work}

TBD

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[t]
    \frametitle{Previous Work}

TBD 
     \begin{columns}
     %\column{0.05\textwidth}
        \column{0.4\textwidth}
        \begin{itemize}
            \item Tree-shaped, then DAG-shaped workflows
            \item Makespan minimization, memory-aware
            \item Mapping only, no adaptive scenarios
        \end{itemize}
        \column{0.4\textwidth}
        \begin{figure}
        \vspace{-1cm}
            \centering
           % \includegraphics[scale=0.15]{diagrams/partitions.png}
            %\caption{Caption }
        \end{figure}
         \column{0.05\textwidth}
    \end{columns}

    ~~~~~~~~~~~~~

    ~~~~~~~~~~~~~~~~~~~~

    \begin{tabular}{|c||c||c|c|}
        \hline
        & Gou et al.~[2] & HetPart1~[3] & HetPart2~[4]  \\
        \hline
        Memory & \cellcolor{pink}homogeneous & \cellcolor{lime} heterogeneous & \cellcolor{lime}heterogeneous\\
        \hline
        Processor speeds & \cellcolor{pink}homogeneous & \cellcolor{pink}homogeneous & \cellcolor{lime}heterogeneous\\
        \hline
    \end{tabular}
   % 
   % \vspace{\fill}

   ~~~~~~~~~~~~~~~~~~~~~~

~~~~~~~~~~~~~~~~~~~~~~~~~~~~

~~~~~~~~~~~~~~~~~~~~

    \footnotesize{[2] Gou, C., Benoit, A., Marchal, L.: Partitioning tree-shaped task graphs for distributed platforms with limited memory. IEEE Trans on Par and DistSystems 31(7), (2020)}

    \footnotesize{[3] Kulagina, S., Meyerhenke, H., Benoit, A.: Mapping Tree-shaped Workflows on Memory-heterogeneous Architectures. HeteroPar 2022}

    \footnotesize{[4] Kulagina, S., Meyerhenke, H., Benoit, A.: Mapping Tree-shaped Workflows on Systems with Different Memory Sizes and Processor Speeds. Wiley's Concurrency and Computation Practice and Experience}

\end{frame}

\end{document}

### 2025-05-16 main.tex
Done on my side
corrected formula of Res and description, committing also the IEEE footnote

@@
 
-\newcommand{\new}[1]{{\color{blue}#1}}
+%\newcommand{\new}[1]{{\color{blue}#1}}
+\newcommand{\new}[1]{{#1}}
 
@@
 
+ \IEEEpubid{\begin{minipage}{\textwidth}\ \\[12pt]
+ © 2025 IEEE. Personal use is permitted, but republication/redistribution requires IEEE permission.
+ See https://www.ieee.org/publications/rights/index.html for more information.
+ The published conference version of this paper appears in the Proceedings of 25th International Symposium on Cluster, Cloud and Internet Computing (CCGrid), IEEE Computer Society.
+\end{minipage}}
+
     \begin{abstract}
@@
 \section{Introduction} %: \skug{Full: 0, Polished: 0}}
-
+    \IEEEpubidadjcol
 %%% CONTEXT %%%
@@
 processors that have become idle after finishing a task.}
+    \IEEEpubidadjcol
 
@@
 
-\smallskip
+\medskip
 \noindent{\bf Compute task ranks. }
@@
 
-\smallskip
+\bigskip
 \noindent  {\bf Task assignment. }
@@
 \noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
-$$Res = availM_j - m_v - \!\!\!\! \sum_{u \in \Pi(v), u\notin T(p_j)}  \!\!\!\!\!\!\!\!\{c_{u,v}\}
-- \sum_{w\in Succ(v)} \!\!\!\!\!\! \{c_{v,w}\}.$$
+$$Res = availM_j - \max{ \{  m_v,  \!\!\!\! \sum_{u \in \Pi(v), u\notin T(p_j)}  \!\!\!\!\!\!\!\!\{c_{u,v}\}
+, \sum_{w\in Succ(v)} \!\!\!\!\!\! \{c_{v,w}\} \} }.$$
 
@@
 already in the memory of~$p_j$. However, the files from the
-other predecessor tasks must be loaded into memory before executing task~$v$,
-as well as the task data of size $m_v$, and the data generated for all successor tasks.
+other predecessor tasks must be loaded into memory before executing task~$v$.
+Depending on what part of $r_u$ was the largest, we need to have enough memory to store
+the task data of size $m_v$, or the data generated for all successor tasks.
 $Res$ then checks whether there is enough memory; if it is negative,
@@
 
+\vspace{-0.09cm}
 \section{Dynamic scenario}
@@
 
+\medskip
 \subsubsection{Workflow instances}
@@
 create larger variants of the workflows above and add the instances created this way to the set.
+
+\medskip
 
@@
 internal composition.}
+
+\medskip
 
@@
 These values align with the findings of Ref.~\cite{lotaru} about small tasks.
+
+\medskip
 
@@
 
+\medskip
+
 \subsubsection{Runtime system}
@@
 
+\medskip
 \subsubsection{Scheduling on the default cluster}
@@
 
+\medskip
 \subsubsection{Scheduling on the memory-constrained cluster}
@@
     \vspace{-0.3cm}
-    \includegraphics[width=1\columnwidth] {images/ms_relations_by_wf_size-constrained-barplot}
+    \includegraphics[width=1.03\columnwidth] {images/ms_relations_by_wf_size-constrained-barplot}
     \caption{Relative makespans on the memory-constrained cluster.
@@
     \centering
-    \includegraphics[width=1\columnwidth] {images/runtimes-logarithmic2}
+  \hspace{-.5cm}  \includegraphics[width=1.01\columnwidth] {images/runtimes-logarithmic2}
     \caption{Average running times (in seconds) of the heuristics \wrt workflow size. The $y$-axis is logarithmic.}

### 2025-05-18 CCGRID Presentation/main.tex
corected Res formula also in the presentation
added pauses

@@
         \item Large applications represented as DAG-shaped workflows
+        \uncover<2,3>{
         \item Heterogeneous execution environment with limited memories
@@
         \item Execute the workflow fast
+         }
     \end{itemize}
@@
     \column{0.39\textwidth}
-    \only<1>{
-     \begin{figure}
-            \centering
-            \includegraphics[scale=0.3]{diagrams/images/CCGrid-muffin.png}
+    \only<2>{
+     \begin{figure}
+            \centering
+            \includegraphics[scale=0.3]{diagrams/images/CCgrid-muffin.png}
 
@@
     }
-    \only<2>{
+    \only<3>{
      \begin{figure}
@@
 
+    \pause
     ~~~~~~~~~~~~~~~~~
@@
         \end{itemize}
+
+        \pause
 
@@
                       \item Communication channel: $rt_{j,j'}$
-                      \item Communication buffer:  $rt_j$
+                      \item Processor:  $rt_j$
                   \end{itemize}  
@@
             \centering
-            \includegraphics[width=.7\textwidth]{diagrams/images/Res-formula}
+            \includegraphics[width=.8\textwidth]{diagrams/images/Res-formula}
         \end{figure}
@@
         \item Start after all communications have finished
+        \pause
         \begin{figure}
@@
     \frametitle{Final Assignment}
-
-    Final assignment to processor
-
-    \begin{itemize}
-        \item Evict files that need to be evicted from memory,
-        \begin{itemize}
-            \item  remove them from pending memory and add to communication buffer
-            \item reduce buffers sizes accordingly
-        \end{itemize}
-        \item New $availM_j$: - eviction, - incoming files, + outgoing files
+    \framesubtitle{Final assignment to processor}
+
+    Previously - no change in the execution environment.
+    \begin{itemize}
+        \item Apply all tentative memory-related changes
+        \item Assign the task
+        \pause
+        \item $availM_j$ after exectution: - eviction, - incoming files, + outgoing files
         \item For each predecessor on another processor $j'$: ready time on communication buffer $rt_{j, j'}$
-
-    \end{itemize}
+    \end{itemize}
+
+    \begin{center}
+        \includegraphics[scale=0.4]{diagrams/images/mem-after}
+    \end{center}
 
@@
          \hskip -0.5cm
-        \includegraphics[scale=0.38]{diagrams/images/success-rates-large2.png}
+        \includegraphics[scale=0.36]{diagrams/images/success-rates-large2.png}
         % \caption{Makespans of HomPart in comparison to HetPart1.}
@@
                 align=center
-            ] at ([xshift=-5cm,yshift=-3.8cm]current page.north east)
+            ] at ([xshift=-5cm,yshift=-4.2cm]current page.north east)
             {HEFTM-* makespans valid, HEFT not \\ HEFTM-BL and BLC : 10 - 35\% worse \\ HEFTM-MM makespans 40\% to 2x worse};
@@
             \centering
-            \hskip -0.5cm
-            \includegraphics[scale=1.25]{diagrams/images/success-rates-tiny.png}
+            \hskip -0.3cm
+            \includegraphics[scale=1.2]{diagrams/images/success-rates-tiny.png}
            % \caption{Makespans of HomPart in comparison to HetPart1.}
@@
             align=center
-        ] at ([xshift=-5.0cm,yshift=-4.5cm]current page.north east)
+        ] at ([xshift=-4.0cm,yshift=-5.5cm]current page.north east)
         {Most makespans invalid, \\
@@
             align=center
-        ] at ([xshift=-4.5cm,yshift=-4.5cm]current page.north east)
+        ] at ([xshift=-4.5cm,yshift=-6.0cm]current page.north east)
         {Runtimes from 1/100 of seconds to minutes\\
@@
 
+    \vspace{-1cm}
+
+    Compare makespans of recomputed schedules vs following the original
+
+    ~~~~
+
     No recomputation: the majority of schedules becomes invalid!
+
+    \pause
 
@@
     \item Valid makespans, only 10-35\% worse than invalid HEFT makespans
-    \item HEFT-\textbf{MM} proves unique resilience in memory-constrained scenario
+    \item HEFTM-\textbf{MM} proves unique resilience in memory-constrained scenario
 \end{itemize}
@@
             \hline
-            Memory very constrained & HEFT-\textbf{MM} \\
+            Memory very constrained & HEFTM-\textbf{MM} \\
             Memory somewhat constrained, need smallest makespan & HEFTM-\textbf{BL} \\

### 2025-05-18 CCGRID Presentation/main.tex
changes

@@
  \begin{columns}
-      \column{0.05\textwidth}
-        \column{0.6\textwidth}
+      \column{0.01\textwidth}
+        \column{0.64\textwidth}
      \begin{itemize}            
@@
              \begin{itemize}
-                \item Execution of 3: $\textcolor{blue}{1000} (c_{1,2})+ \textcolor{blue}{10} (m_3) = 1011$
-                \item Execution of 2: $\textcolor{blue}{1000} (m_2) + \textcolor{blue}{1} (c_{1,2}) = 1001$
+                \item Execution of 3: \\ max ($\textcolor{blue}{1000} (c_{1,2}), \textcolor{blue}{10} (m_3))  + 1 = 1001$
+                \item Execution of 2: \\ max( $\textcolor{blue}{1000} (m_2) , \textcolor{blue}{1} (c_{1,2})) = 1000$
             \end{itemize}
@@
             \begin{itemize}
-                \item Execution of 2: $\textcolor{blue}{1000} (m_2)+ \textcolor{blue}{1} (c_{1,2}) + \textcolor{blue}{1000} (c_{1,3}) = 2001$
+                \item Execution of 2: max($\textcolor{blue}{1000} (m_2), \textcolor{blue}{1} (c_{1,2})) + \textcolor{blue}{1000} (c_{1,3}) = 2000$
             \end{itemize}
@@
     \framesubtitle{Memory-Constrained Cluster: Relative Makespans}
-         \vspace{-0.8cm}
+         \vspace{-0.1cm}
         \begin{figure}
@@
             \hskip -0.8cm
-            \includegraphics[scale=0.48]{diagrams/images/ms_relations_by_wf_size-constrained-barplot.png}
+            \includegraphics[scale=0.35]{diagrams/images/ms_relations_by_wf_size-constrained-barplot.png}
           %  \caption{Runtimes of HetPart2 in comparison to HetPart 1. }
@@
             align=center
-        ] at ([xshift=-4.0cm,yshift=-5.5cm]current page.north east)
+        ] at ([xshift=-4.0cm,yshift=-4.8cm]current page.north east)
         {Most makespans invalid, \\

### 2025-05-21 CCGRID Presentation/main.tex
added 2 pseudo columns on slide 23
KIT affiliation for Svetlana, et al. in footer

@@
 
-\author[S. Kulagina]{\underline{Svetlana Kulagina$^1$}, Anne Benoit$^2$ and Henning Meyerhenke$^{1,3}$}
+\author[S. Kulagina et al.]{\underline{Svetlana Kulagina$^{1,3}$}, Anne Benoit$^2$ and Henning Meyerhenke$^{1,3}$}
 
@@
             {\usebeamerfont{author}\usebeamercolor[fg]{author}\small
-        \underline{Svetlana Kulagina$^1$}, Anne Benoit$^2$ and Henning Meyerhenke$^{1,3}$\par}
+        \underline{Svetlana Kulagina}$^{1,3}$, Anne Benoit$^2$ and Henning Meyerhenke$^{1,3}$\par}
         \vskip1.5em
@@
 
-    %\column{0.05\textwidth}
+    \column{0.05\textwidth}
     \column{0.69\textwidth}
@@
     \end{itemize}
-
+    \column{0.05\textwidth}
 \end{columns}

### 2025-05-21 CCGRID Presentation/main.tex
fixes: 7 workflow -> 7 workflows, HEFT-* valid for all -> HEFT-MM valid for all (slide 20)
changed "worse" to "surplus"

@@
         \begin{itemize}
-            \item HEFTM-\textbf{BL} and HEFTM-\textbf{BLC} produce valid makespans that are only $10$-$35\%$ worse than invalid HEFT makespans
+            \item HEFTM-\textbf{BL} and HEFTM-\textbf{BLC} produce valid makespans with just $10$-$35\%$ surplus over invalid HEFT makespans
             \item HEFTM-\textbf{MM} able to schedule all workflows even under extreme memory constraint
@@
              \item Real-world workflows
-            \item Generated with WFGen [7] from 7 workflow: sizes from 1K to 30K tasks
+            \item Generated with WFGen [7] from 7 workflows: sizes from 1K to 30K tasks
             \item Divided in groups by size
@@
             ] at ([xshift=-5cm,yshift=-4.2cm]current page.north east)
-            {HEFTM-* makespans valid, HEFT not \\ HEFTM-BL and BLC : 10 - 35\% worse \\ HEFTM-MM makespans 40\% to 2x worse};
+            {HEFTM-* makespans valid, HEFT not \\ HEFTM-BL and BLC : 10 - 35\% surplus \\ HEFTM-MM makespans 40\% to 2x surplus};
         \end{tikzpicture}
@@
         {HEFT invalid over 100 tasks \\ HEFTM-BL / HEFTM-BLC invalid over  \\8000 tasks
-        \\ HEFTM-* valid for all workflows };
+        \\ HEFTM-MM valid for all workflows };
     \end{tikzpicture}
@@
     \item Memory-aware, communicate over communication buffers
-    \item Valid makespans, only 10-35\% worse than invalid HEFT makespans
+    \item Valid makespans, only 10-35\% surplus over invalid HEFT makespans
     \item HEFTM-\textbf{MM} proves unique resilience in memory-constrained scenario
@@
         \renewcommand{\arraystretch}{1.3} % increase row height for readability
-        \begin{tabular}{|p{6.5cm}|p{2.5cm}|}
+        \begin{tabular}{|p{6.5cm}|p{2.8cm}|}
             \hline
-            \textbf{Scenario} & \textbf{Winner} \\
+            \textbf{Scenario} & \textbf{Recommendation} \\
             \hline
